---
title: "First article supporting material (after first review)"
description: |
  Crude and age-sex-year adjusted mortality rates and ratios by cause. After first review.
date: last-modified
author: "**"
bibliography: [_lit/refs.bib]
csl: "_lit/american-medical-association.csl" # (Opcional: estilo de citas, p.ej., APA)
toc: true
toc-depth: 5
toc-location: left
toc-float: true
format: 
  html:
    css: [_lib/styles.scss]
    code-fold: true
    embed-resources: true
    theme: 
      - journal #https://quarto.org/docs/output-formats/html-themes-more.html
      #- style.scss
    fig-cap-location: top  
    toc: true
    toc-depth: 5
    toc-location: left
execute:
  error: true        # Show errors in the output
  echo: true         # Show R code in the output
  warning: true      # Show warnings
  message: true      # Show messages from the code
  freeze: auto       # Use cache to speed up rendering if there are no changes
editor: source
execute-dir: project
lightbox: auto
---

```{=html}
<style type="text/css">

.showopt {

background-color: #004c93; color: #FFFFFF; width: 100px; height: 20px; text-align: center; vertical-align: middle !important; float: right; font-family: sans-serif; border-radius: 8px;

}

.showopt:hover {
background-color: #dfe4f2;
color: #004c93;

}

pre.plot {
background-color: white !important;
}

.tablelines table, .tablelines td, .tablelines th {
border: 1px solid black;
}

.centrado {
text-align: center;
}

.table.center {
margin-left:auto;
margin-right:auto;
}

/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/

/* Applies to outputs that are not code other than R*/

pre {
overflow-x: auto !important;
}

pre code {
word-wrap: normal !important;
white-space: pre !important;
}

/*
pre:not(.sourceCode) {
white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence */
font-size: 10px !important;
line-height: 50% !important;
}
body{ /* Normal */
text-align: justify;
}
.superbigimage{
overflow-y:scroll;
height:350px;
white-space: nowrap;
overflow-x: auto;
width:100%;
}
.superbigimage img{
overflow-y: scroll;
overflow-x: hidden;
}
.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}
.pandoc-table { /* Should add !important; but it seems no necessary */
margin-left:auto; /* To center */
margin-right:auto;
border-collapse: collapse;
table-layout: auto;
font-size: 11px;
overflow-y: auto;
max-height:450px !important;
white-space: nowrap;
overflow-x: auto;
width:450px;
}
.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}
.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}
.pandoc-table caption {
text-align: left !important;
font-size: 11px !important;
}

.center-table {
text-align: left !important;
font-size: 9px;
overflow-y:scroll;
height:450px;
overflow-x: scroll;
}

.controlly{
overflow-y:scroll;
height:350px;
overflow-x: scroll;

}
/*2025-03-07*/
.callout-warning,
.callout-error,
.callout-message {
  font-size: 0.7rem !important;  /* Adjust this value as needed */
}

.alert, .message {
    font-size: 0.7em !important;
}

.alert.alert-warning,
.alert.alert-danger,
.message {
    font-size: 0.7em !important;
}
</style>
```
```{css, echo=F}
h1 {
    color: var(--heading-color);
    font-size: 2rem;
    margin-bottom: 1vh;
}

p {
  font-size: 1.1rem;
  line-height: 1.6rem;
}

a {
  color: var(--primary-color);
  text-decoration: none;
  border-bottom: 3px solid transparent;
  font-weight: bold;
  &:hover, &:focus {
      border-bottom: 3px solid currentColor;
  }
}

section {
  margin: 0 auto;
}

.post-meta {
  font-size: 1rem;
  font-style: italic;
  display: block;
  margin-bottom: 4vh;
  color: var(--secondary-color);
}

nav {
  display: flex;
  justify-content: flex-end;
  padding: 20px 0;
}

/*slider switch css */
.theme-switch-wrapper {
  display: flex;
  align-items: center;
  
  em {
    margin-left: 10px;
    font-size: 1rem;
  }
}
.theme-switch {
  display: inline-block;
  height: 34px;
  position: relative;
  width: 60px;
}

.theme-switch input {
  display:none;
}

.slider {
  background-color: #ccc;
  bottom: 0;
  cursor: pointer;
  left: 0;
  position: absolute;
  right: 0;
  top: 0;
  transition: .4s;
}

.slider:before {
  background-color: #fff;
  bottom: 4px;
  content: "";
  height: 26px;
  left: 4px;
  position: absolute;
  transition: .4s;
  width: 26px;
}

input:checked + .slider {
  background-color: #66bb6a;
}

input:checked + .slider:before {
  transform: translateX(26px);
}

.slider.round {
  border-radius: 34px;
}

.slider.round:before {
  border-radius: 50%;
}
```

```{=html}

<script>
function myFunction1() {
var x = document.getElementById("myDIV");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>

<script>
function myFunction2() {
var x = document.getElementById("myDIV2");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
```

<style>
.scrollable-content {
  max-height: 350px;
  overflow-y: auto;
}
pre.scrollable-code {
  max-height: 350px;
  overflow-y: auto;
}

</style>

<br>

# Data Loading and Exploration

## Loading Packages and uniting databases

Proceed to load the necessary packages.

<div class="scrollable-content">
```{r load}
#| message: false
#| include: true
#| warning: false
#| results: "hold"

# --- Bootstrap reticulate con ruta relativa a getwd() ---
suppressPackageStartupMessages(library(reticulate))

# Busca .mamba_root/envs/py311/python.exe desde getwd() hacia padres
find_python_rel <- function(start = getwd(),
                            rel = file.path(".mamba_root","envs","py311","python.exe")) {
  cur <- normalizePath(start, winslash = "/", mustWork = FALSE)
  repeat {
    cand <- normalizePath(file.path(cur, rel), winslash = "/", mustWork = FALSE)
    if (file.exists(cand)) return(cand)
    parent <- dirname(cur)
    if (identical(parent, cur)) return(NA_character_)  # llegó a la raíz
    cur <- parent
  }
}

py <- find_python_rel()

if (is.na(py)) {
  stop("No se encontró Python relativo a getwd() (buscando '.mamba_root/envs/py311/python.exe').\n",
       "Directorio actual: ", getwd())
}

# Forzar ese intérprete
Sys.unsetenv(c("RETICULATE_CONDAENV","RETICULATE_PYTHON_FALLBACK"))
Sys.setenv(RETICULATE_PYTHON = py)
use_python(py, required = TRUE)

py_config()  # verificación

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

#clean enviroment
rm(list = ls()); gc()
file.path(paste0(gsub("/cons","",gsub("cons","",paste0(getwd(),"/cons"))),"data/20241015_out"))

wdpath<-
paste0(gsub("/cons","",gsub("cons","",paste0(getwd(),"/cons"))))
wdpath

envpath<- if(regmatches(wdpath, regexpr("[A-Za-z]+", wdpath))=="G"){"G:/Mi unidad/Alvacast/SISTRAT 2023/"}else{"E:/Mi unidad/Alvacast/SISTRAT 2023/"}
envpath

time_before_dedup2<-Sys.time()

load('G:/My Drive/Alvacast/SISTRAT 2023//data/20241015_out/mort_2025_08_21.Rdata')
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
mexhaz,      # Flexible parametric hazard regression models for survival analysis
tidyverse,   # Collection of packages for data manipulation, visualization, and more (includes dplyr, ggplot2, tidyr, etc.)
janitor,     # Simple tools for examining and cleaning dirty data
tableone,    # Create "Table 1" summaries for descriptive statistics in medical research
cowplot,     # Streamlined plot theme and plot annotations for ggplot2
grid,        # Base R package for low-level graphics functions (used for arranging plots)
rio,         # Simplifies data import/export with a consistent interface
coin,        # Conditional inference procedures for hypothesis testing
kableExtra,  # Enhances 'knitr::kable()' for creating complex tables in R Markdown
epitools,    # Epidemiological tools for data analysis
relsurv,     # Relative survival analysis for population-based cancer studies
survminer,   # Survival analysis and visualization based on 'survival' package
biostat3,    # Biostatistics functions and datasets for teaching and research
tableone,    # (Repeated) Create descriptive summary tables for clinical research
popEpi,      # For SMRs and SIR
metafor,     # For heterogeneity test /Cochrane Q
parallel,    # Parallel computing (for bootstrap)
install = T  # Automatically install packages if not already installed
)

try(rm("HOSP_filter_pl_filt"))

knitr::opts_chunk$set(results = "hold")
```
</div>


## Utility Functions

```{r utility-functions}
# Confidence interval formatting functions
dsr_format <- function(rate, se, phi = 1, factor = 1e3, digits = 2, conf = 0.95) {
z <- qnorm(1 - (1 - conf)/2)
sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
rate*factor,
pmax(0, (rate - z*se*sqrt(phi))*factor),
(rate + z*se*sqrt(phi))*factor)
}

dsr_format_corr <- function(rate, se, phi = 1,
factor = 1e4,       # multiplica la tasa (p. ej. ×100 000)
digits = 2,         # decimales a mostrar
conf   = 0.95) {    # nivel de confianza
z  <- qnorm(1 - (1 - conf) / 2)
se <- se * sqrt(phi)                     # sobredispersión
se_log <- ifelse(rate > 0, se / rate, NA)
L <- rate * exp(-z * se_log)
U <- rate * exp( z * se_log)
sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
rate * factor, L * factor, U * factor)
}

sir_ci_phi_improved <- function(sir_obj, phi, conf.level = 0.95) {
total_obs <- sir_obj$observed
total_exp <- sir_obj$expected
theta <- total_obs / total_exp
z <- qnorm(1 - (1 - conf.level)/2)
phi<- base::pmax(1, phi)
se_log <- sqrt(phi / total_obs)
lci <- theta * exp(-z * se_log)
uci <- theta * exp(z * se_log)
data.frame(
SIR = theta,
CI_low = lci,
CI_high = uci,
phi_used = phi
)
}

extract_phi_by <- function(df, by, rhs_vars) {
by_sym <- rlang::sym(by)
rhs <- base::paste(sprintf("factor(%s)", rhs_vars), collapse = " + ")
fml <- stats::as.formula(base::paste("from0to1 ~", rhs))
df |>
dplyr::group_by(!!by_sym) |>
dplyr::reframe(
phi = {
d <- dplyr::pick(dplyr::everything()) |>
dplyr::filter(pyrs > 0)
if (base::nrow(d) < 3) return(1)  # default conservador
m <- stats::glm(
formula = fml,
offset  = base::log(pyrs),
family  = stats::poisson(),
data    = d
)
dfres <- stats::df.residual(m)
if (dfres <= 0) return(1)
base::sum(stats::residuals(m, type = "pearson")^2) / dfres
}
) |>
dplyr::rename(group = !!by_sym) |>
dplyr::mutate(phi = base::pmax(1, phi))
}
run_dsr_by <- function(lexis_df, weights_df, print_var, digits=2) {
dsr <- popEpi::rate(
data    = lexis_df,
obs     = "from0to1",
pyrs    = "pyrs",
print   = print_var,
adjust  = c("year", "sex", "agegroup"),
weights = weights_df
) |>
dplyr::as_tibble()
phi_tbl <- extract_phi_by(lexis_df, by = print_var, rhs_vars = c("year","sex","agegroup"))
z <- 1.959964
dsr |>
dplyr::rename(group = !!print_var) |>
dplyr::left_join(phi_tbl, by = "group") |>
dplyr::mutate(
phi = dplyr::if_else(is.na(phi) | phi < 1, 1, phi),
rate_1k = rate.adj * 1e3,
sei_log = dplyr::if_else(
rate.adj > 0,
(SE.rate.adj * sqrt(phi)) / rate.adj,
NA_real_
),
rate_1k_L = dplyr::if_else(
rate.adj > 0,
(rate.adj * exp(-z * sei_log)) * 1e3,
NA_real_
),
rate_1k_U = dplyr::if_else(
rate.adj > 0,
(rate.adj * exp(+z * sei_log)) * 1e3,
NA_real_
),
print = dplyr::if_else(
rate.adj > 0,
sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
rate_1k, rate_1k_L, rate_1k_U),
sprintf(paste0("%.", digits, "f (NA–NA)"), rate_1k)
)
)
}
run_dsr_by_agecat <- function(lexis_df, w_year, w_sex, conf.level = 0.95, digits = 2) {
dsr <- popEpi::rate(
data    = lexis_df,
obs     = "from0to1",
pyrs    = "pyrs",
print   = "disch_age_cat",
adjust  = c("year","sex"),
weights = list(year = w_year, sex = w_sex)
) |>
dplyr::as_tibble() |>
dplyr::rename(group = disch_age_cat)
phi_tbl <- extract_phi_by(
lexis_df,
by       = "disch_age_cat",
rhs_vars = c("year","sex")
)
z <- stats::qnorm(1 - (1 - conf.level)/2)
dsr |>
dplyr::left_join(phi_tbl, by = "group") |>
dplyr::mutate(
phi = dplyr::if_else(is.na(phi) | phi < 1, 1, phi),
rate_1k = rate.adj * 1e3,
sei_log = dplyr::if_else(
rate.adj > 0,
(SE.rate.adj * sqrt(phi)) / rate.adj,
NA_real_
),
rate_1k_L = dplyr::if_else(
rate.adj > 0,
(rate.adj * exp(-z * sei_log)) * 1e3,
NA_real_
),
rate_1k_U = dplyr::if_else(
rate.adj > 0,
(rate.adj * exp(+z * sei_log)) * 1e3,
NA_real_
),
print = dplyr::if_else(
rate.adj > 0,
sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
rate_1k, rate_1k_L, rate_1k_U),
sprintf(paste0("%.", digits, "f (NA–NA)"), rate_1k)
)
)
}
run_smr_by <- function(lexis_df, mx_national_clean, print_var) {
popEpi::sir(
coh.data = lexis_df,
coh.obs  = "from0to1",
coh.pyrs = "pyrs",
ref.data = mx_national_clean,
ref.rate = "mx",
print    = print_var,
adjust   = c("year","sex","agegroup")
) |>
dplyr::as_tibble() |>
dplyr::rename(group = !!print_var)
}
sir_ci_phi_profile <- function(sir_tbl, phi, conf.level = 0.95,
sir = "sir", lo = "sir.lo", hi = "sir.hi",
obs = "observed") {
z <- stats::qnorm(1 - (1 - conf.level)/2)
stopifnot(all(c(sir, lo, hi, obs) %in% names(sir_tbl)))
O <- base::as.numeric(sir_tbl[[obs]])
th <- base::as.numeric(sir_tbl[[sir]])
lo0 <- base::as.numeric(sir_tbl[[lo]])
hi0 <- base::as.numeric(sir_tbl[[hi]])
se_u <- (base::log(hi0) - base::log(th)) / z
se_l <- (base::log(th) - base::log(lo0)) / z
se_log <- base::rowMeans(cbind(se_u, se_l), na.rm = TRUE)
phi <- base::as.numeric(phi)
if (length(phi) == 1L) phi <- base::rep(phi, length(th))
phi <- base::pmax(1, phi)
se_log_phi <- se_log * base::sqrt(phi)
se_log_phi[!is.finite(se_log_phi)] <- NA_real_
out <- sir_tbl
out$sir.phi.lo <- base::exp(base::log(th) - z * se_log_phi)
out$sir.phi.hi <- base::exp(base::log(th) + z * se_log_phi)
out
}

assign_cause_cat_vec_corr <- function(diag1, diag2) {
n <- length(diag1)
key3 <- function(x) {
x <- toupper(trimws(as.character(x)))
x[is.na(x) | x == ""] <- NA_character_
substr(x, 1, 3)
}
to_num <- function(k3) {
let <- substr(k3, 1, 1)
num <- suppressWarnings(as.integer(substr(k3, 2, 3)))
valid <- !is.na(let) & !is.na(num) & let %in% LETTERS
out <- rep(NA_integer_, length(k3))
out[valid] <- match(let[valid], LETTERS) * 1000L + num[valid]
out
}
k1 <- key3(diag1)
k2 <- key3(diag2)
n1 <- to_num(k1)
n2 <- to_num(k2)
is_ext1 <- !is.na(n1) & n1 >= 22001 & n1 <= 25098
is_ext2 <- !is.na(n2) & n2 >= 22001 & n2 <= 25098
kext <- ifelse(is_ext1, n1, ifelse(is_ext2, n2, NA_integer_))
out <- rep(NA_character_, n)
out[!is.na(kext) & kext >= 24085 & kext <= 25009] <- "Assaults/ Aggressions / Homicide (X85–Y09)"
out[!is.na(kext) & kext >= 24060 & kext <= 24084] <- "Intentional self-harm (X60–X84)"
out[!is.na(kext) & kext >= 22001 & kext <= 22099] <- "Transport accidents (V01–V99)"
out[!is.na(kext) & kext >= 23000 & kext <= 24059] <- "Other unintentional external causes (W00–X59)"
out[!is.na(kext) & is.na(out)] <- "Other external causes (Y10–Y98)"
na_mask <- is.na(out) & !is.na(n1)
out[na_mask & n1 >= 1000 & n1 <= 2099] <- "Infectious & parasitic (A00–B99)"
out[na_mask & n1 >= 3000 & n1 <= 3096] <- "Malignant neoplasms (C00–C96)"
out[na_mask & n1 >= 5000 & n1 <= 5099] <- "Endocrine & metabolic (E00–E99)"
out[na_mask & n1 >= 6001 & n1 <= 6099] <- "Mental and behavioral (F01–F99)"
out[na_mask & n1 >= 7000 & n1 <= 7099] <- "Nervous system (G00–G99)"
out[na_mask & n1 >= 9000 & n1 <= 9099] <- "Circulatory (I00–I99)"
out[na_mask & n1 >= 10000 & n1 <= 10099] <- "Respiratory (J00–J99)"
out[na_mask & n1 >= 11000 & n1 <= 11093] <- "Digestive (K00–K93)"
out[na_mask & n1 >= 18000 & n1 <= 18099] <- "Symptoms & signs (R00–R99)"
na_mask <- is.na(out) & !is.na(k1)
let1 <- substr(k1[na_mask], 1, 1)
other_letters <- c("D", "H", "L", "M", "N", "O", "P", "Q", "S", "T", "U", "Z")
out[na_mask][let1 %in% other_letters] <- "Other underlying causes (D, H, L, M, N, O, P, Q, S, T, U)"
out[is.na(out)] <- "Unknown/Unclassified"
out
}
cause_levels <- c(
"Infectious & parasitic (A00–B99)",
"Malignant neoplasms (C00–C96)",
"Endocrine & metabolic (E00–E99)",
"Mental and behavioral (F01–F99)",
"Nervous system (G00–G99)",
"Circulatory (I00–I99)",
"Respiratory (J00–J99)",
"Digestive (K00–K93)",
"Symptoms & signs (R00–R99)",
"Other underlying causes (D, H, L, M, N, O, P, Q, S, T, U)",
"Assaults/ Aggressions / Homicide (X85–Y09)",
"Intentional self-harm (X60–X84)",
"Transport accidents (V01–V99)",
"Other unintentional external causes (W00–X59)",
"Other external causes (Y10–Y98)",
"Unknown/Unclassified"
)
flag_external_underlying <- function(cause_cat) {
cc <- as.character(cause_cat)
is_external <- cc %in% c(
"Assaults/ Aggressions / Homicide (X85–Y09)",
"Intentional self-harm (X60–X84)",
"Transport accidents (V01–V99)",
"Other unintentional external causes (W00–X59)",
"Other external causes (Y10–Y98)"
)
dplyr::case_when(
is.na(cc)        ~ NA_character_,
is_external      ~ "External causes",
TRUE             ~ "Underlying causes"
)
}
assign_cause_group2_vec <- function(diag1, diag2) {
cause_cat <- assign_cause_cat_vec_corr(diag1, diag2)
flag_external_underlying(cause_cat)
}
age_to_5y <- function(age_years) 5L * base::floor(age_years / 5)
build_mx_national_by_cause <- function(
deaths_nat_df, pop_df, years_keep, agegroups_keep,
year_col = "ano_def", sex_col = "sexo", age_col = "edad_cant",
diag1_col = "diag1", diag2_col = "diag2",
pop_year_col = "year", pop_agegroup_col = "agegroup",
pop_sex_col = NULL, pop_pop_col = "pop",
age_upper_exclusive = NULL
) {
if (is.null(pop_sex_col)) {
pop_sex_col <- intersect(c("sex_rec","sex"), names(pop_df))[1]
if (is.na(pop_sex_col)) stop("No sex column in pop_df")
}
keep_sorted <- sort(unique(as.integer(agegroups_keep)))
if (is.null(age_upper_exclusive) && length(keep_sorted) >= 2 && median(diff(keep_sorted)) == 5) {
age_upper_exclusive <- max(keep_sorted) + 5L
}
sex_map <- c(`1`="Male",`2`="Female",hombre="Male",male="Male",m="Male",
mujer="Female",female="Female",f="Female")
dt_deaths <- data.table::as.data.table(deaths_nat_df)
dt_pop    <- data.table::as.data.table(pop_df)
dt_deaths <- dt_deaths[get(year_col) %in% years_keep,
.(yr = as.integer(get(year_col)),
sx = tolower(as.character(get(sex_col))),
ag = as.numeric(get(age_col)),
d1 = get(diag1_col),
d2 = get(diag2_col))]
dt_deaths[, sex := sex_map[sx]]
dt_deaths[is.na(sex), sex := NA_character_]
dt_deaths[, sex := factor(sex, levels = c("Male","Female"))]
dt_deaths[, agegroup := keep_sorted[findInterval(ag, keep_sorted, rightmost.closed = TRUE)]]
dt_deaths[findInterval(ag, keep_sorted, rightmost.closed = TRUE) == 0L, agegroup := NA_integer_]
if (!is.null(age_upper_exclusive)) dt_deaths[ag >= age_upper_exclusive, agegroup := NA_integer_]
dt_deaths <- dt_deaths[agegroup %in% keep_sorted & !is.na(sex)]
if (exists("assign_cause_cat_vec_corr", mode = "function")) {
dt_deaths[, cause_cat := assign_cause_cat_vec_corr(d1, d2)]
} else {
dt_deaths[, cause_cat := mapply(assign_cause_cat, d1, d2)]
}
dt_deaths[, cause_cat := factor(cause_cat, levels = cause_levels)]
dt_deaths <- dt_deaths[!is.na(cause_cat)]
deaths_agg <- dt_deaths[, .(deaths = .N), by = .(year = yr, sex, agegroup, cause_cat)]
dt_pop <- dt_pop[get(pop_year_col) %in% years_keep & get(pop_agegroup_col) %in% keep_sorted]
dt_pop[, `:=`(
year     = as.integer(get(pop_year_col)),
agegroup = as.integer(get(pop_agegroup_col)),
sx       = tolower(as.character(get(pop_sex_col))),
pop      = as.numeric(get(pop_pop_col))
)]
dt_pop[, sex := factor(sex_map[sx], levels = c("Male","Female"))]
dt_pop <- dt_pop[!is.na(sex)]
pop_agg <- dt_pop[, .(pop = sum(pop, na.rm = TRUE)), by = .(year, sex, agegroup)]
grid <- data.table::CJ(
year      = unique(pop_agg$year),
sex       = factor(c("Male","Female"), levels = c("Male","Female")),
agegroup  = keep_sorted,
cause_cat = factor(cause_levels, levels = cause_levels)
)
grid <- pop_agg[grid, on = .(year, sex, agegroup)]
grid <- deaths_agg[grid, on = .(year, sex, agegroup, cause_cat)]
grid[is.na(deaths), deaths := 0L]
grid[, mx := deaths / pmax(pop, 1)]
grid[, .(year, sex, agegroup, cause_cat, mx, deaths, pop)]
}
build_cohort_by_cause <- function(clean_df, lexis_df) {
expo <- lexis_df |>
tidytable::summarise(pyrs = sum(pyrs), .by = c(year, sex, agegroup))
years_keep <- sort(unique(expo$year))
ages_keep  <- sort(unique(expo$agegroup))
map_age_to_bins <- function(age_years) {
idx <- findInterval(age_years, ages_keep, rightmost.closed = TRUE)
out <- ages_keep[idx]
out[idx == 0L] <- NA_integer_
out
}
deaths <- clean_df |>
tidytable::filter(status == 1, !is.na(death_date_rec), !is.na(birth_date_rec)) |>
tidytable::mutate(
year = as.integer(format(death_date_rec, "%Y")),
age_years = as.numeric(difftime(death_date_rec, birth_date_rec, units = "days")) / 365.241,
agegroup  = map_age_to_bins(age_years),
sex = factor(stringr::str_to_title(as.character(sex_rec)), levels = c("Male","Female")),
cause_cat = assign_cause_cat_vec_corr(diag1, diag2),
cause_cat = factor(cause_cat, levels = cause_levels)
) |>
tidytable::filter(
year %in% years_keep,
agegroup %in% ages_keep,
!is.na(sex),
!is.na(cause_cat)
) |>
tidytable::summarise(from0to1 = n(), .by = c(year, sex, agegroup, cause_cat))
tidyr::crossing(
expo,
cause_cat = factor(cause_levels, levels = cause_levels)
) |>
tidytable::left_join(deaths, by = c("year","sex","agegroup","cause_cat")) |>
tidytable::mutate(from0to1 = ifelse(is.na(from0to1), 0L, from0to1))
}
extract_phi_by <- function(df,
by,
rhs_vars = c("year","sex","agegroup"),
min_events = 20,
phi_fallback = 1) {
needed <- c(by, rhs_vars, "from0to1", "pyrs")
if (!all(needed %in% names(df))) {
stop("Missing columns: ", paste(setdiff(needed, names(df)), collapse = ", "))
}
rhs <- paste(sprintf("factor(%s)", rhs_vars), collapse = " + ")
fml <- stats::as.formula(paste("from0to1 ~", rhs))
out <- df |>
dplyr::group_by(.data[[by]]) |>
dplyr::summarise(
result = list({
d_all <- dplyr::cur_data_all()
n_events_val <- sum(d_all$from0to1, na.rm = TRUE)
n_rows_val <- nrow(d_all)
d <- d_all |>
dplyr::filter(pyrs > 0) |>
dplyr::mutate(.offset = log(pmax(pyrs, .Machine$double.eps)))
n_rows_used <- nrow(d)
O <- sum(d$from0to1, na.rm = TRUE)
phi_val <- NA_real_
warning_msg <- NA_character_
converged <- FALSE
if (nrow(d) < 3L) {
warning_msg <- sprintf("< 3 rows with pyrs > 0 (only %d)", nrow(d))
} else if (!is.null(min_events) && is.finite(min_events) && O < min_events) {
warning_msg <- sprintf("events (%d) < min_events (%d)", O, min_events)
} else {
res <- tryCatch({
m <- stats::glm(
formula = fml,
family  = stats::poisson(),
data    = d,
offset  = .offset
)
dfres <- stats::df.residual(m)
if (!is.finite(dfres) || dfres <= 0) {
list(phi = NA_real_, converged = FALSE, warning = sprintf("df.residual invalid: %.2f", dfres))
} else {
list(
phi = sum(stats::residuals(m, type = "pearson")^2) / dfres,
converged = TRUE,
warning = NA_character_
)
}
}, error = function(e) {
list(phi = NA_real_, converged = FALSE, warning = substr(e$message, 1, 80))
}, warning = function(w) {
m <- stats::glm(
formula = fml,
family  = stats::poisson(),
data    = d,
offset  = .offset
)
dfres <- stats::df.residual(m)
list(
phi = if (is.finite(dfres) && dfres > 0) sum(stats::residuals(m, type = "pearson")^2) / dfres else NA_real_,
converged = TRUE,
warning = substr(w$message, 1, 80)
)
})
phi_val <- res$phi
converged <- res$converged
warning_msg <- res$warning
}
list(
n_events = n_events_val,
n_rows = n_rows_val,
n_rows_used = n_rows_used,
phi_raw = phi_val,
converged = converged,
warning = warning_msg
)
}),
.groups = "drop"
) |>
tidyr::unnest_wider(result) |>
dplyr::rename(group = 1)
phi_fb <- as.numeric(phi_fallback)[1]
if (!is.finite(phi_fb)) phi_fb <- 1
out |>
dplyr::mutate(
phi = dplyr::if_else(is.finite(phi_raw), pmax(1, phi_raw), phi_fb)
)
}
extract_phi_global <- function(df, rhs_vars = c("year","sex","agegroup")) {
rhs <- base::paste(base::sprintf("factor(%s)", rhs_vars), collapse = " + ")
fml <- stats::as.formula(base::paste("from0to1 ~", rhs))
d <- df |>
dplyr::filter(pyrs > 0)
off <- base::log(base::pmax(d$pyrs, .Machine$double.eps))
m <- stats::glm(fml, family = stats::poisson(), data = d, offset = off)
phi <- base::sum(stats::residuals(m, type = "pearson")^2) / stats::df.residual(m)
base::pmax(1, phi)
}
sir_ci_phi_profile <- function(sir_tbl, phi, conf.level = 0.95,
sir = "sir", lo = "sir.lo", hi = "sir.hi") {
z <- stats::qnorm(1 - (1 - conf.level)/2)
th  <- base::as.numeric(sir_tbl[[sir]])
lo0 <- base::as.numeric(sir_tbl[[lo]])
hi0 <- base::as.numeric(sir_tbl[[hi]])
se_u <- (base::log(hi0) - base::log(th)) / z
se_l <- (base::log(th) - base::log(lo0)) / z
se_log <- base::rowMeans(cbind(se_u, se_l), na.rm = TRUE)
phi <- base::as.numeric(phi)
if (length(phi) == 1L) phi <- base::rep(phi, length(th))
phi <- base::pmax(1, phi)
se_log_phi <- se_log * base::sqrt(phi)
out <- sir_tbl
out$sir.phi.lo <- base::exp(base::log(th) - z * se_log_phi)
out$sir.phi.hi <- base::exp(base::log(th) + z * se_log_phi)
out
}
run_smr_by_cause <- function(coh_cause, mx_national_by_cause, phi_tbl, digits = 2, conf = 0.95) {
smr <- popEpi::sir(
coh.data = coh_cause,
coh.obs  = "from0to1",
coh.pyrs = "pyrs",
ref.data = mx_national_by_cause,
ref.rate = "mx",
print    = "cause_cat",
adjust   = c("year","sex","agegroup","cause_cat")
) |>
dplyr::as_tibble() |>
dplyr::rename(group = cause_cat) |>
dplyr::left_join(phi_tbl, by = "group")
smr_phi <- sir_ci_phi_profile(smr, phi = smr$phi, conf.level = conf)
smr_phi |>
dplyr::mutate(
print = base::sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
sir, sir.phi.lo, sir.phi.hi)
)
}
run_dsr_by_cause <- function(coh_cause, weights_corr, digits = 2, conf = 0.95) {
z <- stats::qnorm(1 - (1 - conf)/2)
dsr <- popEpi::rate(
data    = coh_cause,
obs     = "from0to1",
pyrs    = "pyrs",
print   = "cause_cat",
adjust  = c("year","sex","agegroup"),
weights = weights_corr
) |>
dplyr::as_tibble() |>
dplyr::rename(group = cause_cat)
phi_tbl <- extract_phi_by(
coh_cause,
by       = "cause_cat",
rhs_vars = c("year","sex","agegroup")
)
dsr |>
dplyr::left_join(phi_tbl, by = "group") |>
dplyr::mutate(
phi = dplyr::if_else(is.na(phi) | phi < 1, 1, phi),
rate_1k = rate.adj * 1e3,
sei_log = dplyr::if_else(
rate.adj > 0,
(SE.rate.adj * sqrt(phi)) / rate.adj,
NA_real_
),
rate_1k_L = dplyr::if_else(
rate.adj > 0,
(rate.adj * exp(-z * sei_log)) * 1e3,
NA_real_
),
rate_1k_U = dplyr::if_else(
rate.adj > 0,
(rate.adj * exp(+z * sei_log)) * 1e3,
NA_real_
),
print = dplyr::if_else(
rate.adj > 0,
sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
rate_1k, rate_1k_L, rate_1k_U),
sprintf(paste0("%.", digits, "f (NA–NA)"), rate_1k)
)
)
}


extract_phi_by_sex <- function(df) {
  df |>
    dplyr::group_by(sex) |>
    dplyr::reframe(
      phi = {
        d <- dplyr::pick(dplyr::everything()) |>
          dplyr::filter(pyrs > 0)
        
        if (base::nrow(d) < 3) return(NA_real_)
        
        m <- stats::glm(
          from0to1 ~ factor(agegroup) + factor(year),
          offset = base::log(pyrs),
          family = stats::poisson(),
          data   = d
        )
        
        dfres <- stats::df.residual(m)
        if (dfres <= 0) return(NA_real_)
        
        base::sum(stats::residuals(m, type = "pearson")^2) / dfres
      },
      n_rows    = dplyr::n(),
      n_events  = base::sum(from0to1, na.rm = TRUE),
      pyrs_total = base::sum(pyrs, na.rm = TRUE)
    )
}


se_log_from_ci <- function(est, lo, hi, conf = 0.95) {
  z <- stats::qnorm(1 - (1 - conf)/2)
  
  # proteger contra ceros
  eps <- .Machine$double.eps
  est <- base::pmax(est, eps)
  lo  <- base::pmax(lo,  eps)
  hi  <- base::pmax(hi,  eps)
  
  se_u <- (base::log(hi)  - base::log(est)) / z
  se_l <- (base::log(est) - base::log(lo))  / z
  
  # promedio por elemento (NO mean() global)
  0.5 * (se_u + se_l)
}
se_log_from_rate_se <- function(rate, se_rate, phi = 1) {
  se_rate_phi <- se_rate * base::sqrt(phi)
  se_rate_phi / rate
}
assign_cause_cat_vec_corr <- function(diag1, diag2) {
  n <- length(diag1)
  
  # 1. Cleaner Key Extraction
  key3 <- function(x) {
    x <- toupper(trimws(as.character(x)))
    x[is.na(x) | x == ""] <- NA_character_
    substr(x, 1, 3)
  }
  
  # 2. Vectorized ICD10 to Numeric
  # Maps A->1000 ... U->21000 ... Z->26000 reliably
  to_num <- function(k3) {
    let <- substr(k3, 1, 1)
    num <- suppressWarnings(as.integer(substr(k3, 2, 3)))
    # Ensure valid letter/number combo to avoid NA errors later
    valid <- !is.na(let) & !is.na(num) & let %in% LETTERS
    out <- rep(NA_integer_, length(k3))
    out[valid] <- match(let[valid], LETTERS) * 1000L + num[valid]
    out
  }
  
  k1 <- key3(diag1)
  k2 <- key3(diag2)
  n1 <- to_num(k1)
  n2 <- to_num(k2)
  
  # --- RANGE DEFINITIONS (Numeric) ---
  # V01(22001) - Y98(25098)
  
  # 3. External Cause Logic (Priority)
  # Check if diag1 OR diag2 is an external cause (V-Y)
  is_ext1 <- !is.na(n1) & n1 >= 22001 & n1 <= 25098
  is_ext2 <- !is.na(n2) & n2 >= 22001 & n2 <= 25098
  
  # If diag1 is external, use it. If not, but diag2 is, use diag2.
  kext <- ifelse(is_ext1, n1, ifelse(is_ext2, n2, NA_integer_))
  
  # Initialize Output
  out <- rep(NA_character_, n)
  
  # --- GROUP 1: EXTERNAL CAUSES ---
  # Assault (X85-Y09) -> 24085 - 25009
  out[!is.na(kext) & kext >= 24085 & kext <= 25009] <- "Assaults/ Aggressions / Homicide (X85–Y09)"
  
  # Self-harm (X60-X84) -> 24060 - 24084
  out[!is.na(kext) & kext >= 24060 & kext <= 24084] <- "Intentional self-harm (X60–X84)"
  
  # Transport (V01-V99) -> 22001 - 22099
  out[!is.na(kext) & kext >= 22001 & kext <= 22099] <- "Transport accidents (V01–V99)"
  
  # Other Unintentional (W00-X59) -> 23000 - 24059
  out[!is.na(kext) & kext >= 23000 & kext <= 24059] <- "Other unintentional external causes (W00–X59)"
  
  # Residual External (Any other V, W, X, Y code not caught above)
  out[!is.na(kext) & is.na(out)] <- "Other external causes (Y10–Y98)"
  
  # --- GROUP 2: UNDERLYING CAUSES (diag1) ---
  # Only process if 'out' is still NA (not external)
  
  # Infectious (A00-B99) -> 1000 - 2099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 1000 & n1 <= 2099] <- "Infectious & parasitic (A00–B99)"
  
  # Neoplasms (C00-C96) -> 3000 - 3096 (Strict Malignant)
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 3000 & n1 <= 3096] <- "Malignant neoplasms (C00–C96)"
  
  # Endocrine (E00-E99) -> 5000 - 5099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 5000 & n1 <= 5099] <- "Endocrine & metabolic (E00–E99)"
  
  # Mental (F01-F99) -> 6001 - 6099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 6001 & n1 <= 6099] <- "Mental and behavioral (F01–F99)"
  
  # Nervous (G00-G99) -> 7000 - 7099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 7000 & n1 <= 7099] <- "Nervous system (G00–G99)"
  
  # Circulatory (I00-I99) -> 9000 - 9099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 9000 & n1 <= 9099] <- "Circulatory (I00–I99)"
  
  # Respiratory (J00-J99) -> 10000 - 10099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 10000 & n1 <= 10099] <- "Respiratory (J00–J99)"
  
  # Digestive (K00-K93) -> 11000 - 11093
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 11000 & n1 <= 11093] <- "Digestive (K00–K93)"
  
  # Symptoms (R00-R99) -> 18000 - 18099
  na_mask <- is.na(out) & !is.na(n1)
  out[na_mask & n1 >= 18000 & n1 <= 18099] <- "Symptoms & signs (R00–R99)"
  
  # --- GROUP 3: CATCH-ALL FOR REMAINING ---
  # Corrected to include missing S, T, U, Z codes
  # U = COVID/Special (21000s)
  # S, T = Injury (19000s, 20000s)
  # Z = Health Factors (26000s)
  na_mask <- is.na(out) & !is.na(k1)
  let1 <- substr(k1[na_mask], 1, 1)
  
  # Explicitly added S, T, U, Z here
  other_letters <- c("D", "H", "L", "M", "N", "O", "P", "Q", "S", "T", "U", "Z")
  
  out[na_mask][let1 %in% other_letters] <- "Other underlying causes (D, H, L, M, N, O, P, Q, S, T, U)"
  
  # --- GROUP 4: FINAL RESIDUAL ---
  # If still NA, it means code was invalid or diag1 was missing.
  # Assign a label instead of NA so they aren't dropped from summaries.
  out[is.na(out)] <- "Unknown/Unclassified"
  
  out
}

cause_levels <- c(
  "Infectious & parasitic (A00–B99)",
  "Malignant neoplasms (C00–C96)",
  "Endocrine & metabolic (E00–E99)",
  "Mental and behavioral (F01–F99)",
  "Nervous system (G00–G99)",
  "Circulatory (I00–I99)",
  "Respiratory (J00–J99)",
  "Digestive (K00–K93)",
  "Symptoms & signs (R00–R99)",
  "Other underlying causes (D, H, L, M, N, O, P, Q, S, T, U)", 
  "Assaults/ Aggressions / Homicide (X85–Y09)",
  "Intentional self-harm (X60–X84)",
  "Transport accidents (V01–V99)",
  "Other unintentional external causes (W00–X59)",
  "Other external causes (Y10–Y98)",
  "Unknown/Unclassified"  # <--- ADD THIS LINE
)

# Flag external vs underlying (keeps original 15-category label)
flag_external_underlying <- function(cause_cat) {
  cc <- as.character(cause_cat)
  is_external <- cc %in% c(
    "Assaults/ Aggressions / Homicide (X85–Y09)",
    "Intentional self-harm (X60–X84)",
    "Transport accidents (V01–V99)",
    "Other unintentional external causes (W00–X59)",
    "Other external causes (Y10–Y98)"
  )
  
  dplyr::case_when(
    is.na(cc)        ~ NA_character_,
    is_external      ~ "External causes",
    TRUE             ~ "Underlying causes"
  )
}

# Collapse to 2 cause groups directly from diag1/diag2
assign_cause_group2_vec <- function(diag1, diag2) {
  cause_cat <- assign_cause_cat_vec_corr(diag1, diag2)
  flag_external_underlying(cause_cat)
}

age_to_5y <- function(age_years) 5L * base::floor(age_years / 5)

build_mx_national_by_cause <- function(
    deaths_nat_df, pop_df, years_keep, agegroups_keep,
    year_col = "ano_def", sex_col = "sexo", age_col = "edad_cant",
    diag1_col = "diag1", diag2_col = "diag2",
    pop_year_col = "year", pop_agegroup_col = "agegroup",
    pop_sex_col = NULL, pop_pop_col = "pop",
    age_upper_exclusive = NULL
) {
  # auto-detect pop sex column
  if (is.null(pop_sex_col)) {
    pop_sex_col <- intersect(c("sex_rec","sex"), names(pop_df))[1]
    if (is.na(pop_sex_col)) stop("No sex column in pop_df")
  }
  # precompute age bins
  keep_sorted <- sort(unique(as.integer(agegroups_keep)))
  if (is.null(age_upper_exclusive) && length(keep_sorted) >= 2 && median(diff(keep_sorted)) == 5) {
    age_upper_exclusive <- max(keep_sorted) + 5L
  }
  # sex lookup (vectorized, no recode)
  sex_map <- c(`1`="Male",`2`="Female",hombre="Male",male="Male",m="Male",
               mujer="Female",female="Female",f="Female")
  # convert to data.table ONCE
  dt_deaths <- data.table::as.data.table(deaths_nat_df)
  dt_pop    <- data.table::as.data.table(pop_df)
  # DEATHS: filter early, select minimal cols, compute in-place
  dt_deaths <- dt_deaths[get(year_col) %in% years_keep,
                         .(yr = as.integer(get(year_col)),
                           sx = tolower(as.character(get(sex_col))),
                           ag = as.numeric(get(age_col)),
                           d1 = get(diag1_col),
                           d2 = get(diag2_col))]
  # vectorized sex
  dt_deaths[, sex := sex_map[sx]]
  dt_deaths[is.na(sex), sex := NA_character_]
  dt_deaths[, sex := factor(sex, levels = c("Male","Female"))]
  # vectorized agegroup
  dt_deaths[, agegroup := keep_sorted[findInterval(ag, keep_sorted, rightmost.closed = TRUE)]]
  dt_deaths[findInterval(ag, keep_sorted, rightmost.closed = TRUE) == 0L, agegroup := NA_integer_]
  if (!is.null(age_upper_exclusive)) dt_deaths[ag >= age_upper_exclusive, agegroup := NA_integer_]
  # filter valid rows
  dt_deaths <- dt_deaths[agegroup %in% keep_sorted & !is.na(sex)]
  # cause_cat: vectorized if available, else mapply
  if (exists("assign_cause_cat_vec_corr", mode = "function")) {
    dt_deaths[, cause_cat := assign_cause_cat_vec_corr(d1, d2)]
  } else {
    dt_deaths[, cause_cat := mapply(assign_cause_cat, d1, d2)]
  }
  dt_deaths[, cause_cat := factor(cause_cat, levels = cause_levels)]
  dt_deaths <- dt_deaths[!is.na(cause_cat)]
  # aggregate deaths
  deaths_agg <- dt_deaths[, .(deaths = .N), by = .(year = yr, sex, agegroup, cause_cat)]
  # POP: filter + aggregate
  dt_pop <- dt_pop[get(pop_year_col) %in% years_keep & get(pop_agegroup_col) %in% keep_sorted]
  dt_pop[, `:=`(
    year     = as.integer(get(pop_year_col)),
    agegroup = as.integer(get(pop_agegroup_col)),
    sx       = tolower(as.character(get(pop_sex_col))),
    pop      = as.numeric(get(pop_pop_col))
  )]
  dt_pop[, sex := factor(sex_map[sx], levels = c("Male","Female"))]
  dt_pop <- dt_pop[!is.na(sex)]
  pop_agg <- dt_pop[, .(pop = sum(pop, na.rm = TRUE)), by = .(year, sex, agegroup)]
  # expand grid: pop x cause_levels
  grid <- data.table::CJ(
    year      = unique(pop_agg$year),
    sex       = factor(c("Male","Female"), levels = c("Male","Female")),
    agegroup  = keep_sorted,
    cause_cat = factor(cause_levels, levels = cause_levels)
  )
  # merge pop
  grid <- pop_agg[grid, on = .(year, sex, agegroup)]
  # merge deaths
  grid <- deaths_agg[grid, on = .(year, sex, agegroup, cause_cat)]
  # fill NA deaths with 0, compute rate
  grid[is.na(deaths), deaths := 0L]
  grid[, mx := deaths / pmax(pop, 1)]
  
  grid[, .(year, sex, agegroup, cause_cat, mx, deaths, pop)]
}

build_cohort_by_cause <- function(clean_df, lexis_df) {
  # PY desde Lexis (una vez)
  expo <- lexis_df |>
    tidytable::summarise(pyrs = sum(pyrs), .by = c(year, sex, agegroup))
  
  years_keep <- sort(unique(expo$year))
  ages_keep  <- sort(unique(expo$agegroup))
  
  map_age_to_bins <- function(age_years) {
    idx <- findInterval(age_years, ages_keep, rightmost.closed = TRUE)
    out <- ages_keep[idx]
    out[idx == 0L] <- NA_integer_
    out
  }
  # Deaths by cause (from clean_df; NO from Lexis)
  deaths <- clean_df |>
    tidytable::filter(status == 1, !is.na(death_date_rec), !is.na(birth_date_rec)) |>
    tidytable::mutate(
      year = as.integer(format(death_date_rec, "%Y")),
      age_years = as.numeric(difftime(death_date_rec, birth_date_rec, units = "days")) / 365.241,
      agegroup  = map_age_to_bins(age_years),
      sex = factor(stringr::str_to_title(as.character(sex_rec)), levels = c("Male","Female")),
      cause_cat = assign_cause_cat_vec_corr(diag1, diag2),
      cause_cat = factor(cause_cat, levels = cause_levels)
    ) |>
    tidytable::filter(
      year %in% years_keep,
      agegroup %in% ages_keep,
      !is.na(sex),
      !is.na(cause_cat)
    ) |>
    tidytable::summarise(from0to1 = n(), .by = c(year, sex, agegroup, cause_cat))
  
  tidyr::crossing(
    expo,
    cause_cat = factor(cause_levels, levels = cause_levels)
  ) |>
    tidytable::left_join(deaths, by = c("year","sex","agegroup","cause_cat")) |>
    tidytable::mutate(from0to1 = ifelse(is.na(from0to1), 0L, from0to1))
}
extract_phi_by <- function(df,
                           by,
                           rhs_vars = c("year","sex","agegroup"),
                           min_events = 20,
                           phi_fallback = 1) {
  
  needed <- c(by, rhs_vars, "from0to1", "pyrs")
  if (!all(needed %in% names(df))) {
    stop("Missing columns: ", paste(setdiff(needed, names(df)), collapse = ", "))
  }
  rhs <- paste(sprintf("factor(%s)", rhs_vars), collapse = " + ")
  fml <- stats::as.formula(paste("from0to1 ~", rhs))
  out <- df |>
    dplyr::group_by(.data[[by]]) |>
    dplyr::summarise(
      result = list({
        d_all <- dplyr::cur_data_all()
        n_events_val <- sum(d_all$from0to1, na.rm = TRUE)
        n_rows_val <- nrow(d_all)
        d <- d_all |> 
          dplyr::filter(pyrs > 0) |>
          dplyr::mutate(.offset = log(pmax(pyrs, .Machine$double.eps)))
        n_rows_used <- nrow(d)
        O <- sum(d$from0to1, na.rm = TRUE)
        
        phi_val <- NA_real_
        warning_msg <- NA_character_
        converged <- FALSE
        
        if (nrow(d) < 3L) {
          warning_msg <- sprintf("< 3 rows with pyrs > 0 (only %d)", nrow(d))
        } else if (!is.null(min_events) && is.finite(min_events) && O < min_events) {
          warning_msg <- sprintf("events (%d) < min_events (%d)", O, min_events)
        } else {
          res <- tryCatch({
            m <- stats::glm(
              formula = fml,
              family  = stats::poisson(),
              data    = d,
              offset  = .offset
            )
            dfres <- stats::df.residual(m)
            
            if (!is.finite(dfres) || dfres <= 0) {
              list(phi = NA_real_, converged = FALSE, warning = sprintf("df.residual invalid: %.2f", dfres))
            } else {
              list(
                phi = sum(stats::residuals(m, type = "pearson")^2) / dfres,
                converged = TRUE,
                warning = NA_character_
              )
            }
          }, error = function(e) {
            list(phi = NA_real_, converged = FALSE, warning = substr(e$message, 1, 80))
          }, warning = function(w) {
            # Capture warnings but still return result
            m <- stats::glm(
              formula = fml,
              family  = stats::poisson(),
              data    = d,
              offset  = .offset
            )
            dfres <- stats::df.residual(m)
            list(
              phi = if (is.finite(dfres) && dfres > 0) sum(stats::residuals(m, type = "pearson")^2) / dfres else NA_real_,
              converged = TRUE,
              warning = substr(w$message, 1, 80)
            )
          })
          
          phi_val <- res$phi
          converged <- res$converged
          warning_msg <- res$warning
        }
        
        list(
          n_events = n_events_val,
          n_rows = n_rows_val,
          n_rows_used = n_rows_used,
          phi_raw = phi_val,
          converged = converged,
          warning = warning_msg
        )
      }),
      .groups = "drop"
    ) |>
    tidyr::unnest_wider(result) |>
    dplyr::rename(group = 1)
  
  phi_fb <- as.numeric(phi_fallback)[1]
  if (!is.finite(phi_fb)) phi_fb <- 1
  
  out |>
    dplyr::mutate(
      phi = dplyr::if_else(is.finite(phi_raw), pmax(1, phi_raw), phi_fb)
    )
}
extract_phi_global <- function(df, rhs_vars = c("year","sex","agegroup")) {
  rhs <- base::paste(base::sprintf("factor(%s)", rhs_vars), collapse = " + ")
  fml <- stats::as.formula(base::paste("from0to1 ~", rhs))
  
  d <- df |>
    dplyr::filter(pyrs > 0)
  
  off <- base::log(base::pmax(d$pyrs, .Machine$double.eps))
  
  m <- stats::glm(fml, family = stats::poisson(), data = d, offset = off)
  phi <- base::sum(stats::residuals(m, type = "pearson")^2) / stats::df.residual(m)
  base::pmax(1, phi)
}

sir_ci_phi_profile <- function(sir_tbl, phi, conf.level = 0.95,
                               sir = "sir", lo = "sir.lo", hi = "sir.hi") {
  
  z <- stats::qnorm(1 - (1 - conf.level)/2)
  
  th  <- base::as.numeric(sir_tbl[[sir]])
  lo0 <- base::as.numeric(sir_tbl[[lo]])
  hi0 <- base::as.numeric(sir_tbl[[hi]])
  
  se_u <- (base::log(hi0) - base::log(th)) / z
  se_l <- (base::log(th) - base::log(lo0)) / z
  se_log <- base::rowMeans(cbind(se_u, se_l), na.rm = TRUE)
  
  phi <- base::as.numeric(phi)
  if (length(phi) == 1L) phi <- base::rep(phi, length(th))
  phi <- base::pmax(1, phi)
  
  se_log_phi <- se_log * base::sqrt(phi)
  
  out <- sir_tbl
  out$sir.phi.lo <- base::exp(base::log(th) - z * se_log_phi)
  out$sir.phi.hi <- base::exp(base::log(th) + z * se_log_phi)
  out
}

run_smr_by_cause <- function(coh_cause, mx_national_by_cause, phi_tbl, digits = 2, conf = 0.95) {
  
  smr <- popEpi::sir(
    coh.data = coh_cause,
    coh.obs  = "from0to1",
    coh.pyrs = "pyrs",
    ref.data = mx_national_by_cause,
    ref.rate = "mx",
    print    = "cause_cat",
    adjust   = c("year","sex","agegroup","cause_cat")
  ) |>
    dplyr::as_tibble() |>
    dplyr::rename(group = cause_cat) |>
    dplyr::left_join(phi_tbl, by = "group")
  
  smr_phi <- sir_ci_phi_profile(smr, phi = smr$phi, conf.level = conf)
  
  smr_phi |>
    dplyr::mutate(
      print = base::sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
                            sir, sir.phi.lo, sir.phi.hi)
    )
}
run_dsr_by_cause <- function(coh_cause, weights_corr, digits = 2, conf = 0.95) {
  
  z <- stats::qnorm(1 - (1 - conf)/2)
  
  dsr <- popEpi::rate(
    data    = coh_cause,
    obs     = "from0to1",
    pyrs    = "pyrs",
    print   = "cause_cat",
    adjust  = c("year","sex","agegroup"),
    weights = weights_corr
  ) |>
    dplyr::as_tibble() |>
    dplyr::rename(group = cause_cat)
  
  phi_tbl <- extract_phi_by(
    coh_cause,
    by       = "cause_cat",
    rhs_vars = c("year","sex","agegroup")
  )
  
  dsr |>
    dplyr::left_join(phi_tbl, by = "group") |>
    dplyr::mutate(
      # conservative + stable: if phi missing or <1, use 1
      phi = dplyr::if_else(is.na(phi) | phi < 1, 1, phi),
      
      rate_1k = rate.adj * 1e3,
      
      # don't use pmax(); just compute when rate>0
      sei_log = dplyr::if_else(
        rate.adj > 0,
        (SE.rate.adj * sqrt(phi)) / rate.adj,
        NA_real_
      ),
      
      rate_1k_L = dplyr::if_else(
        rate.adj > 0,
        (rate.adj * exp(-z * sei_log)) * 1e3,
        NA_real_
      ),
      rate_1k_U = dplyr::if_else(
        rate.adj > 0,
        (rate.adj * exp(+z * sei_log)) * 1e3,
        NA_real_
      ),
      
      print = dplyr::if_else(
        rate.adj > 0,
        sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
                rate_1k, rate_1k_L, rate_1k_U),
        sprintf(paste0("%.", digits, "f (NA–NA)"), rate_1k)
      )
    )
}
fmt_interval <- function(x) {
  # "x" is a character like "13.7 (9.5–19.9)" OR numeric vector not expected here
  if (knitr::is_latex_output()) base::gsub("–", "--", x) else x
}
fmt_interval <- function(x) if (knitr::is_latex_output()) base::gsub("–", "--", x) else x
# helper: pick first existing column name from candidates
pick_col <- function(df, candidates) {
  hits <- candidates[candidates %in% base::names(df)]
  if (length(hits) == 0) stop("None of these columns exist: ", base::paste(candidates, collapse=", "))
  hits[1]
}

pairwise_smr_test <- function(smrs, lowers, uppers,
                              alpha = 0.05, a = NULL,
                              ci_level = 0.95,
                              p_adjust = "holm") {
  # allow legacy 'a='
  if (!is.null(a)) alpha <- a
  
  stopifnot(
    length(smrs) == length(lowers),
    length(smrs) == length(uppers),
    ci_level > 0, ci_level < 1
  )
  
  # z corresponding to the CI level (usually 95%)
  z_ci <- qnorm(1 - (1 - ci_level) / 2)
  
  # SE on log scale from log-normal CI
  se_log <- (log(uppers) - log(lowers)) / (2 * z_ci)
  
  n <- length(smrs)
  group_names <- paste("Group", seq_len(n))
  comparisons <- combn(n, 2, simplify = FALSE)
  
  results <- lapply(comparisons, function(pair) {
    i <- pair[1]; j <- pair[2]
    
    diff_log <- log(smrs[i]) - log(smrs[j])
    se_diff  <- sqrt(se_log[i]^2 + se_log[j]^2)
    
    data.frame(
      group1 = group_names[i],
      group2 = group_names[j],
      smr1 = smrs[i],
      smr2 = smrs[j],
      difference = smrs[i] - smrs[j],        # keep your original column
      ratio_1_over_2 = exp(diff_log),        # more appropriate for ratios
      se_diff = se_diff,
      z = abs(diff_log) / se_diff
    )
  }) |> do.call(what = rbind)
  
  results$p_unadj <- 2 * pnorm(-results$z)
  results$p_holm  <- p.adjust(results$p_unadj, method = p_adjust)
  
  results
}
cmr_1k_print <- function(rate, lo, hi, digits = 1) {
  fmt_interval(base::sprintf(
    paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
    rate*1e3, lo*1e3, hi*1e3
  ))
}
prep_dsr <- function(x, section) {
  df <- base::as.data.frame(x)
  df$Section <- section
  df$group   <- base::as.character(df$group)

  keep <- c(
    "Section","group",
    "from0to1","pyrs","phi",
    "rate_1k","rate_1k_L","rate_1k_U","print",
    "rate","rate.lo","rate.hi"
  )
  df <- df[, keep, drop = FALSE]

  # --- format (minimal + auditable)
  df$pyrs <- base::formatC(df$pyrs, format = "f", digits = 1, big.mark = ",")
  df$phi  <- base::round(df$phi, 3)

  # DSR already per 1,000
  df$rate_1k   <- base::round(df$rate_1k, 2)
  df$rate_1k_L <- base::round(df$rate_1k_L, 2)
  df$rate_1k_U <- base::round(df$rate_1k_U, 2)

  # Crude mortality rates (CMR): convert to per 1,000 PY for readability
  df$cmr_1k    <- base::round(df$rate    * 1000, 2)
  df$cmr_1k_L  <- base::round(df$rate.lo * 1000, 2)
  df$cmr_1k_U  <- base::round(df$rate.hi * 1000, 2)

  # LaTeX safety (en-dash)
  if (knitr::is_latex_output()) df$print <- base::gsub("–", "--", df$print)

  out <- df[, c(
    "Section","group","from0to1","pyrs",
    "rate_1k","rate_1k_L","rate_1k_U","print",
    "cmr_1k","cmr_1k_L","cmr_1k_U",
    "phi"
  ), drop = FALSE]

  base::names(out) <- c(
    "Section","Group","Deaths","Person-years",
    "DSR/1,000","DSR L","DSR U","DSR/1,000 (95% CI)",
    "CMR/1,000","CMR L","CMR U",
    "Phi"
  )
  out
}
make_smr_phi <- function(smr_df, dsr_df, digits = 2) {
  smr_df %>%
    dplyr::left_join(dplyr::select(dsr_df, group, phi), by = "group") %>%
    sir_ci_phi_profile(phi = .$phi) %>%
    dplyr::mutate(
      print = base::sprintf(paste0("%.", digits, "f (%.", digits, "f–%.", digits, "f)"),
                            sir, sir.phi.lo, sir.phi.hi)
    )
}
het_dsr <- function(dsr_df) {
  dsr_df |>
    dplyr::mutate(
      phi_used = if ("phi" %in% base::names(dsr_df)) phi else 1,
      yi  = base::log(rate.adj),
      sei = (SE.rate.adj * base::sqrt(phi_used)) / rate.adj
    ) |>
    metafor::rma(yi = yi, sei = sei, method = "FE", data = _)
}

het_smr <- function(smr_phi_df) {
  smr_phi_df |>
    dplyr::mutate(
      yi  = base::log(sir),
      sei = 0.5 * (
        (base::log(sir.phi.hi) - base::log(sir))/1.96 +
        (base::log(sir) - base::log(sir.phi.lo))/1.96
      )
    ) |>
    metafor::rma(yi = yi, sei = sei, method = "FE", data = _)
}
het_row <- function(meta_obj, label) {
  data.frame(
    Analysis = label,
    Q  = base::sprintf("%1.2f", meta_obj$QE),
    df = meta_obj$k - meta_obj$p,
    p  = base::sprintf("%1.3f", meta_obj$QEp),
    I2 = base::sprintf("%1.1f", meta_obj$I2),
    stringsAsFactors = FALSE
  )
}
het_row_strata <- function(meta_obj, contrast, outcome) {
  data.frame(
    Contrast = contrast,
    Outcome  = outcome,
    Q  = base::sprintf("%1.2f", meta_obj$QE),
    df = meta_obj$k - meta_obj$p,
    p  = base::sprintf("%1.3f", meta_obj$QEp),
    I2 = base::sprintf("%1.1f", meta_obj$I2),
    stringsAsFactors = FALSE
  )
}
```




## Data Cleaning

### Missed Admission & Discharge Info

```{r missed-admission-discharge}
cat("Duplicados\n")
SISTRAT23_c1_2010_2022_df_prev1q_sel|>
janitor::get_dupes(hash_key, adm_age_rec2)|>  nrow()

SISTRAT23_c1_2010_2022_df_prev1q_sel|>
(\(df) {
nrow(df)->>before_disc_dup_hash_age_adm_nrow
cat(paste0("1.Number of cases before discarding duplicates in admission age and hash key: ", formatC(nrow(df), big.mark=",")),"\n")
cat(paste0("1.Number of patients before discarding duplicates in admission age and hash key: ", formatC(nrow(distinct(df, hash_key)), big.mark=",")),"\n");
mutate(df, years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |>
filter(is.na(years_in_tr)) ->> first_st_miss_days_in_tr
cat(paste0("Records with unavailable missing days in treatment (eg., currently in treatment): ",nrow(first_st_miss_days_in_tr), "\n"));
mutate(df, years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |>
filter(years_in_tr<0) ->> first_st_neg_days_in_tr;
cat(paste0("Records with negative days in treatment: ",nrow(first_st_neg_days_in_tr), "\n")) ;
mutate(df, years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |>
filter(years_in_tr>3) ->> first_st_more3yrs;
cat(paste0("Records with more than 3 years in treatment: ",nrow(first_st_more3yrs), "\n"))
df
})()|>
mutate(years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year"))|>
filter(!is.na(years_in_tr)|years_in_tr>=0|years_in_tr<=3)|>
group_by(hash_key, adm_age_rec2)|>
slice_max(dit_rec6)|>
ungroup()|>
(\(df) {
nrow(df)->>after_disc_dup_hash_age_adm_nrow
cat(paste0("1.Number of cases after discarding duplicates in admission age and hash key and validating days in treatment: ", formatC(nrow(df), big.mark=",")),"\n")
cat(paste0("1.Number of patients after discarding duplicates in admission age and hash key and validating days in treatment: ", formatC(nrow(distinct(df, hash_key)), big.mark=",")),"\n")
})()

warning("Resultado práctico: con | estás manteniendo todas las filas no-missing,")
warning("independientemente de si years_in_tr está fuera del rango\n")

SISTRAT23_c1_2010_2022_df_prev1q_sel|>
(\(df) {
cat(paste0("1.Number of cases before discarding duplicates in admission age and hash key: ", formatC(nrow(df), big.mark=",")),"\n")
cat(paste0("1.Number of patients before discarding duplicates in admission age and hash key: ", formatC(nrow(distinct(df, hash_key)), big.mark=",")),"\n")
cat(paste0("Records with unavailable missing days in treatment (eg., currently in treatment): ",df |> mutate(years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |>
filter(is.na(years_in_tr)) |> nrow(), "\n"));
cat(paste0("Records with negative days in treatment: ",df |> mutate(years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |>
filter(years_in_tr<0)|>  nrow(), "\n")) ;
cat(paste0("Records with more than 3 years in treatment: ",df |> mutate(years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |>
filter(years_in_tr>3)|> nrow(), "\n"))
df
})()|>
mutate(years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year"))|>
filter(is.na(years_in_tr))|>
nrow()

nrow(SISTRAT23_c1_2010_2022_df_prev1q_sel)-nrow(SISTRAT23_c1_2010_2022_df_prev1q_sel2)

cat("Si dije que descartaba quienes no completaban los ttos.\n")

cat("Discarded rows\n")
before_disc_dup_hash_age_adm_nrow-after_disc_dup_hash_age_adm_nrow
cat("Discarded patients\n")
length(unique(SISTRAT23_c1_2010_2022_df_prev1q_sel$hash_key))-length(unique(SISTRAT23_c1_2010_2022_df_prev1q_sel2$hash_key))
```

## Cohort Construction

```{r cohort-construction}
df0 <- SISTRAT23_c1_2010_2022_df_prev1q_sel2_surv
df1 <- df0 |>
filter(adm_date_rec2 >= "2010-01-01", adm_date_rec2 < "2021-01-01")
df2 <- df1 |>
arrange(hash_key, adm_date_rec2) |>
group_by(hash_key) |>
mutate(n_ttos = n(), tto = row_number(), post_ttos = as.integer(n_ttos > 1)) |>
slice_min(tto) |>
ungroup()
df3 <- df2 |>
filter(adm_age_rec2 >= 18, adm_age_rec2 < 65)

df2 |>
filter(adm_age_rec2 >= 65) |> count(!is.na(death_date))

warning("Muere el 18.5% de los atendidos >64")

audit <- data.frame(
step = c("inicio", "ventana temporal", "primer tratamiento", "edad 18-64"),
cases = c(nrow(df0), nrow(df1), nrow(df2), nrow(df3)),
patients = c(
nrow(distinct(df0, hash_key)),
nrow(distinct(df1, hash_key)),
nrow(distinct(df2, hash_key)),
nrow(distinct(df3, hash_key))
)
)
audit
```

### Exclude Invalid Treatment Compliance

```{r exclude-invalid-compliance}

days_years  <- 365.2425
censor_date <- as.Date("2020-12-31")

df4a_0 <- SISTRAT23_c1_2010_2022_df_prev1q_sel3a_surv
df4a_1 <- df4a_0 |>
  filter(
  !is.na(tr_compliance_rec3),
  tr_compliance_rec3 != "death",
  !grepl("truncated|currently|referral", tr_compliance_rec3)
  )
df4a_2 <- df4a_1 |>
  mutate(disch_age_rec = (dit_rec6 / days_years) + adm_age_rec2)
df4a_3 <- df4a_2 |>
  mutate(
  timesurv = case_when(
  status == 1 ~ lubridate::time_length(interval(adm_date_rec2, death_date), unit = "year"),
  status == 0 ~ lubridate::time_length(interval(adm_date_rec2, censor_date), unit = "year"),
  TRUE ~ NA_real_
  ),
  death_date_rec = case_when(
  status == 1 ~ death_date,
  status == 0 ~ censor_date,
  TRUE ~ as.Date(NA)
  ),
  death_age_rec = as.integer(timesurv + adm_age_rec2)
  )

SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv_rev <- df4a_3

cat("#_#_#_#_#_\nAudit table for 4a\n#_#_#_#_\n")
audit_4a <- data.frame(
step = c(
"start (3a)",
"exclude invalid tr. compliance",
"add discharge age",
"derive survival/death vars"
),
n = c(
nrow(df4a_0),
nrow(df4a_1),
nrow(df4a_2),
nrow(df4a_3)
),
patients = c(
nrow(distinct(df4a_0, hash_key)),
nrow(distinct(df4a_1, hash_key)),
nrow(distinct(df4a_2, hash_key)),
nrow(distinct(df4a_3, hash_key))
)
)
audit_4a
```

### Alternative Processing Path

```{r alternative-processing}
SISTRAT23_c1_2010_2022_df_prev1q_sel2_POST<-SISTRAT23_c1_2010_2022_df_prev1q_sel|>
tidytable::filter(adm_date_rec2>="2010-01-01", adm_date_rec2<"2020-12-31")|>
tidytable::arrange(hash_key, adm_age_rec2)|>
tidytable::filter(adm_date_rec2>="2010-01-01", adm_date_rec2<"2020-12-31")|>
tidytable::group_by(hash_key) |>
tidytable::mutate(tto= tidytable::row_number())|>
tidytable::slice_min(tto)|>
tidytable::ungroup()|>
tidytable::filter(adm_age_rec2>=18, adm_age_rec2<65)|>
tidytable::mutate(post_ttos=ifelse(tto>1, 1, 0))|>
(\(df) {
nrow(df)->>before_disc_dup_hash_age_adm_nrow_POST
cat(paste0("1.Number of cases before discarding duplicates in admission age and hash key: ", formatC(nrow(df), big.mark=",")),"\n")
cat(paste0("1.Number of patients before discarding duplicates in admission age and hash key: ", formatC(nrow(distinct(df, hash_key)), big.mark=",")),"\n")
mutate(df, years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year")) |> filter(is.na(years_in_tr))->>alt_proc_miss_day_tr_curr_in;
cat(paste0("Records with unavailable missing days in treatment (eg., currently in treatment): ", nrow(alt_proc_miss_day_tr_curr_in), "\n"));
mutate(df, years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year"))|> filter(years_in_tr<0) ->> alt_proc_neg_dit
cat(paste0("Records with negative days in treatment: ",nrow(alt_proc_neg_dit), "\n")) ;
mutate(df, years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year"))|> filter(years_in_tr>3) ->> alt_proc_more3yr;
cat(paste0("Records with more than 3 years in treatment: ", nrow(alt_proc_more3yr), "\n"))
df
})()|>
mutate(years_in_tr= time_length(interval(adm_date_rec2, disch_date_rec6), unit="year"))|>
filter(!is.na(years_in_tr)|years_in_tr>=0|years_in_tr<=3)|>
group_by(hash_key, adm_age_rec2)|>
slice_max(dit_rec6)|>
ungroup()|>
(\(df) {
nrow(df)->>after_disc_dup_hash_age_adm_nrow_POST
cat(paste0("1.Number of cases after discarding duplicates in admission age and hash key and validating days in treatment: ", formatC(nrow(df), big.mark=",")),"\n")
cat(paste0("1.Number of patients after discarding duplicates in admission age and hash key and validating days in treatment: ", formatC(nrow(distinct(df, hash_key)), big.mark=",")),"\n")
df
})()

nrow(SISTRAT23_c1_2010_2022_df_prev1q_sel3a_surv)==nrow(SISTRAT23_c1_2010_2022_df_prev1q_sel2_POST)
identical(SISTRAT23_c1_2010_2022_df_prev1q_sel3a_surv,
SISTRAT23_c1_2010_2022_df_prev1q_sel2_POST)

warning("Se queda con la misma cantidad de pacientes, pero no son los mismos episodios\n")

SISTRAT23_c1_2010_2022_df_prev1q_sel2_POST_surv<-
SISTRAT23_c1_2010_2022_df_prev1q_sel2_POST |>
tidytable::left_join(mortality[,c("hashkey", "death_date")], by=c("hash_key"="hashkey"), multiple="first") |>
tidytable::mutate(status=ifelse(is.na(death_date), 0, 1))

cat("\n#_#_#_#_#_\nAhora hacemos el proceso: Discarded ongoing treatments (truncated, death or currently in treatment, o referrals to teratments outside SENDA network).\n#_#_#_#_#_\n")
days_years<- 365.2425
SISTRAT23_c1_2010_2022_df_prev1q_sel4a_POST_surv<-
SISTRAT23_c1_2010_2022_df_prev1q_sel2_POST_surv|>
(\(df) {
nrow(df)->>df4a_nrow_pre_POST
cat(paste0("Discarded (death, no tr. compliance), cases(4a): ", formatC(nrow(tidytable::filter(df, is.na(tr_compliance_rec3) | tr_compliance_rec3=="death"|grepl("truncated|currently|referral", tr_compliance_rec3))), big.mark=",")),"\n")
cat(paste0("Discarded (death, no tr. compliance), patients(4a): ", formatC(nrow(distinct(tidytable::filter(df, is.na(tr_compliance_rec3) | tr_compliance_rec3=="death"|grepl("truncated|currently|referral", tr_compliance_rec3)), hash_key)), big.mark=",")),"\n")
df
})()|>
tidytable::filter(!is.na(tr_compliance_rec3) & tr_compliance_rec3!="death" & !grepl("truncated|currently|referral", tr_compliance_rec3))|>
tidytable::mutate(disch_age_rec= (dit_rec6/365.241)+adm_age_rec2)|>
tidytable::mutate(timesurv= tidytable::case_when(
status==1~ time_length(interval(adm_date_rec2, death_date), unit="year"),
status==0~ time_length(interval(adm_date_rec2, as.Date("2020-12-31")), unit="year")))|>
tidytable::mutate(death_date_rec= tidytable::case_when(
status==1~ death_date,
status==0~ as.Date("2020-12-31")))|>
tidytable::mutate(death_age_rec= as.integer(timesurv+adm_age_rec2))|>
(\(df) {
cat(paste0("4a.Number of cases: ", formatC(nrow(df), big.mark=",")),"\n")
cat(paste0("4a.Number of patients: ", formatC(nrow(distinct(df, hash_key)), big.mark=",")),"\n")
nrow(df) ->>df4a_nrow_post_POST
df
})()
```

### Final Dataset Summary

```{r final-dataset-summary}
cat("Discarded records:\n")
df4a_nrow_pre_POST-df4a_nrow_post_POST
invisible("Ojo, hay más casos que pacientes (+1)\n")

clean_df |>
filter( rn %in% alt_proc_miss_day_tr_curr_in$rn) |> nrow()

invisible("Ninguno persiste de sin días en tto por ttos en curso\n")

clean_df |>
filter( rn %in% alt_proc_neg_dit$rn) |> nrow()
invisible("Ya ninguno persiste de los días de tto negativos\n")

clean_df |>
filter( rn %in% alt_proc_more3yr $rn) |> nrow()
invisible("278 con mpas de 3 años persisten en la base final\n")
```

### Flowchart

```{r mortality-analysis-setup}
neg_tr_d<- table(as.numeric(with(SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv,
difftime(death_date_rec, disch_date_rec6, unit="days"))/ 365.25)<0)[[2]]
over3yrs<-
table(as.numeric(with(SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv,
difftime(disch_date_rec6, adm_date_rec2, unit="days"))/ 365.25)>3)[[2]]

cat("#_#_#_#_#_\ncon las3 anteriores hago este:\n#_#_#_#_")

disch_after_cens_death

SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv |>
dplyr::filter(disch_date_rec6 >= death_date_rec) |>
dplyr::filter(death_date_rec=="2020-12-31", is.na(death_date)) |> nrow()

SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv |>
dplyr::filter(disch_date_rec6 >= death_date_rec) |>
dplyr::filter(!is.na(death_date)) |> nrow()

SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv |>
dplyr::filter(adm_date_rec2 >= disch_date_rec6) |> nrow()

invisible("No me gustó este criterio, pero lo ocupé. Sin querer eliminé 10 muertos\n")

SISTRAT23_c1_2010_2022_df_prev1q_sel4a_surv |>
dplyr::filter(adm_date_rec2 >= disch_date_rec6) |>
dplyr::filter(!is.na(death_date)) |> nrow()

368+3888+150

invisible("Qué pasa si aplicamos el filtro corregido para aprovechar los 0\nDIT en el análisis de sensibilidad\n")

SISTRAT23_c1_2010_2022_df_prev1q_sel4c_surv|>
mutate(disch_date_corr= as.Date(disch_date_corr2))|> 
(\(df) {
cat("Before discarding missing or discharge dates\n")
print(nrow(df))
df
})()  |>
filter(#!is.na(disch_date_corr),
#!is.na(death_date_rec),
#disch_date_corr <= death_date_rec,
adm_date_rec2  >= disch_date_corr) |> nrow()

invisible("En la base de sensibilidad en que imputo con la mediana,\n202 tienen fecha de ingreso posterior o igual a la de egreso, y los elimino\n")

invisible("Pacientes con fechas de ingresos iguales o mayores a la de egreso: 150\n")
invisible("Pacientes en primer tratamiento que terminaron posterior al 2020-12-31, los elimino: 3888.\n")
invisible("Pacientes mueren antes de terminar el primer tto., los elimino: 368\n")

nrow(clean_df)
```


```{r flowchart}

gr<-
  DiagrammeR::grViz("
    digraph flowchart {
      graph [layout = dot, rankdir = TB, nodesep = 0.2, ranksep = 0.3]
    
      # General node styling
      node [fontname = Times, shape = rectangle, fontsize = 17, style = filled, fillcolor = transparent]
    
      # Main flow nodes
      original [label = 'Original Database\\n2010-2022\\n(n = 150,046;\\nPatients = 106,283)', fillcolor = lightgray]
      c1_dataset [label = 'Database\\n(n = 146,012;\\nPatients = 103,612)']
      after_discard [label = 'Database\\n(n = 88,774;\\nPatients = 88,774)']
      after_discard2 [label = 'Database\\n(n = 74,470;\\nPatients = 74,470)']
      final_dataset [label = 'Final Database\\n(n = 70,064;\\nPatients = 70,064)', fillcolor = lightgray]
      
      # Discard nodes (aligned between main flow steps)
      discard_referrals [label = '&#8226;Duplicates in admission age and patient ID kept last (n= 54): 27\\l&#8226;Records with missing days in treatment (e.g., still in treatment\\l  at the dataset extraction date, April 28, 2023): 4,007\\l']
      discard_duplicates [label = '&#8226;Restricted time window of tr. episode (admission\\l  date: 2010-2020) (n=21,434; patients=13,506)\\l&#8226;Kept one episode per patient (younger age) (n=13,506)\\l&#8226;Restricted admission age to 18-64 (n=1,332; patients=1,332)\\l'
      ]
      discard_single [label = '&#8226;Excluded episodes of referrals outside SENDA network: 13,886\\l&#8226;Excluded truncated episodes (not administratively closed): 418\\l'
      ] // &#8226;Discarded (death, no tr. compliance): 14,304
      discard_single2 [label = '&#8226;Discarded patients that finished treatment after 2020 (n=3,888),\\l patients that died before finishing their first treatment (n=368), and \\linconsistent treatment dates (negative or no follow-up periods; n=150): 4,406\\l']

      # Invisible vertices for middle line
      v1 [shape = point, width = 0, style = invis]
      v2 [shape = point, width = 0, style = invis]
      v3 [shape = point, width = 0, style = invis]
      v4 [shape = point, width = 0, style = invis]
    
      # Main flow edges (vertical line)
      original -> v1 [arrowhead = none]
      v1 -> c1_dataset
      c1_dataset -> v2 [arrowhead = none]
      v2 -> after_discard
      after_discard ->v3  [arrowhead = none]
      v3 -> after_discard2
      after_discard2 -> v4 [arrowhead = none]
      v4 -> final_dataset
    
      # Discard connections (from the middle line)
      v1 -> discard_referrals
      v2 -> discard_duplicates
      v3 -> discard_single
      v4 -> discard_single2
    
      # Alignment
      { rank = same; discard_referrals; v1 }
      { rank = same; discard_duplicates; v2 }
      { rank = same; discard_single; v3 }
      { rank = same; discard_single2; v4 }
    }
  ", width = 1100, height = 1550)

base    <- gsub("/cons","", getwd())
fig_dir <- file.path(base, "cons/_figs")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

html_file <- file.path(fig_dir, "_mortality_flowchart_corr.html")
png_file  <- file.path(fig_dir, "_mortality_flowchart_corr.png")

# Guarda el widget
htmlwidgets::saveWidget(gr, html_file, selfcontained = TRUE)

# Captura grande y nítida
webshot::webshot(
  url = html_file,
  file = png_file,
  vwidth = 1100,
  vheight = 1550,
  zoom = 4,      # 1100*4 = 4400 px de ancho aprox
  expand = 20
)

png_file
```

## Corrected main analysis

### 5-Year Bins

```{r sensitivity-5yr-bins}
invisible("Para comparar si hay alguna categoría muy chica en muestra, e inestable\n")

mltper_5x1 <- try(rio::import(paste0(base_path, "mltper_5x1.txt")))
fltper_5x1 <- try(rio::import(paste0(base_path, "fltper_5x1.txt")))
mltper_5x1_filt<-mltper_5x1[between(mltper_5x1$Year,2010,2020),]
fltper_5x1_filt<-fltper_5x1[between(fltper_5x1$Year,2010,2020),]
mltper_5x1_filt$age_rec <- as.numeric(mltper_5x1_filt$Age)
fltper_5x1_filt$age_rec <- as.numeric(fltper_5x1_filt$Age)
mltper_5x1_filt$sex <- "male"
fltper_5x1_filt$sex <- "female"
cons_rate_sex_5x1<-
rbind.data.frame(mltper_5x1_filt, fltper_5x1_filt)[,c("Year","Age","sex", "lx","qx","mx")]
cons_rate_sex_5x1$lambda_p <- -log( 1 - cons_rate_sex_5x1$qx ) / 365.241
cons_rate_sex_5x1<-cons_rate_sex_5x1[which(cons_rate_sex_5x1$Age!="110+"),]

mx_national_raw <- cons_rate_sex_5x1
mx_national_clean <- mx_national_raw |>
  dplyr::mutate(
  agegroup = dplyr::case_when(
  Age == "0"   ~ 0,
  Age == "1-4" ~ 1,
  TRUE         ~ as.numeric(stringr::str_extract(Age, "^[0-9]+"))
  ),
  sex = stringr::str_to_title(sex),
  year = as.integer(Year)
  ) |>
  dplyr::select(year, sex, agegroup, mx) |>
  dplyr::distinct(year, sex, agegroup, .keep_all = TRUE) |>
  dplyr::arrange(year, sex, agegroup)

std_age_breaks <- c(0, 1, seq(5, 85, by = 5))
age_breaks <- sort(unique(mx_national_clean$agegroup))
age_breaks <- age_breaks[age_breaks >= 15]

c_SISTRAT_std_bins <- lexpand(
  clean_df,
  status = status,
  birth = birth_date_rec,
  exit = death_date_rec,
  entry = disch_date_rec6,
  breaks = list(
  per = seq(2010, 2021, by = 1),
  age = age_breaks
  ),
  aggre = list(year = per,
    agegroup = age,
    sex = sex_rec)
)

cat("Agegroups:\n")
sort(unique(c_SISTRAT_std_bins$agegroup))
```

#### SMR with 5-Year Bins

```{r smr-5yr}
sir_final_std <- sir(
  coh.data = c_SISTRAT_std_bins,
  coh.obs = 'from0to1',
  coh.pyrs = 'pyrs',
  ref.data = mx_national_clean,
  ref.rate = 'mx',
  adjust = c('agegroup', 'year', 'sex')
)
print(sir_final_std)

cat("Check differences in age-sex-year strata\n")
keys_coh <- unique(c_SISTRAT_std_bins[, c("year","sex","agegroup")])
keys_ref <- unique(mx_national_clean[, c("year","sex","agegroup")])
stopifnot(
  nrow(dplyr::anti_join(keys_coh, keys_ref, by=c("year","sex","agegroup")))==0
)

sir_ci_phi_improved( sir_final_std, extract_phi(c_SISTRAT_std_bins)) |>
  dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
  SIR,
  CI_low,
  CI_high)) |>
  data.table::data.table()
```

#### DSR with 5-Year Bins

```{r contruct-weights-pop-proj}
proy_ine_com_2010_2020_corr<-
rio::import("https://www.ine.gob.cl/docs/default-source/proyecciones-de-poblacion/cuadros-estadisticos/base-2017/ine_estimaciones-y-proyecciones-2002-2035_base-2017_comunas0381d25bc2224f51b9770a705a434b74.csv?sfvrsn=b6e930a7_3&download=true")|>
  tidyr::pivot_longer(cols = dplyr::starts_with("Poblacion"),
  names_to = "anio",
  values_to = "poblacion")|>
  dplyr::mutate(anio= gsub("Poblacion ","",anio), anio=as.numeric(anio))|>
  dplyr::filter(anio>=2010 & anio<=2020)|>
  dplyr::mutate(agegroup = floor(Edad / 5) * 5) |>
  dplyr::filter(agegroup >= 15 & agegroup <= 75)|>
  dplyr::mutate(sex_rec= ifelse(`Sexo (1=Hombre 2=Mujer)`==2,"female","male"))|>
  group_by(anio, agegroup, sex_rec)|>
  summarise(pop=sum(poblacion, na.rm=T))|>
  ungroup()|>
  rename("year"="anio")

weights_corr <- proy_ine_com_2010_2020_corr |>
  dplyr::group_by(year, sex_rec, agegroup) |>
  dplyr::summarise(pop = sum(pop), .groups = "drop") |>
  dplyr::mutate(weights = pop / sum(pop)) |>
  dplyr::rename(sex = sex_rec) |>
  dplyr::select(year, sex, agegroup, weights)|>
  dplyr::mutate(sex = stringr::str_to_title(sex))

weights_corr <- data.table::as.data.table(weights_corr)[
  , .(year        = as.integer(year),
  agegroup    = as.integer(agegroup),
  sex         = factor(sex, levels = c("Male","Female")),
  weights     = as.numeric(weights))
]

proy_base <- proy_ine_com_2010_2020_corr |>
  dplyr::mutate(
    year = base::as.integer(year),
    agegroup = base::as.numeric(agegroup),
    sex = stringr::str_to_title(sex_rec)   # "Male"/"Female"
  )
w_year <- proy_base |>
  dplyr::summarise(pop = base::sum(pop), .by = year) |>
  dplyr::mutate(w = pop/base::sum(pop)) |>
  {\(d) stats::setNames(d$w, d$year)}()
w_sex <- proy_base |>
  dplyr::summarise(pop = base::sum(pop), .by = sex) |>
  dplyr::mutate(w = pop/base::sum(pop)) |>
  {\(d) stats::setNames(d$w, d$sex)}()


### Constructing weights  ======================

# Named weights vector in the order of levels you pass
make_w_named <- function(df, var, lev) {
  tmp <- df |>
    dplyr::summarise(pop = base::sum(pop), .by = dplyr::all_of(var)) |>
    dplyr::mutate(w = pop / base::sum(pop))
  
  tmp <- tmp[base::match(lev, tmp[[var]]), , drop = FALSE]
  stats::setNames(tmp$w, tmp[[var]])
}
#________________
# Cohort: real levels
#________________
lev_year <- 2010:2020
lev_sex  <- c("Male", "Female")

#________________
# Make weights
#________________
proy_for_w <- proy_base |>
  dplyr::filter(year %in% lev_year)
w_year <- make_w_named(proy_for_w, "year", lev_year)
w_sex  <- make_w_named(proy_for_w, "sex",  lev_sex)

### Constructing weights  ======================
#________________
# Cohort: real levels
#________________
# For 5 -year bins
lev_age_5y  <- base::sort(base::unique(c_SISTRAT_std_bins$agegroup))
lev_year_5y <- base::sort(base::unique(c_SISTRAT_std_bins$year))
proy_5y <- proy_base |>
  dplyr::filter(year %in% lev_year_5y, agegroup %in% lev_age_5y) |>
  dplyr::summarise(pop = base::sum(pop), .by = c(year, agegroup))  # colapsa sex
#________________
# Make weights
#________________
# For 5 -year bins
w_year_5y <- make_w_named(proy_5y, "year", lev_year_5y)
w_age_5y  <- make_w_named(proy_5y, "agegroup", lev_age_5y)
```


```{r dsr-5yr}
dsr_tot <- popEpi::rate(
data    = c_SISTRAT_std_bins,
obs     = "from0to1",
pyrs    = "pyrs",
adjust  = c("year","sex","agegroup"),
weights = weights_corr
)
dsr_tot$rate_1k <- dsr_tot$rate * 1e3

DSR_1k_corr_corr <- mapply(
dsr_format_corr,
dsr_tot$rate.adj,
dsr_tot$SE.rate.adj,
MoreArgs = list(
phi    = max(1,  extract_phi_dir(c_SISTRAT_std_bins)),
factor = 1e3,
digits = 6,
conf   = 0.95))
DSR_1k_corr_corr
```

### By age

```{r setup-age}
# 5y attained-age (15,20,...,75) -> old bins (18,30,45,60)
to_old_agegroup <- function(age5) {
  dplyr::case_when(
    age5 < 30 ~ 18,
    age5 < 45 ~ 30,
    age5 < 60 ~ 45,
    TRUE      ~ 60
  )
}

message("#_#_#_#_#_\n Transform to 15-year bins \n#_#_#_#_#_\n")
c_SISTRAT_std_old <- c_SISTRAT_std_bins |>
  dplyr::mutate(agegroup_old = to_old_agegroup(agegroup)) |>
  dplyr::group_by(year, sex, agegroup_old) |>
  dplyr::summarise(
    pyrs     = base::sum(pyrs),
    from0to1 = base::sum(from0to1),
    .groups  = "drop"
  )
```


```{r dsr-age}
phi_df_age_old <- extract_phi_by_age(c_SISTRAT_std_old |> dplyr::rename(agegroup = agegroup_old))

dsr_age_old <- popEpi::rate(
  data    = c_SISTRAT_std_old,
  obs     = "from0to1",
  pyrs    = "pyrs",
  print   = "agegroup_old",
  adjust  = c("year","sex"),
  weights = list(year = w_year, sex = w_sex)
)

phi_vec_age_old <- base::pmax(1, phi_df_age_old$phi[match(dsr_age_old$agegroup_old, phi_df_age_old$agegroup)])

# dplyr::select(dsr_age_old, rate, rate.lo, rate.hi)|> 
#   mutate_if(is.numeric, ~sprintf("%1.1f",.*1e3))

dsr_age_old<- cbind.data.frame(dplyr::as_tibble(dsr_age_old),phi= phi_vec_age_old) 

DSR_1k_corr_agegr_corr <- mapply(
  dsr_format_corr,
  dsr_age_old$rate.adj,
  dsr_age_old$SE.rate.adj,
  dsr_age_old$phi,
  MoreArgs = list(factor = 1e3, digits = 6, conf = .95)
)

format_interval <- function(s) {
  # Encuentra todas las coincidencias de números decimales
  m <- gregexpr("[0-9]+\\.[0-9]+", s)
  regmatches(s, m) <- lapply(regmatches(s, m), function(nums) {
    formatC(as.numeric(nums), format = "f", digits = 1)
  })
  s
}
# 
# cat("\n#_#_#_#_#_#_#_\nResults of DSRs by age\n#_#_#_#_#_#_#_\n")
# data.table::data.table(as.vector(sapply(DSR_1k_corr_agegr_corr, 
#                                         format_interval)))
```


```{r smr-age}
c_SISTRAT_std_bins_wold <- c_SISTRAT_std_bins |>
  dplyr::mutate(agegroup_old = to_old_agegroup(agegroup))

smr_age_old <- 
popEpi::sir(c_SISTRAT_std_bins_wold, coh.obs = 'from0to1',
                        coh.pyrs = 'pyrs',
                        ref.data = mx_national_clean,
                        ref.rate = mx,
                        print = c("agegroup_old"),
                        adjust = c("year", "sex", "agegroup"), #2025-12-16: the original still adjusts for agegroup
                        test.type = "homogeneity",
                        conf.type = "wald", #conf.type = "wald" usa la aproximación normal de Poisson (la misma lógica que explicamos antes).
                        conf.level = 0.95, EAR = T)
smr_age_old
smr_age_old_tbl <- cbind.data.frame(dplyr::as_tibble(smr_age_old), phi=phi_vec_age_old)

```


### By sex

```{r dsr-sex}
phi_sex_df  <- extract_phi_by_sex(c_SISTRAT_std_bins)

dsr_by_sex <- popEpi::rate(
  data    = c_SISTRAT_std_bins,
  obs     = "from0to1",
  pyrs    = "pyrs",
  print   = "sex",
  adjust  = c("year","agegroup"),
  weights = list(year = w_year_5y, agegroup = w_age_5y)
)

# dplyr::select(dsr_by_sex, rate, rate.lo, rate.hi)|> 
#   mutate_if(is.numeric, ~sprintf("%1.1f",.*1e3))
# #   rate rate.lo rate.hi
# # 1  8.9     8.6     9.3
# # 2  7.0     6.5     7.6

phi_sex_vec <- pmax(1, phi_sex_df$phi[match(dsr_by_sex$sex, phi_sex_df$sex)])
dsr_by_sex <- cbind.data.frame(dplyr::as_tibble(dsr_by_sex),phi= phi_sex_vec) 

DSR_1k_corr_sex_corr <- mapply(
  dsr_format_corr,
  dsr_by_sex $rate.adj,
  dsr_by_sex $SE.rate.adj,
  phi     = pmax(1,phi_sex_vec)[match(dsr_by_sex$sex, phi_sex_df$sex)],
  MoreArgs = list(factor = 1e3, digits = 6, conf = .95)
)
# cat("\n#_#_#_#_#_#_#_\nResults of DSRs by sex\n#_#_#_#_#_#_#_\n")
# data.table::data.table(as.vector(sapply(DSR_1k_corr_sex_corr, 
#                                         format_interval)))
# Male: 13.7 (9.5–19.9)
# Female:   7.9 (6.6–9.5)
```


```{r smr-sex}
smr_sex_corr <- popEpi::sir(
  coh.data = c_SISTRAT_std_bins,
  coh.obs  = "from0to1",
  coh.pyrs = "pyrs",
  ref.data = mx_national_clean,
  ref.rate = "mx",
  print    = "sex",
  adjust   = c("year","sex","agegroup")  # <- 5y internal
)
smr_sex_corr

smr_sex_corr_tbl <- cbind.data.frame(dplyr::as_tibble(smr_sex_corr), phi=phi_sex_vec)

sir_ci_phi_improved( smr_sex_corr, smr_sex_corr_tbl$phi)|> 
  dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)", 
                               SIR, 
                               CI_low, 
                               CI_high)) |> 
  data.table::data.table()
```

```{r summary_sir_dsr_table, results="asis"}

# -------------------------
# SEX (one row per sex)
# -------------------------
sir_sex_dt <- sir_ci_phi_improved(smr_sex_corr, smr_sex_corr_tbl$phi)
sir_sex_print <- fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", sir_sex_dt$SIR, sir_sex_dt$CI_low, sir_sex_dt$CI_high))

sex_sir_map <- data.frame(
    Subgroup = base::as.character(smr_sex_corr_tbl$sex),
    SIR      = sir_sex_print,
    stringsAsFactors = FALSE
)

sex_subcol <- pick_col(dsr_by_sex, c("group","sex"))  # <-- THIS fixes your error

sex_dsr_vec <- fmt_interval(base::as.vector(base::sapply(DSR_1k_corr_sex_corr, format_interval)))

sex_cmr <- fmt_interval(base::paste0(
    base::sprintf("%1.1f", dsr_by_sex$rate*1e3),
    " (", base::sprintf("%1.1f", dsr_by_sex$rate.lo*1e3),
    "–", base::sprintf("%1.1f", dsr_by_sex$rate.hi*1e3), ")"
))

sex_tbl <- data.frame(
    Section = "Sex",
    Subgroup = base::as.character(dsr_by_sex[[sex_subcol]]),
    DSR_1k   = sex_dsr_vec,
    CMR_1k   = sex_cmr,
    Deaths   = dsr_by_sex$from0to1,
    `Person-years` = dsr_by_sex$pyrs,
    stringsAsFactors = FALSE
)

sex_tbl <- base::merge(sex_tbl, sex_sir_map, by = "Subgroup", all.x = TRUE)

# reorder
sex_tbl <- sex_tbl[, c("Section","Subgroup","SIR","DSR_1k","CMR_1k","Deaths","Person.years")]

# -------------------------
# AGE (one row per agegroup_old)
# -------------------------
sir_age_dt <- sir_ci_phi_improved(smr_age_old, smr_age_old_tbl$phi)
sir_age_print <- fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", sir_age_dt$SIR, sir_age_dt$CI_low, sir_age_dt$CI_high))

age_subcol <- pick_col(smr_age_old_tbl, c("agegroup_old","agegroup"))
age_sir_map <- data.frame(
    Subgroup = base::as.character(smr_age_old_tbl[[age_subcol]]),
    SIR      = sir_age_print,
    stringsAsFactors = FALSE
)

age_dsr_vec <- fmt_interval(base::as.vector(base::sapply(DSR_1k_corr_agegr_corr, format_interval)))

age_cmr <- fmt_interval(base::paste0(
    base::sprintf("%1.1f", dsr_age_old$rate*1e3),
    " (", base::sprintf("%1.1f", dsr_age_old$rate.lo*1e3),
    "–", base::sprintf("%1.1f", dsr_age_old$rate.hi*1e3), ")"
))

age_tbl <- data.frame(
    Section = "Age group",
    Subgroup = base::as.character(dsr_age_old$agegroup_old),
    DSR_1k   = age_dsr_vec,
    CMR_1k   = age_cmr,
    Deaths   = dsr_age_old$from0to1,
    `Person-years` = dsr_age_old$pyrs,
    stringsAsFactors = FALSE
)

age_tbl <- base::merge(age_tbl, age_sir_map, by = "Subgroup", all.x = TRUE)
age_tbl <- age_tbl[, c("Section","Subgroup","SIR","DSR_1k","CMR_1k","Deaths","Person.years")]

# -------------------------
# OVERALL (single row)
# -------------------------
sir_overall_dt <- sir_ci_phi_improved(sir_final_std, extract_phi(c_SISTRAT_std_bins))
sir_overall_print <- fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", sir_overall_dt$SIR, sir_overall_dt$CI_low, sir_overall_dt$CI_high))

cmr_overall <- fmt_interval(base::sprintf("%1.1f (%1.1f–%1.1f)",
                                      dsr_tot$rate*1e3, dsr_tot$rate.lo*1e3, dsr_tot$rate.hi*1e3
))
DSR_1k_corr_corr_1dec<-
    mapply(
        dsr_format_corr,                 # FUN
        dsr_tot$rate.adj,           # primer vector (rate)
        dsr_tot$SE.rate.adj,        # segundo vector (se)
        MoreArgs = list(            # argumentos fijos extra
            #phi    = extract_phi_dir(c_SISTRAT_std_bins),
            #opc. conservadora
            phi    = max(1,  extract_phi_dir(c_SISTRAT_std_bins)),
            factor = 1e3,
            digits = 1,
            conf   = 0.95))

overall_tbl <- data.frame(
    Section = "Overall",
    Subgroup = "All",
    SIR = sir_overall_print,
    DSR_1k = fmt_interval(DSR_1k_corr_corr_1dec),
    CMR_1k = cmr_overall,
    Deaths = sir_final_std$observed,
    `Person-years` = sir_final_std$pyrs,
    stringsAsFactors = FALSE
)

# -------------------------
# FINAL TABLE
# -------------------------
final_tbl <- base::rbind(overall_tbl, sex_tbl, age_tbl)

# format PY like your output
final_tbl$`Person.years` <- base::formatC(final_tbl$`Person.years`, format="f", digits=0, big.mark=",")
final_tbl$`Deaths` <- base::formatC(final_tbl$`Deaths`, format="f", digits=0, big.mark=",")

knitr::kable(
    final_tbl,
    caption = "SMR (with overdispersion adjustment) and DSR/CMR per 1,000 person-years.",
    align = c("l","l","l","l","l","r","r")
)
```

### Heterogeneity

```{r het-age-sex}
smr_sex_dat <- sir_ci_phi_improved( smr_sex_corr, smr_sex_corr_tbl$phi) |>
    dplyr::mutate(
        yi  = base::log(SIR),
        sei = se_log_from_ci(SIR, CI_low, CI_high, conf = 0.95)
    )

meta_smr_sex <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = smr_sex_dat
)

smr_age_dat <- sir_ci_phi_improved( smr_age_old, smr_age_old_tbl$phi) |>
    dplyr::mutate(
        yi  = base::log(SIR),
        sei = se_log_from_ci(SIR,CI_low, CI_high, conf = 0.95)
    )

meta_smr_age <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = smr_age_dat
)

dsr_sex_dat <- dsr_by_sex |>
    dplyr::mutate(sex = base::as.character(sex)) |>
    dplyr::mutate(
        yi  = base::log(rate.adj),
        sei = se_log_from_rate_se(rate.adj, SE.rate.adj, phi = phi)
    )

meta_dsr_sex <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = dsr_sex_dat
)

dsr_age_dat <- dsr_age_old |>
    dplyr::mutate(
        yi  = base::log(rate.adj),
        sei = se_log_from_rate_se(rate.adj, SE.rate.adj, phi = phi)
    )

meta_dsr_age <- metafor::rma(yi = yi, sei = sei, method = "FE", data = dsr_age_dat)


het_tbl <- base::rbind(
  het_row(meta_smr_sex, "SMR (SIR) by sex"),
  het_row(meta_smr_age, "SMR (SIR) by age group"),
  het_row(meta_dsr_sex, "DSR by sex"),
  het_row(meta_dsr_age, "DSR by age group")
)

knitr::kable(
  het_tbl,
  caption = "Heterogeneity tests (fixed-effect meta-analysis; Cochran’s Q and I²).",
  align = c("l","r","r","r","r")
)
```


```{r by-age-groups}
dsr_sep <- strcapture("^\\s*([0-9.]+)\\s*\\(\\s*([0-9.]+)\\s*[–-]\\s*([0-9.]+)\\s*\\)\\s*$", 
                      DSR_1k_corr_agegr_corr, 
                      proto = list(estimate = numeric(), 
                                   ci_low = numeric(), 
                                   ci_high = numeric()))

age_labels <- c("18-29", "30-44", "45-59", "60+")

map_groups <- function(df, labels = age_labels) {
  df |>
    dplyr::mutate(
      group1 = ifelse(grepl("^Group", group1), labels[as.integer(sub("Group ", "", group1))], group1),
      group2 = ifelse(grepl("^Group", group2), labels[as.integer(sub("Group ", "", group2))], group2)
    )
}

tab_pairwise <- rbind.data.frame(
  tibble::tibble(
    group1     = "SMR",
    group2     = NA_character_,
    smr1       = NA_real_,
    smr2       = NA_real_,
    difference = NA_real_,
    ratio_1_over_2 = NA_real_,
    se_diff    = NA_real_,
    z          = NA_real_,
    p_unadj    = NA_real_,
    p_holm     = NA_real_
  ),
  pairwise_smr_test(
    smrs   = sir_ci_phi_improved( smr_age_old, smr_age_old_tbl$phi)[,1],
    lowers = sir_ci_phi_improved( smr_age_old, smr_age_old_tbl$phi)[,2],
    uppers = sir_ci_phi_improved( smr_age_old, smr_age_old_tbl$phi)[,3]
  ) |> map_groups(),
  tibble::tibble(
    group1     = "DSR",
    group2     = NA_character_,
    smr1       = NA_real_,
    smr2       = NA_real_,
    difference = NA_real_,
    ratio_1_over_2 = NA_real_,
    se_diff    = NA_real_,
    z          = NA_real_,
    p_unadj    = NA_real_,
    p_holm     = NA_real_
  ),
  pairwise_smr_test(
    smrs   = dsr_sep$estimate,
    lowers = dsr_sep$ci_low,
    uppers = dsr_sep$ci_high
  ) |> map_groups()
) |>
  dplyr::rename(estimate1 = smr1, estimate2 = smr2) |>
  dplyr::mutate(dplyr::across(c(se_diff, z), ~ round(., 2)))

knitr::kable(tab_pairwise, "markdown", caption = "Pairwise comparison, age groups")
```


## By strata


```{r strata}
clean_df2 <- clean_df |>
dplyr::mutate(
setting = dplyr::if_else(res_plan == 1, "Residential", "Ambulatory"),
setting = base::factor(setting, levels = c("Ambulatory","Residential")),
completed = base::as.character(tr_compliance_status),
completed = base::factor(completed, levels = c("Completed","Not completed")),
licit_illicit = base::as.character(prim_sub_licit),
licit_illicit = base::factor(licit_illicit, levels = c("licit","illicit")),
disch_age_cat = base::as.character(disch_age_cat),
disch_age_cat = base::factor(disch_age_cat, levels = c("18-29","30-44","45-59","60+")),
sex = stringr::str_to_title(base::as.character(sex_rec)),
sex = base::factor(sex, levels = c("Male","Female"))
)
c_SISTRAT_5y <- popEpi::lexpand(
data   = clean_df2,
status = status,
birth  = birth_date_rec,
entry  = disch_date_rec6,
exit   = death_date_rec,
breaks = list(
per = base::seq(2010, 2021, by = 1),
age = age_breaks
),
aggre = list(
year    = per,
agegroup= age,
sex     = sex,
setting      = setting,
completed    = completed,
licit_illicit= licit_illicit,
disch_age_cat= disch_age_cat
)
)

warning("disch_age_cat es edad al alta (baseline) → perfecto para “preservar
los grupos antiguos” aunque el Lexis use attained-age 5y internamente para el ajuste.")
dsr_setting   <- run_dsr_by(c_SISTRAT_5y, weights_corr, "setting")
dsr_completed <- run_dsr_by(c_SISTRAT_5y, weights_corr, "completed")
dsr_licit     <- run_dsr_by(c_SISTRAT_5y, weights_corr, "licit_illicit")

```

```{r dsr_one_kable, results="asis"}
tbl_all <- base::rbind(
  prep_dsr(dsr_setting,   "Setting"),
  prep_dsr(dsr_completed, "Completion"),
  prep_dsr(dsr_licit,     "Licit/illicit")#,  prep_dsr(dsr_age,       "Age group")
)

knitr::kable(
  tbl_all,
  caption = "Directly standardized rates (DSR) and crude mortality rates (CMR), per 1,000 person-years, across subgroups.",
  align = c("l","l","r","r","r","r","r","l","r","r","r","r")
)
```


```{r strata-smr}
smr_setting   <- run_smr_by(c_SISTRAT_5y, mx_national_clean, "setting")
smr_completed <- run_smr_by(c_SISTRAT_5y, mx_national_clean, "completed")
smr_licit     <- run_smr_by(c_SISTRAT_5y, mx_national_clean, "licit_illicit")
#smr_agecat    <- run_smr_by(c_SISTRAT_5y, mx_national_clean, "disch_age_cat")
smr_setting_phi <- smr_setting%>%
dplyr::left_join(dplyr::select(dsr_setting, group, phi), by = "group")%>%
sir_ci_phi_profile(phi = .$phi) |>
dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
sir, sir.phi.lo, sir.phi.hi))
smr_completed_phi <- smr_completed%>%
dplyr::left_join(dplyr::select(dsr_completed, group, phi), by = "group")%>%
sir_ci_phi_profile(phi = .$phi) |>
dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
sir, sir.phi.lo, sir.phi.hi))
smr_licit_phi <- smr_licit%>%
dplyr::left_join(dplyr::select(dsr_licit, group, phi), by = "group")%>%
sir_ci_phi_profile(phi = .$phi) |>
dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
sir, sir.phi.lo, sir.phi.hi))
# smr_agecat_phi <- smr_agecat%>%
# dplyr::left_join(subset(run_dsr_by_agecat(c_SISTRAT_5y, w_year, w_sex), select=c("group","phi")),
# by = "group")%>%
# sir_ci_phi_profile(phi = .$phi) |>
# dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
# sir, sir.phi.lo, sir.phi.hi))
```

```{r smr-by-strata}
# Combine all tables into one with a "Category" column
smr_table <- bind_rows(
  smr_setting_phi %>% mutate(Category = "Setting") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print),
  smr_completed_phi %>% mutate(Category = "Treatment Completion") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print),
  smr_licit_phi %>% mutate(Category = "Primary Substance") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print)#,
  # smr_agecat_phi %>% mutate(Category = "Age at Discharge") %>% select(Category, Group = group, Observed = observed, Expected = expected, SMR = print)
) %>%
  mutate(
    Category = ifelse(
      Category != lag(Category, default = ""),
      Category,
      ""
    )
  )

# Create kable
kable(
  smr_table,
  col.names = c("Category", "Group", "Observed", "Expected", "SMR (95% CI)"),
  align = c("l", "l", "r", "r", "r"),
  booktabs = TRUE,
  linesep = ""
) 
```

### Heterogeneity

```{r het-by-strata}
# --- DSR: Treatment Setting ---
meta_dsr_setting <- dsr_setting |>
  dplyr::mutate(
    # Use phi if present, else 1
    phi_used = if("phi" %in% names(dsr_setting)) phi else 1,
    yi  = log(rate.adj),
    sei = (SE.rate.adj * sqrt(phi_used)) / rate.adj
  ) |>
  metafor::rma(yi = yi, sei = sei, method = "FE", data = _)

# --- SMR: Treatment Setting ---
meta_smr_setting <- smr_setting_phi |>
  dplyr::mutate(
    yi  = log(sir),
    # Calculate SE on log scale from the Phi-corrected CIs (averaging upper and lower)
    sei = 0.5 * ((log(sir.phi.hi) - log(sir))/1.96 + (log(sir) - log(sir.phi.lo))/1.96)
  ) |>
  metafor::rma(yi = yi, sei = sei, method = "FE", data = _)

# --- DSR: Compliance ---
meta_dsr_compl <- dsr_completed |>
  dplyr::mutate(
    phi_used = if("phi" %in% names(dsr_completed)) phi else 1,
    yi  = log(rate.adj),
    sei = (SE.rate.adj * sqrt(phi_used)) / rate.adj
  ) |>
  metafor::rma(yi = yi, sei = sei, method = "FE", data = _)

# --- SMR: Compliance ---
meta_smr_compl <- smr_completed_phi |>
  dplyr::mutate(
    yi  = log(sir),
    sei = 0.5 * ((log(sir.phi.hi) - log(sir))/1.96 + (log(sir) - log(sir.phi.lo))/1.96)
  ) |>
  metafor::rma(yi = yi, sei = sei, method = "FE", data = _)

# --- DSR: Licit/Illicit ---
meta_dsr_licit <- dsr_licit |>
  dplyr::mutate(
    phi_used = if("phi" %in% names(dsr_licit)) phi else 1,
    yi  = log(rate.adj),
    sei = (SE.rate.adj * sqrt(phi_used)) / rate.adj
  ) |>
  metafor::rma(yi = yi, sei = sei, method = "FE", data = _)

# --- SMR: Licit/Illicit ---
meta_smr_licit <- smr_licit_phi |>
  dplyr::mutate(
    yi  = log(sir),
    sei = 0.5 * ((log(sir.phi.hi) - log(sir))/1.96 + (log(sir) - log(sir.phi.lo))/1.96)
  ) |>
  metafor::rma(yi = yi, sei = sei, method = "FE", data = _)


het_tbl <- base::rbind(
  het_row_strata(meta_dsr_setting, "Setting",        "DSR"),
  het_row_strata(meta_smr_setting, "Setting",        "SMR"),
  het_row_strata(meta_dsr_compl,   "Compliance",     "DSR"),
  het_row_strata(meta_smr_compl,   "Compliance",     "SMR"),
  het_row_strata(meta_dsr_licit,   "Licit/Illicit",  "DSR"),
  het_row_strata(meta_smr_licit,   "Licit/Illicit",  "SMR")
)

knitr::kable(
  het_tbl,
  caption = "Heterogeneity tests (fixed-effect meta-analysis): Cochran’s Q and I² for subgroup contrasts.",
  align = c("l","l","r","r","r","r")
)
```


## By causes


```{r causes}
years_keep    <- base::sort(base::unique(base::as.integer(c_SISTRAT_std_bins_wold$year)))
agegroups_keep<- base::sort(base::unique(base::as.integer(c_SISTRAT_std_bins_wold$agegroup)))

mx_national_by_cause <- build_mx_national_by_cause(
  deaths_nat_df = mortality_deduplicated,
  pop_df        = proy_ine_com_2010_2020_corr,
  years_keep    = years_keep,
  agegroups_keep= agegroups_keep,
  year_col  = "ano_def",
  sex_col   = "sexo",
  age_col   = "edad_cant",
  diag1_col = "diag1",
  diag2_col = "diag2",
  pop_sex_col = "sex_rec"
)

clean_df3<- left_join(clean_df2,
  dplyr::select(mortality_deduplicated,
  hashkey,
  diag1,
  diag2 ), by=c("hash_key"="hashkey"))
coh_cause <- build_cohort_by_cause(clean_df3, c_SISTRAT_5y)

smr_cause <- popEpi::sir(
  coh.data = coh_cause,
  coh.obs  = "from0to1",
  coh.pyrs = "pyrs",
  ref.data = mx_national_by_cause,
  ref.rate = "mx",
  print    = "cause_cat",
  adjust   = c("year","sex","agegroup","cause_cat")
  ) |>
  dplyr::as_tibble()

warning("In trasnport accidents, in fact were observed 162 and expected 77.5")
mortality_deduplicated |>
  tidytable::mutate(
    icd = stringr::str_to_upper(as.character(diag1)),
    icd = stringr::str_replace_all(icd, "\\.", ""),
    icd3 = stringr::str_extract(icd, "^[A-Z][0-9]{2}"),
    cause_cat = assign_cause_cat_vec_corr(diag1, diag2)
  ) |>
  tidytable::summarise(
  n = .N,
  n_na = sum(is.na(cause_cat)),
  prop_na = mean(is.na(cause_cat))
  )

warning("From 2025-12-30, now there are no NAs")
phi_global <- extract_phi_global(coh_cause, rhs_vars = c("year","sex","agegroup"))
phi_cause <- extract_phi_by(
    coh_cause,
    by = "cause_cat",
    rhs_vars = c("year","sex","agegroup"),
    min_events = 20,
    phi_fallback = phi_global
    )

dsr_cause <- run_dsr_by_cause(coh_cause, weights_corr, digits = 2)
```

```{r summary-by-cause-smr-dsr}
# --- SMR/SIR side (keep only what you want) ---
smr_tbl <- run_smr_by_cause(coh_cause, mx_national_by_cause, phi_cause) |>
  dplyr::mutate(
    obs_exp = base::paste0(observed, "/", base::as.integer(expected)),
    phi     = base::sprintf("%1.2f", phi),
    SMR     = fmt_interval(print)
  ) |>
  dplyr::select(group, obs_exp, phi, SMR)

# --- DSR/CMR side (format + keep) ---

dsr_sep_cause <- strcapture("^\\s*([0-9.]+)\\s*\\(\\s*([0-9.]+)\\s*[–-]\\s*([0-9.]+)\\s*\\)\\s*$", 
                      dsr_cause$print, 
                      proto = list(estimate = numeric(), 
                                   ci_low = numeric(), 
                                   ci_high = numeric()))


dsr_tbl <- cbind.data.frame(dsr_cause, 
    DSR=with(dsr_sep_cause, base::sprintf(
      "%1.1f (%1.1f–%1.1f)",
      estimate, ci_low, ci_high
    ))) |>
  dplyr::mutate(
    DSR = fmt_interval(DSR),
    CMR = fmt_interval(base::sprintf(
      "%1.1f (%1.1f–%1.1f)",
      rate*1e3, rate.lo*1e3, rate.hi*1e3
    ))
  ) |>
  dplyr::select(group, from0to1, DSR, CMR) |>
  dplyr::mutate(from0to1 = base::as.integer(from0to1))  # if it's numeric already, harmless

# --- Join (by group) ---
joint_tbl <- dplyr::left_join(smr_tbl, dsr_tbl, by = "group") |>
  dplyr::transmute(
    Group = base::as.character(group),
    `Obs/Exp` = obs_exp,
    Phi = phi,
    `SMR (95% CI)` = SMR,
    `DSR/1,000 (95% CI)` = DSR,
    `CMR/1,000 (95% CI)` = CMR
  )

knitr::kable(
  joint_tbl,
  caption = "Cause-specific SMR (with phi), DSR and CMR (per 1,000 person-years).",
  align = c("l","l","r","l","r","l","l")
)
```



### By major causes

```{r major-causes-smr-dsr}

external_cats <- c(
"Assaults/ Aggressions / Homicide (X85–Y09)",
"Intentional self-harm (X60–X84)",
"Transport accidents (V01–V99)",
"Other unintentional external causes (W00–X59)",
"Other external causes (Y10–Y98)"
)

clean_df4 <- clean_df2 %>%
left_join(
mortality_deduplicated %>% dplyr::select(hashkey, diag1, diag2),
by = c("hash_key" = "hashkey")
) %>%
mutate(
cause_cat   = assign_cause_cat_vec_corr(diag1, diag2),
is_external = cause_cat %in% external_cats,
status= ifelse(!is.na(death_date),1,0),
death_external   = dplyr::case_when(!is.na(death_date) & is_external ~ 1L, TRUE ~ 0L),
death_underlying = dplyr::case_when(!is.na(death_date) & !is_external ~ 1L, TRUE ~ 0L)
)
age_breaks <- sort(unique(mx_national_clean$agegroup))
age_breaks <- age_breaks[age_breaks >= 15]
c_external <- lexpand(
clean_df4,
status = death_external,
birth  = birth_date_rec,
exit   = death_date_rec,
entry  = disch_date_rec6,
breaks = list(per = seq(2010, 2021, 1),
age = age_breaks),
aggre  = list(year = per, agegroup = age, sex = sex_rec)
)
c_underlying <- lexpand(
clean_df4,
status = death_underlying,
birth  = birth_date_rec,
exit   = death_date_rec,
entry  = disch_date_rec6,
breaks = list(per = seq(2010, 2021, 1),
age = age_breaks),
aggre  = list(year = per, agegroup = age, sex = sex_rec)
)
sum(c_underlying$from0to1)
sum(c_external$from0to1)
1982+1014

ref_mx_agg <- function(mx_df, cats){
mx_df %>%
filter(cause_cat %in% cats) %>%
group_by(year, sex, agegroup) %>%
summarise(
deaths = sum(deaths, na.rm = TRUE),
pop    = dplyr::first(pop),
mx     = deaths / pop,
.groups = "drop"
)
}
ref_external   <- ref_mx_agg(mx_national_by_cause, external_cats)
ref_underlying <- ref_mx_agg(mx_national_by_cause, setdiff(cause_levels, external_cats))
smr_external <- popEpi::sir(
coh.data = c_external, coh.obs = "from0to1", coh.pyrs = "pyrs",
ref.data = ref_external, ref.rate = "mx",
adjust = c("year","sex","agegroup"), EAR = TRUE
)
smr_underlying <- popEpi::sir(
coh.data = c_underlying, coh.obs = "from0to1", coh.pyrs = "pyrs",
ref.data = ref_underlying, ref.rate = "mx",
adjust = c("year","sex","agegroup"), EAR = TRUE
)
dsr_external <- popEpi::rate(
data = c_external, obs = "from0to1", pyrs = "pyrs",
adjust = c("year","sex","agegroup"), weights = weights_corr
)
dsr_underlying <- popEpi::rate(
data = c_underlying, obs = "from0to1", pyrs = "pyrs",
adjust = c("year","sex","agegroup"), weights = weights_corr
)
```

```{r underlying_external_table, results="asis"}
# Profile-based SIR (phi-adjusted) strings
sir_under <- sir_ci_phi_profile(smr_underlying, base::pmax(1, extract_phi_dir(c_underlying)))
sir_ext   <- sir_ci_phi_profile(smr_external,   base::pmax(1, extract_phi_dir(c_external)))

sir_under_print <- fmt_interval(with(sir_under,
  base::sprintf("%.2f (%.2f–%.2f)", sir, sir.phi.lo, sir.phi.hi)
))
sir_ext_print <- fmt_interval(with(sir_ext,
  base::sprintf("%.2f (%.2f–%.2f)", sir, sir.phi.lo, sir.phi.hi)
))

# DSR strings (per 1,000 PY)
dsr_under_print <- fmt_interval(dsr_format_corr(dsr_underlying$rate.adj, dsr_underlying$SE.rate.adj, factor=1e3, digits=1))
dsr_ext_print   <- fmt_interval(dsr_format_corr(dsr_external$rate.adj,   dsr_external$SE.rate.adj,   factor=1e3, digits=1))

# CMR strings (per 1,000 PY)
cmr_under_print <- fmt_interval(with(dsr_underlying, base::sprintf("%.1f (%.1f–%.1f)", rate*1e3, rate.lo*1e3, rate.hi*1e3)))
cmr_ext_print   <- fmt_interval(with(dsr_external,   base::sprintf("%.1f (%.1f–%.1f)", rate*1e3, rate.lo*1e3, rate.hi*1e3)))

tbl <- data.frame(
  Cause_group = c("Underlying causes", "External causes"),
  Phi_dir     = base::round(c(extract_phi_dir(c_underlying), extract_phi_dir(c_external)), 3),
  `Obs/Exp`   = c(
    base::paste0(smr_underlying$observed, "/", base::round(smr_underlying$expected, 0)),
    base::paste0(smr_external$observed,   "/", base::round(smr_external$expected,   0))
  ),
  `SMR (phi-profile 95% CI)` = c(sir_under_print, sir_ext_print),
  `DSR/1,000 (95% CI)`       = c(dsr_under_print, dsr_ext_print),
  `CMR/1,000 (95% CI)`       = c(cmr_under_print, cmr_ext_print),
  stringsAsFactors = FALSE
)

knitr::kable(
  tbl,
  caption = "Summary of mortality metrics for underlying vs external causes (SMR uses profile CI with phi adjustment).",
  align = c("l","r","l","l","l","l"),
  col.names= c("Major cause group", "Phi (raw)", "Obs / Exp", "SMR", "DSR x1000PY", "CMR x1000PY")
)
```

### Prevalence

```{r causes-prev-by-causes}
ref_dist <- mx_national_by_cause |>
dplyr::group_by(cause_cat) |>
dplyr::summarise(
Ref_Deaths = sum(deaths, na.rm = TRUE)
) |>
dplyr::mutate(
Ref_Pct = (Ref_Deaths / sum(Ref_Deaths)) * 100
)
coh_dist <- coh_cause |>
dplyr::group_by(cause_cat) |>
dplyr::summarise(
Study_Deaths = sum(from0to1, na.rm = TRUE)
) |>
dplyr::mutate(
Study_Pct = (Study_Deaths / sum(Study_Deaths)) * 100
)
cause_distribution <- ref_dist |>
dplyr::full_join(coh_dist, by = "cause_cat") |>
dplyr::mutate(
across(c(Ref_Deaths, Study_Deaths), ~tidyr::replace_na(., 0)),
across(c(Ref_Pct, Study_Pct), ~tidyr::replace_na(., 0))
) |>
dplyr::transmute(
Category = cause_cat,
`National Deaths (n)` = formatC(Ref_Deaths, format="d", big.mark=","),
`National (%)` = sprintf("%.1f%%", Ref_Pct),
`Cohort Deaths (n)` = formatC(Study_Deaths, format="d", big.mark=","),
`Cohort (%)` = sprintf("%.1f%%", Study_Pct)
)

knitr::kable(cause_distribution, caption= "Prevalence of ICD-10 cause groups")
```


```{r causes-major-causes-prev}
ref_broad_dist <- mx_national_by_cause |>
dplyr::mutate(
Broad_Group = dplyr::if_else(cause_cat %in% external_cats, "External causes", "Underlying causes")
) |>
dplyr::group_by(Broad_Group) |>
dplyr::summarise(
Ref_Deaths = sum(deaths, na.rm = TRUE)
) |>
dplyr::mutate(
Ref_Pct = (Ref_Deaths / sum(Ref_Deaths)) * 100
)
coh_broad_dist <- tibble::tibble(
Broad_Group = c("External causes", "Underlying causes"),
Study_Deaths = c(1014, 1982)
) |>
dplyr::mutate(
Study_Pct = (Study_Deaths / sum(Study_Deaths)) * 100
)
broad_comparison <- ref_broad_dist |>
dplyr::left_join(coh_broad_dist, by = "Broad_Group") |>
dplyr::transmute(
Group = Broad_Group,
`National Deaths (n)` = formatC(Ref_Deaths, format="d", big.mark=","),
`National (%)` = sprintf("%.1f%%", Ref_Pct),
`Cohort Deaths (n)` = formatC(Study_Deaths, format="d", big.mark=","),
`Cohort (%)` = sprintf("%.1f%%", Study_Pct)
)

knitr::kable(broad_comparison, caption= "Prevalence of major cause groups")
```


```{r causes-pop-dsr}
pop_std <- mx_national_by_cause %>%
dplyr::distinct(year, sex, agegroup, pop)
weights_std <- pop_std %>%
dplyr::group_by(sex, agegroup) %>%
dplyr::summarise(pop = sum(pop), .groups = "drop") %>%
dplyr::mutate(weight = pop / sum(pop)) %>%
dplyr::select(sex, agegroup, weight)

dsr_year_cause <- mx_national_by_cause %>%
dplyr::left_join(weights_std, by = c("sex","agegroup")) %>%
dplyr::group_by(year, cause_cat) %>%
dplyr::summarise(dsr = sum(weight * mx, na.rm = TRUE), .groups = "drop")

mx_pooled <- mx_national_by_cause %>%
dplyr::group_by(sex, agegroup, cause_cat) %>%
dplyr::summarise(deaths = sum(deaths), pop = sum(pop), .groups = "drop") %>%
dplyr::mutate(mx = deaths / pop)

dsr_pooled_cause <- mx_pooled %>%
dplyr::left_join(weights_std, by = c("sex","agegroup")) %>%
dplyr::group_by(cause_cat) %>%
dplyr::summarise(dsr = sum(weight * mx, na.rm = TRUE), .groups = "drop")

dsr_pooled_cause$dsr_print <- sprintf("%1.2f",dsr_pooled_cause$dsr*1e3)
cat("Correct Population DSR
")

knitr::kable(subset(dsr_pooled_cause, select= c("cause_cat", "dsr_print") ), caption= "DSRs at the national level by causes (per 1,000 PY)")
```

## Plot


```{r plot}

c_SISTRAT_std_bins_fot <- lexpand(
clean_df2,
status = status,
birth = birth_date_rec,
exit = death_date_rec,
entry = disch_date_rec6,
breaks = list(
per = seq(2010, 2021, by = 1),
age = age_breaks,
fot = c(0, .0386,.2465, .5, 1, 3, 5, 7, 9, Inf)
),
aggre = list(agegroup = age, year = per, sex = sex_rec, fot= fot)
)
sr_1_std_bins_sex_fot <- popEpi::sir(c_SISTRAT_std_bins_fot, coh.obs = 'from0to1',
                                     coh.pyrs = 'pyrs',
                                     ref.data = mx_national_clean,
                                     ref.rate = "mx",
                                     print = c("sex", "fot"),
                                     adjust = c("agegroup", "sex", "year"),
                                     test.type = "homogeneity",
                                     conf.type = "wald",
                                     conf.level = 0.95, EAR = T)
r2_std_bins_sex_fot <- popEpi::rate( data = c_SISTRAT_std_bins_fot,
                                     obs = from0to1,
                                     pyrs = pyrs,
                                     print = c("sex","fot"),
                                     adjust = c("year", "agegroup"),
                                     weights =  list(year = w_year_5y, agegroup = w_age_5y),
)
rates_std_bins_sex_df_fot <- r2_std_bins_sex_fot %>%
  mutate(
    rate_adj_1k     = rate.adj     * 1e3,
    rate_adj_lo_1k  = rate.adj.lo  * 1e3,
    rate_adj_hi_1k  = rate.adj.hi  * 1e3
  )
p_std_bins_rate2 <- ggplot(rates_std_bins_sex_df_fot, aes(x = fot, y = rate_adj_1k, fill=sex)) +
  geom_ribbon(aes(ymin = rate_adj_lo_1k, ymax = rate_adj_hi_1k, fill=sex),
              alpha = .20) +
  geom_line(aes(color=sex), linewidth = .9) +
  geom_point(size = 2,aes(color=sex, shape=sex)) +
  scale_x_continuous("Years since discharge",
                     breaks = setdiff(round(rates_df_fot$fot,2),c(0.04, 0.25))) +
  scale_y_continuous("Adjusted rate
(deaths ×1,000 PY)",
                     limits = c(0, NA)) +
  theme_minimal(base_family = "serif")+
  theme(axis.title.x = element_blank())+
  scale_colour_manual(
    values = c(Male = "#2C3E8B", Female = "#E69F00")
  ) +
  scale_shape_manual(values = c(Male = 16, Female = 17))+
  scale_fill_manual(
    values = c(Male = alpha("#2C3E8B", 0.25),
               Female = alpha("#E69F00", 0.25))
  )+
  guides(
    fill = guide_legend(title = "Sex"),
    color = guide_legend(title = "Sex"),
    shape = guide_legend(title = "Sex")
  )
p_std_bins_sir2 <- ggplot(sr_1_std_bins_sex_fot, aes(x = fot, y = sir, fill= sex)) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey50") +
  geom_ribbon(aes(ymin = sir.lo, ymax = sir.hi, fill=sex),
              alpha = .20) +
  geom_line(aes(color=sex), linewidth = .9) +
  geom_point(aes(color=sex, shape=sex),size = 2) +
  scale_x_continuous("Years since discharge",
                     breaks = setdiff(round(rates_df_fot$fot,2),c(0.04, 0.25))) +
  theme_minimal(base_family = "serif")+
  theme(axis.title.x = element_blank())+
  scale_y_log10(
    "Adjusted SMR",
    breaks = setdiff(round(rates_df_fot$fot,2),c(0.04, 0.25)),
    labels = scales::number_format(accuracy = 0.1)
  ) +
  scale_colour_manual(
    values = c(Male = "#2C3E8B", Female = "#E69F00")
  ) +
  scale_shape_manual(values = c(Male = 16, Female = 17))+
  scale_fill_manual(
    values = c(Male = alpha("#2C3E8B", 0.25),
               Female = alpha("#E69F00", 0.25))
  )+ theme(legend.position="none")+
  guides(
    fill = guide_legend(title = "Sex"),
    color = guide_legend(title = "Sex"),
    shape = guide_legend(title = "Sex")
  )
legend_shared_std_bins <- ggpubr::get_legend(
  p_std_bins_rate2 +
    theme(legend.position = "bottom") +
    guides(
      fill = guide_legend(title = "Sex"),
      color = guide_legend(title = "Sex"),
      shape = guide_legend(title = "Sex")
    )
)
panels_std_bins <- cowplot::plot_grid(
  p_std_bins_rate2+ theme(legend.position="none"), p_std_bins_sir2+ theme(legend.position="none"),
  ncol             = 1,
  labels           = c("a", "b"),
  label_size       = 14,
  label_fontfamily = "serif",
  align            = "v",
  axis             = "l",
  label_x          = 0,
  label_y          = 1,
  hjust            = -0.1,
  vjust            = 1.2
)
xlab_shared_std_bins <- cowplot::ggdraw() +
  cowplot::draw_label("Years since discharge",
                      fontfamily = "serif",
                      fontface = "plain", size = 12, hjust = 0.5)
final_std_bins_plot <- cowplot::plot_grid(
  panels_std_bins,
  xlab_shared_std_bins,
  legend_shared_std_bins,
  ncol = 1,
  rel_heights = c(1, 0.06, 0.10)
)
print(final_std_bins_plot)
figexp<- 1.5
deinflar_word <- 1/1.07653631284916
ggsave(
  paste0(gsub("/cons","",getwd()), "/cons/_figs/Figure_1_rates_and_SIR_by_fot_updated_postrev.pdf"),
  dpi = 600,
  width = 80 *figexp,
  height = 80 * figexp* (7/6),
  units = "mm",
  device = cairo_pdf
)
ggsave(
  paste0(gsub("/cons","",getwd()), "/cons/_figs/Figure_1_rates_and_SIR_by_fot_updated_postrev.png"),
  dpi = 600,
  width = 80 *figexp*deinflar_word,
  height = 80 * figexp*deinflar_word* (7/6),
  units = "mm"
)
```

## Summary statistics

```{r}
cat("Nearly inmediate deaths\n")
scales::percent((9+22+40+113)/2996, accuracy=.1)

clean_df |>
mutate(fu_years = as.numeric(difftime(death_date_rec, disch_date_rec6, units = "days")) / 365.2425) |>
summarise(
median = sprintf("%1.2f",median(fu_years)),
q25 = sprintf("%1.2f",quantile(fu_years, 0.25)),
q75 = sprintf("%1.2f",quantile(fu_years, 0.75))
)
```

# Sensitivity, last treatment

## Overall, age, sex
 
```{r  sens1-smr-overall}
## SMR, Overall  -----------------------------------------------------------
c_SISTRAT_b_std_bins <- lexpand(
  clean_df_b, 
  status = status, 
  birth = birth_date_rec, 
  exit = death_date_rec, 
  entry = disch_date_rec6,
  breaks = list(
    per = seq(2010, 2021, by = 1), 
    age = age_breaks # <--- This matches the reference keys (15, 20, 25...)
  ),
  aggre = list(year = per, 
               agegroup = age, 
               sex = sex_rec)
)


sir_final_b_std <- popEpi::sir(
  coh.data = c_SISTRAT_b_std_bins,
  coh.obs = 'from0to1',    # Observed deaths column
  coh.pyrs = 'pyrs',       # Person-years column
  ref.data = mx_national_clean,
  ref.rate = 'mx',         # The mortality rate column in reference
  adjust = c('agegroup', 'year', 'sex')#,
#print = 'total'
)
print(sir_final_b_std)

sir_ci_phi_improved( sir_final_b_std, extract_phi(c_SISTRAT_b_std_bins)) |>
  dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
  SIR,
  CI_low,
  CI_high)) |>
  data.table::data.table()
```

```{r  sens1-dsr-overall}
## DSRs  -----------------------------------------------------------
dsr_tot_b <- popEpi::rate(
data    = c_SISTRAT_b_std_bins,
obs     = "from0to1",
pyrs    = "pyrs",
adjust  = c("year","sex","agegroup"),
weights = weights_corr
)
dsr_tot_b$rate_1k <- dsr_tot_b$rate * 1e3
DSR_1k_corr_corr_b <- base::mapply(
  dsr_format_corr,                 # FUN
  dsr_tot_b$rate.adj,           # primer vector (rate)
  dsr_tot_b$SE.rate.adj,        # segundo vector (se)
  MoreArgs = list(            # argumentos fijos extra
    #phi    = extract_phi_dir(c_SISTRAT_std_bins),
    #opc. conservadora
    phi    = base::max(1,  extract_phi_dir(c_SISTRAT_b_std_bins)),
    factor = 1e3,
    digits = 1,
    conf   = 0.95))
#DSR_1k_corr_corr_b
#[1] "12.1 (9.7–15.2)"
```

```{r  sens1-dsr-age}
## DSR w/ 5-yr-bin lexis, unfinished tr.- by old ages -------------------------------------------
### DSR, by age ---------------------------
c_SISTRAT_b_std_old <- c_SISTRAT_b_std_bins |>
  dplyr::mutate(agegroup_old = to_old_agegroup(agegroup)) |>
  dplyr::group_by(year, sex, agegroup_old) |>
  dplyr::summarise(
  pyrs     = base::sum(pyrs),
  from0to1 = base::sum(from0to1),
  .groups  = "drop"
)
dsr_age_old_b <- popEpi::rate(
  data    = c_SISTRAT_b_std_old,
  obs     = "from0to1",
  pyrs    = "pyrs",
  print   = "agegroup_old",
  adjust  = c("year","sex"),
  weights = list(year = w_year, sex = w_sex)
)
#### Etract phi ---------------------------
phi_df_age_old_b <- extract_phi_by_age(c_SISTRAT_b_std_old|> dplyr::rename(agegroup = agegroup_old))
phi_vec_age_old_b <- base::pmax(1, phi_df_age_old_b$phi[match(dsr_age_old_b$agegroup_old, phi_df_age_old_b$agegroup)])
#### Combine phi ---------------------------
dsr_age_old_b<- cbind.data.frame(dplyr::as_tibble(dsr_age_old_b),phi= phi_vec_age_old_b)
# dplyr::select(dsr_age_old_b, rate, rate.lo, rate.hi)|>
# mutate_if(is.numeric, ~sprintf("%1.1f",.*1e3))

DSR_1k_corr_agegr_corr_b <- base::mapply(
  dsr_format_corr,
  dsr_age_old_b$rate.adj,
  dsr_age_old_b$SE.rate.adj,
  dsr_age_old_b$phi,
  MoreArgs = list(factor = 1e3, digits = 6, conf = .95)
  )

```

```{r  sens1-smr-age}
### SMR by old ages ---------------------------
c_SISTRAT_b_std_bins_wold <- c_SISTRAT_b_std_bins |>
dplyr::mutate(agegroup_old = to_old_agegroup(agegroup))

smr_age_old_b <-
  popEpi::sir(c_SISTRAT_b_std_bins_wold, coh.obs = 'from0to1',
    coh.pyrs = 'pyrs',
    ref.data = mx_national_clean,
    ref.rate = mx,
    print = c("agegroup_old"),
    adjust = c("year", "sex", "agegroup"), #2025-12-16: the original still adjusts for agegroup
    test.type = "homogeneity",
    conf.type = "wald", #conf.type = "wald" usa la aproximación normal de Poisson (la misma lógica que explicamos antes).
    conf.level = 0.95, EAR = T)

# sir_ci_phi_improved( smr_age_old_b, dsr_age_old_b$phi)|>
# dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
# SIR,
# CI_low,
# CI_high)) |>
# data.table::data.table()
```

```{r  sens1-dsr-sex}
## DSR w/ 5-yr-bin lexis, unfinished tr.- by sex -------------------------------------------
### DSR by sex--------------------------
phi_sex_df_b  <- extract_phi_by_sex(c_SISTRAT_b_std_bins)
dsr_by_sex_b <- popEpi::rate(
  data    = c_SISTRAT_b_std_bins,
  obs     = "from0to1",
  pyrs    = "pyrs",
  print   = "sex",
  adjust  = c("year","agegroup"),
    weights = list(year = w_year_5y, agegroup = w_age_5y)
  )
# dplyr::select(dsr_by_sex_b, rate, rate.lo, rate.hi)|>
# mutate_if(is.numeric, ~sprintf("%1.1f",.*1e3))

phi_sex_vec_b <- base::pmax(1, phi_sex_df_b$phi[match(dsr_by_sex_b$sex, phi_sex_df_b$sex)])
dsr_by_sex_b <- cbind.data.frame(dplyr::as_tibble(dsr_by_sex_b),phi= phi_sex_vec_b)
DSR_1k_corr_sex_corr_b <- base::mapply(
  dsr_format_corr,
  dsr_by_sex_b $rate.adj,
  dsr_by_sex_b $SE.rate.adj,
  phi     = 1, #not necessary (phis <1)
    MoreArgs = list(factor = 1e3, digits = 6, conf = .95)
  )

# data.table::data.table(as.vector(sapply(DSR_1k_corr_sex_corr_b,
# format_interval)))

```

```{r  sens1-smr-sex}
### SMR by sex--------------------------
smr_sex_corr_b <- popEpi::sir(
coh.data = c_SISTRAT_b_std_bins,
coh.obs  = "from0to1",
coh.pyrs = "pyrs",
ref.data = mx_national_clean,
ref.rate = "mx",
print    = "sex",
adjust   = c("year","sex","agegroup")  # <- 5y internal
)
smr_sex_corr_b

smr_sex_corr_tbl_b <- cbind.data.frame(dplyr::as_tibble(smr_sex_corr_b), phi=phi_sex_vec_b)
# smr_sex_corr_tbl_b

# sir_ci_phi_improved( smr_sex_corr_b, smr_sex_corr_tbl_b$phi)|>
# dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
# SIR,
# CI_low,
# CI_high)) |>
# data.table::data.table()
```


```{r sens1_summary_table, results="asis"}
# -------------------------
# OVERALL
# -------------------------
sir_overall_ci_b <- sir_ci_phi_improved(sir_final_b_std, extract_phi(c_SISTRAT_b_std_bins)) |>
  dplyr::mutate(SMR = fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", SIR, CI_low, CI_high)))

overall_tbl_b <- data.frame(
  Section = "Overall",
  Subgroup = "All",
  `Obs/Exp` = base::paste0(sir_final_b_std$observed[1], "/", base::as.integer(sir_final_b_std$expected[1])),
  `Person-years` = base::formatC(sir_final_b_std$pyrs[1], format="f", digits=0, big.mark=","),
  phi = base::sprintf("%1.2f", base::max(1, extract_phi_dir(c_SISTRAT_b_std_bins))),
  SMR = sir_overall_ci_b$SMR[1],
  DSR_1k = fmt_interval(DSR_1k_corr_corr_b[1]),
  CMR_1k = cmr_1k_print(dsr_tot_b$rate[1], dsr_tot_b$rate.lo[1], dsr_tot_b$rate.hi[1], digits = 1),
  stringsAsFactors = FALSE
)

# -------------------------
# SEX
# -------------------------
smr_sex_corr_b <- smr_sex_corr_b[
  base::match(dsr_by_sex_b$sex, smr_sex_corr_b$sex),
]
sir_sex_ci_b <- sir_ci_phi_improved(smr_sex_corr_b, smr_sex_corr_tbl_b$phi) |>
  dplyr::mutate(SMR = fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", SIR, CI_low, CI_high)))

dsr_sex_str_b <- fmt_interval(base::as.vector(base::sapply(DSR_1k_corr_sex_corr_b, format_interval)))

smr_sex_corr_tbl_b <- smr_sex_corr_tbl_b[
  base::match(dsr_by_sex_b$sex, smr_sex_corr_tbl_b$sex),
]

sex_tbl_b <- data.frame(
  Section = "Sex",
  Subgroup = base::as.character(dsr_by_sex_b$sex),
  `Obs/Exp` = base::paste0(smr_sex_corr_tbl_b$observed, "/", base::as.integer(smr_sex_corr_tbl_b$expected)),
  `Person-years` = base::formatC(dsr_by_sex_b$pyrs, format="f", digits=0, big.mark=","),
  phi = base::sprintf("%1.2f", dsr_by_sex_b$phi),
  SMR = sir_sex_ci_b$SMR,
  DSR_1k = dsr_sex_str_b,
  CMR_1k = cmr_1k_print(dsr_by_sex_b$rate, dsr_by_sex_b$rate.lo, dsr_by_sex_b$rate.hi, digits = 1),
  stringsAsFactors = FALSE
)

# -------------------------
# AGE (old bins)
# -------------------------
sir_age_ci_b <- sir_ci_phi_improved(smr_age_old_b, dsr_age_old_b$phi) |>
  dplyr::mutate(SMR = fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", SIR, CI_low, CI_high)))

dsr_age_str_b <- fmt_interval(base::as.vector(base::sapply(DSR_1k_corr_agegr_corr_b, format_interval)))

age_tbl_b <- data.frame(
  Section = "Age group",
  Subgroup = base::as.character(dsr_age_old_b$agegroup_old),
  `Obs/Exp` = base::paste0(smr_age_old_b$observed, "/", base::as.integer(smr_age_old_b$expected)),
  `Person-years` = base::formatC(dsr_age_old_b$pyrs, format="f", digits=0, big.mark=","),
  phi = base::sprintf("%1.2f", dsr_age_old_b$phi),
  SMR = sir_age_ci_b$SMR,
  DSR_1k = dsr_age_str_b,
  CMR_1k = cmr_1k_print(dsr_age_old_b$rate, dsr_age_old_b$rate.lo, dsr_age_old_b$rate.hi, digits = 1),
  stringsAsFactors = FALSE
)

# -------------------------
# FINAL
# -------------------------
final_tbl_b <- base::rbind(overall_tbl_b, sex_tbl_b, age_tbl_b)

knitr::kable(
  final_tbl_b,
  caption = "Sensitivity 1: Summary of SMR (phi-corrected), DSR and CMR (per 1,000 person-years) overall and by sex/age.",
  align = c("l","l","l","r","r","l","l","l")
)
```

### Heterogeneity

```{r sens1-het-b-age-sex}
smr_sex_dat_b <- sir_ci_phi_improved( smr_sex_corr_b, smr_sex_corr_tbl_b$phi) |>
    dplyr::mutate(
        yi  = base::log(SIR),
        sei = se_log_from_ci(SIR, CI_low, CI_high, conf = 0.95)
    )

meta_smr_sex_b <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = smr_sex_dat_b
)

smr_age_dat_b <- sir_ci_phi_improved( smr_age_old_b, sir_age_ci_b$phi) |>
    dplyr::mutate(
        yi  = base::log(SIR),
        sei = se_log_from_ci(SIR,CI_low, CI_high, conf = 0.95)
    )

meta_smr_age_b <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = smr_age_dat_b
)

dsr_sex_dat_b <- dsr_by_sex_b |>
    dplyr::mutate(sex = base::as.character(sex)) |>
    dplyr::mutate(
        yi  = base::log(rate.adj),
        sei = se_log_from_rate_se(rate.adj, SE.rate.adj, phi = phi)
    )

meta_dsr_sex_b <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = dsr_sex_dat_b
)

dsr_age_dat_b <- dsr_age_old_b |>
    dplyr::mutate(
        yi  = base::log(rate.adj),
        sei = se_log_from_rate_se(rate.adj, SE.rate.adj, phi = phi)
    )

meta_dsr_age_b <- metafor::rma(yi = yi, sei = sei, method = "FE", data = dsr_age_dat_b)


het_tbl_b <- base::rbind(
  het_row(meta_smr_sex_b, "SMR (SIR) by sex"),
  het_row(meta_smr_age_b, "SMR (SIR) by age group"),
  het_row(meta_dsr_sex_b, "DSR by sex"),
  het_row(meta_dsr_age_b, "DSR by age group")
)

knitr::kable(
  het_tbl_b,
  caption = "Heterogeneity tests (fixed-effect meta-analysis; Cochran’s Q and I²).",
  align = c("l","r","r","r","r")
)
```

## Reformat Lexis

```{r sens1-lexis-format}
## By strata, last tr. -----------------------------------------------------
clean_df2_b <- clean_df_b |>
  dplyr::mutate(
  # 0/1 -> labels ( 1 = Residential, 0 = Ambulatory)
  setting = dplyr::if_else(res_plan == 1, "Residential", "Ambulatory"),
  setting = base::factor(setting, levels = c("Ambulatory","Residential")),
  completed = base::as.character(tr_compliance_status),
  completed = base::factor(completed, levels = c("Completed","Not completed")),
  licit_illicit = base::as.character(prim_sub_licit),
  licit_illicit = base::factor(licit_illicit, levels = c("licit","illicit")),
  disch_age_cat = base::as.character(disch_age_cat),
  disch_age_cat = base::factor(disch_age_cat, levels = c("18-29","30-44","45-59","60+")),
  # Sex: asegúrate que sea Male/Female como tus tasas nacionales
  sex = stringr::str_to_title(base::as.character(sex_rec)),
  sex = base::factor(sex, levels = c("Male","Female"))
)
# ---- 2) 5-year attained-age Lexis (keep 2021 as boundary)
c_SISTRAT_5y_b <- popEpi::lexpand(
  data   = clean_df2_b,
  status = status,
  birth  = birth_date_rec,
  entry  = disch_date_rec6,
  exit   = death_date_rec,
  breaks = list(
  per = base::seq(2010, 2021, by = 1),
  age = age_breaks
  ),
  aggre = list(
  year    = per,
  agegroup= age,      # 5y attained-age keys (15,20,25,...)
  sex     = sex,
  # subgroup vars (constantes por individuo; quedan replicadas por split)
  setting      = setting,
  completed    = completed,
  licit_illicit= licit_illicit,
  disch_age_cat= disch_age_cat
  )
)
```


## Strata


### DSR

```{r sens1-format-lexis-strata-b}

dsr_setting_b   <- run_dsr_by(c_SISTRAT_5y_b, weights_corr, "setting")
dsr_completed_b <- run_dsr_by(c_SISTRAT_5y_b, weights_corr, "completed")
dsr_licit_b     <- run_dsr_by(c_SISTRAT_5y_b, weights_corr, "licit_illicit")

```

```{r sens1-strata-dsr_one_kable, results="asis"}
tbl_all_b <- base::rbind(
    prep_dsr(dsr_setting_b,   "Setting"),
    prep_dsr(dsr_completed_b, "Completion"),
    prep_dsr(dsr_licit_b,     "Licit/illicit")#,  prep_dsr(dsr_age,       "Age group")
)

knitr::kable(
    tbl_all_b,
    caption = "Directly standardized rates (DSR) and crude mortality rates (CMR), per 1,000 person-years, across subgroups.",
    align = c("l","l","r","r","r","r","r","l","r","r","r","r")
)
```

### SMR

```{r sens1-strata-smr, results="asis"}
smr_setting_b   <- run_smr_by(c_SISTRAT_5y_b, mx_national_clean, "setting")
smr_completed_b <- run_smr_by(c_SISTRAT_5y_b, mx_national_clean, "completed")
smr_licit_b     <- run_smr_by(c_SISTRAT_5y_b, mx_national_clean, "licit_illicit")


smr_setting_phi_b <- smr_setting_b%>%
    dplyr::left_join(dplyr::select(dsr_setting_b, group, phi), by = "group")%>%
    sir_ci_phi_profile(phi = .$phi) |>
    dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
                                  sir, sir.phi.lo, sir.phi.hi))
smr_completed_phi_b <- smr_completed_b%>%
    dplyr::left_join(dplyr::select(dsr_completed_b, group, phi), by = "group")%>%
    sir_ci_phi_profile(phi = .$phi) |>
    dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
                                  sir, sir.phi.lo, sir.phi.hi))
smr_licit_phi_b <- smr_licit_b%>%
    dplyr::left_join(dplyr::select(dsr_licit_b, group, phi), by = "group")%>%
    sir_ci_phi_profile(phi = .$phi) |>
    dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
                                  sir, sir.phi.lo, sir.phi.hi))
# Combine all tables into one with a "Category" column
smr_table_b <- bind_rows(
    smr_setting_phi_b %>% mutate(Category = "Setting") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print),
    smr_completed_phi_b %>% mutate(Category = "Treatment Completion") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print),
    smr_licit_phi_b %>% mutate(Category = "Primary Substance") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print)#,
    # smr_agecat_phi %>% mutate(Category = "Age at Discharge") %>% select(Category, Group = group, Observed = observed, Expected = expected, SMR = print)
) %>%
    dplyr::mutate(Category = ifelse( Category != lag(Category, default = ""), 
            Category,
            ""))

# Create kable
kable(
    smr_table_b,
    col.names = c("Category", "Group", "Observed", "Expected", "SMR (95% CI)"),
    align = c("l", "l", "r", "r", "r"),
    booktabs = TRUE,
    linesep = ""
) 
```


### Heterogeneity

```{r bootstrap_phi_smr_and_het_b, results="asis"}
# --- 1) Run SMRs ----------------------------------------------------------
smr_setting_b   <- run_smr_by(c_SISTRAT_5y_b, mx_national_clean, "setting")
smr_completed_b <- run_smr_by(c_SISTRAT_5y_b, mx_national_clean, "completed")
smr_licit_b     <- run_smr_by(c_SISTRAT_5y_b, mx_national_clean, "licit_illicit")

# --- 2) Phi-corrected SMRs (profile CI) ----------------------------------
smr_setting_phi_b   <- make_smr_phi(smr_setting_b,   dsr_setting_b)
smr_completed_phi_b <- make_smr_phi(smr_completed_b, dsr_completed_b)
smr_licit_phi_b     <- make_smr_phi(smr_licit_b,     dsr_licit_b)

# --- 3) Heterogeneity helper (DSR + SMR) ---------------------------------


# --- 4) Fit meta models ---------------------------------------------------
meta_dsr_setting_b <- het_dsr(dsr_setting_b)
meta_smr_setting_b <- het_smr(smr_setting_phi_b)

meta_dsr_compl_b   <- het_dsr(dsr_completed_b)
meta_smr_compl_b   <- het_smr(smr_completed_phi_b)

meta_dsr_licit_b   <- het_dsr(dsr_licit_b)
meta_smr_licit_b   <- het_smr(smr_licit_phi_b)

# --- 5) Print quick heterogeneity table ----------------------------------
het_tbl_b <- base::rbind(
  het_row_strata(meta_dsr_setting_b, "Setting",       "DSR"),
  het_row_strata(meta_smr_setting_b, "Setting",       "SMR"),
  het_row_strata(meta_dsr_compl_b,   "Compliance",    "DSR"),
  het_row_strata(meta_smr_compl_b,   "Compliance",    "SMR"),
  het_row_strata(meta_dsr_licit_b,   "Licit/Illicit", "DSR"),
  het_row_strata(meta_smr_licit_b,   "Licit/Illicit", "SMR")
)

knitr::kable(
  het_tbl_b,
  caption = "Heterogeneity tests (fixed-effect meta-analysis): Cochran’s Q and I² (phi-corrected).",
  align = c("l","l","r","r","r","r")
)
```


# Sensitivity, unfinished treatments

## Overall, age, sex

```{r  sens2-smr-overall}
## SMR, Overall  -----------------------------------------------------------
c_SISTRAT_c_std_bins <- popEpi::lexpand(
  clean_df_c,
  status = status,
  birth = birth_date_rec,
  exit = death_date_rec,
  entry = disch_date_corr,
  breaks = list(
  per = seq(2010, 2021, by = 1),
  age = age_breaks # <--- This matches the reference keys (15, 20, 25...)
  ),
    aggre = list(year = per,
    agegroup = age,
    sex = sex_rec)
)

sir_final_c_std <- popEpi::sir(
  coh.data = c_SISTRAT_c_std_bins,
  coh.obs = 'from0to1',    # Observed deaths column
  coh.pyrs = 'pyrs',       # Person-years column
  ref.data = mx_national_clean,
  ref.rate = 'mx',         # The mortality rate column in reference
  adjust = c('agegroup', 'year', 'sex')#,
#print = 'total'
)
print(sir_final_c_std)

sir_ci_phi_improved( sir_final_c_std, extract_phi(c_SISTRAT_c_std_bins)) |>
  dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
  SIR,
  CI_low,
  CI_high)) |>
  data.table::data.table()
```

```{r  sens2-dsr-overall}
## DSRs  -----------------------------------------------------------
dsr_tot_c <- popEpi::rate(
data    = c_SISTRAT_c_std_bins,
obs     = "from0to1",
pyrs    = "pyrs",
adjust  = c("year","sex","agegroup"),
weights = weights_corr
)
dsr_tot_c$rate_1k <- dsr_tot_c$rate * 1e3
DSR_1k_corr_corr_c <- base::mapply(
  dsr_format_corr,                 # FUN
  dsr_tot_c$rate.adj,           # primer vector (rate)
  dsr_tot_c$SE.rate.adj,        # segundo vector (se)
  MoreArgs = list(            # argumentos fijos extra
    #phi    = extract_phi_dir(c_SISTRAT_std_bins),
    #opc. conservadora
    phi    = base::max(1,  extract_phi_dir(c_SISTRAT_c_std_bins)),
    factor = 1e3,
    digits = 1,
    conf   = 0.95))
# DSR_1k_corr_corr_c
# [1] "11.2 (10.0–12.5)"
```

```{r  sens2-dsr-age}
## DSR w/ 5-yr-bin lexis, unfinished tr.- by old ages -------------------------------------------
### DSR, by age ---------------------------
c_SISTRAT_c_std_old <- c_SISTRAT_c_std_bins |>
dplyr::mutate(agegroup_old = to_old_agegroup(agegroup)) |>
dplyr::group_by(year, sex, agegroup_old) |>
dplyr::summarise(
pyrs     = base::sum(pyrs),
from0to1 = base::sum(from0to1),
.groups  = "drop"
)
dsr_age_old_c <- popEpi::rate(
data    = c_SISTRAT_c_std_old,
obs     = "from0to1",
pyrs    = "pyrs",
print   = "agegroup_old",
adjust  = c("year","sex"),
weights = list(year = w_year, sex = w_sex)
)
#### Etract phi ---------------------------
phi_df_age_old_c <- extract_phi_by_age(c_SISTRAT_c_std_old|> dplyr::rename(agegroup = agegroup_old))
phi_vec_age_old_c <- base::pmax(1, phi_df_age_old_c$phi[match(dsr_age_old_c$agegroup_old, phi_df_age_old_c$agegroup)])
#### Combine phi ---------------------------
dsr_age_old_c<- cbind.data.frame(dplyr::as_tibble(dsr_age_old_c),phi= phi_vec_age_old_c)
# dplyr::select(dsr_age_old_c, rate, rate.lo, rate.hi)|>
# mutate_if(is.numeric, ~sprintf("%1.1f",.*1e3))

DSR_1k_corr_agegr_corr_c <- base::mapply(
  dsr_format_corr,
  dsr_age_old_c$rate.adj,
  dsr_age_old_c$SE.rate.adj,
  dsr_age_old_c$phi,
  MoreArgs = list(factor = 1e3, digits = 6, conf = .95)
  )

# data.table::data.table(as.vector(sapply(DSR_1k_corr_agegr_corr_c,
# format_interval)))

```

```{r  sens2-smr-age}
### SMR by old ages ---------------------------
c_SISTRAT_c_std_bins_wold <- c_SISTRAT_c_std_bins |>
dplyr::mutate(agegroup_old = to_old_agegroup(agegroup))

smr_age_old_c <-
  popEpi::sir(c_SISTRAT_c_std_bins_wold, coh.obs = 'from0to1',
    coh.pyrs = 'pyrs',
    ref.data = mx_national_clean,
    ref.rate = mx,
    print = c("agegroup_old"),
    adjust = c("year", "sex", "agegroup"), #2025-12-16: the original still adjusts for agegroup
    test.type = "homogeneity",
    conf.type = "wald", #conf.type = "wald" usa la aproximación normal de Poisson (la misma lógica que explicamos antes).
    conf.level = 0.95, EAR = T)
# sir_ci_phi_improved( smr_age_old_c, dsr_age_old_c$phi)|>
#   dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
#   SIR,
#   CI_low,
#   CI_high)) |>
#   data.table::data.table()
```

```{r  sens2-dsr-sex}
## DSR w/ 5-yr-bin lexis, unfinished tr.- by sex -------------------------------------------
### DSR by sex--------------------------
phi_sex_df_c  <- extract_phi_by_sex(c_SISTRAT_c_std_bins)
dsr_by_sex_c <- popEpi::rate(
  data    = c_SISTRAT_c_std_bins,
  obs     = "from0to1",
  pyrs    = "pyrs",
  print   = "sex",
  adjust  = c("year","agegroup"),
  weights = list(year = w_year_5y, agegroup = w_age_5y)
  )
# dplyr::select(dsr_by_sex_c, rate, rate.lo, rate.hi)|>
# mutate_if(is.numeric, ~sprintf("%1.1f",.*1e3))

phi_sex_vec_c <- base::pmax(1, phi_sex_df_c$phi[match(dsr_by_sex_c$sex, phi_sex_df_c$sex)])
dsr_by_sex_c <- cbind.data.frame(dplyr::as_tibble(dsr_by_sex_c),phi= phi_sex_vec_c)
DSR_1k_corr_sex_corr_c <- base::mapply(
dsr_format_corr,
dsr_by_sex_c $rate.adj,
dsr_by_sex_c $SE.rate.adj,
phi     = 1, #not necessary (phis <1)
MoreArgs = list(factor = 1e3, digits = 6, conf = .95)
)
# 
# data.table::data.table(as.vector(sapply(DSR_1k_corr_sex_corr_c,
# format_interval)))
```

```{r  sens2-smr-sex}
### SMR by sex--------------------------
smr_sex_corr_c <- popEpi::sir(
coh.data = c_SISTRAT_c_std_bins,
coh.obs  = "from0to1",
coh.pyrs = "pyrs",
ref.data = mx_national_clean,
ref.rate = "mx",
print    = "sex",
adjust   = c("year","sex","agegroup")  # <- 5y internal
)
smr_sex_corr_c

smr_sex_corr_tbl_c <- cbind.data.frame(dplyr::as_tibble(smr_sex_corr_c), phi=phi_sex_vec_c)
# smr_sex_corr_tbl_c
# 
# sir_ci_phi_improved( smr_sex_corr_c, smr_sex_corr_tbl_c$phi)|>
#   dplyr::mutate(print= sprintf("%.2f (%.2f–%.2f)",
#   SIR,
#   CI_low,
#   CI_high)) |>
#   data.table::data.table()
```


```{r sens2_summary_table, results="asis"}
# -------------------------
# OVERALL
# -------------------------
sir_overall_ci_c <- sir_ci_phi_improved(sir_final_c_std, extract_phi(c_SISTRAT_c_std_bins)) |>
  dplyr::mutate(SMR = fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", SIR, CI_low, CI_high)))

overall_tbl_c <- data.frame(
  Section = "Overall",
  Subgroup = "All",
  `Obs/Exp` = base::paste0(sir_final_c_std$observed[1], "/", base::as.integer(sir_final_c_std$expected[1])),
  `Person-years` = base::formatC(sir_final_c_std$pyrs[1], format="f", digits=0, big.mark=","),
  phi = base::sprintf("%1.2f", base::max(1, extract_phi_dir(c_SISTRAT_c_std_bins))),
  SMR = sir_overall_ci_c$SMR[1],
  DSR_1k = fmt_interval(DSR_1k_corr_corr_c[1]),
  CMR_1k = cmr_1k_print(dsr_tot_c$rate[1], dsr_tot_c$rate.lo[1], dsr_tot_c$rate.hi[1], digits = 1),
  stringsAsFactors = FALSE
)

# -------------------------
# SEX
# -------------------------
smr_sex_corr_c <- smr_sex_corr_c[
  base::match(dsr_by_sex_c$sex, smr_sex_corr_c$sex),
]
sir_sex_ci_c <- sir_ci_phi_improved(smr_sex_corr_c, smr_sex_corr_tbl_c$phi) |>
  dplyr::mutate(SMR = fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", SIR, CI_low, CI_high)))

dsr_sex_str_c <- fmt_interval(base::as.vector(base::sapply(DSR_1k_corr_sex_corr_c, format_interval)))

smr_sex_corr_tbl_c <- smr_sex_corr_tbl_c[
  base::match(dsr_by_sex_c$sex, smr_sex_corr_tbl_c$sex),
]

sex_tbl_c <- data.frame(
  Section = "Sex",
  Subgroup = base::as.character(dsr_by_sex_c$sex),
  `Obs/Exp` = base::paste0(smr_sex_corr_tbl_c$observed, "/", base::as.integer(smr_sex_corr_tbl_c$expected)),
  `Person-years` = base::formatC(dsr_by_sex_c$pyrs, format="f", digits=0, big.mark=","),
  phi = base::sprintf("%1.2f", dsr_by_sex_c$phi),
  SMR = sir_sex_ci_c$SMR,
  DSR_1k = dsr_sex_str_c,
  CMR_1k = cmr_1k_print(dsr_by_sex_c$rate, dsr_by_sex_c$rate.lo, dsr_by_sex_c$rate.hi, digits = 1),
  stringsAsFactors = FALSE
)

# -------------------------
# AGE (old bins)
# -------------------------
sir_age_ci_c <- sir_ci_phi_improved(smr_age_old_c, dsr_age_old_c$phi) |>
  dplyr::mutate(SMR = fmt_interval(base::sprintf("%.2f (%.2f–%.2f)", SIR, CI_low, CI_high)))

dsr_age_str_c <- fmt_interval(base::as.vector(base::sapply(DSR_1k_corr_agegr_corr_c, format_interval)))

age_tbl_c <- data.frame(
  Section = "Age group",
  Subgroup = base::as.character(dsr_age_old_c$agegroup_old),
  `Obs/Exp` = base::paste0(smr_age_old_c$observed, "/", base::as.integer(smr_age_old_c$expected)),
  `Person-years` = base::formatC(dsr_age_old_c$pyrs, format="f", digits=0, big.mark=","),
  phi = base::sprintf("%1.2f", dsr_age_old_c$phi),
  SMR = sir_age_ci_c$SMR,
  DSR_1k = dsr_age_str_c,
  CMR_1k = cmr_1k_print(dsr_age_old_c$rate, dsr_age_old_c$rate.lo, dsr_age_old_c$rate.hi, digits = 1),
  stringsAsFactors = FALSE
)

# -------------------------
# FINAL
# -------------------------
final_tbl_c <- base::rbind(overall_tbl_c, sex_tbl_c, age_tbl_c)

knitr::kable(
  final_tbl_c,
  caption = "Sensitivity 2: Summary of SMR (phi-corrected), DSR and CMR (per 1,000 person-years) overall and by sex/age.",
  align = c("l","l","l","r","r","l","l","l")
)
```


### Heterogeneity


```{r sens2-het-c-age-sex}
smr_sex_dat_c <- sir_ci_phi_improved( smr_sex_corr_c, smr_sex_corr_tbl_c$phi) |>
    dplyr::mutate(
        yi  = base::log(SIR),
        sei = se_log_from_ci(SIR, CI_low, CI_high, conf = 0.95)
    )

meta_smr_sex_c <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = smr_sex_dat_c
)

smr_age_dat_c <- sir_ci_phi_improved( smr_age_old_c, sir_age_ci_c$phi) |>
    dplyr::mutate(
        yi  = base::log(SIR),
        sei = se_log_from_ci(SIR,CI_low, CI_high, conf = 0.95)
    )

meta_smr_age_c <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = smr_age_dat_c
)

dsr_sex_dat_c <- dsr_by_sex_c |>
    dplyr::mutate(sex = base::as.character(sex)) |>
    dplyr::mutate(
        yi  = base::log(rate.adj),
        sei = se_log_from_rate_se(rate.adj, SE.rate.adj, phi = phi)
    )

meta_dsr_sex_c <- metafor::rma(
    yi = yi,
    sei = sei,
    method = "FE",
    data = dsr_sex_dat_c
)

dsr_age_dat_c <- dsr_age_old_c |>
    dplyr::mutate(
        yi  = base::log(rate.adj),
        sei = se_log_from_rate_se(rate.adj, SE.rate.adj, phi = phi)
    )

meta_dsr_age_c <- metafor::rma(yi = yi, sei = sei, method = "FE", data = dsr_age_dat_c)


het_tbl_c <- base::rbind(
  het_row(meta_smr_sex_c, "SMR (SIR) by sex"),
  het_row(meta_smr_age_c, "SMR (SIR) by age group"),
  het_row(meta_dsr_sex_c, "DSR by sex"),
  het_row(meta_dsr_age_c, "DSR by age group")
)

knitr::kable(
  het_tbl_c,
  caption = "Heterogeneity tests (fixed-effect meta-analysis; Cochran’s Q and I²).",
  align = c("l","r","r","r","r")
)
```


## Reformat Lexis

```{r sens2-lexis-format}
## By strata, unfinished tr. -----------------------------------------------------
clean_df2_c <- clean_df_c |>
dplyr::mutate(
# 0/1 -> labels ( 1 = Residential, 0 = Ambulatory)
setting = dplyr::if_else(res_plan == 1, "Residential", "Ambulatory"),
setting = base::factor(setting, levels = c("Ambulatory","Residential")),
completed = base::as.character(tr_compliance_status),
completed = base::factor(completed, levels = c("Completed","Not completed")),
licit_illicit = base::as.character(prim_sub_licit),
licit_illicit = base::factor(licit_illicit, levels = c("licit","illicit")),
disch_age_cat = base::as.character(disch_age_cat),
disch_age_cat = base::factor(disch_age_cat, levels = c("18-29","30-44","45-59","60+")),
# Sex: asegúrate que sea Male/Female como tus tasas nacionales
sex = stringr::str_to_title(base::as.character(sex_rec)),
sex = base::factor(sex, levels = c("Male","Female"))
)
# ---- 2) 5-year attained-age Lexis (keep 2021 as boundary)
table(clean_df2_c$tr_compliance_status, exclude=NULL)
# Completed Not completed          <NA>
#     19213         51047         13522
c_SISTRAT_5y_c <- popEpi::lexpand(
data   = clean_df2_c|>
dplyr::mutate(tr_compliance_status=
ifelse(is.na(tr_compliance_status),"Unfinished", tr_compliance_status)),
status = status,
birth  = birth_date_rec,
entry  = disch_date_corr,
exit   = death_date_rec,
breaks = list(
per = base::seq(2010, 2021, by = 1),
age = age_breaks
),
aggre = list(
year    = per,
agegroup= age,      # 5y attained-age keys (15,20,25,...)
sex     = sex,
# subgroup vars (constantes por individuo; quedan replicadas por split)
setting      = res_plan,
completed    = tr_compliance_status,
licit_illicit= licit_illicit,
disch_age_cat= disch_age_cat
)
)

warning("2025-12-30= I chose to separate into Unfinished and Tr. completed
")
c_SISTRAT_5y_c <- c_SISTRAT_5y_c |>
dplyr::mutate(
setting = factor(setting, levels = c(0, 1),
labels = c("Ambulatory", "Residential"))
)
```


## Strata

### DSR

```{r sens2-format-lexis-strata-b}
warning("disch_age_cat es edad al alta (baseline) → perfecto para “preservar
los grupos antiguos” aunque el Lexis use attained-age 5y internamente para el ajuste.")
dsr_setting_c   <- run_dsr_by(c_SISTRAT_5y_c, weights_corr, "setting")
dsr_completed_c <- run_dsr_by(c_SISTRAT_5y_c, weights_corr, "completed")
dsr_licit_c     <- run_dsr_by(c_SISTRAT_5y_c, weights_corr, "licit_illicit")

```

```{r sens2-strata-dsr_one_kable, results="asis"}
tbl_all_c <- base::rbind(
    prep_dsr(dsr_setting_c,   "Setting"),
    prep_dsr(dsr_completed_c, "Completion"),
    prep_dsr(dsr_licit_c,     "Licit/illicit")#,  prep_dsr(dsr_age,       "Age group")
)

knitr::kable(
    tbl_all_c,
    caption = "Directly standardized rates (DSR) and crude mortality rates (CMR), per 1,000 person-years, across subgroups.",
    align = c("l","l","r","r","r","r","r","l","r","r","r","r")
)
```

### SMR

```{r sens2-strata-smr, results="asis"}
smr_setting_c   <- run_smr_by(c_SISTRAT_5y_c, mx_national_clean, "setting")
smr_completed_c <- run_smr_by(c_SISTRAT_5y_c, mx_national_clean, "completed")
smr_licit_c     <- run_smr_by(c_SISTRAT_5y_c, mx_national_clean, "licit_illicit")


smr_setting_phi_c <- smr_setting_c%>%
    dplyr::left_join(dplyr::select(dsr_setting_c, group, phi), by = "group")%>%
    sir_ci_phi_profile(phi = .$phi) |>
    dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
                                  sir, sir.phi.lo, sir.phi.hi))
smr_completed_phi_c <- smr_completed_c%>%
    dplyr::left_join(dplyr::select(dsr_completed_c, group, phi), by = "group")%>%
    sir_ci_phi_profile(phi = .$phi) |>
    dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
                                  sir, sir.phi.lo, sir.phi.hi))
smr_licit_phi_c <- smr_licit_c%>%
    dplyr::left_join(dplyr::select(dsr_licit_c, group, phi), by = "group")%>%
    sir_ci_phi_profile(phi = .$phi) |>
    dplyr::mutate(print = sprintf(paste0("%.", 2, "f (%.", 2, "f–%.", 2, "f)"),
                                  sir, sir.phi.lo, sir.phi.hi))
# Combine all tables into one with a "Category" column
smr_table_c <- bind_rows(
    smr_setting_phi_c %>% mutate(Category = "Setting") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print),
    smr_completed_phi_c %>% mutate(Category = "Treatment Completion") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print),
    smr_licit_phi_c %>% mutate(Category = "Primary Substance") %>% dplyr::select(Category, Group = group, Observed = observed, Expected = expected, SMR = print)#,
) %>% mutate(Category = ifelse( Category != lag(Category, default = ""),
            Category,
            ""))

# Create kable
kable(
    smr_table_c,
    col.names = c("Category", "Group", "Observed", "Expected", "SMR (95% CI)"),
    align = c("l", "l", "r", "r", "r"),
    booktabs = TRUE,
    linesep = ""
) 
```


### Heterogeneity

```{r bootstrap_phi_smr_and_het_c, results="asis"}
# --- 1) Run SMRs ----------------------------------------------------------
smr_setting_c   <- run_smr_by(c_SISTRAT_5y_c, mx_national_clean, "setting")
smr_completed_c <- run_smr_by(c_SISTRAT_5y_c, mx_national_clean, "completed")
smr_licit_c     <- run_smr_by(c_SISTRAT_5y_c, mx_national_clean, "licit_illicit")

# --- 2) Phi-corrected SMRs (profile CI) ----------------------------------
smr_setting_phi_c   <- make_smr_phi(smr_setting_c,   dsr_setting_c)
smr_completed_phi_c <- make_smr_phi(smr_completed_c, dsr_completed_c)
smr_licit_phi_c     <- make_smr_phi(smr_licit_c,     dsr_licit_c)

# --- 3) Heterogeneity helper (DSR + SMR) ---------------------------------
het_dsr <- function(dsr_df) {
  dsr_df |>
    dplyr::mutate(
      phi_used = if ("phi" %in% base::names(dsr_df)) phi else 1,
      yi  = base::log(rate.adj),
      sei = (SE.rate.adj * base::sqrt(phi_used)) / rate.adj
    ) |>
    metafor::rma(yi = yi, sei = sei, method = "FE", data = _)
}

het_smr <- function(smr_phi_df) {
  smr_phi_df |>
    dplyr::mutate(
      yi  = base::log(sir),
      sei = 0.5 * (
        (base::log(sir.phi.hi) - base::log(sir))/1.96 +
        (base::log(sir) - base::log(sir.phi.lo))/1.96
      )
    ) |>
    metafor::rma(yi = yi, sei = sei, method = "FE", data = _)
}

# --- 4) Fit meta models ---------------------------------------------------
meta_dsr_setting_c <- het_dsr(dsr_setting_c)
meta_smr_setting_c <- het_smr(smr_setting_phi_c)

meta_dsr_compl_c   <- het_dsr(dsr_completed_c)
meta_smr_compl_c   <- het_smr(smr_completed_phi_c)

meta_dsr_licit_c   <- het_dsr(dsr_licit_c)
meta_smr_licit_c   <- het_smr(smr_licit_phi_c)

# --- 5) Print quick heterogeneity table ----------------------------------

het_tbl_c <- base::rbind(
  het_row_strata(meta_dsr_setting_c, "Setting",       "DSR"),
  het_row_strata(meta_smr_setting_c, "Setting",       "SMR"),
  het_row_strata(meta_dsr_compl_c,   "Compliance",    "DSR"),
  het_row_strata(meta_smr_compl_c,   "Compliance",    "SMR"),
  het_row_strata(meta_dsr_licit_c,   "Licit/Illicit", "DSR"),
  het_row_strata(meta_smr_licit_c,   "Licit/Illicit", "SMR")
)

knitr::kable(
  het_tbl_c,
  caption = "Heterogeneity tests (fixed-effect meta-analysis): Cochran’s Q and I² (phi-corrected).",
  align = c("l","l","r","r","r","r")
)
```


# Session info

```{r session-info}
#| echo: true
#| error: true
#| message: true
#| paged.print: true

message(paste0("R library: ", Sys.getenv("R_LIBS_USER")))
message(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))

cat("quarto version: "); quarto::quarto_version()

sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
) %>% 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('R packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}")))

```

```{r session-info-python}
#| echo: true
#| error: true
#| message: true
#| paged.print: true
#| class-output: center-table
#| eval: true

cat("Python version\n")
reticulate::py_config()

reticulate::py_list_packages() %>% 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Python packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}"))) 

```

Save

```{r save-project}
#| echo: true
#| error: true
#| warning: true
#| message: true
#| paged.print: true
#| results: "hold"
#| eval: true
# 
# wdpath <- paste0(gsub("/cons","",gsub("cons","",paste0(getwd(),"/cons"))))
# paste0(getwd(),"/cons")
# file.path(paste0(wdpath,"data/20241015_out"))
# 
# # Save
# rdata_path <- file.path(wdpath, "data/20241015_out", paste0("mort2_", format(Sys.time(), "%Y_%m_%d"), ".Rdata"))
# 
# save.image(rdata_path)
# cat("Saved in:", rdata_path)
# 
# # Logic for saving history/encryption (from script)
# if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
#   # password <- Sys.getenv("PASSWORD_SECRET")
# } else {
#   if (interactive()) {
#     utils::savehistory(tempfile())
#     utils::loadhistory()
#   }
# }

```

```

```