---
title: "SENDAs Agreement 1 Update 2010-2022 (Prediction, step 2)"
description: |
  This step involves exploring the behavior of the predictors and best subset selection algorithms.
date: last-modified
author: "Andrés González Santa Cruz"
bibliography: [_lit/refs.bib]
csl: "_lit/american-medical-association.csl" # (Opcional: estilo de citas, p.ej., APA)
toc: true
toc-depth: 5
toc-location: left
toc-float: true
format: 
  html:
    css: [_lib/styles.scss]
    code-fold: true
    embed-resources: true
    theme: 
      - journal #https://quarto.org/docs/output-formats/html-themes-more.html
      #- style.scss
    fig-cap-location: top  
    toc: true
    toc-depth: 5
    toc-location: left
execute:
  error: true        # Show errors in the output
  echo: true         # Show R code in the output
  warning: true      # Show warnings
  message: true      # Show messages from the code
  freeze: auto       # Use cache to speed up rendering if there are no changes
editor: source
execute-dir: project
lightbox: auto
---

```{=html}
<style type="text/css">

.showopt {

background-color: #004c93; color: #FFFFFF; width: 100px; height: 20px; text-align: center; vertical-align: middle !important; float: right; font-family: sans-serif; border-radius: 8px;

}

.showopt:hover {
background-color: #dfe4f2;
color: #004c93;

}

pre.plot {
background-color: white !important;
}

.tablelines table, .tablelines td, .tablelines th {
border: 1px solid black;
}

.centrado {
text-align: center;
}

.table.center {
margin-left:auto;
margin-right:auto;
}

/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/

/* Applies to outputs that are not code other than R*/

pre {
overflow-x: auto !important;
}

pre code {
word-wrap: normal !important;
white-space: pre !important;
}

/*
pre:not(.sourceCode) {
white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence */
font-size: 10px !important;
line-height: 50% !important;
}
body{ /* Normal */
text-align: justify;
}
.superbigimage{
overflow-y:scroll;
height:350px;
white-space: nowrap;
overflow-x: auto;
width:100%;
}
.superbigimage img{
overflow-y: scroll;
overflow-x: hidden;
}
.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}
.pandoc-table { /* Should add !important; but it seems no necessary */
margin-left:auto; /* To center */
margin-right:auto;
border-collapse: collapse;
table-layout: auto;
font-size: 11px;
overflow-y: auto;
max-height:450px !important;
white-space: nowrap;
overflow-x: auto;
width:450px;
}
.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}
.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}
.pandoc-table caption {
text-align: left !important;
font-size: 11px !important;
}

.center-table {
text-align: left !important;
font-size: 9px;
overflow-y:scroll;
height:450px;
overflow-x: scroll;
}

.controlly{
overflow-y:scroll;
height:350px;
overflow-x: scroll;

}
/*2025-03-07*/
.callout-warning,
.callout-error,
.callout-message {
  font-size: 0.7rem !important;  /* Adjust this value as needed */
}

.alert, .message {
    font-size: 0.7em !important;
}

.alert.alert-warning,
.alert.alert-danger,
.message {
    font-size: 0.7em !important;
}
</style>
```
```{css, echo=F}
#| label: css-format

h1 {
    color: var(--heading-color);
    font-size: 2rem;
    margin-bottom: 1vh;
}

p {
  font-size: 1.1rem;
  line-height: 1.6rem;
}

a {
  color: var(--primary-color);
  text-decoration: none;
  border-bottom: 3px solid transparent;
  font-weight: bold;
  &:hover, &:focus {
      border-bottom: 3px solid currentColor;
  }
}

section {
  margin: 0 auto;
}

.post-meta {
  font-size: 1rem;
  font-style: italic;
  display: block;
  margin-bottom: 4vh;
  color: var(--secondary-color);
}

nav {
  display: flex;
  justify-content: flex-end;
  padding: 20px 0;
}

/*slider switch css */
.theme-switch-wrapper {
  display: flex;
  align-items: center;
  
  em {
    margin-left: 10px;
    font-size: 1rem;
  }
}
.theme-switch {
  display: inline-block;
  height: 34px;
  position: relative;
  width: 60px;
}

.theme-switch input {
  display:none;
}

.slider {
  background-color: #ccc;
  bottom: 0;
  cursor: pointer;
  left: 0;
  position: absolute;
  right: 0;
  top: 0;
  transition: .4s;
}

.slider:before {
  background-color: #fff;
  bottom: 4px;
  content: "";
  height: 26px;
  left: 4px;
  position: absolute;
  transition: .4s;
  width: 26px;
}

input:checked + .slider {
  background-color: #66bb6a;
}

input:checked + .slider:before {
  transform: translateX(26px);
}

.slider.round {
  border-radius: 34px;
}

.slider.round:before {
  border-radius: 50%;
}
```

```{=html}

<script>
function myFunction1() {
var x = document.getElementById("myDIV");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>

<script>
function myFunction2() {
var x = document.getElementById("myDIV2");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
```

<style>
.scrollable-content {
  max-height: 350px;
  overflow-y: auto;
}
pre.scrollable-code {
  max-height: 350px;
  overflow-y: auto;
}
.superbigimage {
  overflow-x: scroll;
  white-space: nowrap;
}
.superbigimage img, 
.superbigimage svg {
  max-width: none;
  height: auto;
}
</style>

<br>

# Data Loading and Exploration

## Loading Packages and uniting databases

<div class="scrollable-content">
```{r}
#| label: load
#| message: false
#| include: true
#| warning: false
#| results: "hold"

# invisible("Only run from Ubuntu")
# if (!(Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv"))) {
#   if(Sys.info()["sysname"]!="Windows"){
#     Sys.setenv(RETICULATE_PYTHON = "/home/fondecytacc/.pyenv/versions/3.11.5/bin/python")
#   }
# }

#clean enviroment
rm(list = ls()); gc()

time_before_dedup2<-Sys.time()

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

# Busca .mamba_root/envs/py311/python.exe desde getwd() hacia padres
find_python_rel <- function(start = getwd(),
                            rel = file.path(".mamba_root","envs","py311","python.exe")) {
  cur <- normalizePath(start, winslash = "/", mustWork = FALSE)
  repeat {
    cand <- normalizePath(file.path(cur, rel), winslash = "/", mustWork = FALSE)
    if (file.exists(cand)) return(cand)
    parent <- dirname(cur)
    if (identical(parent, cur)) return(NA_character_)  # llegó a la raíz
    cur <- parent
  }
}
# --- Bootstrap reticulate con ruta relativa a getwd() ---
if(Sys.info()["sysname"]!="Windows"){
  Sys.setenv(RETICULATE_PYTHON = "usr/bin/python3")
  reticulate::py_config()
} else {
  py <- find_python_rel()
  if (is.na(py)) {
    stop("No se encontró Python relativo a getwd() (buscando '.mamba_root/envs/py311/python.exe').\n",
         "Directorio actual: ", getwd())
  }
  # Forzar ese intérprete
  Sys.unsetenv(c("RETICULATE_CONDAENV","RETICULATE_PYTHON_FALLBACK"))
  Sys.setenv(RETICULATE_PYTHON = py)
  reticulate::use_python(py, required=T)
  reticulate::py_config()  # verificación
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# --- Load pred1_ndp_2026_01_05.Rdata (simple, project-root aware, no functions) ---

# 1) Locate the file anywhere under the project root (current WD with a trailing "/cons" trimmed)
project_root <- sub("(/)?cons/?$", "", normalizePath(getwd(), winslash = "/", mustWork = FALSE))
fname        <- "pred1_ndp_2026_01_05.Rdata"

hits <- list.files(project_root, pattern = paste0("^", fname, "$"),
                   recursive = TRUE, full.names = TRUE, ignore.case = TRUE)

if (!length(hits)) stop("File not found: ", fname, "\nSearched under: ", project_root)
path <- hits[1]  # take first match

message("Loading: ", path)

# 2) Load into a dedicated environment (keeps .GlobalEnv clean)
env_pred <- new.env(parent = emptyenv())
loaded_objects <- load(path, envir = env_pred)

message("Loaded ", length(loaded_objects), " object(s): ", paste(loaded_objects, collapse = ", "))

# 3) Quick sanity check for each loaded object
for (obj_name in loaded_objects) {
  obj <- get(obj_name, envir = env_pred)
  message("\n--- ", obj_name, " ---")
  if (is.data.frame(obj)) {
    message("Rows: ", nrow(obj), " | Cols: ", ncol(obj))
  } else {
    message("Class: ", paste(class(obj), collapse = ", "), " | Length/Size: ", length(obj))
  }
  #print(str(obj))
}

# 4) Extract all loaded objects to global environment
for (obj_name in loaded_objects) {
  assign(obj_name, get(obj_name, envir = env_pred), envir = .GlobalEnv)
}

# Clean up temporary environment
rm(env_pred)
# (Optional) If you want it in the global env, uncomment:
# list2env(env_prev1y, envir = .GlobalEnv)
```
</div>

<div class="scrollable-content">
```{r}
#| label: setup
#| results: "hold"
#https://github.com/rstudio/renv/issues/544
#renv falls back to copying rather than symlinking, which is evidently very slow in this configuration.
renv::settings$use.cache(FALSE)

#only use explicit dependencies (in DESCRIPTION)
renv::settings$snapshot.type("implicit")

#check if rstools is installed
if(Sys.info()["sysname"]=="Windows"){
try(installr::install.Rtools(check_r_update=F))
}

check_quarto_version <- function(required = "1.7.29", comparator = c("ge","gt","le","lt","eq")) {
  comparator <- match.arg(comparator)
  current <- package_version(paste(unlist(quarto::quarto_version()), collapse = "."))
  req     <- package_version(required)

  ok <- switch(comparator,
               ge = current >= req,
               gt = current >  req,
               le = current <= req,
               lt = current <  req,
               eq = current == req)

  if (!ok) {
    stop(sprintf("Quarto version check failed: need %s %s (installed: %s).",
                 comparator, required, current), call. = FALSE)
  }
  invisible(TRUE)
}

check_quarto_version("1.7.29", "ge") 

#change repository to CL
local({
  r <- getOption("repos")
  r["CRAN"] <- "https://cran.dcc.uchile.cl/"
  options(repos=r)
})

if(!require(pacman)){install.packages("pacman");require(pacman)}
if(!require(pak)){install.packages("pak");require(pak)}

pacman::p_unlock(lib.loc = .libPaths()) #para no tener problemas reinstalando paquetes

if(Sys.info()["sysname"]=="Windows"){
if (getRversion() != "4.4.1") { stop("Requires R version 4.4.1; Actual: ", getRversion()) }
}

#check docker
check_docker_running <- function() {
  # Try running 'docker info' to check if Docker is running
  system("docker info", intern = TRUE, ignore.stderr = TRUE)
}
if(Sys.info()["sysname"]=="Windows"){
  install_docker <- function() {
    # Open the Docker Desktop download page in the browser for installation
    browseURL("https://www.docker.com/products/docker-desktop")
  }
  # Main logic
  if (inherits(try(check_docker_running(), silent = TRUE), "try-error")) {
    liftr::install_docker()
  } else {
    message("Docker is running.")
  }
}

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#PACKAGES#######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

unlink("*_cache", recursive=T)

# ----------------------------------------------------------------------
# 2. Use a single pak::pkg_install() call for most CRAN packages
# ----------------------------------------------------------------------

paks <-
  c(#"git", 
    # To connect to github
    "gh", #interface for  GitHub API from R
    #
    "gitcreds", # manages Git credentials (usernames, passwords, tokens)
    #
    "usethis", # simplifies common project setup tasks for R developers
    # Package to bring packages in development
    "devtools",
    # Package administration
    "renv",
    # To manipulate data
    "knitr", "pander", "DT",
    # Join
    "fuzzyjoin", "RecordLinkage",
    # For tables
    "tidyverse", "janitor",
    # For contingency tables
    "kableExtra",
    # For connections with python
    "reticulate",
    # To manipulate big data
    "polars", "sqldf",
    # To bring big databases
    "nanoparquet",
    # Interface for R and RStudio in R
    "installr", "rmarkdown", "quarto", "yaml", #"rstudioapi",
    # Time handling
    "clock",
    # Combine plots
    "ggpubr",
    # Parallelized iterative processing
    "furrr",
    # Work like a tibble with a data.table database
    "tidytable",
    # Split database into training and testing
    "caret",
    # Impute missing data
    "missRanger", "mice",
    # To modularize tasks
    "job",
    # For PhantomJS install checks
    "webshot"
  )

# dplyr
# janitor
# reshape2
# tidytable
# arrow
# boot
# broom
# car
# caret
# data.table
# DiagrammeR
# DiagrammeRsvg
# dplyr
# epiR
# epitools
# ggplot2
# glue
# htmlwidgets
# knitr
# lubridate
# naniar
# parallel
# polycor
# pROC
# psych
# readr
# rio
# rsvg
# scales
# stringr
# tableone
# rmarkdown
# biostat3
# codebook
# finalfit
# Hmisc
# kableExtra
# knitr
# devtools
# tidyr
# stringi
# stringr
# muhaz
# sqldf
# compareGroups
# survminer
# lubridate
# ggfortify
# car
# fuzzyjoin
# compareGroups
# caret
# job
# htmltools
# nanoparquet
# ggpubr
# polars
# installr
# clock
# pander
# reshape
# mice
# missRanger
# VIM
# withr
# biostat3
# broom
# glue
# finalfit
# purrr
# sf


# pak::pkg_install(paks)

pak::pak_sitrep()
# pak::sysreqs_check_installed(unique(unlist(paks)))
#pak::lockfile_create(unique(unlist(paks)),  "dependencies_duplicates24.lock", dependencies=T)
#pak::lockfile_install("dependencies_duplicates24.lock")
#https://rdrr.io/cran/pak/man/faq.html
#pak::cache_delete()

library(tidytable)
library(ggplot2)
library(readr)
library(tableone)
library(survivalmodels)
#renv::install("patchwork@1.2.0")
library(survex)
library(SemiMarkov)
library(flexsurv)
library(rms)
library(survidm)
library(caret)

# library(shapr)#https://norskregnesentral.github.io/shapr/
# library(survidm)#https://sci-hub.ee/10.1002/bimj.201700200
# library(SemiMarkov)#https://www.degruyterbrill.com/document/doi/10.1515/ijb-2020-0083/html?lang=en
# #https://www.jclinepi.com/article/S0895-4356(24)00142-2/fulltext

# ----------------------------------------------------------------------
# 3. Activate polars code completion (safe to try even if it fails)
# ----------------------------------------------------------------------
#try(polars_code_completion_activate())

# ----------------------------------------------------------------------
# 4. BPMN from GitHub (not on CRAN, so install via devtools if missing)
# ----------------------------------------------------------------------
if (!requireNamespace("bpmn", quietly = TRUE)) {
  devtools::install_github("bergant/bpmn")
}

# ----------------------------------------------------------------------
# 5. PhantomJS Check (use webshot if PhantomJS is missing)
# ----------------------------------------------------------------------
# if (!webshot::is_phantomjs_installed()) {
#   webshot::install_phantomjs()
# }

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#FUNCTIONS######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
# NO MORE DEBUGS
options(error = NULL)        # si antes tenías options(error = recover) o browser)
options(browserNLdisabled = FALSE)


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#NAs are replaced with "" in knitr kable
options(knitr.kable.NA = '')

pander::panderOptions('big.mark', ',')
pander::panderOptions('decimal.mark', '.')

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#

#to format rows in bold
format_cells <- function(df, rows ,cols, value = c("italics", "bold", "strikethrough")){

  # select the correct markup
  # one * for italics, two ** for bold
  map <- setNames(c("*", "**", "~~"), c("italics", "bold", "strikethrough"))
  markup <- map[value]  

  for (r in rows){
    for(c in cols){

      # Make sure values are not factors
      df[[c]] <- as.character( df[[c]])

      # Update formatting
      df[r, c] <- ifelse(nchar(df[r, c])==0,"",paste0(markup, gsub(" ", "", df[r, c]), markup))
    }
  }

  return(df)
}
#To produce line breaks in messages and warnings
knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger" style="font-size: small !important;">',
           gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning" style="font-size: small !important;">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('<div class="message" style="font-size: small !important;">',
           gsub('##', '\n', x),
           '</div>', sep = '\n')
   }
)

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#

sum_dates <- function(x){
  cbind.data.frame(
    min= as.Date(min(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01"),
    p001= as.Date(quantile(unclass(as.Date(x)), .001, na.rm=T), origin = "1970-01-01"),
    p005= as.Date(quantile(unclass(as.Date(x)), .005, na.rm=T), origin = "1970-01-01"),
    p025= as.Date(quantile(unclass(as.Date(x)), .025, na.rm=T), origin = "1970-01-01"),
    p25= as.Date(quantile(unclass(as.Date(x)), .25, na.rm=T), origin = "1970-01-01"),
    p50= as.Date(quantile(unclass(as.Date(x)), .5, na.rm=T), origin = "1970-01-01"),
    p75= as.Date(quantile(unclass(as.Date(x)), .75, na.rm=T), origin = "1970-01-01"),
    p975= as.Date(quantile(unclass(as.Date(x)), .975, na.rm=T), origin = "1970-01-01"),
    p995= as.Date(quantile(unclass(as.Date(x)), .995, na.rm=T), origin = "1970-01-01"),
    p999= as.Date(quantile(unclass(as.Date(x)), .999, na.rm=T), origin = "1970-01-01"),
    max= as.Date(max(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01")
  )
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

# Define the function adapted for Polars
sum_dates_polars <- function(df, date_col) {
  # Create the list of quantiles
  quantiles <- c(0.001, 0.005, 0.025, 0.25, 0.5, 0.75, 0.975, 0.995, 0.999)
  # Create expressions to calculate min and max
  expr_list <- list(
    pl$col(date_col)$min()$alias("min"),
    pl$col(date_col)$max()$alias("max")
  )
  # Add expressions for quantiles
  for (q in quantiles) {
    expr_list <- append(expr_list, pl$col(date_col)$quantile(q)$alias(paste0("p", sub("\\.", "", as.character(q)))))
  }
  # Apply the expressions and return a DataFrame with the results
  df$select(expr_list)
}

# Custom function for sampling with a seed
sample_n_with_seed <- function(data, size, seed) {
  set.seed(seed)
  dplyr::sample_n(data, size)
}

# Function to get the most frequent value 
most_frequent <- function(x) { 
  uniq_vals <- unique(x)
  freq_vals <- sapply(uniq_vals, function(val) sum(x == val))
  most_freq <- uniq_vals[which(freq_vals == max(freq_vals))]
  
  if (length(most_freq) == 1) {
    return(most_freq)
  } else {
    return(NA)
  }
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#CONFIG #######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

options(scipen=2) #display numbers rather scientific number

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

# Define the function first
#joins these values with semicolons and optionally truncates the result if it exceeds a specified width.
toString2 <- function(x, width = NULL, ...) {
    string <- paste(x, collapse = "; ")
    if (missing(width) || is.null(width) || width == 0) 
        return(string)
    if (width < 0) 
        stop("'width' must be positive")
    if (nchar(string, type = "w") > width) {
        width <- max(6, width)
        string <- paste0(substr(string, 1, width - 3), "...")
    }
    string
}
normalize_txt <- function(x) {
  x|>
    stringi::stri_trans_general("Latin-ASCII")|>
    tolower()|>
    trimws()
}

pair_str <- function(main, sub) {
  main <- as.character(main)
  sub  <- as.character(sub)
  both_na <- is.na(main) & is.na(sub)

  out <- paste0(
    tidytable::coalesce(main, "NA"),
    "::",
    tidytable::coalesce(sub,  "NA")
  )

  out[both_na] <- NA_character_
  out
}

# ── Helpers ────────────────────────────────────────────────────────────
mode_pick_int <- function(x){
  x <- x[!is.na(x)]
  if(length(x)==0) return(NA_integer_)
  tx <- sort(table(x), decreasing = TRUE)
  as.integer(names(tx)[1L])
}

subkey_to_label <- function(x){
  y <- gsub("_"," ", tolower(x))
  y <- gsub("amphetamine type stimulants","amphetamine-type stimulants", y)
  y <- gsub("tranquilizers hypnotics","tranquilizers/hypnotics", y)
  y
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# ── Encryption ────────────────────────────────────────────────────────

# Pack: [24-byte nonce][ciphertext] -> hex
encrypt_chr <- function(x, key) {
    vapply(x, function(v) {
        if (is.na(v)) return(NA_character_)
        ct <- sodium::data_encrypt(charToRaw(as.character(v)), key)
        nonce <- attr(ct, "nonce")                 # 24 bytes
        sodium::bin2hex(c(nonce, ct))              # pack nonce + ct
    }, FUN.VALUE = character(1))
}

# Unpack: hex -> [nonce|ciphertext], restore attr(nonce), then decrypt
decrypt_chr <- function(x, key) {
    vapply(x, function(v) {
        if (is.na(v)) return(NA_character_)
        buf <- sodium::hex2bin(v)
        if (length(buf) < 25L) stop("Ciphertext too short or corrupted.")
        nonce <- buf[1:24]
        ct    <- buf[-(1:24)]
        attr(ct, "nonce") <- nonce
        rawToChar(sodium::data_decrypt(ct, key))
    }, FUN.VALUE = character(1))
}

# Function to encrypt a character vector
encrypt_column <- function(x, key) {
  sapply(x, function(item) {
    if (is.na(item) || item == "") {
      return(NA_character_)
    }
    encrypted_raw <- sodium::data_encrypt(charToRaw(item), key)
    base64enc::base64encode(encrypted_raw)  # Convert to base64 for storage
  }, USE.NAMES = FALSE)
}

decrypt_column <- function(x, key) {
  sapply(x, function(item) {
    if (is.na(item)) return(NA_character_)
    encrypted_raw <- base64enc::base64decode(item)
    rawToChar(sodium::data_decrypt(encrypted_raw, key))
  }, USE.NAMES = FALSE)
}


is_stata_ok <- function(x) {
  nchar(x) <= 32 & grepl("^[A-Za-z][A-Za-z0-9_]*$", x)
}


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# ── Tidy RMS ────────────────────────────────────────────────────────

tidy_cph <- function(model) {
  if (!inherits(model, "cph")) stop("Model must be an rms::cph object")
  
  # run summary
  s <- summary(model)
  df_s <- as.data.frame(unclass(s))
  df_s$rown <- rownames(s)
  
  # find rows that are "Hazard Ratio"
  hr_rows <- trimws(df_s$rown) == "Hazard Ratio"
  
  # build tidy tibble
  tibble::tibble(
    term     = trimws(rownames(s)[which(hr_rows) - 1L]), # previous row = variable name
    estimate = df_s$Effect[hr_rows],                     # Hazard Ratio
    conf.low = df_s$`Lower 0.95`[hr_rows],
    conf.high= df_s$`Upper 0.95`[hr_rows],
    p.value  = df_s$P[hr_rows]
  )
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# --- Step 1: Define File Paths ---

# SET THIS to the correct path of your secret text file
secret_file_path <- "C:/Users/andre/Documents/secret.txt"
# This locates the user-level .Renviron file automatically
renviron_path <- file.path(Sys.getenv("HOME"), ".Renviron")
# --- Step 2: Read the Secret Key from the .txt file ---
# Read the first line of the file and trim any whitespace
api_key <- trimws(readLines(secret_file_path, n = 1, warn = FALSE))
# --- Step 3: Check if the key is already in .Renviron ---
# Read all existing lines from .Renviron (if it exists)
if (file.exists(renviron_path)) {
  existing_lines <- readLines(renviron_path, warn = FALSE)
} else {
  existing_lines <- c()
}
# Check if a line for OPENAI_API_KEY already exists
key_exists <- any(grepl("^OPENAI_API_KEY=", existing_lines))
# --- Step 4: Write to .Renviron if the key isn't already there ---
if (!key_exists) {
  # Format the line to be added
  line_to_add <- sprintf('OPENAI_API_KEY="%s"', api_key)
  # Append the line to the .Renviron file, adding a newline before it if the file isn't empty
  cat(line_to_add, "\n", file = renviron_path, append = TRUE)
  message("✅ Success! OPENAI_API_KEY has been added to your .Renviron file.")
  message("Please restart your R session for the change to take effect.")
} else {
  message("ℹ️ The OPENAI_API_KEY already exists in your .Renviron file. No changes were made.")
}
options(.gander_chat = ellmer::chat_openai(model = "o1-mini"))
```
</div>

<br>

::: {.callout-note}
**Aims**

- To identify key prognostic and predictive factors associated with the risk of, and time to, the first readmission to treatment among patients aged 18-64 undergoing initial treatment for Substance Use Disorder (SUD) in Chile between 2010 and 2020.

- To identify key prognostic and predictive factors associated with the risk of, and time to, all-cause mortality among the same cohort of patients undergoing initial treatment for SUD in Chile.

- To predict the risk and timing of first treatment readmission and all-cause mortality simultaneously, treating these outcomes within a semi-competing risks framework. This will allow for the estimation of cumulative incidence for each event in the presence of the other for patients aged 18-64 who initiated SUD treatment in Chile between 2010 and 2020.

This study has three primary objectives. First, we aim to identify key prognostic and predictive factors associated with the risk and timing of two critical outcomes: (i) first readmission to treatment and (ii) all-cause mortality. Second, using a semi-competing risks framework, we will develop a model to simultaneously predict these two outcomes, accounting for the fact that death precludes the possibility of readmission. Finally, this model will be used to estimate the probability of various patient trajectories over time following initial treatment. The study population includes all patients aged 18-64 who began treatment for Substance Use Disorder (SUD) in Chile between 2010 and 2020.
:::

<br>

## 0. Prepare data

In this step, we imported the center identifiers and then applied a correction procedure across all imputed datasets to ensure consistency in event-time definitions. Specifically, we converted time and event variables to numeric and integer formats, identified cases where death occurred before or at the readmission endpoint (competing risks), and adjusted follow‑up times and event indicators accordingly by censoring readmission in those situations. After cleaning up temporary variables, we added the center ID as a factor to each dataset. We also excluded `high_uncertainty_imps`, which encompassed predictors with were not reliable. Finally, we exported the corrected datasets both individually as `.parquet` files named as `corrected_datasets_nondum` (nondum= not-dummified) and collectively as a single `.Rds` object, creating a clean and reproducible output ready for downstream analyses.

```{r}
#| label: prepare-data
#| echo: true
#| error: false
#| warning: true
#| message: true
#| paged.print: true
#| results: "hold"
#| eval: true

id_centro_vec<- 
rio::import(paste0(file.path(wdpath, "data/20241015_out/pred1"),"/id_centro.Rds"))

# 1. Apply Correction to Each Dataset in the List
corrected_datasets_nondum <- map(filled_datasets, function(df) {
  # Ensure columns are numeric for comparison
  df <- df|> 
    mutate(
      readmit_time = as.numeric(readmit_time_from_disch_m),
      death_time   = as.numeric(death_time_from_disch_m),
      # Ensure events are 0/1 integers
      readmit_event = as.integer(as.character(readmit_event)), 
      death_event   = as.integer(as.character(death_event))
    )
  # --- THE CORRECTION LOGIC ---
  df <- df|>
    dplyr::mutate(
      # condition: Patient died (death_event==1) AND death happened before (or at) the readmission endpoint
      is_competing_risk = (death_event == 1) & (death_time <= readmit_time),
      # 1. Update Time: If competing risk, cut follow-up at death time. Otherwise keep original.
      readmit_time_from_disch_m = dplyr::if_else(is_competing_risk, death_time, readmit_time),
      # 2. Update Event: If competing risk, censor readmission (set to 0). Otherwise keep original.
      readmit_event = dplyr::if_else(is_competing_risk, 0, readmit_event)
    )|>
    # Remove temporary helper columns if you want to keep the dataset clean
    dplyr::select(-readmit_time, -death_time, -is_competing_risk)|>
    dplyr::select(any_of(
      setdiff(c(num_vars, cat_vars, "readmit_event", "death_event", "readmit_time_from_disch_m","death_time_from_disch_m"), high_uncertainty_imps)
    ))|>
    dplyr::mutate(center_id= factor(id_centro_vec))
  return(df)
})
```


```{r, eval=F}
#| label: prepare-data-exp

# 2. Iterate and Export
# iwalk automatically gives you the object (.x) and its index (.y)
purrr::iwalk(corrected_datasets_nondum, function(df, i) {
  # Construct the full file path
  # e.g. .../data/20241015_out/imputation_1.parquet
  file_name <- file.path(output_dir, paste0("imputation_nondum_", i, ".parquet"))
  # Export using rio
  rio::export(df, file_name)
  message(paste("Saved:", file_name))
})

#Export list
rio::export(corrected_datasets_nondum, paste0(file.path(wdpath, "data/20241015_out/pred1"),"/processed_datasets_nondum.Rds"))
#corrected_datasets_nondum<- rio::import(paste0(file.path(wdpath, "data/20241015_out/pred1"),"/processed_datasets_nondum.Rds"))
```

# Split database

Assuming that `corrected_datasets_nondum` has not been modified since its last update on February 21, the order of the relevant rows in the dataset should remain unchanged (n=88,504). Therefore, the train–test split should be identical to the one used in Python. For this reason, we loaded `death_split_seed2125_test20.parquet` to partition observations into training and test sets.

```{r}
#| label: import-parquet
split2125<- 
nanoparquet::read_parquet(paste0(getwd(), "/cons/_out/comb_split_seed2125_test20.parquet"))
```

We generated a defensive reproducibility check with automatic triggers that stop execution if something changed of the 80/20 split (row order, split mismatch, unexpected counts, etc.).

```{r}
#| label: format-python-split

df <- cbind.data.frame(corrected_datasets_nondum[[1]], split2125)|> 
  dplyr::filter(is_train==T)

message(paste0("The training database is composed of ", nrow(df)," observations"))

expected_n_train <- 70803  # replace with known expected value

if (nrow(df) != expected_n_train) {
  stop("❌ Inconsistency detected: Training sample size does not match expected value.")
} else {
  message("✔ Training sample size verified.")
}

expected_foreign_n <- 462

observed_foreign_n <- df |>
  dplyr::summarise(n = sum(national_foreign == 1, na.rm = TRUE)) |>
  dplyr::pull(n)

if (observed_foreign_n != expected_foreign_n) {
  stop(paste0("❌ Inconsistency detected: national_foreign count is ",
              observed_foreign_n,
              " but expected ",
              expected_foreign_n))
} else {
  message("✔ national_foreign count verified.")
}


expected_age_mean <- 35.79

if (abs(mean(df$adm_age_rec3, na.rm = TRUE) - expected_age_mean) > 0.01) {
  stop("❌ Age distribution changed — possible split mismatch.")
}
```

```{r}
#| label: format-db-to-python
get_mode_numeric <- function(x) {
  x_no_na <- x[!base::is.na(x)]
  if (base::length(x_no_na) == 0) return(NA_real_)
  tab <- base::table(x_no_na)
  base::as.numeric(base::names(tab)[base::which.max(tab)])
}

preprocess_features_robust <- function(
  df,
  recode_first_sub_used_to_other = FALSE,  # TRUE solo para dataset 1
  do_dummies = TRUE
) {
  # Fuerza data.frame para evitar errores de data.table en subsetting
  df_clean <- df |> base::as.data.frame(stringsAsFactors = FALSE)|> 
    cbind.data.frame(split2125)|> 
    dplyr::filter(is_train==T)|> 
    #dplyr::select(- center_id)
    dplyr::mutate(center_id = suppressWarnings(as.numeric(center_id)))

  get_mode_numeric <- function(x) {
    x_no_na <- x[!base::is.na(x)]
    if (base::length(x_no_na) == 0) return(NA_real_)
    tab <- base::table(x_no_na)
    base::as.numeric(base::names(tab)[base::which.max(tab)])
  }

  # Recode opcional (solo imputation_nodum_1 / dataset 1)
  if (isTRUE(recode_first_sub_used_to_other) && "first_sub_used" %in% base::names(df_clean)) {
    to_other <- c(
      "hallucinogens",
      "opioids",
      "amphetamine-type stimulants",
      "tranquilizers/hypnotics",
      "inhalants",
      "others"
    )

    df_clean[["first_sub_used_original"]] <- df_clean[["first_sub_used"]]

    s <- df_clean[["first_sub_used"]]|>
      base::as.character() |>
      base::trimws() |>
      base::tolower()

    s[s %in% to_other] <- "other"
    df_clean[["first_sub_used"]] <- base::factor(s)
  }

  # 1) Ordinal encoding
  ordered_mappings <- list(
    tenure_status_household = c(
      "illegal settlement" = 4,
      "stays temporarily with a relative" = 3,
      "others" = 2,
      "renting" = 1,
      "owner/transferred dwellings/pays dividends" = 0
    ),
    urbanicity_cat = c(
      "1.Rural" = 2,
      "2.Mixed" = 1,
      "3.Urban" = 0
    ),
    evaluacindelprocesoteraputico = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_consumo = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_fam = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_relinterp = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_ocupacion = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_sm = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_fisica = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    eva_transgnorma = c("logro minimo" = 2, "logro intermedio" = 1, "logro alto" = 0),
    prim_sub_freq_rec = c(
      "1.≤1 day/wk" = 0,
      "2.2–6 days/wk" = 1,
      "3.Daily" = 2
    ),
    ed_attainment_corr = c(
      "3-Completed primary school or less" = 2,
      "2-Completed high school or less" = 1,
      "1-Completed higher education" = 0
    )
  )

  for (col in base::names(ordered_mappings)) {
    if (col %in% base::names(df_clean)) {
      values <- df_clean[[col]] |>
        base::as.character() |>
        base::trimws()

      mapped <- ordered_mappings[[col]][values] |>
        base::unname() |>
        base::as.numeric()

      n_missing <- base::sum(base::is.na(mapped))
      if (n_missing > 0) {
        if (n_missing == base::nrow(df_clean)) {
          base::message(base::sprintf("WARNING: Mapping failed completely for '%s'.", col))
        }
        mode_val <- get_mode_numeric(mapped)
        if (!base::is.na(mode_val)) {
          mapped[base::is.na(mapped)] <- mode_val
        }
      }

      df_clean[[col]] <- mapped
    }
  }

  # 2) Fijar categoría de referencia
  dummy_reference <- c(
    sex_rec = "man",
    plan_type_corr = "pg-pab",
    marital_status_rec = "married/cohabiting",
    cohabitation = "alone",
    sub_dep_icd10_status = "hazardous consumption",
    tr_outcome = "completion",
    adm_motive = "spontaneous consultation",
    tipo_de_vivienda_rec2 = "formal housing",
    occupation_condition_corr24 = "employed",
    any_violence = "0.No domestic violence/sex abuse",
    first_sub_used = "marijuana"
  )

  for (col in base::names(dummy_reference)) {
    if (col %in% base::names(df_clean)) {
      values <- df_clean[[col]] |>
        base::as.character() |>
        base::trimws()

      ref <- dummy_reference[[col]]
      cats <- base::unique(values)

      if (ref %in% cats) {
        levs <- c(ref, cats[cats != ref])
        df_clean[[col]] <- base::factor(values, levels = levs, ordered = FALSE)
      } else {
        base::message(base::sprintf("WARNING: Reference '%s' not found in %s", ref, col))
        df_clean[[col]] <- values
      }
    }
  }

  if (!isTRUE(do_dummies)) {
    return(df_clean)
  }

  # 3) One-hot (k-1, equivalente a drop_first=True)
  cat_cols <- base::names(df_clean)[
    base::vapply(
      df_clean,
      FUN = function(x) base::is.character(x) || base::is.factor(x),
      FUN.VALUE = logical(1)
    )
  ]

  if (base::length(cat_cols) == 0) {
    return(df_clean)
  }

  df_cat <- df_clean |>
    dplyr::mutate(dplyr::across(dplyr::all_of(cat_cols), base::as.factor))

  cat_df <- df_cat |>
    dplyr::select(dplyr::all_of(cat_cols)) |>
    base::as.data.frame(stringsAsFactors = FALSE)

  # Critical fix: keep all rows (no implicit row drop in model.matrix)
  cat_df[] <- lapply(cat_df, function(x) {
    x <- base::as.factor(x)
    x <- base::addNA(x, ifany = TRUE)
    lev <- base::levels(x)
    lev[base::is.na(lev)] <- "__MISSING__"
    base::levels(x) <- lev
    x
  })

  mm <- stats::model.matrix(
    stats::as.formula("~ ."),
    data = cat_df
  ) |>
    base::as.data.frame(check.names = FALSE)

  if ("(Intercept)" %in% base::names(mm)) {
    mm <- dplyr::select(mm, -`(Intercept)`)
  }

  non_cat_df <- df_cat |>
    dplyr::select(-dplyr::all_of(cat_cols))

  dplyr::bind_cols(non_cat_df, mm)
}

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# Process datasets [[1]]..[[5]] and replace in py_corrected_datasets_nondum
#The Quarto LSP server crashed 5 times in the last 3 minutes. The server will not be restarted. See the output for more information.
idx <- base::seq_len(base::min(5, base::length(corrected_datasets_nondum)))

py_corrected_datasets_nondum <- list()
py_corrected_datasets_nondum[idx] <- 
  corrected_datasets_nondum[idx] |> 
  purrr::imap(
    \(x, i) {
      x |>
        preprocess_features_robust(
          recode_first_sub_used_to_other = T,#(i == 1), # only dataset 1
          do_dummies = TRUE
        ) |>
        janitor::clean_names()
    }
  )

```

# Proportionallity

```{r}
#| label: nonprop-def-models2

# Define the list of predictors based on your provided list
predictors <- c(
  "adm_age_rec3", "porc_pobr", "dit_m", "sex_rec", "tenure_status_household",
  "cohabitation", "sub_dep_icd10_status", "any_violence",
  "prim_sub_freq_rec","tr_outcome", "adm_motive", "first_sub_used",
  "primary_sub_mod", "tipo_de_vivienda_rec2", "national_foreign", 
  "occupation_condition_corr24", "marital_status_rec", "urbanicity_cat", 
  "ed_attainment_corr", "evaluacindelprocesoteraputico", "eva_consumo", 
  "eva_fam", "eva_relinterp", "eva_ocupacion", "eva_sm", "eva_fisica", 
  "eva_transgnorma", "ethnicity", "dg_psiq_cie_10_instudy", 
  "dg_psiq_cie_10_dg", "dx_f3_mood", "dx_f6_personality", 
  "dx_f_any_severe_mental", "any_phys_dx", "polysubstance_strict"
)

# Create the formulas
formula_readmit <- as.formula(paste("Surv(readmit_time_from_disch_m, readmit_event) ~", paste(predictors, collapse = " + "),"+ strata(plan_type_corr)"))

formula_death <- as.formula(paste("Surv(death_time_from_disch_m, death_event) ~", paste(predictors, collapse = " + "),"+ strata(plan_type_corr)"))
```


### Residuals, readmission

```{r}
#| label: nonprop2-test-coxzph-readm

# --- READMISSION MODEL ---
cox_readmit <- coxph(formula_readmit, data = df)
ph_test_readmit <- cox.zph(cox_readmit)

ph_test_readmit$table |>
  data.table::data.table(keep.rownames = TRUE) |>
  dplyr::arrange(desc(chisq)) |>
  dplyr::mutate(
    p_fmt = ifelse(p < 0.001, "<0.001", sprintf("%.3f", p)),
    chi_fmt = sprintf("χ²(%d)= %.2f, p%s", df, chisq, ifelse(p < 0.001, "<0.001", paste0("=", p_fmt)))) |>
  dplyr::select(covariable = rn, `χ²(gl), p` = chi_fmt)|>
  knitr::kable(caption="Schofeld residuals, readmission")


# 1. Identify violators (you already have this part)
violations_readm <- ph_test_readmit$table |> 
  data.table::data.table(keep.rownames = T) |> 
  dplyr::mutate(p = as.numeric(p)) |> 
  dplyr::filter(p < 0.05) |> 
  dplyr::pull(rn) |> 
  setdiff("GLOBAL")

# 2. Generate the combined plot
# ggcoxzph creates a list of plots if you pass multiple variables
combined_plots <- survminer::ggcoxzph(ph_test_readmit, var = violations_readm)

# Open a PDF device 
pdf("_figs/ph_test_violations_readm.pdf", width = 16*1.5, height = 12*1.5) 
# Draw the combined plot 
print(combined_plots) 
# Close the device 
dev.off()

png(
  filename = "_figs/ph_test_violations_readm.png",
  width  = 16 * 1.5,
  height = 12 * 1.5,
  units  = "in",
  res    = 300
)
print(combined_plots) 

dev.off()
```

### Residuals, death

```{r}
#| label: nonprop3-test-coxzph

library(survminer)

# --- DEATH MODEL ---
cox_death <- coxph(formula_death, data = df)
ph_test_death <- cox.zph(cox_death)

ph_test_death$table|>
  data.table::data.table(keep.rownames = TRUE)|>
  dplyr::arrange(desc(chisq))|>
  dplyr::mutate(
    p_fmt = ifelse(p < 0.001, "<0.001", sprintf("%.3f", p)),
    chi_fmt = sprintf("χ²(%d)= %.2f, p%s", df, chisq, ifelse(p < 0.001, "<0.001", paste0("=", p_fmt))))|>
  dplyr::select(covariable = rn, `χ²(gl), p` = chi_fmt)|>
  knitr::kable(caption="Schofeld residuals, death")

# 1. Identify violators (you already have this part)
violations <- ph_test_death$table|> 
  data.table::data.table(keep.rownames = T)|> 
  dplyr::mutate(p = as.numeric(p))|> 
  dplyr::filter(p < 0.05)|> 
  dplyr::pull(rn)|> 
  setdiff("GLOBAL")

# 2. Generate the combined plot
# ggcoxzph creates a list of plots if you pass multiple variables
combined_plots <- survminer::ggcoxzph(ph_test_death, var = violations)

# Open a PDF device 
pdf("_figs/ph_test_violations_death.pdf", width = 16*1.5, height = 12*1.5) 
# Draw the combined plot 
print(combined_plots) 
# Close the device 
dev.off()
png(
  filename = "_figs/ph_test_violations_death.png",
  width  = 16 * 1.5,
  height = 12 * 1.5,
  units  = "in",
  res    = 300
)
print(combined_plots) 

dev.off()
```

We should:

 - Stratify `tr_outcome` (Treat the baseline hazard differently for each outcome) for both outcomes

- Stratify by Admission motive (`adm_motive`) have very different risk profiles that evolve differently over time, for time to readmission.

- Add time-interaction in Days in treatment (`dit_m`) has a decaying association ([ function(x, t, ...) x * log(t + 1) ]) for time to readmission.

-  Add time-interaction for `any_phys_dx` (Model the decaying association [ function(x, t, ...) x * log(t + 1) ]) for time to death.

```{r}
#| label: nonprop5-log-log

message("We had to handle ties")

fit_stratified <- survfit(Surv(readmit_time_from_disch_m, readmit_event) ~ tr_outcome, data = dplyr::filter(df, tr_outcome!="other")|> 
  dplyr::mutate(readmit_time_from_disch_m =  ifelse(readmit_time_from_disch_m == 0, 0.001, readmit_time_from_disch_m)))

p <- ggsurvplot(fit_stratified, 
           fun = "cloglog", 
           conf.int = TRUE,
           title = "LML Plot for Treatment Outcome (Readmission)",
           xlab = "log(Time from discharge)",
           ylab = "log(-log(Survival))",
           # Provide clean names here in the correct order
           legend.labs = c("Adm Reasons", "Rule Violation", "Completion", "Dropout", "Referral"),
           legend.title = "Treatment Outcome"
)
p$plot <- p$plot +
  geom_point(
    aes(shape = strata, color = strata),
    size = 2
  ) +
  scale_color_manual(
    values = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
    name = "Treatment\nOutcome",
    labels = c("Adm Reasons", "Rule Violation", "Completion", "Dropout", "Referral")
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3), 
    name = "Treatment\nOutcome",
    labels = c("Adm Reasons", "Rule Violation", "Completion", "Dropout", "Referral")
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    text = element_text(family = "Arial")
  )
p$plot


p_esp_readm <- ggsurvplot(
  fit_stratified,
  fun = "cloglog",
  conf.int = TRUE,
  censor = FALSE,   # oculta símbolos de censura
  xlab = "log(Tiempo desde el alta)",
  ylab = "log(-log(Supervivencia))",
  legend.labs = c("Razones\nadministrativas", 
                  "Violación\nde reglas", 
                  "Finalización", 
                  "Abandono", 
                  "Derivación"),
  legend.title = "Resultado\ndel tratamiento",
  palette = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
  ggtheme = theme_minimal(base_family = "Arial")
)

p_esp_readm$plot <- p_esp_readm$plot +
    geom_point(
    aes(shape = strata, color = strata),
    size = 2
  ) +
  scale_color_manual(
    values = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3), 
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  theme(
    text = element_text(family = "Arial"),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.title = element_blank()
  )
ggsave(
  filename = "_figs/supervivencia_cloglog_readmision.png",
  plot = p_esp_readm$plot,
  dpi = 300,
  width = 8,
  height = 6,
  units = "in"
)


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

fit_stratified_death <- survfit(Surv(death_time_from_disch_m, death_event) ~ tr_outcome, data = df |> dplyr::filter(tr_outcome!="other") |> 
  dplyr::mutate(death_time_from_disch_m =  ifelse(death_time_from_disch_m == 0, 0.001, death_time_from_disch_m)))


p <- ggsurvplot(
  fit_stratified_death,
  fun = "cloglog",
  conf.int = TRUE,
  censor = FALSE,
  xlab = "log(Time from discharge)",
  ylab = "log(-log(Survival))",
  legend.labs = c("Adm Reasons", "Rule Violation", "Completion", "Dropout", "Referral"),
  legend.title = "Treatment\nOutcome",
  palette = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"), # escala azul-gris distinguible
  #RColorBrewer::brewer.pal(5, "Greys"),
  ggtheme = theme_minimal(base_family = "Arial") # tipografía Arial
)
p$plot <- p$plot +
  geom_point(
    aes(shape = strata, color = strata),
    size = 2
  ) +
  scale_color_manual(
    values = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
    name = "Treatment\nOutcome",
    labels = c("Adm Reasons", "Rule Violation", "Completion", "Dropout", "Referral")
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3), 
    name = "Treatment\nOutcome",
    labels = c("Adm Reasons", "Rule Violation", "Completion", "Dropout", "Referral")
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    text = element_text(family = "Arial")
  )
p$plot

p_esp_death <- ggsurvplot(
  fit_stratified_death,
  fun = "cloglog",
  conf.int = TRUE,
  censor = FALSE,   # oculta símbolos de censura
  xlab = "log(Tiempo desde el alta)",
  ylab = "log(-log(Supervivencia))",
  legend.labs = c("Razones\nadministrativas", 
                  "Violación\nde reglas", 
                  "Finalización", 
                  "Abandono", 
                  "Derivación"),
  legend.title = "Resultado\ndel tratamiento",
  palette = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
  ggtheme = theme_minimal(base_family = "Arial")
)

p_esp_death$plot <- p_esp_death$plot +
    geom_point(
    aes(shape = strata, color = strata),
    size = 2
  ) +
  scale_color_manual(
    values = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3), 
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  theme(
    text = element_text(family = "Arial"),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.title = element_blank()
  )
ggsave(
  filename = "_figs/supervivencia_cloglog_fallecimiento.png",
  plot = p_esp_death$plot,
  dpi = 300,
  width = 8,
  height = 6,
  units = "in"
)
```

We see that treatment outcome has different baseline shapes.

# Session info

```{r}
#| label: session-info
#| echo: true
#| error: true
#| message: true
#| paged.print: true
message(paste0("R library: ", Sys.getenv("R_LIBS_USER")))
message(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))
message(paste0("Editor context: ", path))

cat("quarto version: "); quarto::quarto_version()

sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
)|> 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('R packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}")))
```


```{r}
#| label: session-info-python
#| echo: true
#| error: true
#| message: true
#| paged.print: true
#| class-output: center-table

reticulate::py_list_packages()|> 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Python packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}"))) 
```

Save

```{r}
#| label: save
#| echo: true
#| error: true
#| warning: true
#| message: true
#| paged.print: true
#| results: "hold"
#| eval: true

wdpath<-
paste0(gsub("/cons","",gsub("cons","",paste0(getwd(),"/cons"))))
envpath<- if(regmatches(wdpath, regexpr("[A-Za-z]+", wdpath))=="G"){"G:/Mi unidad/Alvacast/SISTRAT 2023/"}else{"E:/Mi unidad/Alvacast/SISTRAT 2023/"}

paste0(getwd(),"/cons")
file.path(paste0(wdpath,"data/20241015_out"))
file.path(paste0(envpath,"data/20241015_out"))

# Save
rdata_path <- file.path(wdpath, "data/20241015_out", paste0("pred2_ndp_", format(Sys.time(), "%Y_%m_%d"), ".Rdata"))

save.image(rdata_path)
cat("Saved in:",
    rdata_path)

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
  password <- Sys.getenv("PASSWORD_SECRET")
} else {
  if (interactive()) {
    utils::savehistory(tempfile())
    Sys.setenv(PASSWORD_SECRET = readLines(paste0(wdpath, "secret.txt"), warn = FALSE))
    utils::loadhistory()
  }
  Sys.setenv(PASSWORD_SECRET = readLines(paste0(wdpath, "secret.txt"), warn = FALSE))
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
save.image(paste0(rdata_path,".enc"))

# Encriptar el archivo en el mismo lugar
httr2::secret_encrypt_file(path = paste0(rdata_path,".enc"), key = "PASSWORD_SECRET")

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
cat("Copy renv lock into cons folder\n")

if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
  message("Running on RStudio Server or inside Docker. Folder copy skipped.")

} else {
    
  source_folder <- 
  destination_folder <- paste0(wdpath,"cons/renv")
  
  # Copy the folder recursively
    file.copy(paste0(wdpath,"renv.lock"), paste0(wdpath,"cons/renv.lock"), overwrite = TRUE)
  
  message("Renv lock copy performed.")
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
time_after_dedup2<-Sys.time()

paste0("Time in markdown: ");time_after_dedup2-time_before_dedup2
```