---
title: "SENDAs Agreement 1 Update 2010-2022 (Prediction, step 2)"
description: |
  This step involves exploring the behavior of the predictors and best subset selection algorithms.
date: last-modified
author: "Andrés González Santa Cruz"
bibliography: [_lit/refs.bib]
csl: "_lit/american-medical-association.csl" # (Opcional: estilo de citas, p.ej., APA)
toc: true
toc-location: left
toc-float: true
format: 
  html:
    css: [_lib/styles.scss]
    code-fold: true
    embed-resources: true
    theme: 
      - journal #https://quarto.org/docs/output-formats/html-themes-more.html
      #- style.scss
    fig-cap-location: top  
    toc: true
    toc-location: left
execute:
  error: true        # Show errors in the output
  echo: true         # Show R code in the output
  warning: true      # Show warnings
  message: true      # Show messages from the code
  freeze: auto       # Use cache to speed up rendering if there are no changes
editor: source
execute-dir: project
lightbox: auto
---
```{=html}
<style type="text/css">
.showopt {
background-color: #004c93; color: #FFFFFF; width: 100px; height: 20px; text-align: center; vertical-align: middle !important; float: right; font-family: sans-serif; border-radius: 8px;
}
.showopt:hover {
background-color: #dfe4f2;
color: #004c93;
}
pre.plot {
background-color: white !important;
}
.tablelines table, .tablelines td, .tablelines th {
border: 1px solid black;
}
.centrado {
text-align: center;
}
.table.center {
margin-left:auto;
margin-right:auto;
}
/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/
/* Applies to outputs that are not code other than R*/
pre {
overflow-x: auto !important;
}
pre code {
word-wrap: normal !important;
white-space: pre !important;
}
/*
pre:not(.sourceCode) {
white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence */
font-size: 10px !important;
line-height: 50% !important;
}
body{ /* Normal */
text-align: justify;
}
.superbigimage{
overflow-y:scroll;
height:350px;
white-space: nowrap;
overflow-x: auto;
width:100%;
}
.superbigimage img{
overflow-y: scroll;
overflow-x: hidden;
}
.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}
.pandoc-table { /* Should add !important; but it seems no necessary */
margin-left:auto; /* To center */
margin-right:auto;
border-collapse: collapse;
table-layout: auto;
font-size: 11px;
overflow-y: auto;
max-height:450px !important;
white-space: nowrap;
overflow-x: auto;
width:450px;
}
.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}
.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}
.pandoc-table caption {
text-align: left !important;
font-size: 11px !important;
}
.center-table {
text-align: left !important;
font-size: 9px;
overflow-y:scroll;
height:450px;
overflow-x: scroll;
}
.controlly{
overflow-y:scroll;
height:350px;
overflow-x: scroll;
}
/*2025-03-07*/
.callout-warning,
.callout-error,
.callout-message {
  font-size: 0.7rem !important;  /* Adjust this value as needed */
}
.alert, .message {
    font-size: 0.7em !important;
}
.alert.alert-warning,
.alert.alert-danger,
.message {
    font-size: 0.7em !important;
}
</style>
```
```{css, echo=F}
#| label: css-format
h1 {
    color: var(--heading-color);
    font-size: 2rem;
    margin-bottom: 1vh;
}
p {
  font-size: 1.1rem;
  line-height: 1.6rem;
}
a {
  color: var(--primary-color);
  text-decoration: none;
  border-bottom: 3px solid transparent;
  font-weight: bold;
  &:hover, &:focus {
      border-bottom: 3px solid currentColor;
  }
}
section {
  margin: 0 auto;
}
.post-meta {
  font-size: 1rem;
  font-style: italic;
  display: block;
  margin-bottom: 4vh;
  color: var(--secondary-color);
}
nav {
  display: flex;
  justify-content: flex-end;
  padding: 20px 0;
}
/*slider switch css */
.theme-switch-wrapper {
  display: flex;
  align-items: center;
  
  em {
    margin-left: 10px;
    font-size: 1rem;
  }
}
.theme-switch {
  display: inline-block;
  height: 34px;
  position: relative;
  width: 60px;
}
.theme-switch input {
  display:none;
}
.slider {
  background-color: #ccc;
  bottom: 0;
  cursor: pointer;
  left: 0;
  position: absolute;
  right: 0;
  top: 0;
  transition: .4s;
}
.slider:before {
  background-color: #fff;
  bottom: 4px;
  content: "";
  height: 26px;
  left: 4px;
  position: absolute;
  transition: .4s;
  width: 26px;
}
input:checked + .slider {
  background-color: #66bb6a;
}
input:checked + .slider:before {
  transform: translateX(26px);
}
.slider.round {
  border-radius: 34px;
}
.slider.round:before {
  border-radius: 50%;
}
```
```{=html}
<script>
function myFunction1() {
var x = document.getElementById("myDIV");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
<script>
function myFunction2() {
var x = document.getElementById("myDIV2");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
```
<style>
.scrollable-content {
  max-height: 350px;
  overflow-y: auto;
}
pre.scrollable-code {
  max-height: 350px;
  overflow-y: auto;
}
.superbigimage {
  overflow-x: scroll;
  white-space: nowrap;
}
.superbigimage img, 
.superbigimage svg {
  max-width: none;
  height: auto;
}
</style>
<br>

# Data Loading and Exploration

## Loading Packages and uniting databases

<div class="scrollable-content">
```{r}
#| label: load
#| message: false
#| include: true
#| warning: false
#| results: "hold"
# invisible("Only run from Ubuntu")
# if (!(Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv"))) {
#   if(Sys.info()["sysname"]!="Windows"){
#     Sys.setenv(RETICULATE_PYTHON = "/home/fondecytacc/.pyenv/versions/3.11.5/bin/python")
#   }
# }
#clean enviroment
rm(list = ls()); gc()
time_before_dedup2<-Sys.time()
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# Busca .mamba_root/envs/py311/python.exe desde getwd() hacia padres
# find_python_rel <- function(start = getwd(),
#                            rel = file.path(".mamba_root","envs","py311","python.exe")) {
#  cur <- normalizePath(start, winslash = "/", mustWork = FALSE)
#  repeat {
#    cand <- normalizePath(file.path(cur, rel), winslash = "/", mustWork = FALSE)
#    if (file.exists(cand)) return(cand)
#    parent <- dirname(cur)
#    if (identical(parent, cur)) return(NA_character_)  # llegó a la raíz
#    cur <- parent
#  }
# }
# --- Bootstrap reticulate con ruta relativa a getwd() ---
# if(Sys.info()["sysname"]!="Windows"){
#  Sys.setenv(RETICULATE_PYTHON = "usr/bin/python3")
#  reticulate::py_config()
# } else {
#  py <- find_python_rel()
#  if (is.na(py)) {
#    stop("No se encontró Python relativo a getwd() (buscando '.mamba_root/envs/py311/python.exe').\n",
#         "Directorio actual: ", getwd())
#  }
#  # Forzar ese intérprete
#  Sys.unsetenv(c("RETICULATE_CONDAENV","RETICULATE_PYTHON_FALLBACK"))
#  Sys.setenv(RETICULATE_PYTHON = py)
#  #reticulate::use_python(py, required=T)
#  reticulate::py_config()  # verificación
# }
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# --- Load pred1_ndp_2026_01_05.Rdata (simple, project-root aware, no functions) ---
# 1) Locate the file anywhere under the project root (current WD with a trailing "/cons" trimmed)
project_root <- sub("(/)?cons/?$", "", normalizePath(getwd(), winslash = "/", mustWork = FALSE))
fname        <- "pred1_ndp_2026_01_05.Rdata"
hits <- list.files(project_root, pattern = paste0("^", fname, "$"),
                   recursive = TRUE, full.names = TRUE, ignore.case = TRUE)
if (!length(hits)) stop("File not found: ", fname, "\nSearched under: ", project_root)
path <- hits[1]  # take first match
message("Loading: ", path)
# 2) Load into a dedicated environment (keeps .GlobalEnv clean)
env_pred <- new.env(parent = emptyenv())
loaded_objects <- load(path, envir = env_pred)
message("Loaded ", length(loaded_objects), " object(s): ", paste(loaded_objects, collapse = ", "))
# 3) Quick sanity check for each loaded object
for (obj_name in loaded_objects) {
  obj <- get(obj_name, envir = env_pred)
  message("\n--- ", obj_name, " ---")
  if (is.data.frame(obj)) {
    message("Rows: ", nrow(obj), " | Cols: ", ncol(obj))
  } else {
    message("Class: ", paste(class(obj), collapse = ", "), " | Length/Size: ", length(obj))
  }
  #print(str(obj))
}
# 4) Extract all loaded objects to global environment
for (obj_name in loaded_objects) {
  assign(obj_name, get(obj_name, envir = env_pred), envir = .GlobalEnv)
}
# Clean up temporary environment
rm(env_pred)
# (Optional) If you want it in the global env, uncomment:
# list2env(env_prev1y, envir = .GlobalEnv)
```
</div>

<div class="scrollable-content">
```{r}
#| label: setup
#| results: "hold"
#renv falls back to copying rather than symlinking, which is evidently very slow in this configuration.
renv::settings$use.cache(FALSE)
#only use explicit dependencies (in DESCRIPTION)
renv::settings$snapshot.type("implicit")
#check if rstools is installed
if(Sys.info()["sysname"]=="Windows"){
try(installr::install.Rtools(check_r_update=F))
}
check_quarto_version <- function(required = "1.7.29", comparator = c("ge","gt","le","lt","eq")) {
  comparator <- match.arg(comparator)
  current <- package_version(paste(unlist(quarto::quarto_version()), collapse = "."))
  req     <- package_version(required)
  ok <- switch(comparator,
               ge = current >= req,
               gt = current >  req,
               le = current <= req,
               lt = current <  req,
               eq = current == req)
  if (!ok) {
    stop(sprintf("Quarto version check failed: need %s %s (installed: %s).",
                 comparator, required, current), call. = FALSE)
  }
  invisible(TRUE)
}
check_quarto_version("1.7.29", "ge") 
#change repository to CL
local({
  r <- getOption("repos")
  r["CRAN"] <- "https://cran.dcc.uchile.cl/"
  options(repos=r)
})
if(!require(pacman)){install.packages("pacman");require(pacman)}
if(!require(pak)){install.packages("pak");require(pak)}
pacman::p_unlock(lib.loc = .libPaths()) #para no tener problemas reinstalando paquetes
if(Sys.info()["sysname"]=="Windows"){
if (getRversion() != "4.4.1") { stop("Requires R version 4.4.1; Actual: ", getRversion()) }
}
#check docker
check_docker_running <- function() {
  # Try running 'docker info' to check if Docker is running
  system("docker info", intern = TRUE, ignore.stderr = TRUE)
}
if(Sys.info()["sysname"]=="Windows"){
  install_docker <- function() {
    # Open the Docker Desktop download page in the browser for installation
    browseURL("https://www.docker.com/products/docker-desktop")
  }
  # Main logic
  if (inherits(try(check_docker_running(), silent = TRUE), "try-error")) {
    liftr::install_docker()
  } else {
    message("Docker is running.")
  }
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#PACKAGES#######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
unlink("*_cache", recursive=T)
pak::pak_sitrep()
# pak::sysreqs_check_installed(unique(unlist(paks)))
#pak::lockfile_create(unique(unlist(paks)),  "dependencies_duplicates24.lock", dependencies=T)
#pak::lockfile_install("dependencies_duplicates24.lock")
#https://rdrr.io/cran/pak/man/faq.html
#pak::cache_delete()
library(tidytable)
library(ggplot2)
library(readr)
library(tableone)
library(survivalmodels)
#renv::install("patchwork@1.2.0")
library(rms)
library(survidm)
library(caret)
library(survival)
library(caret)
library(dplyr)
library(survex)
library(pec)
library(future)
library(future.apply)
library(parallel)
library(mice)
library(riskRegression)
library(tidyr)
library(doFuture)
library(vcd)
library(survAUC)

#flexsurv
#if(!require(compareCstat)){install.pacakges("compareCstat")}
# library(shapr)#https://norskregnesentral.github.io/shapr/
# library(SemiMarkov)#https://www.degruyterbrill.com/document/doi/10.1515/ijb-2020-0083/html?lang=en
# #https://www.jclinepi.com/article/S0895-4356(24)00142-2/fulltext
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# 3. Activate polars code completion (safe to try even if it fails)
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
#try(polars_code_completion_activate())
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# 4. BPMN from GitHub (not on CRAN, so install via devtools if missing)
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
if (!requireNamespace("bpmn", quietly = TRUE)) {
  devtools::install_github("bergant/bpmn")
}
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# 5. PhantomJS Check (use webshot if PhantomJS is missing)
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# if (!webshot::is_phantomjs_installed()) {
#   webshot::install_phantomjs()
# }
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#FUNCTIONS######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
# NO MORE DEBUGS
options(error = NULL)        # si antes tenías options(error = recover) o browser)
options(browserNLdisabled = FALSE)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#NAs are replaced with "" in knitr kable
options(knitr.kable.NA = '')
pander::panderOptions('big.mark', ',')
pander::panderOptions('decimal.mark', '.')
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#to format rows in bold
format_cells <- function(df, rows ,cols, value = c("italics", "bold", "strikethrough")){
  # select the correct markup
  # one * for italics, two ** for bold
  map <- setNames(c("*", "**", "~~"), c("italics", "bold", "strikethrough"))
  markup <- map[value]  
  for (r in rows){
    for(c in cols){
      # Make sure values are not factors
      df[[c]] <- as.character( df[[c]])
      # Update formatting
      df[r, c] <- ifelse(nchar(df[r, c])==0,"",paste0(markup, gsub(" ", "", df[r, c]), markup))
    }
  }
  return(df)
}
#To produce line breaks in messages and warnings
knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger" style="font-size: small !important;">',
           gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning" style="font-size: small !important;">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('<div class="message" style="font-size: small !important;">',
           gsub('##', '\n', x),
           '</div>', sep = '\n')
   }
)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
sum_dates <- function(x){
  cbind.data.frame(
    min= as.Date(min(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01"),
    p001= as.Date(quantile(unclass(as.Date(x)), .001, na.rm=T), origin = "1970-01-01"),
    p005= as.Date(quantile(unclass(as.Date(x)), .005, na.rm=T), origin = "1970-01-01"),
    p025= as.Date(quantile(unclass(as.Date(x)), .025, na.rm=T), origin = "1970-01-01"),
    p25= as.Date(quantile(unclass(as.Date(x)), .25, na.rm=T), origin = "1970-01-01"),
    p50= as.Date(quantile(unclass(as.Date(x)), .5, na.rm=T), origin = "1970-01-01"),
    p75= as.Date(quantile(unclass(as.Date(x)), .75, na.rm=T), origin = "1970-01-01"),
    p975= as.Date(quantile(unclass(as.Date(x)), .975, na.rm=T), origin = "1970-01-01"),
    p995= as.Date(quantile(unclass(as.Date(x)), .995, na.rm=T), origin = "1970-01-01"),
    p999= as.Date(quantile(unclass(as.Date(x)), .999, na.rm=T), origin = "1970-01-01"),
    max= as.Date(max(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01")
  )
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# Define the function adapted for Polars
sum_dates_polars <- function(df, date_col) {
  # Create the list of quantiles
  quantiles <- c(0.001, 0.005, 0.025, 0.25, 0.5, 0.75, 0.975, 0.995, 0.999)
  # Create expressions to calculate min and max
  expr_list <- list(
    pl$col(date_col)$min()$alias("min"),
    pl$col(date_col)$max()$alias("max")
  )
  # Add expressions for quantiles
  for (q in quantiles) {
    expr_list <- append(expr_list, pl$col(date_col)$quantile(q)$alias(paste0("p", sub("\\.", "", as.character(q)))))
  }
  # Apply the expressions and return a DataFrame with the results
  df$select(expr_list)
}
# Custom function for sampling with a seed
sample_n_with_seed <- function(data, size, seed) {
  set.seed(seed)
  dplyr::sample_n(data, size)
}
# Function to get the most frequent value 
most_frequent <- function(x) { 
  uniq_vals <- unique(x)
  freq_vals <- sapply(uniq_vals, function(val) sum(x == val))
  most_freq <- uniq_vals[which(freq_vals == max(freq_vals))]
  
  if (length(most_freq) == 1) {
    return(most_freq)
  } else {
    return(NA)
  }
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#CONFIG #######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
options(scipen=2) #display numbers rather scientific number
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# Define the function first
#joins these values with semicolons and optionally truncates the result if it exceeds a specified width.
toString2 <- function(x, width = NULL, ...) {
    string <- paste(x, collapse = "; ")
    if (missing(width) || is.null(width) || width == 0) 
        return(string)
    if (width < 0) 
        stop("'width' must be positive")
    if (nchar(string, type = "w") > width) {
        width <- max(6, width)
        string <- paste0(substr(string, 1, width - 3), "...")
    }
    string
}
normalize_txt <- function(x) {
  x|>
    stringi::stri_trans_general("Latin-ASCII")|>
    tolower()|>
    trimws()
}
pair_str <- function(main, sub) {
  main <- as.character(main)
  sub  <- as.character(sub)
  both_na <- is.na(main) & is.na(sub)
  out <- paste0(
    tidytable::coalesce(main, "NA"),
    "::",
    tidytable::coalesce(sub,  "NA")
  )
  out[both_na] <- NA_character_
  out
}
# ── Helpers ────────────────────────────────────────────────────────────
mode_pick_int <- function(x){
  x <- x[!is.na(x)]
  if(length(x)==0) return(NA_integer_)
  tx <- sort(table(x), decreasing = TRUE)
  as.integer(names(tx)[1L])
}
subkey_to_label <- function(x){
  y <- gsub("_"," ", tolower(x))
  y <- gsub("amphetamine type stimulants","amphetamine-type stimulants", y)
  y <- gsub("tranquilizers hypnotics","tranquilizers/hypnotics", y)
  y
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# ── Encryption ────────────────────────────────────────────────────────
# Pack: [24-byte nonce][ciphertext] -> hex
encrypt_chr <- function(x, key) {
    vapply(x, function(v) {
        if (is.na(v)) return(NA_character_)
        ct <- sodium::data_encrypt(charToRaw(as.character(v)), key)
        nonce <- attr(ct, "nonce")                 # 24 bytes
        sodium::bin2hex(c(nonce, ct))              # pack nonce + ct
    }, FUN.VALUE = character(1))
}
# Unpack: hex -> [nonce|ciphertext], restore attr(nonce), then decrypt
decrypt_chr <- function(x, key) {
    vapply(x, function(v) {
        if (is.na(v)) return(NA_character_)
        buf <- sodium::hex2bin(v)
        if (length(buf) < 25L) stop("Ciphertext too short or corrupted.")
        nonce <- buf[1:24]
        ct    <- buf[-(1:24)]
        attr(ct, "nonce") <- nonce
        rawToChar(sodium::data_decrypt(ct, key))
    }, FUN.VALUE = character(1))
}
# Function to encrypt a character vector
encrypt_column <- function(x, key) {
  sapply(x, function(item) {
    if (is.na(item) || item == "") {
      return(NA_character_)
    }
    encrypted_raw <- sodium::data_encrypt(charToRaw(item), key)
    base64enc::base64encode(encrypted_raw)  # Convert to base64 for storage
  }, USE.NAMES = FALSE)
}
decrypt_column <- function(x, key) {
  sapply(x, function(item) {
    if (is.na(item)) return(NA_character_)
    encrypted_raw <- base64enc::base64decode(item)
    rawToChar(sodium::data_decrypt(encrypted_raw, key))
  }, USE.NAMES = FALSE)
}
is_stata_ok <- function(x) {
  nchar(x) <= 32 & grepl("^[A-Za-z][A-Za-z0-9_]*$", x)
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# ── Tidy RMS ────────────────────────────────────────────────────────
tidy_cph <- function(model) {
  if (!inherits(model, "cph")) stop("Model must be an rms::cph object")
  
  # run summary
  s <- summary(model)
  df_s <- as.data.frame(unclass(s))
  df_s$rown <- rownames(s)
  
  # find rows that are "Hazard Ratio"
  hr_rows <- trimws(df_s$rown) == "Hazard Ratio"
  
  # build tidy tibble
  tibble::tibble(
    term     = trimws(rownames(s)[which(hr_rows) - 1L]), # previous row = variable name
    estimate = df_s$Effect[hr_rows],                     # Hazard Ratio
    conf.low = df_s$`Lower 0.95`[hr_rows],
    conf.high= df_s$`Upper 0.95`[hr_rows],
    p.value  = df_s$P[hr_rows]
  )
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
find_latest_file <- function(project_root, prefix) {
  # Get all files matching the pattern
  pattern <- paste0(prefix, "_\\d{8}_\\d{4}\\.csv$")
  files <- list.files(
    path = project_root,
    pattern = pattern,
    full.names = TRUE
  )
  if (length(files) == 0) {
    stop(paste("No files found matching pattern:", pattern))
  }
  # Extract timestamps from filenames (format: YYYYMMDD_HHMM)
  extract_timestamp <- function(filename) {
    matches <- regmatches(
      basename(filename),
      gregexpr("\\d{8}_\\d{4}", basename(filename))
    )
    if (length(matches[[1]]) == 0) {
      return(NA)
    }
    timestamp_str <- matches[[1]][length(matches[[1]])]
    return(timestamp_str)
  }
  timestamps <- sapply(files, extract_timestamp)
  # Remove any files where timestamp extraction failed
  valid_idx <- !is.na(timestamps)
  files <- files[valid_idx]
  timestamps <- timestamps[valid_idx]
  if (length(files) == 0) {
    stop("No valid timestamps found in filenames")
  }
  # Sort by timestamp (descending) and return the latest
  latest_idx <- order(timestamps, decreasing = TRUE)[1]
  return(files[latest_idx])
}

```
</div>

```{r}
# Instala 'here' si no lo tienes: if(!require(here)) install.packages("here")

# Define la raíz del proyecto sin importar el OS
# 'here' siempre apunta a la carpeta que contiene el archivo .Rproj
root <- here::here()

# Si necesitas la ruta "padre" de /cons (como haces en tu script):
project_root <- gsub("/cons$", "", root) 

# Definir carpetas de salida de forma relativa
data_out <- file.path(project_root, "data", "20241015_out")
figs_out <- file.path(root, "_figs")
```

<br>

::: {.callout-note}
**Aims**
- To identify key prognostic and predictive factors associated with the risk of, and time to, the first readmission to treatment among patients aged 18-64 undergoing initial treatment for Substance Use Disorder (SUD) in Chile between 2010 and 2020.
- To identify key prognostic and predictive factors associated with the risk of, and time to, all-cause mortality among the same cohort of patients undergoing initial treatment for SUD in Chile.
- To predict the risk and timing of first treatment readmission and all-cause mortality simultaneously, treating these outcomes within a semi-competing risks framework. This will allow for the estimation of cumulative incidence for each event in the presence of the other for patients aged 18-64 who initiated SUD treatment in Chile between 2010 and 2020.
This study has three primary objectives. First, we aim to identify key prognostic and predictive factors associated with the risk and timing of two critical outcomes: (i) first readmission to treatment and (ii) all-cause mortality. Second, using a semi-competing risks framework, we will develop a model to simultaneously predict these two outcomes, accounting for the fact that death precludes the possibility of readmission. Finally, this model will be used to estimate the probability of various patient trajectories over time following initial treatment. The study population includes all patients aged 18-64 who began treatment for Substance Use Disorder (SUD) in Chile between 2010 and 2020.
:::
<br>

## 0. Prepare data

In this step, we imported the center identifiers and then applied a correction procedure across all imputed datasets to ensure consistency in event-time definitions. Specifically, we converted time and event variables to numeric and integer formats, identified cases where death occurred before or at the readmission endpoint (competing risks), and adjusted follow‑up times and event indicators accordingly by censoring readmission in those situations. After cleaning up temporary variables, we added the center ID as a factor to each dataset. We also excluded `high_uncertainty_imps`, which encompassed predictors with were not reliable. Finally, we exported the corrected datasets both individually as `.parquet` files named as `corrected_datasets_nondum` (nondum= not-dummified) and collectively as a single `.Rds` object, creating a clean and reproducible output ready for downstream analyses.

```{r}
#| label: prepare-data
#| echo: true
#| error: false
#| warning: true
#| message: true
#| paged.print: true
#| results: "hold"
#| eval: true

id_centro_vec <- rio::import(paste0(data_out, "/pred1/id_centro.Rds"))

# 1. Apply Correction to Each Dataset in the List
corrected_datasets_nondum <- map(filled_datasets, function(df) {
  # Ensure columns are numeric for comparison
  df <- df|> 
    mutate(
      readmit_time = as.numeric(readmit_time_from_disch_m),
      death_time   = as.numeric(death_time_from_disch_m),
      # Ensure events are 0/1 integers
      readmit_event = as.integer(as.character(readmit_event)), 
      death_event   = as.integer(as.character(death_event))
    )
  # --- THE CORRECTION LOGIC ---
  df <- df|>
    dplyr::mutate(
      # condition: Patient died (death_event==1) AND death happened before (or at) the readmission endpoint
      is_competing_risk = (death_event == 1) & (death_time <= readmit_time),
      # 1. Update Time: If competing risk, cut follow-up at death time. Otherwise keep original.
      readmit_time_from_disch_m = dplyr::if_else(is_competing_risk, death_time, readmit_time),
      # 2. Update Event: If competing risk, censor readmission (set to 0). Otherwise keep original.
      readmit_event = dplyr::if_else(is_competing_risk, 0, readmit_event)
    )|>
    # Remove temporary helper columns if you want to keep the dataset clean
    dplyr::select(-readmit_time, -death_time, -is_competing_risk)|>
    dplyr::select(any_of(
      setdiff(c(num_vars, cat_vars, "readmit_event", "death_event", "readmit_time_from_disch_m","death_time_from_disch_m"), high_uncertainty_imps)
    ))|>
    dplyr::mutate(center_id= factor(id_centro_vec))
  return(df)
})
```
```{r }
#| label: prepare-data-exp
#| eval: False

# 2. Iterate and Export
# iwalk automatically gives you the object (.x) and its index (.y)
purrr::iwalk(corrected_datasets_nondum, function(df, i) {
  # Construct the full file path
  # e.g. .../data/20241015_out/imputation_1.parquet
  file_name <- file.path(output_dir, paste0("imputation_nondum_", i, ".parquet"))
  # Export using rio
  rio::export(df, file_name)
  message(paste("Saved:", file_name))
})
#Export list
rio::export(corrected_datasets_nondum, paste0(file.path(wdpath, "data/20241015_out/pred1"),"/processed_datasets_nondum.Rds"))
#corrected_datasets_nondum<- rio::import(paste0(file.path(wdpath, "data/20241015_out/pred1"),"/processed_datasets_nondum.Rds"))
```

# Split database

Assuming that `corrected_datasets_nondum` has not been modified since its last update on February 21, the order of the relevant rows in the dataset should remain unchanged (n=88,504). Therefore, the train–test split should be identical to the one used in Python. For this reason, we loaded `death_split_seed2125_test20.parquet` to partition observations into training and test sets.

```{r}
#| label: import-parquet
split2125<- 
nanoparquet::read_parquet(paste0(project_root, "/cons/_out/readm_split_seed2125_test20_mar26.parquet"))

```

```{r}
# 2. Filter the dataset list
corrected_datasets_nondum_filt <- corrected_datasets_nondum |> 
  purrr::map(\(df) df |> 
    dplyr::mutate(esto= dplyr::case_when(grepl("adm reasons",tr_outcome) & death_event==1 & death_time_from_disch_m <= 0.23~"artifact", grepl("other", tr_outcome)~"other", T~NA_character_)) |> 
    dplyr::filter(is.na(esto)) |> 
    dplyr::select(-esto))

corrected_datasets_nondum[[1]] |> 
  dplyr::mutate(esto= dplyr::case_when(grepl("adm reasons",tr_outcome) & death_event==1 & death_time_from_disch_m <= 0.23~"artifact", grepl("other", tr_outcome)~"other", T~NA_character_)) |> 
  janitor::tabyl(esto) |> 
  knitr::kable("markdown", caption="Distribution of Excluded Observations")
#      esto     n     percent valid_percent
#  artifact   137 0.001547953     0.3892045
#     other   215 0.002429269     0.6107955
#      <NA> 88152 0.996022779            NA    
# 
nrow(corrected_datasets_nondum_filt[[1]])          
```


```{r}
# 1. Read Python's split file
py_split <- split2125
cat("Python split:\n")
cat("Total rows:", nrow(py_split), "\n")
cat("Train:", sum(py_split$is_train), "\n")
cat("Test:", sum(!py_split$is_train), "\n\n")

# 2. Check your R data dimensions
cat("Your R data:\n")
cat("Rows in corrected_datasets_nondum_filt[[1]]:", nrow(corrected_datasets_nondum_filt[[1]]), "\n")

# 3. Create matching signature in R (combine death_time + age)
r_data <- corrected_datasets_nondum_filt[[1]] %>%
  mutate(
    row_id = row_number(),  # 1-based row number
    # Adjust column names if yours are different!
    death_time_r = round(death_time_from_disch_m, 2),  # or your death time column
    adm_age_r = adm_age_rec3,  # or your age column
    # Create signature for matching
    signature = paste0("d", death_time_r, "_a", adm_age_r)
  )

# 4. Create same signature in Python data
py_split$signature <- paste0("d", py_split$death_time_from_disch_m, "_a", py_split$adm_age_rec3)

# 5. Find discrepancies
in_py_not_r <- py_split %>% filter(!signature %in% r_data$signature)
in_r_not_py <- r_data %>% filter(!signature %in% py_split$signature)

cat("\n--- MISMATCHES ---\n")
cat("In Python but NOT in R:", nrow(in_py_not_r), "rows\n")
cat("In R but NOT in Python:", nrow(in_r_not_py), "rows\n")

# 6. Show examples of missing rows
if(nrow(in_py_not_r) > 0) {
  cat("\nFirst 10 rows in Python but missing in R:\n")
  print(head(in_py_not_r[, c("row_id", "is_train", "death_time_from_disch_m", "adm_age_rec3")], 10))
}

if(nrow(in_r_not_py) > 0) {
  cat("\nFirst 10 rows in R but missing in Python:\n")
  print(head(in_r_not_py[, c("row_id", "death_time_r", "adm_age_r")], 10))
}

# 7. Check if row_id ranges match
cat("\n--- ROW ID RANGES ---\n")
cat("Python row_id range:", min(py_split$row_id), "-", max(py_split$row_id), "\n")
cat("R row_id range:", min(r_data$row_id), "-", max(r_data$row_id), "\n")
```

We generated a defensive reproducibility check with automatic triggers that stop execution if something changed of the 80/20 split (row order, split mismatch, unexpected counts, etc.).

```{r}
#| label: format-python-split
df <- cbind.data.frame(corrected_datasets_nondum_filt[[1]], dplyr::select(split2125, is_train))|> 
  dplyr::filter(is_train==T)
message(paste0("The training database is composed of ", nrow(df)," observations"))
expected_n_train <- 70521   # replace with known expected value
if (nrow(df) != expected_n_train) {
  stop("❌ Inconsistency detected: Training sample size does not match expected value.")
} else {
  message("✔ Training sample size verified.")
}

observed_foreign_n <- df |>
  dplyr::summarise(n = sum(national_foreign == 1, na.rm = TRUE)) |>
  dplyr::pull(n)
if (observed_foreign_n != 453) {
  stop(paste0("❌ Inconsistency detected: national_foreign count is ",
              observed_foreign_n,
              " but expected ",
              expected_foreign_n))
} else {
  message("✔ national_foreign count verified.")
}
if (abs(mean(df$adm_age_rec3, na.rm = TRUE) -  35.7256) > 0.01) {
  stop("❌ Age distribution changed — possible split mismatch.")
}
```

To ensure consistency between our R and Python modeling environments, we implemented a harmonized feature preprocessing pipeline. First, we isolated and preserved our target and metadata variables (such as readmission and death events, time-to-event tracking, and center identifiers) to ensure they remained unaltered during transformations. We then generated **dummy variables** (one-hot encoding with k-1 categories) for nominal and ordinal features (such as housing status, urbanicity, clinical evaluations, and educational attainment). Crucially, we explicitly defined the **reference category** for each variable (e.g., 'marihuana' for the primary substance `primary_sub_mod` and first substance `first_sub_used`, and 'pg-pab' for the treatment plan type) to guarantee that both environments dropped the exact same baseline category. 

To overcome native differences in how R and Python name generated dummy variables, we implemented a **custom string harmonization** step in R to perfectly replicate Python's snake_case format. Finally, to handle high cardinality and sparse dummy variables within the first_sub_used feature, we **collapsed low-frequency substances** (such as opioids, hallucinogens, inhalants, tranquilizers, and amphetamines) into a single, aggregated first_sub_used_other indicator. This combination of strict reference matching, naming harmonization, and dimensionality reduction guaranteed a uniform input shape of exactly 62 identically named columns (56 features and 6 metadata variables) across all datasets in both programming environments.

```{r}
#| label: discard-great-databases
# 1. Define your threshold in MB
threshold_mb <- 50 
# 2. Identify large objects that ARE NOT in your keep list
heavy_to_remove <- Filter(function(x) {
  # Calculate size in MB
  obj_size <- object.size(get(x)) / (1024^2)
  # Logic: Is it heavy AND is it NOT one of my protected objects?
  obj_size > threshold_mb && !(x %in% c("corrected_datasets_nondum_filt", "py_corrected_datasets"))
}, ls())
# Excepting the databases of focus
heavy_to_remove<- setdiff(heavy_to_remove , c("py_corrected_datasets", "corrected_datasets_nondum_filt"))
# 3. Remove them and clean up
if (length(heavy_to_remove) > 0) {
  rm(list = heavy_to_remove)
  message("Removed: ", paste(heavy_to_remove, collapse = ", "))
} else {
  message("No objects found above the threshold (excluding protected datasets).")
}
# 4. Release memory to Windows
gc()
```

```{r}
#| label: format-db-to-python
#dg_psiq_cie_10_dg: This looks like a broader psychiatric diagnosis. Is it categorical or already a dummy?= this is the diagnosis of drug consumption
#ehtnicity is ok
#national_foreign is ok
# ==========================================
# 1. DEFINE TARGET COLUMNS TO ENFORCE
# ==========================================
meta_cols <- c(
  "readmit_event", "death_event", 
  "readmit_time_from_disch_m", "death_time_from_disch_m", 
  "center_id", "row_id"
)
cox_columns <- c(
  # 1. Numerical and natural binary variables (Unchanged)
  "adm_age_rec3", "porc_pobr", "dit_m", "national_foreign", 
  "ethnicity", "dg_psiq_cie_10_instudy", "dg_psiq_cie_10_dg", 
  "dx_f3_mood", "dx_f6_personality", "dx_f_any_severe_mental", 
  "any_phys_dx", "polysubstance_strict",
  # 2. Original categorical dummies (Already existing in your vector)
  "sex_rec_woman", "cohabitation_family_of_origin", 
  "cohabitation_with_couple_children", "cohabitation_others", 
  "sub_dep_icd10_status_drug_dependence", 
  "any_violence_1_domestic_violence_sex_abuse", "tr_outcome_referral", 
  "tr_outcome_dropout", "tr_outcome_adm_discharge_rule_violation_undet", 
  "tr_outcome_adm_discharge_adm_reasons", #"tr_outcome_other", 
  "adm_motive_sanitary_sector", "adm_motive_another_sud_facility_fonodrogas_senda_previene", 
  "adm_motive_justice_sector", "adm_motive_other", "first_sub_used_alcohol", 
  "first_sub_used_cocaine_paste", "first_sub_used_cocaine_powder", "first_sub_used_other",
  "primary_sub_mod_cocaine_paste", "primary_sub_mod_cocaine_powder", #"primary_sub_mod_marijuana",
   "primary_sub_mod_alcohol", "primary_sub_mod_others", 
  "tipo_de_vivienda_rec2_other_unknown", "plan_type_corr_pg_pr", 
  "plan_type_corr_m_pr", "plan_type_corr_pg_pai", "plan_type_corr_m_pai", 
  "occupation_condition_corr24_unemployed", "occupation_condition_corr24_inactive", 
  "marital_status_rec_single", "marital_status_rec_separated_divorced_annulled_widowed", 
  # 3. NEW Dummies for ordinal variables
  # (Automatically excluding your chosen reference categories) 
  # Tenure status (Ref: owner/transferred dwellings/pays dividends)
  "tenure_status_household_renting", 
  "tenure_status_household_others", 
  "tenure_status_household_stays_temporarily_with_a_relative", 
  "tenure_status_household_illegal_settlement",
  # Urbanicity (Reference: 3.Urban)
  "urbanicity_cat_2_mixed", 
  "urbanicity_cat_1_rural",
  # Clinical evaluation variables (Reference: logro alto / high achievement)
  "evaluacindelprocesoteraputico_logro_intermedio", "evaluacindelprocesoteraputico_logro_minimo",
  "eva_consumo_logro_intermedio", "eva_consumo_logro_minimo",
  "eva_fam_logro_intermedio", "eva_fam_logro_minimo",
  "eva_relinterp_logro_intermedio", "eva_relinterp_logro_minimo",
  "eva_ocupacion_logro_intermedio", "eva_ocupacion_logro_minimo",
  "eva_sm_logro_intermedio", "eva_sm_logro_minimo",
  "eva_fisica_logro_intermedio", "eva_fisica_logro_minimo",
  "eva_transgnorma_logro_intermedio", "eva_transgnorma_logro_minimo",
  # Substance use frequency (Reference: 1.≤1 day/wk)
  "prim_sub_freq_rec_2_2_6_days_wk", 
  "prim_sub_freq_rec_3_daily",
  # Educational attainment (Reference: 1-More than high school)
  "ed_attainment_corr_2_completed_high_school_or_less", 
  "ed_attainment_corr_3_completed_primary_school_or_less"
)
python_columns <- cox_columns
# ==========================================
# 2. HELPER FUNCTION
# ==========================================
get_mode_numeric <- function(x) {
  x_no_na <- x[!is.na(x)]
  if (!length(x_no_na)) return(NA_real_)
  tab <- table(x_no_na)
  as.numeric(names(tab)[which.max(tab)])
}
# ==========================================
# 3. MAIN PREPROCESSING FUNCTION
# ==========================================
# Keep meta_cols and cox_columns exactly as you already defined above.
# Compatibility alias if anything else still expects python_columns:
python_columns <- cox_columns
get_mode_numeric <- function(x) {
  x_no_na <- x[!is.na(x)]
  if (!length(x_no_na)) return(NA_real_)
  tab <- table(x_no_na)
  as.numeric(names(tab)[which.max(tab)])
}
DUMMY_REFERENCE <- c(
  sex_rec = "man",
  plan_type_corr = "pg-pab",
  marital_status_rec = "married/cohabiting",
  cohabitation = "alone",
  sub_dep_icd10_status = "hazardous consumption",
  tr_outcome = "completion",
  adm_motive = "spontaneous consultation",
  tipo_de_vivienda_rec2 = "formal housing",
  occupation_condition_corr24 = "employed",
  any_violence = "0.No domestic violence/sex abuse",
  first_sub_used = "marijuana",
  primary_sub_mod = "marijuana",
  tenure_status_household = "owner/transferred dwellings/pays dividends",
  urbanicity_cat = "3.Urban",
  evaluacindelprocesoteraputico = "logro alto",
  eva_consumo = "logro alto",
  eva_fam = "logro alto",
  eva_relinterp = "logro alto",
  eva_ocupacion = "logro alto",
  eva_sm = "logro alto",
  eva_fisica = "logro alto",
  eva_transgnorma = "logro alto",
  prim_sub_freq_rec = "1.≤1 day/wk",
  ed_attainment_corr = "1-More than high school"
)
apply_ordered_mappings <- function(df, ordered_mappings = NULL) {
  if (is.null(ordered_mappings) || !length(ordered_mappings)) return(df)
  for (col in intersect(names(ordered_mappings), names(df))) {
    raw_values <- trimws(as.character(df[[col]]))
    mapped <- as.numeric(unname(ordered_mappings[[col]][raw_values]))
    failed <- is.na(mapped) & !is.na(raw_values) & nzchar(raw_values)
    if (any(failed)) warning(paste("Mapping failed for some values in:", col))
    mode_val <- get_mode_numeric(mapped)
    mapped[is.na(mapped)] <- mode_val
    df[[col]] <- mapped
  }
  df
}
set_reference_levels <- function(df, refs = DUMMY_REFERENCE) {
  for (col in intersect(names(refs), names(df))) {
    values <- trimws(as.character(df[[col]]))
    ref <- refs[[col]]
    cats <- unique(values)
    df[[col]] <- if (ref %in% cats) {
      factor(values, levels = c(ref, cats[cats != ref]), ordered = FALSE)
    } else {
      factor(values, ordered = FALSE)
    }
  }
  df
}
make_dummies <- function(df) {
  cat_cols <- names(df)[vapply(df, function(x) is.character(x) || is.factor(x), logical(1))]
  if (!length(cat_cols)) return(df)
  cat_df <- as.data.frame(df[cat_cols], stringsAsFactors = FALSE)
  cat_df[] <- lapply(cat_df, function(x) {
    x <- factor(x)
    x <- addNA(x, ifany = TRUE)
    lev <- levels(x)
    lev[is.na(lev)] <- "__MISSING__"
    levels(x) <- paste0("_", lev)
    x
  })
  mm <- model.matrix(~ ., data = cat_df) |> as.data.frame(check.names = FALSE)
  mm[["(Intercept)"]] <- NULL
  cbind(df[setdiff(names(df), cat_cols)], mm)
}
preprocess_features_robust <- function(
  df,
  split_info = split2125,
  recode_first_sub_used_to_other = FALSE,
  do_dummies = TRUE,
  ordered_mappings = NULL
) {
  if (!"is_train" %in% names(split_info)) stop("split_info must include `is_train`.")
  if (nrow(df) != nrow(split_info)) stop("`df` and `split_info` must have the same number of rows.")
  df_clean <- df |>
    as.data.frame(stringsAsFactors = FALSE) |>
    dplyr::bind_cols(dplyr::select(as.data.frame(split_info), is_train)) |>
    dplyr::filter(is_train) |>
    dplyr::select(-is_train)
  if ("center_id" %in% names(df_clean)) {
    df_clean$center_id <- suppressWarnings(as.numeric(df_clean$center_id))
  }
  if (isTRUE(recode_first_sub_used_to_other) && "first_sub_used" %in% names(df_clean)) {
    to_other <- c(
      "hallucinogens", "opioids", "amphetamine-type stimulants",
      "tranquilizers/hypnotics", "inhalants", "others"
    )
    s <- df_clean[["first_sub_used"]] |> as.character() |> trimws() |> tolower()
    s[s %in% to_other] <- "other"
    df_clean[["first_sub_used"]] <- factor(s)
  }
  df_clean <- df_clean |>
    apply_ordered_mappings(ordered_mappings = ordered_mappings) |>
    set_reference_levels(refs = DUMMY_REFERENCE)
  if (!isTRUE(do_dummies)) return(df_clean)
  make_dummies(df_clean)
}
collapse_first_sub_used_other <- function(df) {
  cols_to_group <- c(
    "first_sub_used_opioids", "first_sub_used_others",
    "first_sub_used_hallucinogens", "first_sub_used_inhalants",
    "first_sub_used_tranquilizers_hypnotics",
    "first_sub_used_amphetamine_type_stimulants"
  )
  cols_present <- intersect(cols_to_group, names(df))
  df$first_sub_used_other <- if (length(cols_present)) {
    as.integer(rowSums(df[cols_present], na.rm = TRUE) > 0)
  } else {
    0L
  }
  df
}
idx <- seq_len(min(5, length(corrected_datasets_nondum_filt)))
all_required_cols <- c(meta_cols, cox_columns)
py_corrected_datasets <- purrr::map(
  corrected_datasets_nondum_filt[idx],
  \(x) {
    x_proc <- x |>
      preprocess_features_robust(
        split_info = split2125,
        recode_first_sub_used_to_other = FALSE,
        do_dummies = TRUE,
        ordered_mappings = NULL  # pass list(...) if you want ordinal encoding enabled
      ) |>
      janitor::clean_names() |>
      collapse_first_sub_used_other()
    missing_cols <- setdiff(all_required_cols, names(x_proc))
    dplyr::select(x_proc, dplyr::any_of(all_required_cols))
  }
)
glimpse(py_corrected_datasets[[1]])
```

```{r}
# Define the list of predictors based on your provided list
predictors <- c(
  # Numeric, Ordinal, and Boolean features
  "adm_age_rec3", "porc_pobr", "dit_m", "national_foreign", 
  "ethnicity", "dg_psiq_cie_10_instudy", 
  "dg_psiq_cie_10_dg", "dx_f3_mood", "dx_f6_personality", 
  "dx_f_any_severe_mental", "any_phys_dx", "polysubstance_strict",
  
  # Dummy encoded features
  "sex_rec_woman", 
  "cohabitation_family_of_origin", "cohabitation_with_couple_children", "cohabitation_others", 
  "sub_dep_icd10_status_drug_dependence", 
  "any_violence_1_domestic_violence_sex_abuse", 
  "tr_outcome_referral", "tr_outcome_dropout", "tr_outcome_adm_discharge_rule_violation_undet", "tr_outcome_adm_discharge_adm_reasons", 
  #"tr_outcome_other",  #eliminated because of the artifact issue
  "adm_motive_sanitary_sector", "adm_motive_another_sud_facility_fonodrogas_senda_previene", "adm_motive_justice_sector", "adm_motive_other", 
  "first_sub_used_alcohol", "first_sub_used_cocaine_paste", "first_sub_used_cocaine_powder", "first_sub_used_other", 
  "primary_sub_mod_cocaine_paste", "primary_sub_mod_cocaine_powder", "primary_sub_mod_alcohol", "primary_sub_mod_others", 
  "tipo_de_vivienda_rec2_other_unknown", 
  #"plan_type_corr_pg_pr", "plan_type_corr_m_pr", "plan_type_corr_pg_pai", "plan_type_corr_m_pai", 
  "occupation_condition_corr24_unemployed", "occupation_condition_corr24_inactive", 
  "marital_status_rec_single", "marital_status_rec_separated_divorced_annulled_widowed", 
  # NEW Dummies for ordinal variables
  # (Automatically excluding your chosen reference categories) 
  # Tenure status (Reference: owner/transferred dwellings/pays dividends)
  "tenure_status_household_renting", 
  "tenure_status_household_others", 
  "tenure_status_household_stays_temporarily_with_a_relative", 
  "tenure_status_household_illegal_settlement",
  # Urbanicity (Reference: 3.Urban)
  "urbanicity_cat_2_mixed", 
  "urbanicity_cat_1_rural",
  # Clinical evaluation variables (Reference: logro alto / high achievement)
  "evaluacindelprocesoteraputico_logro_intermedio", "evaluacindelprocesoteraputico_logro_minimo",
  "eva_consumo_logro_intermedio", "eva_consumo_logro_minimo",
  "eva_fam_logro_intermedio", "eva_fam_logro_minimo",
  "eva_relinterp_logro_intermedio", "eva_relinterp_logro_minimo",
  "eva_ocupacion_logro_intermedio", "eva_ocupacion_logro_minimo",
  "eva_sm_logro_intermedio", "eva_sm_logro_minimo",
  "eva_fisica_logro_intermedio", "eva_fisica_logro_minimo",
  "eva_transgnorma_logro_intermedio", "eva_transgnorma_logro_minimo",
  # Substance use frequency (Reference: 1.≤1 day/wk)
  "prim_sub_freq_rec_2_2_6_days_wk", 
  "prim_sub_freq_rec_3_daily",
  # Educational attainment (Reference: 1-More than high school)
  "ed_attainment_corr_2_completed_high_school_or_less", 
  "ed_attainment_corr_3_completed_primary_school_or_less"
)
# Create the formulas
formula_readmit <- as.formula(paste("Surv(readmit_time_from_disch_m, readmit_event) ~", paste(predictors, collapse = " + "),"+ strata(plan_type_strata)"))
formula_death <- as.formula(paste("Surv(death_time_from_disch_m, death_event) ~", paste(predictors, collapse = " + "),"+ strata(plan_type_strata)"))
```

## Correlation

```{r}
#| label: corr
#| warning: false
df <- py_corrected_datasets[[1]]
is_num <- function(x) is.numeric(x) || is.integer(x)
is_cat <- function(x) is.factor(x) || is.character(x) || is.logical(x)
cramers_v <- function(x, y) {
  tab <- table(x, y)
  if (nrow(tab) < 2 || ncol(tab) < 2) return(NA_real_)
  as.numeric(vcd::assocstats(tab)$cramer)
}
eta_num_cat <- function(num, cat) {
  d <- data.frame(num = num, cat = as.factor(cat))
  d <- d[stats::complete.cases(d), , drop = FALSE]
  if (nrow(d) < 3 || nlevels(d$cat) < 2) return(NA_real_)
  sqrt(max(summary(stats::lm(num ~ cat, data = d))$r.squared, 0))
}
mixed_assoc <- function(x, y) {
  d <- data.frame(x = x, y = y)
  d <- d[stats::complete.cases(d), , drop = FALSE]
  if (nrow(d) < 3) return(NA_real_)
  if (is_num(d$x) && is_num(d$y)) return(abs(stats::cor(d$x, d$y, method = "spearman")))
  if (is_cat(d$x) && is_cat(d$y)) return(abs(cramers_v(d$x, d$y)))
  if (is_num(d$x) && is_cat(d$y)) return(abs(eta_num_cat(d$x, d$y)))
  if (is_cat(d$x) && is_num(d$y)) return(abs(eta_num_cat(d$y, d$x)))
  NA_real_
}
vars <- names(df)
pairs <- utils::combn(vars, 2, simplify = FALSE)
pairs_gt_02 <- purrr::map_dfr(pairs, function(v) {
  a <- mixed_assoc(df[[v[1]]], df[[v[2]]])
  tibble::tibble(var1 = v[1], var2 = v[2], assoc = a)
}) |>
  dplyr::filter(!is.na(assoc), assoc > 0.20) |>
  dplyr::arrange(dplyr::desc(assoc))
pairs_gt_02 |> 
  dplyr::filter(assoc>=.9)
rm("df"); rm("series_dummies"); rm("series_collapsed")
```

There are no highly correlated variables.

## GVIF

```{r}
#| label: gvif
py_corrected_datasets <- purrr::map(
  py_corrected_datasets,
  \(df) df |>
    dplyr::mutate(
      plan_type_strata = dplyr::case_when(
        plan_type_corr_pg_pr  == 1 ~ "pg-pr",
        plan_type_corr_m_pr   == 1 ~ "m-pr",
        plan_type_corr_pg_pai == 1 ~ "pg-pai",
        plan_type_corr_m_pai  == 1 ~ "m-pai",
        TRUE ~ "pg-pab"
      ),
      plan_type_strata = as.factor(plan_type_strata),
      death_time_from_disch_m = dplyr::if_else(death_time_from_disch_m <= 0, 1e-6, death_time_from_disch_m),
      readmit_time_from_disch_m = dplyr::if_else(readmit_time_from_disch_m <= 0, 1e-6, readmit_time_from_disch_m)
    )
)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#Large generalized VIF (as calculated by rms::vif()) suggests that there might be large off-diagonal elements in the variance-covariance matrix of the coefficient estimates
cox_readmit <- coxph(  formula_readmit,  data = py_corrected_datasets[[1]] )
cox_death <- coxph(formula_death, data = py_corrected_datasets[[1]] )
 rms::vif(cox_death) |> 
  as.data.frame() |> 
  tibble::rownames_to_column("Variable") |> 
   dplyr::rename(Gvif = 2) |>
  dplyr::arrange(dplyr::desc(dplyr::pick(2)[[1]]))|> 
    knitr::kable("markdown", caption= "Death")
 rms::vif(cox_readmit) |> 
  as.data.frame() |> 
  tibble::rownames_to_column("Variable") |>
  dplyr::rename(Gvif = 2) |>
  dplyr::arrange(dplyr::desc(dplyr::pick(2)[[1]])) |> 
  knitr::kable("markdown", caption= "Readmission")
```

Variables `evaluacindelprocesoteraputico_logro_minimo` was highly collinear.

```{r}
#| label: update-model-to-gvif
formula_readmit_updated <- update(
  formula_readmit,
  . ~ . - evaluacindelprocesoteraputico_logro_minimo
)
formula_death_updated <- update(
  formula_death,
  . ~ . - evaluacindelprocesoteraputico_logro_minimo
)
```

### Functions for comparison

```{r}
#| label: EVALUATE-MI-CV

doFuture::registerDoFuture()  # Register with foreach
options(future.globals.maxSize = 2 * 1024^3)  # 2 GiB
#HELPER FUNCTION: Evaluates a Cox model with multiple imputation and cross-validation, computing C-index, Uno's C, and IBS at specified time points. Handles parallel processing and ensures reproducibility with seeds. Validates input schema and provides detailed error messages for missing variables.
# Define these at TOP LEVEL, outside evaluate_mi_cv()
to01 <- function(x) {
  if (is.factor(x)) x <- as.character(x)
  x <- suppressWarnings(as.numeric(x))
  if (all(is.na(x) | x %in% c(0,1))) return(as.integer(x))
  if (all(is.na(x) | x %in% c(1,2))) return(as.integer(x - 1))
  stop("event_col must be binary 0/1 (or 1/2).")
}
ibs_ipcw_train <- function(df_train, df_test, surv_mat, times,
                            time_col, event_col, eps = 1e-8, g_min = 0.05) {
  ev_train <- to01(df_train[[event_col]])
  sf_cens  <- survival::survfit(
    Surv(df_train[[time_col]], 1 - ev_train) ~ 1
  )
  # ── KEY FIX: find the last time G(t) >= g_min in THIS fold's training KM ──
  safe_times <- sf_cens$time[sf_cens$surv >= g_min]
  tau_ipcw   <- if (length(safe_times)) max(safe_times) else max(times)
  # Restrict integration to times <= tau_ipcw
  keep  <- times <= tau_ipcw
  if (sum(keep) < 2L) return(NA_real_)   # not enough support
  times <- times[keep]
  surv_mat <- surv_mat[, keep, drop = FALSE]
  G_at <- function(tt) {
    g <- summary(sf_cens, times = tt, extend = TRUE)$surv
    pmax(g, g_min)   # not eps, to avoid extreme weights and infinite variance in IPCW
    #pmax(g, eps)
  }
  Tt <- as.numeric(df_test[[time_col]])
  Dt <- to01(df_test[[event_col]])
  bs_t <- vapply(seq_along(times), function(j) {
    t0 <- times[j]
    S0 <- surv_mat[, j]
    Y0 <- as.numeric(Tt > t0)
    w  <- ifelse(
      Tt <= t0 & Dt == 1L,
      1 / G_at(pmax(Tt - 1e-8, 0)),
      ifelse(Tt > t0, 1 / G_at(t0), 0)
    )
    mean(w * (Y0 - S0)^2, na.rm = TRUE)
  }, numeric(1))
  as.numeric(
    sum(diff(times) * (head(bs_t, -1) + tail(bs_t, -1)) / 2) /
      (max(times) - min(times))
  )
}

evaluate_mi_cv <- function(formula, imputed_list, time_col, event_col, cv_strata_col = "plan_type_strata", k = 5, seed = 2125, n_eval_times = 100, n_repeats=20, cpus=16, verbose = FALSE, log_every = 100L) {
  #horizons at 1,3,5 years
  horizons_month <- c(12, 36, 60)
  
  t_start <- Sys.time()
  H <- length(horizons_month)
  M <- length(imputed_list)
  r <- as.numeric(n_repeats)
  cat(sprintf("Starting Evaluation: %d Imputations x %d Folds = %d Total Models\n", M, k, M*k*r))
  cat("------------------------------------------------------------\n")
  validate_imputed_schema <- function(formula, imputed_list, time_col, event_col, cv_strata_col) {
    fml <- if (inherits(formula, "formula")) formula else stats::as.formula(formula)
    required <- unique(c(all.vars(fml), time_col, event_col, cv_strata_col))
    missing_by_imp <- lapply(imputed_list, function(df) setdiff(required, names(df)))
    bad <- which(lengths(missing_by_imp) > 0L)
    if (length(bad)) {
      detail <- paste(
        vapply(bad, function(i) {
          sprintf("imputation %d missing: %s", i, paste(missing_by_imp[[i]], collapse = ", "))
        }, character(1)),
        collapse = "\n"
      )
      stop(
        sprintf("Input schema mismatch.\nRequired vars: %s\n%s",
                paste(required, collapse = ", "), detail),
        call. = FALSE
      )
    }
    invisible(required)
  }
  validate_imputed_schema(formula, imputed_list, time_col, event_col, cv_strata_col)
  
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  # PARALLEL CLUSTER SETUP
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  # use
  n_cores <- max(1L, min(cpus, parallel::detectCores(logical = TRUE) - 16L))# Leave some cores free for other processes
  seeds <- seed + seq_len(n_repeats) - 1L# Generate a unique seed for each repeat, ensuring reproducibility across runs
  cat(sprintf("Booting up %d CPU cores using future backend...\n", n_cores))
  
  # Use multisession (works on Windows + Positron)
  plan(multisession, workers = n_cores)
  on.exit(plan(sequential), add = TRUE)  # Cleanup
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  # PRE-FLIGHT CHECKS / DEFINE FOLDS & METADATA
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  stopifnot("Row counts across imputed datasets do not match!" = 
              all(sapply(imputed_list, nrow) == nrow(imputed_list[[1]])))
  
  tau <- as.numeric(quantile(imputed_list[[1]][[time_col]], 0.90, na.rm = TRUE))
  c_tau <- tau * 0.95
  cat(paste0("\n#:_:_:_:_:_\nTau: ", round(tau,2), "; C-tau: ", c_tau,"\n#:_:_:_:_:_\n"))
  eval_times <- sort(unique(c(seq(1e-6, tau, length.out = n_eval_times), horizons_month)))
  set.seed(seed)
  cv_strata <- interaction(imputed_list[[1]][[event_col]],
                          imputed_list[[1]][[cv_strata_col]], drop = TRUE)
  folds_by_repeat <- lapply(seq_len(n_repeats), function(r) {
    set.seed(seeds[r])
    caret::createFolds(cv_strata, k = k, list = TRUE)
  }) 
  folds <- folds_by_repeat[[1]]
  
  cv_strata_table <- as.data.frame(table(cv_strata), stringsAsFactors = FALSE)
  names(cv_strata_table) <- c("cv_strata", "n")
  fold_sizes <- vapply(folds, length, integer(1))
  fold_id <- integer(length(cv_strata))
  for (ii in seq_along(folds)) fold_id[folds[[ii]]] <- ii
  cv_strata_by_fold <- as.data.frame(table(fold = fold_id, cv_strata = cv_strata), stringsAsFactors = FALSE)
  # Result matrices
  c_array    <- array(NA_real_, dim = c(n_repeats, M, k))
  unoC_array <- array(NA_real_, dim = c(n_repeats, M, k))
  ibs_array  <- array(NA_real_, dim = c(n_repeats, M, k))
  ibs_ipcw_array <- array(NA_real_, dim = c(n_repeats, M, k))   # ← NEW
  #storage of RMSTs and Concordance at specific horizons (1,3,5 years)
  c_horizon  <- array(NA_real_, dim = c(n_repeats, M, k, H),
                      dimnames = list(NULL, NULL, NULL, paste0("t", horizons_month)))                   
  # Build a grid of all tasks so the parallel workers can grab them. 
  # checklist of all 25 models (5 imputations x 5 folds) x repetitions
  task_grid <- expand.grid(r = seq_len(n_repeats), m = seq_len(M), i = seq_len(k))
  log_every <- as.integer(log_every)
  if (!is.finite(log_every) || log_every < 1L) log_every <- 100L
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  # THE TRIPLE LOOP: Repeats -> Imputations -> Folds
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  # Single parallel execution via future.apply
  all_results <- future.apply::future_lapply(
    X = seq_len(nrow(task_grid)),
    FUN = function(row) {
      r <- task_grid$r[row]
      m <- task_grid$m[row]
      i <- task_grid$i[row]
      
      # Use pre-computed folds
      test_idx <- folds_by_repeat[[r]][[i]]
      df_train <- imputed_list[[m]][-test_idx, , drop = FALSE]
      df_test  <- imputed_list[[m]][test_idx, , drop = FALSE]
      n_test_before <- nrow(df_test)
      
      # Filter FIRST before fitting model and predicting      
      local_formula <- as.formula(paste(deparse(formula), collapse = " "))
      ## Don't change environment - let it inherit naturally
      #environment(local_formula) <- environment()      

      # Parse formula and extract ALL strata variables
      local_formula <- as.formula(paste(deparse(formula), collapse = " "))
      formula_char <- paste(deparse(local_formula), collapse = " ")
      # Find all strata(var) terms and extract variable names
      strata_matches <- gregexpr("strata\\([^)]+\\)", formula_char)
      if (strata_matches[[1]][1] != -1) {
        strata_terms <- regmatches(formula_char, strata_matches)[[1]]
        strata_vars <- gsub("strata\\(([^)]+)\\)", "\\1", strata_terms)
        strata_vars <- trimws(strata_vars)
        
        # Filter test set by each strata variable
        for (sv in strata_vars) {
          if (sv %in% names(df_train)) {
            train_levels <- unique(df_train[[sv]])
            before_nrow <- nrow(df_test)
            df_test <- df_test[df_test[[sv]] %in% train_levels, , drop = FALSE]
            dropped <- before_nrow - nrow(df_test)
            if (verbose && dropped > 0) {
              message(sprintf("   -> Filtered %d rows: strata '%s' not in training", dropped, sv))
            }
          }
        }
      }
      # Check if we have enough data left
      if (nrow(df_test) < 10) {
        warning(sprintf("Imp %d Fold %d: Only %d rows after strata filtering, returning NA", m, i, nrow(df_test)))
        return(list(r = r, m = m, i = i, c_uno = NA_real_, unoC = NA_real_,
                   ibs = NA_real_, ibs_ipcw = NA_real_, c_horizons = rep(NA_real_, H)))
      }
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # 1. GENERATE THE ABSOLUTE SURVIVAL MATRIX
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # NOW fit model and predict (after filtering)
      # Fit model and predict
      cv_model <- coxph(local_formula, data = df_train, x = TRUE, y = TRUE, model = TRUE)
      surv_mat <- pec::predictSurvProb(cv_model, newdata = df_test, times = eval_times)         
      if (is.vector(surv_mat)) {
          surv_mat <- matrix(surv_mat, nrow = nrow(df_test), ncol = length(eval_times))
      }
      stopifnot("Survival matrix dimensions do not match test data!" = nrow(surv_mat) == nrow(df_test))
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # 2. CALCULATE ABSOLUTE RISK (RMST) 
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # Matrix multiplication (%*%) perfectly calculates the area under the step-curve (rectandular integration; not trapezoidal)
      intervals <- diff(eval_times)
      df_test$rmst_score <- as.numeric(surv_mat[, -ncol(surv_mat)] %*% intervals)
      # CRITICAL FIX: concordance() expects Higher Score = Higher Risk.
      # Higher RMST = Lower Risk. We must NEGATE it.
      df_test$risk_score <- -df_test$rmst_score 
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # METRIC 1: Uno's C-Index 
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # We define a slightly shorter tau (95%) for the C-index calculation
      # to prevent the IPCW weights from hitting divide-by-zero at the tail.
      c_uno_val <- tryCatch(
              survival::concordance(
              Surv(df_test[[time_col]], df_test[[event_col]]) ~ risk_score, 
              data = df_test, 
              timewt = "n/G2",
              ymax = c_tau,
              reverse = TRUE #to flip direction.  You are passing a Risk Score (High = Bad) but telling the function it's a Protective Score (High = Good).#RMST negation + reverse=TRUE = valid //I'm giving you a score where high = bad, but you expect high = good, so please invert your interpretation."
              )$concordance,
              error = function(e) {
              message("Fold ", i, " C-index Error: ", e$message)
              NA_real_
              }
          )    
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # METRIC 2: Integrated Brier Score
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      y_true_test <- Surv(df_test[[time_col]], df_test[[event_col]])
      ibs_val <- survex::integrated_brier_score(
          y_true = y_true_test,
          surv   = surv_mat,
          times  = eval_times
      )
      # ── NEW: IPCW IBS using training censoring distribution ──
      ibs_ipcw_val <- tryCatch(
        ibs_ipcw_train(df_train, df_test, surv_mat, eval_times,
                       time_col, event_col),
        error = function(e) { message("IBS-IPCW error: ", e$message); NA_real_ }
      )
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # METRIC 3: survAUC::UnoC (WITH DIAGNOSTICS)
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      ev_train <- to01(df_train[[event_col]])
      ev_test  <- to01(df_test[[event_col]])
      ok <- is.finite(df_test$risk_score) &
          is.finite(df_test[[time_col]]) &
          is.finite(ev_test) &
          complete.cases(surv_mat)
      # Censoring distribution diagnostics (Version 1)
      sf_cens <- survival::survfit(Surv(df_train[[time_col]], 1 - ev_train) ~ 1)
      g_ctau  <- summary(sf_cens, times = c_tau, extend = TRUE)$surv
      tau_uno <- c_tau
      if (!is.finite(g_ctau) || g_ctau <= 1e-6) {
        tt <- sf_cens$time[sf_cens$surv > 1e-6]
        tau_uno <- if (length(tt)) min(c_tau, max(tt)) else NA_real_
      }
      # Before calling UnoC, check event count at tau_uno
      n_evt_tau <- sum(ev_test[ok] == 1L & df_test[[time_col]][ok] <= tau_uno)
      # Break ties for survAUC (Version 2)
      lp_u <- rank(df_test$risk_score[ok], ties.method = "average")
      # Initialize u
      u <- NA_real_
      if (n_evt_tau < 50) {
        message(sprintf("    → WARNING: Only %d events <= tau_uno. UnoC may be unstable.", n_evt_tau))
      } else {
        # Attempt UnoC with FULL error reporting
        u <- tryCatch({
          withCallingHandlers(
            survAUC::UnoC(
              Surv(df_train[[time_col]], ev_train),
              Surv(df_test[[time_col]][ok], ev_test[ok]),
              lp_u,
              time = tau_uno
            ),
            warning = function(w) {
              message(sprintf("    → UnoC WARNING: %s", w$message))
              invokeRestart("muffleWarning")
            }
          )
        }, error = function(e) {
          message(sprintf("    → UnoC ERROR: %s", e$message))
          NA_real_
        })
      }
      # Now u is always defined
      unoC_val <- if (is.finite(u)) u else NA_real_
      if (verbose && (row == 1L || row %% log_every == 0L)) {
        surv_f <- surv_mat[is.finite(surv_mat)]
        surv_min <- if (length(surv_f)) min(surv_f) else NA_real_
        surv_max <- if (length(surv_f)) max(surv_f) else NA_real_
        in01 <- length(surv_f) > 0L && all(surv_f >= 0 & surv_f <= 1)
        rmst_min <- if (any(is.finite(df_test$rmst_score))) min(df_test$rmst_score, na.rm = TRUE) else NA_real_
        rmst_max <- if (any(is.finite(df_test$rmst_score))) max(df_test$rmst_score, na.rm = TRUE) else NA_real_
        message(sprintf(
          "Rep %d Imp %d Fold %d | surv=[%.4f, %.4f] in01=%s NA=%d rowNA=%d/%d | RMST=[%.2f, %.2f] | G(c_tau)=%.3g tau_uno=%.3f evt<=tau=%d UnoC=%s",
          r, m, i,
          surv_min, surv_max, in01,
          sum(is.na(surv_mat)), sum(!complete.cases(surv_mat)), nrow(surv_mat),
          rmst_min, rmst_max,
          g_ctau, tau_uno, n_evt_tau, format(unoC_val, digits = 6)
        ))
      }
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # METRIC 4: RMSTs and Uno's C for time horizons
      #_#_#_#_#_#_#_#_#_#_#_#_#_
      # Full diagnostic output (Version 1)
      # message(sprintf(
      #   "Rep %d | Imp %d | Fold %d: G(c_tau)=%.3g tau_uno=%.3f events<=tau=%d UnoC=%s",
      #   r, m, i, g_ctau, tau_uno, n_evt_tau, as.character(unoC_val)
      # ))
      c_horizons_vec <- rep(NA_real_, H)
      for (h in seq_along(horizons_month)) {
        t0 <- horizons_month[h]
        idx <- which(eval_times <= t0)
        if (length(idx) < 2) next
        rmst_t0 <- as.numeric(surv_mat[, idx[-length(idx)], drop = FALSE] %*% diff(eval_times[idx]))
        risk_t0 <- -rmst_t0  # ← defined but not used!
        c_horizons_vec[h] <- tryCatch(
          survival::concordance(
            Surv(df_test[[time_col]], df_test[[event_col]]) ~ risk_t0,
            timewt = "n/G2", ymax = t0, reverse = TRUE
          )$concordance,
          error = function(e) NA_real_
        )
      }
      list(
        r = r, m = m, i = i, c_uno = c_uno_val, unoC = unoC_val,
        ibs = ibs_val, ibs_ipcw = ibs_ipcw_val,
        c_horizons = c_horizons_vec
      )
    },
    future.seed = TRUE,
    future.packages = c("survival", "pec", "survex", "survAUC", "dplyr")
  )
    # Aggregation AFTER the parallel loop
    for (res in all_results) {
      c_array[res$r, res$m, res$i]    <- res$c_uno
      unoC_array[res$r, res$m, res$i] <- res$unoC
      ibs_array[res$r, res$m, res$i]  <- res$ibs
      ibs_ipcw_array[res$r, res$m, res$i] <- res$ibs_ipcw
      c_horizon[res$r, res$m, res$i, ] <- res$c_horizons
    }
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  # AGGREGATE & POOL
  #_#_#_#_#_#_#_#_#_#_#_#_#_----------------
  rep_mean <- function(arr3) apply(arr3, 1, function(x) mean(x, na.rm = TRUE))
  c_rep <- rep_mean(c_array)
  unoC_rep <- rep_mean(unoC_array)
  ibs_rep <- rep_mean(ibs_array)
  ibs_ipcw_rep <- rep_mean(ibs_ipcw_array)
  ci95_c <- quantile(c_rep, c(0.025, 0.975), na.rm = TRUE)
  ci95_unoC <- quantile(unoC_rep, c(0.025, 0.975), na.rm = TRUE)
  ci95_ibs <- quantile(ibs_rep, c(0.025, 0.975), na.rm = TRUE)
  ci95_ibs_ipcw <- quantile(ibs_ipcw_rep, c(0.025, 0.975), na.rm = TRUE)
  
  final_pooled_ibs_ipcw <- mean(ibs_ipcw_rep, na.rm = TRUE)
  final_pooled_c <- mean(c_rep, na.rm = TRUE)
  final_pooled_unoC <- mean(unoC_rep, na.rm = TRUE)
  final_pooled_ibs <- mean(ibs_rep, na.rm = TRUE)
  h_rep <- sapply(seq_len(H), function(h) {
    apply(c_horizon[,,,h], 1, function(x) {
      x <- x[is.finite(x)]
      if (length(x) == 0L) NA_real_ else mean(x)
    })
  })
  rmst_h_summary <- data.frame(
    horizon_month = horizons_month,
    c_mean = colMeans(h_rep, na.rm = TRUE),
    ci95_lower = apply(h_rep, 2, quantile, probs = 0.025, na.rm = TRUE),
    ci95_upper = apply(h_rep, 2, quantile, probs = 0.975, na.rm = TRUE),
    valid_repeats = colSums(is.finite(h_rep))
  )
  cat("\n============================================================\n")
  cat("FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)\n")
  cat("============================================================\n")
  cat(sprintf("Global Evaluation Time (tau): %.2f months\n", tau))
  cat(sprintf("Pooled Uno's C-Index: %.4f (95%% CI %.4f-%.4f)\n",
            final_pooled_c, ci95_c[1], ci95_c[2]))
  cat(sprintf("Pooled UnoC (survAUC::UnoC): %.4f (95%% CI %.4f-%.4f)\n",
              final_pooled_unoC, ci95_unoC[1], ci95_unoC[2]))
  cat(sprintf("Pooled IBS: %.4f (95%% CI %.4f-%.4f)\n",
              final_pooled_ibs, ci95_ibs[1], ci95_ibs[2]))
  cat(sprintf("Pooled IBS (IPCW-train): %.4f (95%% CI %.4f-%.4f)\n",
              final_pooled_ibs_ipcw, ci95_ibs_ipcw[1], ci95_ibs_ipcw[2]))               
  cat("RMST-based Uno C-index by horizon:\n")
    for (hh in seq_len(nrow(rmst_h_summary))) {
      cat(sprintf("  t=%d m: C=%.4f (95%% CI %.4f-%.4f, valid repeats=%d)\n",
                  rmst_h_summary$horizon_month[hh],
                  rmst_h_summary$c_mean[hh],
                  rmst_h_summary$ci95_lower[hh],
                  rmst_h_summary$ci95_upper[hh],
                  rmst_h_summary$valid_repeats[hh]))
    }
  cat("============================================================\n")
  
  si <- utils::sessionInfo()
  other_pkgs <- if (is.null(si$otherPkgs)) character(0) else names(si$otherPkgs)
  
  t_end <- Sys.time()
  elapsed_sec <- as.numeric(difftime(t_end, t_start, units = "secs"))
  cat(sprintf("\nTotal runtime: %.2f sec (%.2f min)\n", elapsed_sec, elapsed_sec / 60))
  return(invisible(list(
    raw_c_array = c_array,
    raw_unoC_array = unoC_array,
    raw_ibs_array = ibs_array,
    raw_ibs_ipcw_array = ibs_ipcw_array, 
    ci95_pooled_c = unname(ci95_c),
    ci95_pooled_unoC = unname(ci95_unoC),
    ci95_pooled_ibs = unname(ci95_ibs),
    ci95_pooled_ibs_ipcw = unname(ci95_ibs_ipcw),
    repeat_values = list(c = c_rep, unoC = unoC_rep,
                         ibs = ibs_rep, ibs_ipcw = ibs_ipcw_rep),
    reproducibility = list(
      call = match.call(),
      seed = seed,
      k = k,
      M = M,
      time_col = time_col,
      event_col = event_col,
      formula = paste(deparse(formula), collapse = " "),
      tau = tau,
      c_tau = c_tau,                            # NEW
      n_eval_times = n_eval_times,
      eval_times = eval_times,
      fold_sizes = fold_sizes,
      rmst_unoC_by_horizon = c_horizon,
      rmst_unoC_summary = rmst_h_summary,
      cv_strata_table = cv_strata_table,
      cv_strata_by_fold = cv_strata_by_fold,
      timestamp = as.character(Sys.time()),
      R_version = R.version.string,
      attached_packages = other_pkgs
    )
  )))
}

#survex::integrated_brier_score uses the test-set censoring distribution for IPCW weights. Your new ibs_ipcw_train uses the training-set censoring KM, which is the methodologically correct approach for out-of-sample evaluation 

library(survival)
library(rms)
library(pec)
library(future)
library(future.apply)
library(dplyr)

#' Bootstrap OOB IBS Evaluation for Cox Models with Multiple Imputation
#' 
#' Fully corrected version addressing:
#' - Missing helper functions
#' - Undefined variables
#' - Stratified model handling
#' - Robust error handling
#' - Proper event value extraction

cph_evaluate_boot_oob_mi_corrected <- function(
    formula,
    imputed_list,
    time_col,
    event_col,
    tau = NULL,
    B = 200,
    n_eval_times = 100,
    seed = 2125,
    cpus = 8,
    verbose = TRUE,
    min_oob_n = 10,
    min_oob_events = 3,
    min_valid_imputations = NULL,
    id_col = NULL
) {
    # =========================================================================
    # HELPER FUNCTIONS (previously missing)
    # =========================================================================
    is_pos_int <- function(x) {
        is.numeric(x) && length(x) == 1 && !is.na(x) && x > 0 && x == as.integer(x)
    }
    # =========================================================================
    # 1) INPUT VALIDATION
    # =========================================================================
    t_start <- Sys.time()
    if (!inherits(formula, "formula")) {
        stop("formula must be a formula.")
    }
    if (!is.list(imputed_list) || length(imputed_list) == 0L) {
        stop("imputed_list must be a non-empty list.")
    }
    if (!is.character(time_col) || length(time_col) != 1L) {
        stop("time_col must be a single column name.")
    }
    if (!is.character(event_col) || length(event_col) != 1L) {
        stop("event_col must be a single column name.")
    }
    if (!is_pos_int(B)) stop("B must be a positive integer.")
    if (!is_pos_int(min_oob_n)) stop("min_oob_n must be a positive integer.")
    if (!is_pos_int(min_oob_events)) stop("min_oob_events must be a positive integer.")
    if (!is_pos_int(n_eval_times)) stop("n_eval_times must be a positive integer.")
    if (!is_pos_int(cpus)) stop("cpus must be a positive integer.")
    
    B <- as.integer(B)
    min_oob_n <- as.integer(min_oob_n)
    min_oob_events <- as.integer(min_oob_events)
    n_eval_times <- as.integer(n_eval_times)
    cpus <- as.integer(cpus)
    # Convert to data.frames
    df_list <- lapply(imputed_list, as.data.frame)
    M <- length(df_list)
    # Check consistent row counts
    n_by_imp <- vapply(df_list, nrow, integer(1))
    if (length(unique(n_by_imp)) != 1L) {
        stop("All imputations must have same number of rows.")
    }
    n <- n_by_imp[1]
    # Extract model variables from formula
    model_vars <- unique(all.vars(formula))
    required_cols <- unique(c(model_vars, time_col, event_col))
    # Check all required columns exist
    for (m in seq_len(M)) {
        miss <- setdiff(required_cols, names(df_list[[m]]))
        if (length(miss) > 0L) {
            stop(sprintf("Imputation %d is missing columns: %s", 
                        m, paste(miss, collapse = ", ")))
        }
    }
    # Check ID column consistency
    if (!is.null(id_col)) {
        if (!is.character(id_col) || length(id_col) != 1L) {
            stop("id_col must be NULL or a single column name.")
        }
        if (!all(vapply(df_list, function(d) id_col %in% names(d), logical(1)))) {
            stop("id_col not found in at least one imputation.")
        }
        ref_id <- df_list[[1]][[id_col]]
        same_order <- vapply(df_list, function(d) identical(d[[id_col]], ref_id), logical(1))
        if (!all(same_order)) {
            stop("Row order differs across imputations (checked with id_col).")
        }
    }
    # Set default for min_valid_imputations
    if (is.null(min_valid_imputations)) {
        min_valid_imputations <- M
    }
    if (!is_pos_int(min_valid_imputations) || min_valid_imputations > M) {
        stop("min_valid_imputations must be an integer between 1 and M.")
    }
    min_valid_imputations <- as.integer(min_valid_imputations)
    # =========================================================================
    # 2) EXTRACT EVENT VALUE AND CREATE INDICATORS
    # =========================================================================
    # Auto-detect event value from first imputation
    event_values <- unique(df_list[[1]][[event_col]])
    event_values <- event_values[!is.na(event_values)]
    if (length(event_values) == 0) {
        stop("No non-missing event values found.")
    }
    # Use highest value as event (convention: 0=censored, 1=event, or 1=censored, 2=event)
    event_value <- max(event_values)
    if (verbose && length(event_values) > 2) {
        warning("More than 2 unique event values detected. Using max as event: ", event_value)
    }
    # Create event indicators (1 = event, 0 = censored)
    event_indicator <- lapply(df_list, function(d) {
        as.integer(d[[event_col]] == event_value)
    })
    # =========================================================================
    # 3) SET UP EVALUATION TIMES AND FORMULAS
    # =========================================================================
    # Determine tau if not provided
    if (is.null(tau)) {
        all_times <- unlist(lapply(df_list, function(d) d[[time_col]]), use.names = FALSE)
        tau <- as.numeric(stats::quantile(all_times, probs = 0.90, na.rm = TRUE))
        if (verbose) {
            message("Auto-selected tau (90th percentile): ", round(tau, 2))
        }
    }
    if (!is.numeric(tau) || length(tau) != 1L || !is.finite(tau) || tau <= 0) {
        stop("tau must be a finite positive scalar.")
    }
    # Evaluation time grid
    eval_times <- seq(1e-6, tau, length.out = n_eval_times)
    fallback_times <- rev(eval_times[eval_times < tau])
    # Censoring formula for pec
    cens_fml <- stats::as.formula(sprintf("Hist(%s, %s) ~ 1", time_col, event_col))
    # Convert strata() to strat() for rms::cph
    model_formula_txt <- paste(deparse(formula), collapse = " ")
    model_formula_txt <- gsub("\\bstrata\\(", "strat(", model_formula_txt)
    model_formula <- stats::as.formula(model_formula_txt, env = environment(formula))
    # Check for stratification
    has_strata <- grepl("strat\\(", model_formula_txt)
    if (verbose && has_strata) {
        message("Stratified model detected: ", 
                "ensure stratification variable has consistent levels across imputations.")
    }
    model_name <- "CPH_Model"
    # =========================================================================
    # 4) IBS EXTRACTION FUNCTION (robust version)
    # =========================================================================
    extract_ibs <- function(pe_obj, t_value, model_name) {
        # Validate pec object
        if (!inherits(pe_obj, "pec")) {
            return(list(value = NA_real_, err = "invalid_pec_object"))
        }
        # Extract CRPS at specific time
        cr <- tryCatch(
            suppressWarnings(pec::crps(pe_obj, times = t_value)),
            error = function(e) return(NULL)
        )
        if (is.null(cr)) {
            return(list(value = NA_real_, err = "crps_extraction_failed"))
        }
        # Extract value for our model
        val <- NA_real_
        if (is.matrix(cr) || is.data.frame(cr)) {
            if (model_name %in% rownames(cr)) {
                val <- as.numeric(cr[model_name, 1])
            } else if (nrow(cr) > 0) {
                val <- as.numeric(cr[1, 1])
                if (verbose) {
                    warning("Model name '", model_name, "' not found, using first available.")
                }
            }
        } else {
            val <- as.numeric(cr[1])
        }
        if (!is.finite(val)) {
            return(list(value = NA_real_, err = "non_finite_crps"))
        }
        list(value = val, err = "")
    }
    # =========================================================================
    # 5) PARALLEL SETUP
    # =========================================================================
    detected <- suppressWarnings(parallel::detectCores(logical = TRUE))
    if (!is.numeric(detected) || length(detected) != 1L || !is.finite(detected) || detected < 1L) {
        detected <- 1L
    }
    n_cores <- as.integer(max(1L, min(cpus, B, max(1L, detected - 1L))))
    old_plan <- future::plan()
    on.exit(future::plan(old_plan), add = TRUE)
    if (n_cores > 1L) {
        future::plan(future::multisession, workers = n_cores)
    } else {
        future::plan(future::sequential)
    }
    if (verbose) {
        cat(sprintf("Starting OOB bootstrap + MI: B=%d, M=%d\n", B, M))
        cat(sprintf("Evaluating IBS up to tau = %.4f\n", tau))
        cat(sprintf("Using %d worker(s)...\n", n_cores))
        cat("------------------------------------------------------------\n")
    }
    # =========================================================================
    # 6) PARALLEL BOOTSTRAP LOOP
    # =========================================================================
    boot_results <- future.apply::future_lapply(
        X = seq_len(B),
        FUN = function(b) {
            row_ibs <- rep(NA_real_, M)
            row_fails <- rep("", M)
            split_fail <- ""
            # Bootstrap sample
            idx_train <- sample.int(n = n, size = n, replace = TRUE)
            idx_test <- setdiff(seq_len(n), unique(idx_train))
            # Check OOB size
            if (length(idx_test) < min_oob_n) {
                split_fail <- "too_few_oob_rows"
                return(list(ibs = row_ibs, fails = row_fails, split_fail = split_fail))
            }
            # Process each imputation
            for (m in seq_len(M)) {
                # Check OOB events
                oob_events <- sum(event_indicator[[m]][idx_test] == 1L, na.rm = TRUE)
                if (oob_events < min_oob_events) {
                    row_fails[m] <- "too_few_oob_events"
                    next
                }
                # Check events before tau
                n_evt_tau <- sum(
                    event_indicator[[m]][idx_test] == 1L & 
                    df_list[[m]][idx_test, time_col] <= tau,
                    na.rm = TRUE
                )
                if (n_evt_tau < 1L) {
                    row_fails[m] <- "no_events_before_tau"
                    next
                }
                # Split data
                df_train <- df_list[[m]][idx_train, , drop = FALSE]
                df_test <- df_list[[m]][idx_test, , drop = FALSE]
                # Handle factor levels
                fchar_cols <- intersect(
                    model_vars,
                    names(df_train)[vapply(df_train, function(x) is.factor(x) || is.character(x), logical(1))]
                )
                bad_col <- NA_character_
                if (length(fchar_cols) > 0L) {
                    for (fc in fchar_cols) {
                        tr_chr <- as.character(df_train[[fc]])
                        te_chr <- as.character(df_test[[fc]])
                        tr_levels <- unique(tr_chr[!is.na(tr_chr)])
                        unseen <- !is.na(te_chr) & !(te_chr %in% tr_levels)
                        if (any(unseen)) {
                            bad_col <- fc
                            break
                        }
                        df_train[[fc]] <- factor(tr_chr, levels = tr_levels)
                        df_test[[fc]] <- factor(te_chr, levels = tr_levels)
                    }
                }
                if (!is.na(bad_col)) {
                    row_fails[m] <- paste0("unseen_factor_level:", bad_col)
                    next
                }
                # Fit model and evaluate
                res <- tryCatch({
                    fit <- rms::cph(
                        model_formula,
                        data = df_train,
                        x = TRUE,
                        y = TRUE,
                        surv = TRUE,
                        time.inc = tau
                    )
                    # Ensure data reference for pec
                    fit$call$data <- quote(df_train)
                    # Calculate prediction error
                    pec_args <- list(
                        object = stats::setNames(list(fit), model_name),
                        data = df_test,
                        formula = cens_fml,
                        splitMethod = "none",
                        times = eval_times,
                        cens.model = "marginal",
                        exact = TRUE,
                        verbose = FALSE
                    )
                    pe <- do.call(pec::pec, pec_args)
                    ibs_res <- extract_ibs(pe, tau, model_name = model_name)
                    # Try fallback times if needed
                    if (is.na(ibs_res$value) && length(fallback_times) > 0L) {
                        for (tt in fallback_times) {
                            ibs_try <- extract_ibs(pe, tt, model_name = model_name)
                            if (!is.na(ibs_try$value)) {
                                ibs_res <- ibs_try
                                break
                            }
                        }
                    }
                    if (is.na(ibs_res$value)) {
                        list(value = NA_real_, err = ibs_res$err)
                    } else {
                        list(value = ibs_res$value, err = "")
                    }
                }, error = function(e) {
                    list(value = NA_real_, err = paste0("cph_or_pec_error:", conditionMessage(e)))
                })
                row_ibs[m] <- res$value
                if (nzchar(res$err)) row_fails[m] <- res$err
            }
            list(ibs = row_ibs, fails = row_fails, split_fail = split_fail)
        },
        future.seed = seed,
        future.packages = c("survival", "pec", "rms")
    )
    # =========================================================================
    # 7) RESULTS AGGREGATION
    # =========================================================================
    results_matrix <- do.call(rbind, lapply(boot_results, `[[`, "ibs"))
    fail_matrix <- do.call(rbind, lapply(boot_results, `[[`, "fails"))
    split_fail_reason <- vapply(boot_results, `[[`, "", "split_fail")
    # Pool across imputations for each bootstrap
    boot_pooled <- apply(results_matrix, 1, function(x) {
        valid <- x[is.finite(x)]
        if (length(valid) < min_valid_imputations) NA_real_ else mean(valid)
    })
    boot_pooled <- boot_pooled[is.finite(boot_pooled)]
    # Check for complete failure
    if (length(boot_pooled) < 2L) {
        cat("\n[!] CRITICAL FAILURE: Almost all bootstrap replicates failed.\n")
        cat("\n--- TOP REASONS FOR FAILURE ---\n")
        all_fails <- c(
            as.vector(fail_matrix[fail_matrix != ""]), 
            split_fail_reason[split_fail_reason != ""]
        )
        if (length(all_fails) > 0L) {
            print(sort(table(all_fails), decreasing = TRUE))
        } else {
            cat("No explicit errors captured.\n")
        }
        cat("-------------------------------\n")
        stop("Too few valid bootstrap replicates to estimate IBS or confidence intervals.")
    }
    # Final statistics
    pooled_ibs <- mean(boot_pooled)
    ci <- stats::quantile(boot_pooled, probs = c(0.025, 0.975), na.rm = TRUE, names = FALSE)
    # =========================================================================
    # 8) OUTPUT
    # =========================================================================
    if (verbose) {
        cat("\n============================================================\n")
        cat("FINAL POOLED RESULTS (OOB Bootstrap + MI)\n")
        cat("============================================================\n")
        cat(sprintf("Pooled OOB IBS: %.4f (95%% CI: %.4f - %.4f)\n", 
                   pooled_ibs, ci[1], ci[2]))
        cat(sprintf("Valid bootstrap replicates: %d / %d\n", 
                   length(boot_pooled), B))
        cat(sprintf("Event value used: %s\n", as.character(event_value)))
        cat("============================================================\n")
        cat(sprintf("Total runtime: %.2f min\n", 
                   as.numeric(difftime(Sys.time(), t_start, units = "mins"))))
    }
    invisible(list(
        pooled_ibs = pooled_ibs,
        ci_95 = c(lower = ci[1], upper = ci[2]),
        boot_pooled = boot_pooled,
        imputation_means = colMeans(results_matrix, na.rm = TRUE),
        raw_matrix = results_matrix,
        fail_matrix = fail_matrix,
        split_fail_reason = split_fail_reason,
        event_value = event_value,
        reproducibility = list(
            call = match.call(),
            seed = seed,
            B = B,
            M = M,
            tau = tau,
            min_oob_n = min_oob_n,
            min_oob_events = min_oob_events,
            min_valid_imputations = min_valid_imputations,
            event_value = event_value,
            n_eval_times = n_eval_times,
            n_cores = n_cores,
            valid_boot = length(boot_pooled),
            has_strata = has_strata
        )
    ))
}


#' Calibration Analysis and Plotting for Cox Models
#' 
#' Performs calibration analysis using riskRegression methods and generates
#' calibration plots (predicted vs observed risk) at multiple time horizons.

#' Calibration Analysis for Cox Models (Corrected Version)
#' 
#' Addresses Gemini audit findings:
#' 1. Generic variable extraction from formula
#' 2. No look-ahead bias using summary(survfit, times = ...)
#' 3. Proper confidence intervals from survfit
#' 4. Correct terminology (ECE not ICI)
#' 5. Explicit namespace usage


calibrate_cph_corrected <- function(
        formula,
        data,
        times = c(12, 36, 60),  # Months
        seed = 2125L,
        n_imputations = NULL
) {
    set.seed(seed)
    
    # =========================================================================
    # FORMULA PARSING
    # =========================================================================
    surv_call <- formula[[2]]
    if (surv_call[[1]] != as.name("Surv")) {
        stop("Formula must have a Surv() response variable.")
    }
    
    time_var <- as.character(surv_call[[2]])
    event_var <- as.character(surv_call[[3]])
    
    message("Detected time variable: ", time_var)
    message("Detected event variable: ", event_var)
    
    # =========================================================================
    # DATA PREPARATION (Time conversion removed)
    # =========================================================================
    if (is.list(data) && !is.data.frame(data)) {
        df_list <- data
        if (!is.null(n_imputations)) {
            df_list <- df_list[1:min(n_imputations, length(df_list))]
        }
    } else {
        df_list <- list(data)
    }
    
    M <- length(df_list)
    results_all <- list()
    plots_data <- list()
    
    # =========================================================================
    # MAIN LOOP
    # =========================================================================
    for (i in seq_len(M)) {
        message("Processing imputation ", i, "/", M, "...", appendLF = FALSE)
        d <- df_list[[i]]
        
        # Fit Cox model
        fit <- survival::coxph(formula, data = d, x = TRUE, y = TRUE)
        
        for (j in seq_along(times)) {
            t_months <- times[j]
            
            #_#_#_#_#_#_#_#_#_#_#_#_#_------------------------
            # 1. PREDICTED RISK (NATIVE STRATA HANDLING)
            # Temporarily set follow-up time to target horizon for all patients
            #_#_#_#_#_#_#_#_#_#_#_#_#_------------------------
            d_temp <- d
            d_temp[[time_var]] <- t_months
            
            # Predict expected cumulative hazard (handles all strata dynamically)
            cum_haz <- stats::predict(fit, newdata = d_temp, type = "expected")
            
            # Risk = 1 - S(t) = 1 - exp(-Lambda(t))
            pred_risk <- 1 - exp(-cum_haz)
            pred_risk[is.na(pred_risk)] <- 0
            
            # Create Deciles
            deciles <- dplyr::ntile(pred_risk, 10)
            
            #_#_#_#_#_#_#_#_#_#_#_#_#_------------------------
            # 2. OBSERVED RISK (KAPLAN-MEIER BY DECILE)
            #_#_#_#_#_#_#_#_#_#_#_#_#_------------------------
            calib_summary <- lapply(1:10, function(dec) {
                idx <- which(deciles == dec)
                if (length(idx) < 10) return(NULL)
                
                mean_pred <- mean(pred_risk[idx], na.rm = TRUE)
                d_sub <- d[idx, ]
                
                # Unstratified KM for actual observed event rate in this decile
                km_formula <- stats::as.formula(
                    paste("survival::Surv(", time_var, ", ", event_var, ") ~ 1")
                )
                km_fit <- survival::survfit(km_formula, data = d_sub)
                
                # Extract survival at exact time
                km_sum <- base::summary(km_fit, times = t_months, extend = TRUE)
                
                obs_risk <- 1 - km_sum$surv[1]
                obs_upper <- 1 - km_sum$lower[1]
                obs_lower <- 1 - km_sum$upper[1]
                
                if (is.na(obs_lower)) obs_lower <- obs_risk
                if (is.na(obs_upper)) obs_upper <- obs_risk
                
                data.frame(
                    imputation = i,
                    time_months = t_months,
                    decile = dec,
                    mean_predicted = mean_pred,
                    observed = obs_risk,
                    observed_lower = obs_lower,
                    observed_upper = obs_upper,
                    n_patients = length(idx),
                    stringsAsFactors = FALSE
                )
            })
            calib_df <- do.call(rbind, calib_summary)
            if (!is.null(calib_df)) {
                plots_data[[length(plots_data) + 1]] <- calib_df
                ece <- stats::weighted.mean(
                    abs(calib_df$mean_predicted - calib_df$observed),
                    w = calib_df$n_patients
                )
                ici <- tryCatch({
                    smooth_obs <- stats::loess(observed ~ mean_predicted, 
                                               data = calib_df, 
                                               weights = n_patients)$fitted
                    stats::weighted.mean(abs(calib_df$mean_predicted - smooth_obs),
                                         w = calib_df$n_patients)
                }, error = function(e) NA)

                km_form_overall <- stats::as.formula(
                    paste("survival::Surv(", time_var, ", ", event_var, ") ~ 1")
                )
                km_overall <- survival::survfit(km_form_overall, data = d)
                km_sum_overall <- base::summary(km_overall, times = t_months, extend = TRUE)
                obs_risk_overall <- 1 - km_sum_overall$surv[1]
                exp_risk_overall <- mean(pred_risk, na.rm = TRUE)
                # E/O Ratio: 1.0 is perfect. < 1.0 means under-prediction.
                eo_ratio <- exp_risk_overall / obs_risk_overall
                
                results_all[[length(results_all) + 1]] <- data.frame(
                    imputation = i,
                    time_months = t_months,
                    ece = ece,
                    ici = ici,
                    eo_ratio = eo_ratio,
                    mean_pred_risk = exp_risk_overall,
                    n_events = sum(d[[time_var]] <= t_months & d[[event_var]] == 1, na.rm = TRUE),
                    total_n = nrow(d),
                    stringsAsFactors = FALSE
                )
            }
        }
        message(" done")
    }
    
    # =========================================================================
    # POOLING AND PLOTTING
    # =========================================================================
    by_imp <- dplyr::bind_rows(results_all)
    plots_df <- dplyr::bind_rows(plots_data)
    
    pooled_curves <- plots_df %>%
        dplyr::group_by(time_months, decile) %>%
        dplyr::summarise(
            mean_predicted = mean(mean_predicted),
            observed_pooled = mean(observed),
            observed_lower = mean(observed_lower, na.rm = TRUE),
            observed_upper = mean(observed_upper, na.rm = TRUE),
            n_total = sum(n_patients),
            .groups = "drop"
        )
    
    pooled_summary <- by_imp %>%
        dplyr::group_by(time_months) %>%
        dplyr::summarise(
            ece_mean = mean(ece, na.rm = TRUE),
            ece_sd = stats::sd(ece, na.rm = TRUE),
            ici_mean = mean(ici, na.rm = TRUE),
            eo_mean = mean(eo_ratio, na.rm = TRUE),
            eo_sd = stats::sd(eo_ratio, na.rm = TRUE),
            n_imp = dplyr::n(),
            .groups = "drop"
        )
    
    plot_list <- list()
    for (t in times) {
        p_data <- pooled_curves %>% dplyr::filter(time_months == t)
        
        p <- ggplot2::ggplot(p_data, ggplot2::aes(x = mean_predicted, y = observed_pooled)) +
            ggplot2::geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
            ggplot2::geom_errorbar(
                ggplot2::aes(ymin = observed_lower, ymax = observed_upper),
                width = 0.005, alpha = 0.6, color = "steelblue"
            ) +
            ggplot2::geom_point(size = 3, color = "steelblue", fill = "white", shape = 21) +
            ggplot2::geom_line(ggplot2::aes(group = 1), color = "steelblue", alpha = 0.5) +
            ggplot2::labs(
                title = paste(t, "Months"),
                subtitle = paste0("ECE = ", formatC(pooled_summary$ece_mean[pooled_summary$time_months == t], digits = 4),
                                  " | ICI = ", formatC(pooled_summary$ici_mean[pooled_summary$time_months == t], digits = 4)),
                x = "Predicted Probability",
                y = "Observed Proportion"
            ) +
            ggplot2::coord_equal(
                xlim = c(0, max(p_data$mean_predicted, na.rm = TRUE) * 1.1),
                ylim = c(0, max(p_data$observed_pooled, na.rm = TRUE) * 1.1)
            ) +
            ggplot2::theme_minimal()
        
        plot_list[[paste0("month_", t)]] <- p
    }
    
    combined_plot <- ggplot2::ggplot(pooled_curves, ggplot2::aes(x = mean_predicted, y = observed_pooled)) +
        ggplot2::geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
        ggplot2::geom_errorbar(ggplot2::aes(ymin = observed_lower, ymax = observed_upper), width = 0.005, alpha = 0.5, color = "steelblue") +
        ggplot2::geom_point(ggplot2::aes(color = as.factor(time_months)), size = 2) +
        ggplot2::geom_line(ggplot2::aes(group = time_months, color = as.factor(time_months)), alpha = 0.5) +
        ggplot2::facet_wrap(~time_months, labeller = ggplot2::label_both) +
        ggplot2::labs(title = "Calibration Plots (ECE/ICI)", x = "Predicted", y = "Observed") +
        ggplot2::coord_equal() +
        ggplot2::theme_minimal() +
        ggplot2::theme(legend.position = "none")
    
    structure(list(
        by_imputation = by_imp,
        pooled_summary = pooled_summary,
        calibration_curves = pooled_curves,
        plots = plot_list,
        combined_plot = combined_plot,
        variable_names = list(time = time_var, event = event_var),
        n_imputations = M
    ), class = "cph_calibration_corrected")
}

#' Updated Print Method (Removed misleading thresholds)
#' Updated Print Method (Includes E/O Ratio)
print.cph_calibration_corrected <- function(x, ...) {
    cat("Cox Model Calibration Analysis (Corrected)\n")
    cat("==========================================\n")
    cat("Time variable:", x$variable_names$time, "\n")
    cat("Event variable:", x$variable_names$event, "\n")
    cat("Imputations:", x$n_imputations, "\n\n")
    
    cat("Pooled Results:\n")
    print(x$pooled_summary %>% 
              dplyr::select(time_months, ece_mean, ici_mean, eo_mean, n_imp), 
          row.names = FALSE)
    
    cat("\nNote: ECE = Expected Calibration Error (binned)\n")
    cat("      ICI = Integrated Calibration Index (smoothed)\n")
    cat("      E/O = Expected / Observed Ratio (1.0 is perfect, < 1.0 means under-prediction)\n")
    cat("      Lower values for ECE/ICI are better. Interpret alongside visual plots.\n")
    
    invisible(x)
}


compare_mi_models_d1 <- function(formula_full, formula_reduced, imputed_list) {
  # 1. Validate that the input is a list of dataframes
  if (!is.list(imputed_list) || length(imputed_list) == 0) {
    stop("imputed_list must be a list of dataframes.")
  }
  cat("Fitting the FULL model on", length(imputed_list), "imputations...\n")
  fits_full <- lapply(imputed_list, function(df) {
    survival::coxph(formula_full, data = df)
  })
  cat("Fitting the REDUCED model on", length(imputed_list), "imputations...\n")
  fits_reduced <- lapply(imputed_list, function(df) {
    survival::coxph(formula_reduced, data = df)
  })
  # 2. Convert to MIRA objects (Multiple Imputation Repeated Analyses)
  mira_full    <- mice::as.mira(fits_full)
  mira_reduced <- mice::as.mira(fits_reduced)
  # 3. Execute the Multivariate Wald Test (D1)
  cat("Executing the Wald Test (D1)...\n")
  comparison <- mice::D1(mira_full, mira_reduced)
  cat("\n============================================================\n")
  cat("MODEL COMPARISON RESULT (D1 Statistic)\n")
  cat("============================================================\n")
  print(comparison)
  cat("============================================================\n")
  # Quick interpretation printed to console
  p_val <- comparison$p.value[1]
}
compute_mi_aicc <- function(formula, imputed_list) {
    # Adjust the model in each imputation
    fits <- lapply(imputed_list, function(df) {
        survival::coxph(formula, data = df)
    })
    # Extract metrics from each model
    results <- lapply(fits, function(fit) {
        aic_val <- stats::AIC(fit)
        k <- length(fit$coefficients) # No. of parameters (df)
        n_events <- fit$nevent        # Effective sample size (no. of events)
        # Calcular la corrección para muestras pequeñas
        correction <- (2 * k * (k + 1)) / (n_events - k - 1)
        aicc_val <- aic_val + correction
        c(aic = aic_val, correction = correction, aicc = aicc_val, k = k, n_events = n_events)
    })
    # As dataframe
    res_df <- as.data.frame(do.call(rbind, results))
    mean_aic <- mean(res_df$aic)
    mean_aicc <- mean(res_df$aicc)
    mean_correction <- mean(res_df$correction)
    cat("\n============================================================\n")
    cat("AICc ANALYSIS FOR MULTIPLY IMPUTED DATA\n")
    cat("============================================================\n")
    cat(sprintf("Mean AIC:  %.2f\n", mean_aic))
    cat(sprintf("Mean AICc: %.2f\n", mean_aicc))
    cat(sprintf("Correction added (Penalty): %.6f\n", mean_correction))
    cat(sprintf("Parameters (k): %d | Effective Sample Size (Events): %d\n", 
                res_df$k[1], res_df$n_events[1]))
    cat("============================================================\n")
    return(list(mean_aicc = mean_aicc, full_results = res_df))
}
plot_schoenfeld_residuals <- function(formula, variable, df, span = 0.50) {
    
    df <- base::as.data.frame(base::lapply(base::as.data.frame(df), base::as.vector))
    .GlobalEnv$.tmp_df <- df
    base::on.exit(base::rm(".tmp_df", envir = .GlobalEnv))
    
    fit <- survival::coxph(formula, data = .tmp_df)
    res <- stats::residuals(fit, type = "scaledsch")
    time <- base::as.numeric(base::rownames(res))
    
    # Filter range (10th to 90th percentile)
    t_range <- stats::quantile(time, c(0.1, 0.9), na.rm = TRUE)
    valid <- time >= t_range[1] & time <= t_range[2]
    
    time_clean <- time[valid]
    res_clean <- res[valid, variable]
    
    z2 <- stats::loess(res_clean ~ time_clean, span = span)
    
    # THE FIX: Create an ordering index based on time
    ord <- base::order(time_clean)
    
    # Apply the ordering index to both x and y when plotting
    graphics::plot(time_clean[ord], stats::fitted(z2)[ord], type = "l", col = "blue", lwd = 2,
         main = base::paste("Sch. res. –", variable),
         xlab = "Time", ylab = "Residuals")
         
    # stats::supsmu() handles sorting internally, so it's safe as-is
    graphics::lines(stats::supsmu(time_clean, res_clean), lty = 2, col = "darkgray")
    graphics::abline(h = 0, col = "red", lty = 3)
}
```

### Evaluate models

```{r}
#| label: update-gvif-cindex-ibs

total_cpus <- parallel::detectCores()
cpus <- ifelse(total_cpus > 30, 16, 7)

results_readmit <- evaluate_mi_cv(
  formula = formula_readmit,
  imputed_list = py_corrected_datasets,
  time_col = "readmit_time_from_disch_m",
  event_col = "readmit_event",
  k = length(py_corrected_datasets),
  n_repeats = 2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6067 (95% CI 0.6066-0.6068)
# Pooled UnoC (survAUC::UnoC): 0.6679 (95% CI 0.6477-0.6882)
# Pooled IBS: 0.1599 (95% CI 0.1598-0.1599)
# Pooled IBS (IPCW-train): 0.3475 (95% CI 0.3474-0.3476)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6312 (95% CI 0.6308-0.6315, valid repeats=2)
#   t=36 m: C=0.6153 (95% CI 0.6152-0.6153, valid repeats=2)
#   t=60 m: C=0.6099 (95% CI 0.6099-0.6100, valid repeats=2)
# ============================================================
results_death <- evaluate_mi_cv(
  formula = formula_death,
  imputed_list = py_corrected_datasets,
  time_col = "death_time_from_disch_m",
  event_col = "death_event",
  k = length(py_corrected_datasets),
  n_repeats = 2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7456 (95% CI 0.7455-0.7456)
# Pooled UnoC (survAUC::UnoC): 0.7923 (95% CI 0.7618-0.8228)
# Pooled IBS: 0.0351 (95% CI 0.0351-0.0351)
# Pooled IBS (IPCW-train): 0.0886 (95% CI 0.0885-0.0887)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7581 (95% CI 0.7578-0.7583, valid repeats=2)
#   t=36 m: C=0.7634 (95% CI 0.7626-0.7641, valid repeats=2)
#   t=60 m: C=0.7624 (95% CI 0.7618-0.7630, valid repeats=2)
# ============================================================
results_readmit_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_readmit,
  imputed_list = py_corrected_datasets,
  time_col = "readmit_time_from_disch_m",
  event_col = "readmit_event",
  tau = 99, # Your maximum time horizon
  B = 2, # Standard for stable 95% CIs
  seed = 2125,
  cpus = cpus,
  verbose = TRUE
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1582 - 0.1589)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_death,
  imputed_list = py_corrected_datasets,
  time_col = "death_time_from_disch_m",
  event_col = "death_event",
  tau = 107, # Your maximum time horizon
  B = 2, # Standard for stable 95% CIs
  seed = 2125,
  cpus = cpus,
  verbose = TRUE
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0357 (95% CI: 0.0356 - 0.0359)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
cal_readmit <-
  calibrate_cph_corrected(
    formula = formula_readmit, # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    n_imputations = 5
  )
print(cal_readmit)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00298  0.00199   0.990     5
# 2          36  0.00436  0.00349   0.987     5
# 3          60  0.00505  0.00400   0.983     5
# 4          96  0.00482  0.00373   0.975     5
cal_death <-
  calibrate_cph_corrected(
    formula = formula_death, # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    n_imputations = 5
  )
print(cal_death)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00178  0.00173    1.01     5
# 2          36  0.00276  0.00268    1.01     5
# 3          60  0.00319  0.00300    1.02     5
# 4          96  0.00357  0.00322    1.06     5
gc()
```

```{r}
#| label: update-gvif-cindex-ibs2-noncolinear

cox_readmit2 <- coxph(
  formula_readmit_updated,
  data = py_corrected_datasets[[1]]
)
cox_death2 <- coxph(
  formula_death_updated, 
  data = py_corrected_datasets[[1]])
results_readmit2 <- evaluate_mi_cv(
  formula = formula_readmit_updated, 
  imputed_list = py_corrected_datasets, 
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  k = length(py_corrected_datasets),
  n_repeats=2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6067 (95% CI 0.6066-0.6069)
# Pooled UnoC (survAUC::UnoC): 0.6686 (95% CI 0.6480-0.6891)
# Pooled IBS: 0.1599 (95% CI 0.1598-0.1599)
# Pooled IBS (IPCW-train): 0.3475 (95% CI 0.3474-0.3476)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6313 (95% CI 0.6309-0.6316, valid repeats=2)
#   t=36 m: C=0.6154 (95% CI 0.6153-0.6154, valid repeats=2)
#   t=60 m: C=0.6100 (95% CI 0.6099-0.6100, valid repeats=2)
# ============================================================
results_readmit2_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_readmit_updated,
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  tau = 99,                # Evaluate up to 60 months
  B = 2,                  # Small number for testing
  cpus = 16
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1582 - 0.1589)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death2 <- evaluate_mi_cv(
  formula = formula_death_updated, 
  imputed_list = py_corrected_datasets, 
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  k = length(py_corrected_datasets),
  n_repeats=2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7456 (95% CI 0.7456-0.7457)
# Pooled UnoC (survAUC::UnoC): 0.7919 (95% CI 0.7611-0.8228)
# Pooled IBS: 0.0351 (95% CI 0.0351-0.0351)
# Pooled IBS (IPCW-train): 0.0886 (95% CI 0.0885-0.0887)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7581 (95% CI 0.7578-0.7584, valid repeats=2)
#   t=36 m: C=0.7634 (95% CI 0.7627-0.7642, valid repeats=2)
#   t=60 m: C=0.7625 (95% CI 0.7619-0.7631, valid repeats=2)
# ============================================================
results_death2_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_death_updated,
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  tau = 107,                # Evaluate up to 60 months
  B = 2,                  # Small number for testing
  cpus = 16
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0357 (95% CI: 0.0356 - 0.0359)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
aicc_readmit_upd <- compute_mi_aicc(formula_readmit_updated, py_corrected_datasets)
#Mean AICc: 282681.18
aicc_readmit <- compute_mi_aicc(formula_readmit, py_corrected_datasets)
#Mean AICc: 282683.12
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_readmit_upd$mean_aic - aicc_readmit$mean_aic))
#Difference in Mean AIC: -1.94

compare_mi_models_d1(
    formula_full = formula_readmit,           # This one has 'logro_minimo'
    formula_reduced = formula_readmit_updated,# This one does not
    imputed_list = py_corrected_datasets
)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test  statistic df1 df2 dfcom   p.value           riv
#  1 ~~ 2 0.07629859   1   4 15182 0.7960619 0.00009791482
# ============================================================

aicc_death_upd <- compute_mi_aicc(formula_death_updated, py_corrected_datasets)
#Mean AICc: 53930.03
aicc_death <- compute_mi_aicc(formula_death, py_corrected_datasets)
#Mean AICc: 53932.12
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_death_upd$mean_aic - aicc_death$mean_aic))
#Difference in Mean AIC: -2.08
compare_mi_models_d1(
    formula_full = formula_death,           # This one has 'logro_minimo'
    formula_reduced = formula_death_updated,# This one does not
    imputed_list = py_corrected_datasets
)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test   statistic df1 df2 dfcom   p.value          riv
#  1 ~~ 2 0.003735574   1   4  2974 0.9541961 0.0002646455
# ============================================================

cal_readmit2<- 
calibrate_cph_corrected(
    formula = formula_readmit_updated,  # Original formula with months variable
    data = py_corrected_datasets,      # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    n_imputations = 5
)
print(cal_readmit2)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00297  0.00206   0.990     5
# 2          36  0.00434  0.00343   0.987     5
# 3          60  0.00506  0.00401   0.983     5
# 4          96  0.00503  0.00381   0.975     5
cal_death2<- 
calibrate_cph_corrected(
    formula = formula_death_updated,  # Original formula with months variable
    data = py_corrected_datasets,      # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    n_imputations = 5
)
print(cal_death2)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00176  0.00171    1.01     5
# 2          36  0.00276  0.00268    1.01     5
# 3          60  0.00321  0.00301    1.02     5
# 4          96  0.00360  0.00327    1.06     5
```

## Proportionallity

### Residuals, readmission

```{r}
#| label: nonprop2-test-coxzph-readm
# --- READMISSION MODEL ---
ph_test_readmit <- cox.zph(cox_readmit2)
ph_test_readmit$table |>
  data.table::data.table(keep.rownames = TRUE) |>
  dplyr::arrange(desc(chisq)) |>
  dplyr::mutate(
    p_fmt = ifelse(p < 0.001, "<0.001", sprintf("%.3f", p)),
    chi_fmt = sprintf(
      "χ²(%d)= %.2f, p%s",
      df,
      chisq,
      ifelse(p < 0.001, "<0.001", paste0("=", p_fmt))
    )
  ) |>
  dplyr::select(covariable = rn, `χ²(gl), p` = chi_fmt) |>
  knitr::kable(caption = "Schofeld residuals, readmission")
# 1. Identify violators (you already have this part)
violations_readm <- ph_test_readmit$table |>
  data.table::data.table(keep.rownames = T) |>
  dplyr::mutate(p = as.numeric(p)) |>
  dplyr::filter(p < 0.05) |>
  dplyr::pull(rn) |>
  setdiff("GLOBAL")
# 2. Generate the combined plot
# ggcoxzph creates a list of plots if you pass multiple variables
combined_plots <- survminer::ggcoxzph(ph_test_readmit, var = violations_readm)
# Open a PDF device
pdf(paste0(project_root,"/cons/","_figs/ph_test_violations_readm.pdf"), 
  width = 16 * 1.5, 
  height = 12 * 1.5)
# Draw the combined plot
print(combined_plots)
# Close the device
dev.off()
png(
  filename = paste0(project_root,"/cons/","_figs/ph_test_violations_readm.png"),
  width = 16 * 1.5,
  height = 12 * 1.5,
  units = "in",
  res=300
)
print(combined_plots)
dev.off()

combined_plots
```


```{r}
#| label: nonprop2-5-test-coxzph-readm-spec

#fixed windows based on data density rather than percentiles, and show raw residuals as points

par(mfrow=c(2,2))
plot_schoenfeld_residuals(
    formula = formula_readmit_updated,
    variable = "tr_outcome_adm_discharge_rule_violation_undet",
    df = py_corrected_datasets[[1]]
)
plot_schoenfeld_residuals(
    formula = formula_readmit_updated,
    variable = "adm_motive_justice_sector",
    df = py_corrected_datasets[[1]]
)
plot_schoenfeld_residuals(
    formula = formula_readmit_updated,
    variable = "dit_m",
    df = py_corrected_datasets[[1]]
)
```

The protective effect of "rule violation undetermined" diminishes over time. Initially associated with higher risk (or less protection), later becomes protective or neutral. Complex non-monotonic pattern. The effect of justice sector referral changes direction multiple times - initially neutral, then protective, then risk factor, then strongly protective again. Despite small scale, this shows the effect of dit_m is not constant - changes from positive to negative association over time.

### Residuals, death

```{r}
#| label: nonprop3-test-coxzph-death
# --- DEATH MODEL ---
ph_test_death <- cox.zph(cox_death2)
ph_test_death$table|>
  data.table::data.table(keep.rownames = TRUE)|>
  dplyr::arrange(desc(chisq))|>
  dplyr::mutate(
    p_fmt = ifelse(p < 0.001, "<0.001", sprintf("%.3f", p)),
    chi_fmt = sprintf("χ²(%d)= %.2f, p%s", df, chisq, ifelse(p < 0.001, "<0.001", paste0("=", p_fmt))))|>
  dplyr::select(covariable = rn, `χ²(gl), p` = chi_fmt)|>
  knitr::kable(caption="Schofeld residuals, death")
# 1. Identify violators (you already have this part)
violations <- ph_test_death$table|> 
  data.table::data.table(keep.rownames = T)|> 
  dplyr::mutate(p = as.numeric(p))|> 
  dplyr::filter(p < 0.05)|> 
  dplyr::pull(rn)|> 
  setdiff("GLOBAL")
# 2. Generate the combined plot
# ggcoxzph creates a list of plots if you pass multiple variables
combined_plots <- survminer::ggcoxzph(ph_test_death, var = violations)
# Open a PDF device 
pdf(paste0(project_root,"/cons/","_figs/ph_test_violations_death.pdf"), 
    width = 16*1.5, 
    height = 12*1.5) 
# Draw the combined plot 
print(combined_plots) 
# Close the device 
dev.off()
png(
  filename = paste0(project_root,"/cons/","_figs/ph_test_violations_death.png"),
  width  = 16 * 1.5,
  height = 12 * 1.5,
  units  = "in",
  res = 300
)
print(combined_plots) 
dev.off()
combined_plots
```

We decided to stratify by treatment outcome for mortality by  `tr_outcome_adm_discharge_rule_violation_undet`.


```{r}
#| label: nonprop3-5-test-coxzph-death-spec
par(mfrow=c(1,1))
plot_schoenfeld_residuals(
    formula = formula_death_updated,
    variable = "tr_outcome_adm_discharge_rule_violation_undet",
    df = py_corrected_datasets[[1]]
)
```


We suspect an administrative artifact behind the concentration and violation of proportionallity of hazards


```{r}
#| label: nonprop3-5-test-coxzph-death-artifact

# Create time intervals
time_breaks <- c(0, 5, 10, 20, 60, 120)
time_labels <- c("(0,5]", "(5,10]", "(10,20]", "(20,60]", "(60,120]")

# Cut the time variable
time_cut <- cut(py_corrected_datasets[[1]]$death_time_from_disch_m, 
                breaks = time_breaks, 
                labels = time_labels)

# Get variable names dynamically
var1 <- "tr_outcome_adm_discharge_rule_violation_undet"
var2 <- "tr_outcome_adm_discharge_adm_reasons"

# Create tables
total_table <- table(time_cut)
var1_table <- table(dplyr::pull(py_corrected_datasets[[1]], var1), time_cut)
var2_table <- table(dplyr::pull(py_corrected_datasets[[2]], var2), time_cut)

# Build data frame adaptively
result_df <- data.frame(
  Time_interval = time_labels,
  Total = as.numeric(total_table)
)

# Add var1 levels (excluding first level as reference)
for(i in 2:nrow(var1_table)) {
  col_name <- paste0(var1, "_", rownames(var1_table)[i])
  result_df[[col_name]] <- as.numeric(var1_table[i, ])
}

# Add var2 levels (excluding first level as reference)
for(i in 2:nrow(var2_table)) {
  col_name <- paste0(var2, "_", rownames(var2_table)[i])
  result_df[[col_name]] <- as.numeric(var2_table[i, ])
}

# Generate markdown table
knitr::kable(result_df, 
      format = "markdown",
      col.names = c("Intervals", "total", "Rule violation", "Adm reasons"),
      caption = "Distribution of death times by discharge outcome categories",
      align = c("l", rep("r", ncol(result_df) - 1)))

# Table: Distribution of death times by discharge outcome categories
# 
# |Intervals | total| Rule violation| Adm reasons|
# |:---------|-----:|--------------:|-----------:|
# |(0,5]     |  6159|            256|         155|
# |(5,10]    |  1464|             45|          30|
# |(10,20]   |  5695|            174|         159|
# |(20,60]   | 25526|           1158|         579|
# |(60,120]  | 29412|           3269|           7|
```


```{r}
#| label: nonprop5-log-log

message("We had to handle ties")
df_surv <- py_corrected_datasets[[1]] %>%
  dplyr::mutate(
    tr_outcome_strata = case_when(
      tr_outcome_referral == 1 ~ "referral",
      tr_outcome_dropout == 1 ~ "dropout",
      tr_outcome_adm_discharge_rule_violation_undet == 1 ~ "rule violation/undet",
      tr_outcome_adm_discharge_adm_reasons == 1 ~ "adm reasons",
      #tr_outcome_other == 1 ~ "other",
      T~ "completion"
    ),
    tr_outcome_strata = factor(
      tr_outcome_strata,
      levels = c(
        "completion",
        "referral",
        "dropout",
        "rule violation/undet",
        "adm reasons"#,
        #"other"
      )
    )
  )
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
fit_stratified_readmit <- survfit(
  Surv(readmit_time_from_disch_m, readmit_event) ~ tr_outcome_strata,
  data = df_surv |> dplyr::filter(tr_outcome_strata!="other"))
fit_stratified_death <- survfit(
  Surv(death_time_from_disch_m, death_event) ~ tr_outcome_strata,
  data = df_surv |> dplyr::filter(tr_outcome_strata!="other"))
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# 1. Define tus colores en un vector claro
mis_colores <- c("#000000", "#545454", "#808080", "#A9A9A9", "#D3D3D3")
p <-survminer::ggsurvplot(
  fit_stratified_readmit, 
  fun = "cloglog", 
  conf.int = FALSE,              # Saca esto si se ve muy "sucio" o amontonado
  palette = mis_colores,         # AQUÍ se definen los colores
  title = "LML Plot: Proportional Hazards Check",
  xlab = "log(Time)",
  ylab = "log(-log(Survival))",
  legend.title = "Treatment Outcome",
  legend.labs = c("Completion", "Referral", "Dropout", "Violation", "Adm Reasons"),
  ggtheme = theme_bw()           # Fondo blanco limpio para tesis
)
p$plot <- p$plot + 
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 2)) # Organiza la leyenda en 2 filas si es muy larga
p$plot
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# 1. Define tus colores en un vector claro
mis_colores <- c("#000000", "#545454", "#808080", "#A9A9A9", "#D3D3D3")
p_death_eng <-survminer::ggsurvplot(
  fit_stratified_death, 
  fun = "cloglog", 
  conf.int = FALSE,              # Saca esto si se ve muy "sucio" o amontonado
  palette = mis_colores,         # AQUÍ se definen los colores
  title = "LML Plot: Proportional Hazards Check",
  xlab = "log(Time)",
  ylab = "log(-log(Survival))",
  legend.title = "Treatment Outcome",
  legend.labs = c("Completion", "Referral", "Dropout", "Violation", "Adm Reasons"),
  ggtheme = theme_bw()           # Fondo blanco limpio para tesis
)
p_death_eng$plot <- p_death_eng$plot + 
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 2)) # Organiza la leyenda en 2 filas si es muy larga
p_death_eng$plot
```

```{r}
#| label: nonprop5-log-log-esp
# Readmission
p_esp_readm <- survminer::ggsurvplot(
  fit_stratified_readmit,
  fun = "cloglog",
  conf.int = TRUE,
  censor = FALSE,
  xlab = "log(Tiempo desde el alta)",
  ylab = "log(-log(Supervivencia))",
  legend.labs = c("Razones\nadministrativas",
                  "Violación\nde reglas",
                  "Finalización",
                  "Abandono",
                  "Derivación"),
  legend.title = "Resultado\ndel tratamiento",
  palette = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
  ggtheme = theme_minimal(base_family = "Arial")
)
p_esp_readm$plot <- p_esp_readm$plot +
  geom_point(aes(shape = strata, color = strata), size = 2) +
  scale_color_manual(
    values = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3),
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  theme(
    text = element_text(family = "Arial"),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.title = element_blank()
  )
# Death
p_esp_death <- survminer::ggsurvplot(
  fit_stratified_death,
  fun = "cloglog",
  conf.int = TRUE,
  censor = FALSE,
  xlab = "log(Tiempo desde el alta)",
  ylab = "log(-log(Supervivencia))",
  legend.labs = c("Razones\nadministrativas",
                  "Violación\nde reglas",
                  "Finalización",
                  "Abandono",
                  "Derivación"),
  legend.title = "Resultado\ndel tratamiento",
  palette = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
  ggtheme = theme_minimal(base_family = "Arial")
)
p_esp_death$plot <- p_esp_death$plot +
  geom_point(aes(shape = strata, color = strata), size = 2) +
  scale_color_manual(
    values = c("#000000", "#505050", "#808080", "#B0B0B0", "#D9D9D9"),
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  scale_shape_manual(
    values = c(16, 17, 15, 18, 3),
    name = "Resultado\ndel tratamiento",
    labels = c("Razones\nadministrativas", "Violación\nde reglas", "Finalización", "Abandono", "Derivación")
  ) +
  theme(
    text = element_text(family = "Arial"),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.title = element_blank()
  ) +
  labs(y = NULL) # remove Y label from readmission panel
# Combine with A) and B) in bold
plot_cloglog_comb <- (p_esp_readm$plot + p_esp_death$plot) +
  patchwork::plot_layout(ncol = 2, guides = "collect") +
  patchwork::plot_annotation(
    tag_levels = "A",
    tag_suffix = ")",
    theme = theme(plot.tag = element_text(face = "bold", family = "Arial"))
  ) &
  theme(legend.position = "bottom")
ggsave(
  filename = paste0(project_root, "/cons/", "_figs/supervivencia_cloglog_comb.png"),
  plot = plot_cloglog_comb,
  dpi = 300,
  width = 12*.75,
  height = 6*.75,
  units = "in"
)
```

We see that treatment outcome has different baseline shapes.


### Correction of strata

```{r}
#| label: update-model-to-nph
# 1. Clean the Readmission Formula (Subtract the 1 violator)
formula_readmit_updated2 <- update(
  formula_readmit_updated,
  . ~ . - tr_outcome_adm_discharge_rule_violation_undet
)
# 2. Clean the Death Formula (Subtract both violators)
formula_death_updated2 <- update(
  formula_death_updated,
  . ~ . - tr_outcome_adm_discharge_rule_violation_undet # - tr_outcome_adm_discharge_adm_reasons
)
# 1. READMISSION FORMULA
# Safely collapse the formula into a single text string first
formula_readmit_updated2 <- as.formula(
  gsub(
    "strata\\(plan_type_strata\\)",
    "strata(plan_type_strata)+ strata(tr_outcome_adm_discharge_rule_violation_undet)",
    paste(deparse(formula_readmit_updated2), collapse = " ")
  )
)
# 2. DEATH FORMULA
# Safely collapse the formula into a single text string first
formula_death_updated2 <- as.formula(
  gsub(
    "strata\\(plan_type_strata\\)",
    "strata(plan_type_strata) +strata(tr_outcome_adm_discharge_rule_violation_undet)",
    paste(deparse(formula_death_updated2), collapse = " ")
  )
)

results_readmit3 <- evaluate_mi_cv(
  formula = formula_readmit_updated2,
  imputed_list = py_corrected_datasets,
  time_col = "readmit_time_from_disch_m",
  event_col = "readmit_event",
  k = length(py_corrected_datasets),
  n_repeats = 2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6066 (95% CI 0.6064-0.6068)
# Pooled UnoC (survAUC::UnoC): 0.6674 (95% CI 0.6473-0.6876)
# Pooled IBS: 0.1598 (95% CI 0.1597-0.1598)
# Pooled IBS (IPCW-train): 0.3474 (95% CI 0.3473-0.3475)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6317 (95% CI 0.6313-0.6321, valid repeats=2)
#   t=36 m: C=0.6152 (95% CI 0.6150-0.6153, valid repeats=2)
#   t=60 m: C=0.6098 (95% CI 0.6097-0.6099, valid repeats=2)
# ============================================================
results_death3 <- evaluate_mi_cv(
  formula = formula_death_updated2,
  imputed_list = py_corrected_datasets,
  time_col = "death_time_from_disch_m",
  event_col = "death_event",
  k = length(py_corrected_datasets),
  n_repeats = 2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7449 (95% CI 0.7449-0.7449)
# Pooled UnoC (survAUC::UnoC): 0.7885 (95% CI 0.7529-0.8241)
# Pooled IBS: 0.0350 (95% CI 0.0350-0.0350)
# Pooled IBS (IPCW-train): 0.0883 (95% CI 0.0881-0.0884)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7643 (95% CI 0.7634-0.7653, valid repeats=2)
#   t=36 m: C=0.7638 (95% CI 0.7629-0.7648, valid repeats=2)
#   t=60 m: C=0.7618 (95% CI 0.7612-0.7624, valid repeats=2)
# ============================================================
results_readmit3_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_readmit_updated2,
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "readmit_time_from_disch_m",
  event_col = "readmit_event",
  tau = 99, # Evaluate up to 60 months
  B = 2, # Small number for testing
  cpus = cpus # Number of CPU cores to use
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1582 - 0.1589)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death3_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_death_updated2,
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "death_time_from_disch_m",
  event_col = "death_event",
  tau = 107, # Evaluate up to 60 months
  B = 2, # Small number for testing
  cpus = cpus
)
# ===========================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0356 (95% CI: 0.0355 - 0.0358)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
aicc_readmit_upd2 <- compute_mi_aicc(
  formula_readmit_updated2,
  py_corrected_datasets
)
#Mean AICc: 272549.16
cat(sprintf(
  "\nDifference in Mean AIC: %.2f\n",
  aicc_readmit_upd2$mean_aic - aicc_readmit_upd$mean_aic
))
aicc_death_upd2 <- compute_mi_aicc(
  formula_death_updated2,
  py_corrected_datasets
)
#5Mean AICc: 51496.24
cat(sprintf(
  "\nDifference in Mean AIC: %.2f\n",
  aicc_death_upd2$mean_aic - aicc_death_upd$mean_aic
))
#Difference in Mean AIC: -2433.79

#compare_mi_models_d1(formula_readmit_updated2, formula_readmit_updated, py_corrected_datasets)
# Error en mitml::testModels(fit1, fit0, method = "D1", df.com = dfcom):
#   The 'model' and 'null.model' appear not to be nested or include the same set of parameters.
# Called from: mitml::testModels(fit1, fit0, method = "D1", df.com = dfcom)

compare_mi_models_d1(
  formula_readmit_updated,
  update(
    formula_readmit_updated,
    . ~ . - tr_outcome_adm_discharge_rule_violation_undet
  ),
  py_corrected_datasets
)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test statistic df1 df2 dfcom     p.value          riv
#  1 ~~ 2  51.07488   1   4 15183 0.002028026 0.0001081492
# ============================================================
compare_mi_models_d1(
  formula_death_updated,
  update(
    formula_death_updated,
    . ~ . -
      tr_outcome_adm_discharge_rule_violation_undet -
      tr_outcome_adm_discharge_adm_reasons
  ),
  py_corrected_datasets
)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test statistic df1      df2 dfcom      p.value          riv
#  1 ~~ 2   25.7481   2 2972.522  2975 8.194034e-12 0.0003485494
# ============================================================

cal_readmit3 <-
  calibrate_cph_corrected(
    formula = formula_readmit_updated2, # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    n_imputations = 5
  )
print(cal_readmit3)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00327  0.00225   0.990     5
# 2          36  0.00433  0.00359   0.987     5
# 3          60  0.00508  0.00386   0.984     5
# 4          96  0.00495  0.00331   0.976     5
cal_death3 <-
  calibrate_cph_corrected(
    formula = formula_death_updated2, # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    n_imputations = 5
  )
print(cal_death3)
#plot(cal_death2$combined_plot)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00143  0.00135    1.01     5
# 2          36  0.00271  0.00260    1.01     5
# 3          60  0.00323  0.00297    1.02     5
# 4          96  0.00492  0.00460    1.07     5


```

# SHAP criteria

```{r}
#| label: shap-criteria-setup

predictor_tiers <- readxl::read_excel(
  file.path(project_root, "cons", "_out",
            "xgb9_dual_predictor_analysis_no_clinical_20260306_1821.xlsx"),
  sheet = "Predictors"
) |>
  dplyr::filter(!grepl("MINOR",importance_tier))


predictor_tiers |> knitr::kable(
  caption = "Predictor tiers based on SHAP importance from XGBoost models"
)
```

```{r}
#| label: shap-criteria

gc()

# # Linear time interaction
# coxph(Surv(time, death) ~ var + var:time + other_covariates, data = df)
# # Log-time interaction
# coxph(Surv(time, death) ~ var + var:log(time) + other_covariates, data = df)
# # Step function at 1 year (365 days)
# df$late_period <- as.numeric(df$time > 365)
# coxph(Surv(time, death) ~ var + var:late_period + other_covariates, data = df)
# # Using splines (requires splines package)
# library(splines)
# coxph(Surv(time, death) ~ var + var:ns(time, df=3) + other_covariates, data = df)
vars_shap_r <- c(
  "primary_sub_mod_cocaine_paste",
  "adm_age_rec3",
  "porc_pobr",
  "sex_rec_woman",
  #"plan_type_corr_pg_pr", # Mantenido en el código como strata()
  #"plan_type_corr_m_pr",  # Mantenido en el código como strata()
  "ethnicity",
  "dit_m",
  #"eva_consumo",
  "eva_consumo_logro_intermedio",
  "eva_consumo_logro_minimo",
  #"ed_attainment_corr",
  "ed_attainment_corr_2_completed_high_school_or_less",
  "ed_attainment_corr_3_completed_primary_school_or_less",
  "sub_dep_icd10_status_drug_dependence",
  "occupation_condition_corr24_unemployed",
  "polysubstance_strict",
  "primary_sub_mod_cocaine_powder",
  #"evaluacindelprocesoteraputico",
  "evaluacindelprocesoteraputico_logro_intermedio",
  "evaluacindelprocesoteraputico_logro_minimo", 
  "dg_psiq_cie_10_dg",
  "tr_outcome_adm_discharge_rule_violation_undet",
  #"eva_sm",
  "eva_sm_logro_intermedio",
  "eva_sm_logro_minimo",   
  
  # --- NO LONGER IN TOP 20 MORTALITY (SHAP) ---
  #"adm_motive_justice_sector",
  #"tenure_status_household_others",
  #"tenure_status_household_renting",
  #"tenure_status_household_stays_temporarily_with_a_relative",
  #"tenure_status_household_illegal_settlement",
  #"tr_outcome_referral",
  #"eva_ocupacion_logro_intermedio",
  #"eva_ocupacion_logro_minimo",   
  #"eva_transgnorma_logro_intermedio",
  #"eva_transgnorma_logro_minimo",    
  
  #"prim_sub_freq_rec",
  "prim_sub_freq_rec_2_2_6_days_wk",
  "prim_sub_freq_rec_3_daily",  
  
  # --- NO LONGER IN TOP 20 MORTALITY  (SHAP) ---
  #"cohabitation_with_couple_children",
  
  "primary_sub_mod_alcohol"
  
  # --- NO LONGER IN TOP 20 MORTALITY (SHAP) ---
  #"eva_relinterp_logro_intermedio",
  #"eva_relinterp_logro_minimo",  
  #"tr_outcome_dropout",
  #"dx_f6_personality",
  #"any_phys_dx",
  #"any_violence_1_domestic_violence_sex_abuse",
  #"adm_motive_sanitary_sector",
  #"eva_fam_logro_intermedio",
  #"eva_fam_logro_minimo"
)

vars_shap_d <- c(
  "adm_age_rec3",
  "primary_sub_mod_cocaine_paste",
  #"prim_sub_freq_rec", 
  "prim_sub_freq_rec_2_2_6_days_wk",
  "prim_sub_freq_rec_3_daily",
  
  # --- NO LONGER IN TOP 20 MORTALITY (SHAP 18:22) ---
  #"primary_sub_mod_cocaine_powder",
  
  "occupation_condition_corr24_unemployed",
  "any_phys_dx",
  #"eva_ocupacion",
  "eva_ocupacion_logro_intermedio",
  "eva_ocupacion_logro_minimo",  
  "first_sub_used_alcohol",
  "sex_rec_woman",
  #"eva_fisica"
  "eva_fisica_logro_intermedio",
  "eva_fisica_logro_minimo"
)
formula_shap_readmit <- as.formula(
  paste("Surv(readmit_time_from_disch_m, readmit_event) ~", 
        paste(vars_shap_r, collapse = " + "), "+ strata(plan_type_strata)")
)
formula_shap_death <- as.formula(
  paste("Surv(death_time_from_disch_m, death_event) ~", 
        paste(vars_shap_d, collapse = " + "), "+ strata(plan_type_strata)")
)
dplyr::bind_rows(
  rms::vif(survival::coxph(formula_shap_death, data = py_corrected_datasets[[1]])) |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    dplyr::rename(Gvif = 2) |>
    dplyr::arrange(dplyr::desc(dplyr::pick(2)[[1]])) |>
    dplyr::filter(Gvif > 10) |>
    dplyr::mutate(Outcome = "Death"),
  rms::vif(survival::coxph(formula_shap_readmit, data = py_corrected_datasets[[1]])) |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    dplyr::rename(Gvif = 2) |>
    dplyr::filter(Gvif > 10) |>
    dplyr::mutate(Outcome = "Readmission")
) |>
  knitr::kable("markdown", caption = "GVIF for SHAP-based Cox models, using one imputed dataset")

cat("Removing collineal variable\n")
formula_shap_readmit_clean <- as.formula(
  paste(gsub("evaluacindelprocesoteraputico_logro_minimo \\+\\s*","",paste(deparse(formula_shap_readmit), collapse = " "))))

results_readmit_shap <- evaluate_mi_cv(
  formula = formula_shap_readmit_clean, 
  imputed_list = py_corrected_datasets, 
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  k = length(py_corrected_datasets),
  n_repeats=2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6057 (95% CI 0.6056-0.6058)
# Pooled UnoC (survAUC::UnoC): 0.6620 (95% CI 0.6379-0.6862)
# Pooled IBS: 0.1600 (95% CI 0.1600-0.1600)
# Pooled IBS (IPCW-train): 0.3480 (95% CI 0.3479-0.3481)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6312 (95% CI 0.6310-0.6315, valid repeats=2)
#   t=36 m: C=0.6139 (95% CI 0.6138-0.6139, valid repeats=2)
#   t=60 m: C=0.6087 (95% CI 0.6086-0.6088, valid repeats=2)
# ============================================================
#Total runtime: 148.87 sec (2.48 min)

results_death_shap <- evaluate_mi_cv(
  formula = formula_shap_death, 
  imputed_list = py_corrected_datasets, 
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  k = length(py_corrected_datasets),
  n_repeats=2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7367 (95% CI 0.7366-0.7369)
# Pooled UnoC (survAUC::UnoC): 0.8059 (95% CI 0.8026-0.8093)
# Pooled IBS: 0.0356 (95% CI 0.0356-0.0356)
# Pooled IBS (IPCW-train): 0.0903 (95% CI 0.0902-0.0904)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7377 (95% CI 0.7375-0.7379, valid repeats=2)
#   t=36 m: C=0.7495 (95% CI 0.7492-0.7499, valid repeats=2)
#   t=60 m: C=0.7500 (95% CI 0.7497-0.7502, valid repeats=2)
# ============================================================

results_readmit_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_shap_readmit_clean,
    imputed_list = py_corrected_datasets, # List of data.frames
    time_col = "readmit_time_from_disch_m", 
    event_col = "readmit_event",
    tau = 99,                # Evaluate up to 60 months
    B = 2,                  # Small number for testing
    cpus = cpus
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1584 - 0.1586)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_shap_death,
    imputed_list = py_corrected_datasets, # List of data.frames
    time_col = "death_time_from_disch_m", 
    event_col = "death_event",
    tau = 107,                # Evaluate up to 60 months
    B = 2,                  # Small number for testing
    cpus = cpus
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0363 (95% CI: 0.0362 - 0.0363)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================

aicc_readmit_shap <- compute_mi_aicc(formula_shap_readmit_clean, py_corrected_datasets)
#282719.30
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_readmit_shap$mean_aic - aicc_readmit_upd2$mean_aic))
#Difference in Mean AIC: 10170.15
aicc_death_shap <- compute_mi_aicc(formula_shap_death, py_corrected_datasets)
#54321.99
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_death_shap$mean_aic - aicc_death_upd2$mean_aic))
#Difference in Mean AIC: 2825.75
cat("Comparing readmission w/o collinear var and only one strata(plan)\n")
compare_mi_models_d1(formula_readmit_updated,formula_shap_readmit_clean, py_corrected_datasets)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test statistic df1      df2 dfcom      p.value        riv
#  1 ~~ 2  2.705153  42 11680.04 15183 1.835904e-08 0.05960578
# ============================================================

compare_mi_models_d1(formula_death_updated,formula_shap_death, py_corrected_datasets)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test statistic df1      df2 dfcom     p.value        riv
#  1 ~~ 2  9.059786  52 2901.688  2975 6.69787e-63 0.04231255
# ============================================================

cal_readmit_shap<- 
    calibrate_cph_corrected(
        formula = formula_shap_readmit_clean,  # Original formula with months variable
        data = py_corrected_datasets,      # Original data (months)
        times = c(12, 36, 60, 96),
        seed = 2125,
        n_imputations = 5
    )
print(cal_readmit_shap)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00367  0.00245   0.991     5
# 2          36  0.00635  0.00384   0.988     5
# 3          60  0.00691  0.00478   0.984     5
# 4          96  0.00708  0.00434   0.976     5

cal_death_shap<- 
    calibrate_cph_corrected(
        formula = formula_shap_death,  # Original formula with months variable
        data = py_corrected_datasets,      # Original data (months)
        times = c(12, 36, 60, 96),
        seed = 2125,
        n_imputations = 5  
    )
print(cal_death_shap)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00151  0.00148    1.01     5
# 2          36  0.00230  0.00203    1.02     5
# 3          60  0.00327  0.00305    1.03     5
# 4          96  0.00336  0.00312    1.06     5
```


```{r}
#| label: shap-criteria-nph

cox.zph(coxph(formula_shap_death, data= py_corrected_datasets[[1]]))
cox.zph(coxph(formula_shap_readmit_clean, data= py_corrected_datasets[[1]]))
#tr_outcome_adm_discharge_rule_violation_undet          58.6456  1 1.9e-14

```


```{r}
#| label: shap-criteria-additional-strata

# Residuals: discharge to readmission
res <- resid(coxph(formula_shap_readmit_clean, data= py_corrected_datasets[[1]]), "scaledsch")
time <- as.numeric(dimnames(res)[[1]])
z2 <- loess(res[,"tr_outcome_adm_discharge_rule_violation_undet"] ~ time, span=0.50)

plot(time, fitted(z2), type="l", col="blue",
     main="Sch. res. – Completion (Violation of tules of the center)",
     xlab="Time", ylab="Residuals")
lines(supsmu(time, res[,"tr_outcome_adm_discharge_rule_violation_undet"]), lty=2, col="darkgray")


formula_shap_readmit_clean_updated <- update(
  formula_shap_readmit_clean,
  . ~ . - tr_outcome_adm_discharge_rule_violation_undet + strat(tr_outcome_adm_discharge_rule_violation_undet)
)
```

```{r}
#| label: shap-criteria-additional-strata2

results_readmit_shap2 <- evaluate_mi_cv(
  formula = formula_shap_readmit_clean_updated, 
  imputed_list = py_corrected_datasets, 
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  k = length(py_corrected_datasets),
  n_repeats=2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6057 (95% CI 0.6056-0.6058)
# Pooled UnoC (survAUC::UnoC): 0.6620 (95% CI 0.6379-0.6862)
# Pooled IBS: 0.1600 (95% CI 0.1600-0.1600)
# Pooled IBS (IPCW-train): 0.3480 (95% CI 0.3479-0.3481)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6312 (95% CI 0.6310-0.6315, valid repeats=2)
#   t=36 m: C=0.6139 (95% CI 0.6138-0.6139, valid repeats=2)
#   t=60 m: C=0.6087 (95% CI 0.6086-0.6088, valid repeats=2)
# ============================================================
results_death_shap <- evaluate_mi_cv(
  formula = formula_shap_death, 
  imputed_list = py_corrected_datasets, 
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  k = length(py_corrected_datasets),
  n_repeats=2
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7367 (95% CI 0.7366-0.7369)
# Pooled UnoC (survAUC::UnoC): 0.8059 (95% CI 0.8026-0.8093)
# Pooled IBS: 0.0356 (95% CI 0.0356-0.0356)
# Pooled IBS (IPCW-train): 0.0903 (95% CI 0.0902-0.0904)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7377 (95% CI 0.7375-0.7379, valid repeats=2)
#   t=36 m: C=0.7495 (95% CI 0.7492-0.7499, valid repeats=2)
#   t=60 m: C=0.7500 (95% CI 0.7497-0.7502, valid repeats=2)
# ============================================================
results_readmit_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_shap_readmit_clean,
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  tau = 99,                # Evaluate up to 60 months
  B = 2,                  # Small number for testing
  cpus = cpus
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1584 - 0.1586)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = formula_shap_death,
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  tau = 107,                # Evaluate up to 60 months
  B = 2,                  # Small number for testing
  cpus = cpus
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0363 (95% CI: 0.0362 - 0.0363)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
aicc_readmit_shap <- compute_mi_aicc(formula_shap_readmit_clean, py_corrected_datasets)
#Mean AICc: 282719.30
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_readmit_shap$mean_aic - aicc_readmit_upd2$mean_aic))
#Difference in Mean AIC: 10170.15
aicc_death_shap <- compute_mi_aicc(formula_shap_death, py_corrected_datasets)
#Mean AICc: 54321.99
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_death_shap$mean_aic - aicc_death_upd2$mean_aic))
#Difference in Mean AIC: 2825.75

cal_readmit_shap2<- 
    calibrate_cph_corrected(
        formula = formula_shap_readmit2,  # Original formula with months variable
        data = py_corrected_datasets,      # Original data (months)
        times = c(12, 36, 60, 96),
        seed = 2125,
        n_imputations = 5
    )
print(cal_readmit_shap2)
#  time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00369  0.00281   0.991     5
# 2          36  0.00646  0.00449   0.988     5
# 3          60  0.00736  0.00500   0.985     5
# 4          96  0.00718  0.00405   0.977     5

cal_death_shap2<- 
    calibrate_cph_corrected(
        formula = formula_shap_death,  # Original formula with months variable
        data = py_corrected_datasets,      # Original data (months)
        times = c(12, 36, 60, 96),
        seed = 2125,
        n_imputations = 5  
    )
print(cal_death_shap2)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00151  0.00148    1.01     5
# 2          36  0.00230  0.00203    1.02     5
# 3          60  0.00327  0.00305    1.03     5
# 4          96  0.00336  0.00312    1.06     5
```

With stratification, individuals in different strata have different baseline hazards, and therefore different predicted survival probabilities. This creates some discrimination purely from the stratum membership.

For the Wald test (𝐷1) to be valid, the ‘Reduced’ model must be a perfect subset of the ‘Full’ model. You cannot change the underlying structure (strata).

## Interactions

```{r}
#| label: shap-criteria-int

gglobal_df <- read_csv(
  find_latest_file(
    file.path(project_root, "cons", "_out"),
    "xgb11_dual_interactions_global_top_enriched"),
  show_col_types = FALSE
)
time_dep_df <- read_csv(
  find_latest_file(
    file.path(project_root, "cons", "_out"), 
    "xgb11_dual_interactions_time_dependent"),
  show_col_types = FALSE
)

# 2. Merge on unique keys
merged_df <- global_df %>%
  left_join(time_dep_df, by = c("outcome", "main_feature", "interactor"))

# 3. Apply the Effect-Size & Stability Rule
cox_discriminative_interactions <- merged_df %>%
  filter(
    # Ignore FDR completely. Use joint clinical/linear magnitude:
    abs_corr >= 0.10,
    !is.na(smd_binarized) & abs(smd_binarized) >= 0.15,
    
    # Enforce standard Cox Proportional Hazards assumptions:
    time_dependent_flag == FALSE,
    direction_flip_flag == FALSE
  ) %>%
  select(
    outcome, 
    main_feature, 
    interactor, 
    abs_corr, 
    smd_binarized, 
    direction_q4_q1
  ) %>%
  arrange(outcome, desc(abs_corr))

# Print results
cat("Number of highly discriminative candidates for Cox model:", nrow(cox_discriminative_interactions), "\n")
knitr::kable(cox_discriminative_interactions, "markdown", caption = "Highly discriminative interactions for Cox models based on Effect-Size & Stability Rule")


```

```{r}
#| label: shap-criteria-int2

# ====================================================================
# 1. MODELO DE REINGRESO (READMISSION)
# ====================================================================

# Variables continuas y ordinales retenidas empíricamente
vars_cont_ord_readmit <- c(
  "adm_age_rec3",          # Continua: Edad al ingreso
  "porc_pobr",             # Continua: Índice de pobreza
  "dit_m",                 # Continua: Tiempo en tratamiento
  
  # Ordinales (codificadas como indicadoras/dummies):
  "ed_attainment_corr_2_completed_high_school_or_less",
  "ed_attainment_corr_3_completed_primary_school_or_less",
  "eva_consumo_logro_intermedio",
  "eva_consumo_logro_minimo",
  "evaluacindelprocesoteraputico_logro_intermedio",
  "evaluacindelprocesoteraputico_logro_minimo",
  "eva_sm_logro_intermedio",
  "eva_sm_logro_minimo",
  "prim_sub_freq_rec_2_2_6_days_wk",
  "prim_sub_freq_rec_3_daily"
)

# Interacciones seleccionadas clínica y estadísticamente
interactions_readmit <- c(
  "primary_sub_mod_cocaine_paste:polysubstance_strict",
  "eva_consumo_logro_intermedio:tr_outcome_referral",
  "eva_consumo_logro_minimo:tr_outcome_referral",
  "eva_consumo_logro_intermedio:plan_type_corr_pg_pr",
  "eva_consumo_logro_minimo:plan_type_corr_pg_pr"
)


# ====================================================================
# 2. MODELO DE MORTALIDAD (DEATH)
# ====================================================================

# Variables continuas y ordinales retenidas empíricamente
vars_cont_ord_death <- c(
  "adm_age_rec3",          # Continua: Edad al ingreso
  
  # Ordinales (codificadas como indicadoras/dummies):
  "prim_sub_freq_rec_2_2_6_days_wk",
  "prim_sub_freq_rec_3_daily",
  "eva_ocupacion_logro_intermedio",
  "eva_ocupacion_logro_minimo",
  "eva_fisica_logro_intermedio",
  "eva_fisica_logro_minimo"
)

# Interacciones seleccionadas clínica y estadísticamente
interactions_death <- c(
  "primary_sub_mod_cocaine_paste:polysubstance_strict",
  "cohabitation_with_couple_children:adm_age_rec3",
  "occupation_condition_corr24_unemployed:adm_age_rec3",
  "sex_rec_woman:occupation_condition_corr24_inactive",
  "primary_sub_mod_cocaine_paste:first_sub_used_alcohol"
)


```
## Functional forms

```{r}
#| label: shap-criteria-func-form


```


# Session info
```{r}
#| label: session-info
#| echo: true
#| error: true
#| message: true
#| paged.print: true
message(paste0("R library: ", Sys.getenv("R_LIBS_USER")))
message(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))
message(paste0("Editor context: ", path))
cat("quarto version: "); quarto::quarto_version()
sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
)|> 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('R packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}")))
```
```{r}
#| label: session-info-python
#| echo: true
#| error: true
#| message: true
#| paged.print: true
#| class-output: center-table
reticulate::py_list_packages()|> 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Python packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}"))) 
```
Save
```{r}
#| label: save
#| echo: true
#| error: true
#| warning: true
#| message: true
#| paged.print: true
#| results: "hold"
#| eval: true
wdpath<-
paste0(gsub("/cons","",gsub("cons","",paste0(getwd(),"/cons"))))
envpath<- if(regmatches(wdpath, regexpr("[A-Za-z]+", wdpath))=="G"){"G:/Mi unidad/Alvacast/SISTRAT 2023/"}else{"E:/Mi unidad/Alvacast/SISTRAT 2023/"}
paste0(getwd(),"/cons")
file.path(paste0(wdpath,"data/20241015_out"))
file.path(paste0(envpath,"data/20241015_out"))
# Save
rdata_path <- file.path(wdpath, "data/20241015_out", paste0("pred2_ndp_", format(Sys.time(), "%Y_%m_%d"), ".Rdata"))
save.image(rdata_path)
cat("Saved in:",
    rdata_path)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
  password <- Sys.getenv("PASSWORD_SECRET")
} else {
  if (interactive()) {
    utils::savehistory(tempfile())
    Sys.setenv(PASSWORD_SECRET = readLines(paste0(wdpath, "secret.txt"), warn = FALSE))
    utils::loadhistory()
  }
  Sys.setenv(PASSWORD_SECRET = readLines(paste0(wdpath, "secret.txt"), warn = FALSE))
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
save.image(paste0(rdata_path,".enc"))
# Encriptar el archivo en el mismo lugar
httr2::secret_encrypt_file(path = paste0(rdata_path,".enc"), key = "PASSWORD_SECRET")
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
cat("Copy renv lock into cons folder\n")
if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
  message("Running on RStudio Server or inside Docker. Folder copy skipped.")
} else {
    
  source_folder <- 
  destination_folder <- paste0(wdpath,"cons/renv")
  
  # Copy the folder recursively
    file.copy(paste0(wdpath,"renv.lock"), paste0(wdpath,"cons/renv.lock"), overwrite = TRUE)
  
  message("Renv lock copy performed.")
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
time_after_dedup2<-Sys.time()



```