---
title: "SENDAs Agreement 1 Update 2010-2022 (Prediction, step 2.2, shap-informed models)"
description: |
  This module develops and validates optimized Cox Proportional Hazards models for 
  post-discharge outcomes (readmission and all-cause mortality) in SUD patients (2010–2020). 
  The selection of variables is grounded in SHAP importance tiers from XGBoost models, 
  bridging machine learning discovery with survival analysis. Key refinements include 
  the management of Non-Proportional Hazards via stratification, the identification of 
  highly discriminative interaction terms through an Effect-Size & Stability Rule, and 
  the modeling of non-linear effects using Restricted Cubic Splines.
date: last-modified
author: "Andrés González Santa Cruz"
bibliography: [_lit/refs.bib]
csl: "_lit/american-medical-association.csl" # (Opcional: estilo de citas, p.ej., APA)
toc: true
toc-location: left
toc-float: true
format: 
  html:
    css: [_lib/styles.scss]
    code-fold: true
    embed-resources: true
    theme: 
      - journal #https://quarto.org/docs/output-formats/html-themes-more.html
      #- style.scss
    fig-cap-location: top  
    toc: true
    toc-location: left
execute:
  error: true        # Show errors in the output
  echo: true         # Show R code in the output
  warning: true      # Show warnings
  message: true      # Show messages from the code
  freeze: auto       # Use cache to speed up rendering if there are no changes
editor: source
execute-dir: project
lightbox: auto
---
```{=html}
<style type="text/css">
.showopt {
background-color: #004c93; color: #FFFFFF; width: 100px; height: 20px; text-align: center; vertical-align: middle !important; float: right; font-family: sans-serif; border-radius: 8px;
}
.showopt:hover {
background-color: #dfe4f2;
color: #004c93;
}
pre.plot {
background-color: white !important;
}
.tablelines table, .tablelines td, .tablelines th {
border: 1px solid black;
}
.centrado {
text-align: center;
}
.table.center {
margin-left:auto;
margin-right:auto;
}
/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/
/* Applies to outputs that are not code other than R*/
pre {
overflow-x: auto !important;
}
pre code {
word-wrap: normal !important;
white-space: pre !important;
}
/*
pre:not(.sourceCode) {
white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence */
font-size: 10px !important;
line-height: 50% !important;
}
body{ /* Normal */
text-align: justify;
}
.superbigimage{
overflow-y:scroll;
height:350px;
white-space: nowrap;
overflow-x: auto;
width:100%;
}
.superbigimage img{
overflow-y: scroll;
overflow-x: hidden;
}
.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}
.pandoc-table { /* Should add !important; but it seems no necessary */
margin-left:auto; /* To center */
margin-right:auto;
border-collapse: collapse;
table-layout: auto;
font-size: 11px;
overflow-y: auto;
max-height:450px !important;
white-space: nowrap;
overflow-x: auto;
width:450px;
}
.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}
.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}
.pandoc-table caption {
text-align: left !important;
font-size: 11px !important;
}
.center-table {
text-align: left !important;
font-size: 9px;
overflow-y:scroll;
height:450px;
overflow-x: scroll;
}
.controlly{
overflow-y:scroll;
height:350px;
overflow-x: scroll;
}
/*2025-03-07*/
.callout-warning,
.callout-error,
.callout-message {
  font-size: 0.7rem !important;  /* Adjust this value as needed */
}
.alert, .message {
    font-size: 0.7em !important;
}
.alert.alert-warning,
.alert.alert-danger,
.message {
    font-size: 0.7em !important;
}
</style>
```
```{css, echo=F}
#| label: css-format
h1 {
    color: var(--heading-color);
    font-size: 2rem;
    margin-bottom: 1vh;
}
p {
  font-size: 1.1rem;
  line-height: 1.6rem;
}
a {
  color: var(--primary-color);
  text-decoration: none;
  border-bottom: 3px solid transparent;
  font-weight: bold;
  &:hover, &:focus {
      border-bottom: 3px solid currentColor;
  }
}
section {
  margin: 0 auto;
}
.post-meta {
  font-size: 1rem;
  font-style: italic;
  display: block;
  margin-bottom: 4vh;
  color: var(--secondary-color);
}
nav {
  display: flex;
  justify-content: flex-end;
  padding: 20px 0;
}
/*slider switch css */
.theme-switch-wrapper {
  display: flex;
  align-items: center;
  
  em {
    margin-left: 10px;
    font-size: 1rem;
  }
}
.theme-switch {
  display: inline-block;
  height: 34px;
  position: relative;
  width: 60px;
}
.theme-switch input {
  display:none;
}
.slider {
  background-color: #ccc;
  bottom: 0;
  cursor: pointer;
  left: 0;
  position: absolute;
  right: 0;
  top: 0;
  transition: .4s;
}
.slider:before {
  background-color: #fff;
  bottom: 4px;
  content: "";
  height: 26px;
  left: 4px;
  position: absolute;
  transition: .4s;
  width: 26px;
}
input:checked + .slider {
  background-color: #66bb6a;
}
input:checked + .slider:before {
  transform: translateX(26px);
}
.slider.round {
  border-radius: 34px;
}
.slider.round:before {
  border-radius: 50%;
}
```
```{=html}
<script>
function myFunction1() {
var x = document.getElementById("myDIV");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
<script>
function myFunction2() {
var x = document.getElementById("myDIV2");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
```
<style>
.scrollable-content {
  max-height: 350px;
  overflow-y: auto;
}
pre.scrollable-code {
  max-height: 350px;
  overflow-y: auto;
}
.superbigimage {
  overflow-x: scroll;
  white-space: nowrap;
}
.superbigimage img, 
.superbigimage svg {
  max-width: none;
  height: auto;
}
</style>
<br>

# Data Loading and Exploration

## Loading Packages and uniting databases

<div class="scrollable-content">
```{r}
#| label: load
#| message: false
#| include: true
#| warning: false
#| results: "hold"
# invisible("Only run from Ubuntu")
# if (!(Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv"))) {
#   if(Sys.info()["sysname"]!="Windows"){
#     Sys.setenv(RETICULATE_PYTHON = "/home/fondecytacc/.pyenv/versions/3.11.5/bin/python")
#   }
# }

unlink("_freeze", recursive = TRUE)
unlink("*_cache", recursive = TRUE)

#clean enviroment
rm(list = ls()); gc()
time_before_dedup2<-Sys.time()
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# Busca .mamba_root/envs/py311/python.exe desde getwd() hacia padres
# find_python_rel <- function(start = getwd(),
#                            rel = file.path(".mamba_root","envs","py311","python.exe")) {
#  cur <- normalizePath(start, winslash = "/", mustWork = FALSE)
#  repeat {
#    cand <- normalizePath(file.path(cur, rel), winslash = "/", mustWork = FALSE)
#    if (file.exists(cand)) return(cand)
#    parent <- dirname(cur)
#    if (identical(parent, cur)) return(NA_character_)  # llegó a la raíz
#    cur <- parent
#  }
# }
# --- Bootstrap reticulate con ruta relativa a getwd() ---
# if(Sys.info()["sysname"]!="Windows"){
#  Sys.setenv(RETICULATE_PYTHON = "usr/bin/python3")
#  reticulate::py_config()
# } else {
#  py <- find_python_rel()
#  if (is.na(py)) {
#    stop("No se encontró Python relativo a getwd() (buscando '.mamba_root/envs/py311/python.exe').\n",
#         "Directorio actual: ", getwd())
#  }
#  # Forzar ese intérprete
#  Sys.unsetenv(c("RETICULATE_CONDAENV","RETICULATE_PYTHON_FALLBACK"))
#  Sys.setenv(RETICULATE_PYTHON = py)
#  #reticulate::use_python(py, required=T)
#  reticulate::py_config()  # verificación
# }
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# --- Load pred1_ndp_2026_01_05.Rdata (simple, project-root aware, no functions) ---
# 1) Locate the file anywhere under the project root (current WD with a trailing "/cons" trimmed)
project_root <- sub("(/)?cons/?$", "", normalizePath(getwd(), winslash = "/", mustWork = FALSE))
fname        <- "pred21_ndp_2026_03_25.Rdata"
hits <- list.files(project_root, pattern = paste0("^", fname, "$"),
                   recursive = TRUE, full.names = TRUE, ignore.case = TRUE)
if (!length(hits)) stop("File not found: ", fname, "\nSearched under: ", project_root)
path <- hits[1]  # take first match
message("Loading: ", path)
# 2) Load into a dedicated environment (keeps .GlobalEnv clean)
env_pred <- new.env(parent = emptyenv())
loaded_objects <- load(path, envir = env_pred)
message("Loaded ", length(loaded_objects), " object(s): ", paste(loaded_objects, collapse = ", "))
# 3) Quick sanity check for each loaded object
for (obj_name in loaded_objects) {
  obj <- get(obj_name, envir = env_pred)
  message("\n--- ", obj_name, " ---")
  if (is.data.frame(obj)) {
    message("Rows: ", nrow(obj), " | Cols: ", ncol(obj))
  } else {
    message("Class: ", paste(class(obj), collapse = ", "), " | Length/Size: ", length(obj))
  }
  #print(str(obj))
}
# 4) Extract all loaded objects to global environment
for (obj_name in loaded_objects) {
  assign(obj_name, get(obj_name, envir = env_pred), envir = .GlobalEnv)
}
# Clean up temporary environment
rm(env_pred)
# (Optional) If you want it in the global env, uncomment:
# list2env(env_prev1y, envir = .GlobalEnv)
```
</div>

<div class="scrollable-content">
```{r}
#| label: setup
#| results: "hold"
#renv falls back to copying rather than symlinking, which is evidently very slow in this configuration.
renv::settings$use.cache(FALSE)
#only use explicit dependencies (in DESCRIPTION)
renv::settings$snapshot.type("implicit")
#check if rstools is installed
if(Sys.info()["sysname"]=="Windows"){
try(installr::install.Rtools(check_r_update=F))
}
check_quarto_version <- function(required = "1.7.29", comparator = c("ge","gt","le","lt","eq")) {
  comparator <- match.arg(comparator)
  current <- package_version(paste(unlist(quarto::quarto_version()), collapse = "."))
  req     <- package_version(required)
  ok <- switch(comparator,
               ge = current >= req,
               gt = current >  req,
               le = current <= req,
               lt = current <  req,
               eq = current == req)
  if (!ok) {
    stop(sprintf("Quarto version check failed: need %s %s (installed: %s).",
                 comparator, required, current), call. = FALSE)
  }
  invisible(TRUE)
}
check_quarto_version("1.7.29", "ge") 
#change repository to CL
local({
  r <- getOption("repos")
  r["CRAN"] <- "https://cran.dcc.uchile.cl/"
  options(repos=r)
})
if(!require(pacman)){install.packages("pacman");require(pacman)}
if(!require(pak)){install.packages("pak");require(pak)}
pacman::p_unlock(lib.loc = .libPaths()) #para no tener problemas reinstalando paquetes
if(Sys.info()["sysname"]=="Windows"){
if (getRversion() != "4.4.1") { stop("Requires R version 4.4.1; Actual: ", getRversion()) }
}
#check docker
check_docker_running <- function() {
  # Try running 'docker info' to check if Docker is running
  system("docker info", intern = TRUE, ignore.stderr = TRUE)
}
if(Sys.info()["sysname"]=="Windows"){
  install_docker <- function() {
    # Open the Docker Desktop download page in the browser for installation
    browseURL("https://www.docker.com/products/docker-desktop")
  }
  # Main logic
  if (inherits(try(check_docker_running(), silent = TRUE), "try-error")) {
    liftr::install_docker()
  } else {
    message("Docker is running.")
  }
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#PACKAGES#######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
unlink("*_cache", recursive=T)
pak::pak_sitrep()
# pak::sysreqs_check_installed(unique(unlist(paks)))
#pak::lockfile_create(unique(unlist(paks)),  "dependencies_duplicates24.lock", dependencies=T)
#pak::lockfile_install("dependencies_duplicates24.lock")
#https://rdrr.io/cran/pak/man/faq.html
#pak::cache_delete()
library(tidytable)
library(ggplot2)
library(readr)
library(tableone)
library(survivalmodels)
#renv::install("patchwork@1.2.0")
library(rms)
library(survidm)
library(caret)
library(survival)
library(caret)
library(dplyr)
library(survex)
library(pec)
library(future)
library(future.apply)
library(parallel)
library(mice)
library(riskRegression)
library(tidyr)
library(doFuture)
library(vcd)
library(survAUC)

#flexsurv
#if(!require(compareCstat)){install.pacakges("compareCstat")}
# library(shapr)#https://norskregnesentral.github.io/shapr/
# library(SemiMarkov)#https://www.degruyterbrill.com/document/doi/10.1515/ijb-2020-0083/html?lang=en
# #https://www.jclinepi.com/article/S0895-4356(24)00142-2/fulltext
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# 3. Activate polars code completion (safe to try even if it fails)
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
#try(polars_code_completion_activate())
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# 4. BPMN from GitHub (not on CRAN, so install via devtools if missing)
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
if (!requireNamespace("bpmn", quietly = TRUE)) {
  devtools::install_github("bergant/bpmn")
}
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# 5. PhantomJS Check (use webshot if PhantomJS is missing)
#_#_#_#_#_#_#_#_#_#_#_#_#_-----------------------------
# if (!webshot::is_phantomjs_installed()) {
#   webshot::install_phantomjs()
# }
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#FUNCTIONS######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
# NO MORE DEBUGS
options(error = NULL)        # si antes tenías options(error = recover) o browser)
options(browserNLdisabled = FALSE)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#NAs are replaced with "" in knitr kable
options(knitr.kable.NA = '')
pander::panderOptions('big.mark', ',')
pander::panderOptions('decimal.mark', '.')
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
#to format rows in bold
format_cells <- function(df, rows ,cols, value = c("italics", "bold", "strikethrough")){
  # select the correct markup
  # one * for italics, two ** for bold
  map <- setNames(c("*", "**", "~~"), c("italics", "bold", "strikethrough"))
  markup <- map[value]  
  for (r in rows){
    for(c in cols){
      # Make sure values are not factors
      df[[c]] <- as.character( df[[c]])
      # Update formatting
      df[r, c] <- ifelse(nchar(df[r, c])==0,"",paste0(markup, gsub(" ", "", df[r, c]), markup))
    }
  }
  return(df)
}
#To produce line breaks in messages and warnings
knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger" style="font-size: small !important;">',
           gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning" style="font-size: small !important;">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('<div class="message" style="font-size: small !important;">',
           gsub('##', '\n', x),
           '</div>', sep = '\n')
   }
)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#
sum_dates <- function(x){
  cbind.data.frame(
    min= as.Date(min(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01"),
    p001= as.Date(quantile(unclass(as.Date(x)), .001, na.rm=T), origin = "1970-01-01"),
    p005= as.Date(quantile(unclass(as.Date(x)), .005, na.rm=T), origin = "1970-01-01"),
    p025= as.Date(quantile(unclass(as.Date(x)), .025, na.rm=T), origin = "1970-01-01"),
    p25= as.Date(quantile(unclass(as.Date(x)), .25, na.rm=T), origin = "1970-01-01"),
    p50= as.Date(quantile(unclass(as.Date(x)), .5, na.rm=T), origin = "1970-01-01"),
    p75= as.Date(quantile(unclass(as.Date(x)), .75, na.rm=T), origin = "1970-01-01"),
    p975= as.Date(quantile(unclass(as.Date(x)), .975, na.rm=T), origin = "1970-01-01"),
    p995= as.Date(quantile(unclass(as.Date(x)), .995, na.rm=T), origin = "1970-01-01"),
    p999= as.Date(quantile(unclass(as.Date(x)), .999, na.rm=T), origin = "1970-01-01"),
    max= as.Date(max(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01")
  )
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# Define the function adapted for Polars
sum_dates_polars <- function(df, date_col) {
  # Create the list of quantiles
  quantiles <- c(0.001, 0.005, 0.025, 0.25, 0.5, 0.75, 0.975, 0.995, 0.999)
  # Create expressions to calculate min and max
  expr_list <- list(
    pl$col(date_col)$min()$alias("min"),
    pl$col(date_col)$max()$alias("max")
  )
  # Add expressions for quantiles
  for (q in quantiles) {
    expr_list <- append(expr_list, pl$col(date_col)$quantile(q)$alias(paste0("p", sub("\\.", "", as.character(q)))))
  }
  # Apply the expressions and return a DataFrame with the results
  df$select(expr_list)
}
# Custom function for sampling with a seed
sample_n_with_seed <- function(data, size, seed) {
  set.seed(seed)
  dplyr::sample_n(data, size)
}
# Function to get the most frequent value 
most_frequent <- function(x) { 
  uniq_vals <- unique(x)
  freq_vals <- sapply(uniq_vals, function(val) sum(x == val))
  most_freq <- uniq_vals[which(freq_vals == max(freq_vals))]
  
  if (length(most_freq) == 1) {
    return(most_freq)
  } else {
    return(NA)
  }
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#CONFIG #######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
options(scipen=2) #display numbers rather scientific number
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# Define the function first
#joins these values with semicolons and optionally truncates the result if it exceeds a specified width.
toString2 <- function(x, width = NULL, ...) {
    string <- paste(x, collapse = "; ")
    if (missing(width) || is.null(width) || width == 0) 
        return(string)
    if (width < 0) 
        stop("'width' must be positive")
    if (nchar(string, type = "w") > width) {
        width <- max(6, width)
        string <- paste0(substr(string, 1, width - 3), "...")
    }
    string
}
normalize_txt <- function(x) {
  x|>
    stringi::stri_trans_general("Latin-ASCII")|>
    tolower()|>
    trimws()
}
pair_str <- function(main, sub) {
  main <- as.character(main)
  sub  <- as.character(sub)
  both_na <- is.na(main) & is.na(sub)
  out <- paste0(
    tidytable::coalesce(main, "NA"),
    "::",
    tidytable::coalesce(sub,  "NA")
  )
  out[both_na] <- NA_character_
  out
}
# ── Helpers ────────────────────────────────────────────────────────────
mode_pick_int <- function(x){
  x <- x[!is.na(x)]
  if(length(x)==0) return(NA_integer_)
  tx <- sort(table(x), decreasing = TRUE)
  as.integer(names(tx)[1L])
}
subkey_to_label <- function(x){
  y <- gsub("_"," ", tolower(x))
  y <- gsub("amphetamine type stimulants","amphetamine-type stimulants", y)
  y <- gsub("tranquilizers hypnotics","tranquilizers/hypnotics", y)
  y
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# ── Encryption ────────────────────────────────────────────────────────
# Pack: [24-byte nonce][ciphertext] -> hex
encrypt_chr <- function(x, key) {
    vapply(x, function(v) {
        if (is.na(v)) return(NA_character_)
        ct <- sodium::data_encrypt(charToRaw(as.character(v)), key)
        nonce <- attr(ct, "nonce")                 # 24 bytes
        sodium::bin2hex(c(nonce, ct))              # pack nonce + ct
    }, FUN.VALUE = character(1))
}
# Unpack: hex -> [nonce|ciphertext], restore attr(nonce), then decrypt
decrypt_chr <- function(x, key) {
    vapply(x, function(v) {
        if (is.na(v)) return(NA_character_)
        buf <- sodium::hex2bin(v)
        if (length(buf) < 25L) stop("Ciphertext too short or corrupted.")
        nonce <- buf[1:24]
        ct    <- buf[-(1:24)]
        attr(ct, "nonce") <- nonce
        rawToChar(sodium::data_decrypt(ct, key))
    }, FUN.VALUE = character(1))
}
# Function to encrypt a character vector
encrypt_column <- function(x, key) {
  sapply(x, function(item) {
    if (is.na(item) || item == "") {
      return(NA_character_)
    }
    encrypted_raw <- sodium::data_encrypt(charToRaw(item), key)
    base64enc::base64encode(encrypted_raw)  # Convert to base64 for storage
  }, USE.NAMES = FALSE)
}
decrypt_column <- function(x, key) {
  sapply(x, function(item) {
    if (is.na(item)) return(NA_character_)
    encrypted_raw <- base64enc::base64decode(item)
    rawToChar(sodium::data_decrypt(encrypted_raw, key))
  }, USE.NAMES = FALSE)
}
is_stata_ok <- function(x) {
  nchar(x) <= 32 & grepl("^[A-Za-z][A-Za-z0-9_]*$", x)
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
# ── Tidy RMS ────────────────────────────────────────────────────────
tidy_cph <- function(model) {
  if (!inherits(model, "cph")) stop("Model must be an rms::cph object")
  
  # run summary
  s <- summary(model)
  df_s <- as.data.frame(unclass(s))
  df_s$rown <- rownames(s)
  
  # find rows that are "Hazard Ratio"
  hr_rows <- trimws(df_s$rown) == "Hazard Ratio"
  
  # build tidy tibble
  tibble::tibble(
    term     = trimws(rownames(s)[which(hr_rows) - 1L]), # previous row = variable name
    estimate = df_s$Effect[hr_rows],                     # Hazard Ratio
    conf.low = df_s$`Lower 0.95`[hr_rows],
    conf.high= df_s$`Upper 0.95`[hr_rows],
    p.value  = df_s$P[hr_rows]
  )
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
find_latest_file <- function(project_root, prefix) {
  # Get all files matching the pattern
  pattern <- paste0(prefix, "_\\d{8}_\\d{4}\\.csv$")
  files <- list.files(
    path = project_root,
    pattern = pattern,
    full.names = TRUE
  )
  if (length(files) == 0) {
    stop(paste("No files found matching pattern:", pattern))
  }
  # Extract timestamps from filenames (format: YYYYMMDD_HHMM)
  extract_timestamp <- function(filename) {
    matches <- regmatches(
      basename(filename),
      gregexpr("\\d{8}_\\d{4}", basename(filename))
    )
    if (length(matches[[1]]) == 0) {
      return(NA)
    }
    timestamp_str <- matches[[1]][length(matches[[1]])]
    return(timestamp_str)
  }
  timestamps <- sapply(files, extract_timestamp)
  # Remove any files where timestamp extraction failed
  valid_idx <- !is.na(timestamps)
  files <- files[valid_idx]
  timestamps <- timestamps[valid_idx]
  if (length(files) == 0) {
    stop("No valid timestamps found in filenames")
  }
  # Sort by timestamp (descending) and return the latest
  latest_idx <- order(timestamps, decreasing = TRUE)[1]
  return(files[latest_idx])
}

```
</div>

```{r}
# Instala 'here' si no lo tienes: if(!require(here)) install.packages("here")

# Define la raíz del proyecto sin importar el OS
# 'here' siempre apunta a la carpeta que contiene el archivo .Rproj
root <- here::here()

# Si necesitas la ruta "padre" de /cons (como haces en tu script):
project_root <- gsub("/cons$", "", root) 

# Definir carpetas de salida de forma relativa
data_out <- file.path(project_root, "data", "20241015_out")
figs_out <- file.path(root, "_figs")
```

<br>

::: {.callout-note}
**Aims**
- To identify key prognostic and predictive factors associated with the risk of, and time to, the first readmission to treatment among patients aged 18-64 undergoing initial treatment for Substance Use Disorder (SUD) in Chile between 2010 and 2020.
- To identify key prognostic and predictive factors associated with the risk of, and time to, all-cause mortality among the same cohort of patients undergoing initial treatment for SUD in Chile.
- To predict the risk and timing of first treatment readmission and all-cause mortality simultaneously, treating these outcomes within a semi-competing risks framework. This will allow for the estimation of cumulative incidence for each event in the presence of the other for patients aged 18-64 who initiated SUD treatment in Chile between 2010 and 2020.
This study has three primary objectives. First, we aim to identify key prognostic and predictive factors associated with the risk and timing of two critical outcomes: (i) first readmission to treatment and (ii) all-cause mortality. Second, using a semi-competing risks framework, we will develop a model to simultaneously predict these two outcomes, accounting for the fact that death precludes the possibility of readmission. Finally, this model will be used to estimate the probability of various patient trajectories over time following initial treatment. The study population includes all patients aged 18-64 who began treatment for Substance Use Disorder (SUD) in Chile between 2010 and 2020.
:::
<br>

```{r}
#| label: shap-criteria-setup


gc()

total_cpus <- parallel::detectCores()
cpus <- ifelse(total_cpus > 30, 16, 7)


predictor_tiers <- readxl::read_excel(
  file.path(project_root, "cons", "_out",
            "xgb9_dual_predictor_analysis_no_clinical_20260306_1821.xlsx"),
  sheet = "Predictors"
) |>
  dplyr::filter(!grepl("MINOR",importance_tier))


predictor_tiers |> knitr::kable(
  caption = "Predictor tiers based on SHAP importance from XGBoost models"
)
```

```{r}
#| label: shap-criteria

gc()

# # Linear time interaction
# coxph(Surv(time, death) ~ var + var:time + other_covariates, data = df)
# # Log-time interaction
# coxph(Surv(time, death) ~ var + var:log(time) + other_covariates, data = df)
# # Step function at 1 year (365 days)
# df$late_period <- as.numeric(df$time > 365)
# coxph(Surv(time, death) ~ var + var:late_period + other_covariates, data = df)
# # Using splines (requires splines package)
# library(splines)
# coxph(Surv(time, death) ~ var + var:ns(time, df=3) + other_covariates, data = df)
vars_shap_r <- c(
  "primary_sub_mod_cocaine_paste",
  "adm_age_rec3",
  "porc_pobr",
  "sex_rec_woman",
  #"plan_type_corr_pg_pr", # Mantenido en el código como strata()
  #"plan_type_corr_m_pr",  # Mantenido en el código como strata()
  "ethnicity",
  "dit_m",
  #"eva_consumo",
  "eva_consumo_logro_intermedio",
  "eva_consumo_logro_minimo",
  #"ed_attainment_corr",
  "ed_attainment_corr_2_completed_high_school_or_less",
  "ed_attainment_corr_3_completed_primary_school_or_less",
  "sub_dep_icd10_status_drug_dependence",
  "occupation_condition_corr24_unemployed",
  "polysubstance_strict",
  "primary_sub_mod_cocaine_powder",
  #"evaluacindelprocesoteraputico",
  "evaluacindelprocesoteraputico_logro_intermedio",
  "evaluacindelprocesoteraputico_logro_minimo", 
  "dg_psiq_cie_10_dg",
  "tr_outcome_adm_discharge_rule_violation_undet",
  #"eva_sm",
  "eva_sm_logro_intermedio",
  "eva_sm_logro_minimo",   
  
  # --- NO LONGER IN TOP 20 MORTALITY (SHAP) ---
  #"adm_motive_justice_sector",
  #"tenure_status_household_others",
  #"tenure_status_household_renting",
  #"tenure_status_household_stays_temporarily_with_a_relative",
  #"tenure_status_household_illegal_settlement",
  #"tr_outcome_referral",
  #"eva_ocupacion_logro_intermedio",
  #"eva_ocupacion_logro_minimo",   
  #"eva_transgnorma_logro_intermedio",
  #"eva_transgnorma_logro_minimo",    
  
  #"prim_sub_freq_rec",
  "prim_sub_freq_rec_2_2_6_days_wk",
  "prim_sub_freq_rec_3_daily",  
  
  # --- NO LONGER IN TOP 20 MORTALITY  (SHAP) ---
  #"cohabitation_with_couple_children",
  
  "primary_sub_mod_alcohol"
  
  # --- NO LONGER IN TOP 20 MORTALITY (SHAP) ---
  #"eva_relinterp_logro_intermedio",
  #"eva_relinterp_logro_minimo",  
  #"tr_outcome_dropout",
  #"dx_f6_personality",
  #"any_phys_dx",
  #"any_violence_1_domestic_violence_sex_abuse",
  #"adm_motive_sanitary_sector",
  #"eva_fam_logro_intermedio",
  #"eva_fam_logro_minimo"
)

vars_shap_d <- c(
  "adm_age_rec3",
  "primary_sub_mod_cocaine_paste",
  #"prim_sub_freq_rec", 
  "prim_sub_freq_rec_2_2_6_days_wk",
  "prim_sub_freq_rec_3_daily",
  
  # --- NO LONGER IN TOP 20 MORTALITY (SHAP 18:22) ---
  #"primary_sub_mod_cocaine_powder",
  
  "occupation_condition_corr24_unemployed",
  "any_phys_dx",
  #"eva_ocupacion",
  "eva_ocupacion_logro_intermedio",
  "eva_ocupacion_logro_minimo",  
  "first_sub_used_alcohol",
  "sex_rec_woman",
  #"eva_fisica"
  "eva_fisica_logro_intermedio",
  "eva_fisica_logro_minimo"
)
formula_shap_readmit <- as.formula(
  paste("Surv(readmit_time_from_disch_m, readmit_event) ~", 
        paste(vars_shap_r, collapse = " + "), "+ strata(plan_type_strata)")
)
formula_shap_death <- as.formula(
  paste("Surv(death_time_from_disch_m, death_event) ~", 
        paste(vars_shap_d, collapse = " + "), "+ strata(plan_type_strata)")
)
dplyr::bind_rows(
  rms::vif(survival::coxph(formula_shap_death, data = py_corrected_datasets[[1]])) |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    dplyr::rename(Gvif = 2) |>
    dplyr::arrange(dplyr::desc(dplyr::pick(2)[[1]])) |>
    dplyr::filter(Gvif > 10) |>
    dplyr::mutate(Outcome = "Death"),
  rms::vif(survival::coxph(formula_shap_readmit, data = py_corrected_datasets[[1]])) |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    dplyr::rename(Gvif = 2) |>
    dplyr::filter(Gvif > 10) |>
    dplyr::mutate(Outcome = "Readmission")
) |>
  knitr::kable("markdown", caption = "GVIF for SHAP-based Cox models, using one imputed dataset")

cat("Removing collineal variable\n")
formula_shap_readmit_clean <- as.formula(
  paste(gsub("evaluacindelprocesoteraputico_logro_minimo \\+\\s*","",paste(deparse(formula_shap_readmit), collapse = " "))))

results_readmit_shap <- evaluate_mi_cv_safe_strata(
  formula = formula_shap_readmit_clean, 
  imputed_list = py_corrected_datasets, 
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  k = 5,
  n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6057 (95% CI 0.6056-0.6058)
# Pooled UnoC (survAUC::UnoC): 0.6620 (95% CI 0.6379-0.6862)
# Pooled IBS: 0.1600 (95% CI 0.1600-0.1600)
# Pooled IBS (IPCW-train): 0.3480 (95% CI 0.3479-0.3481)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6312 (95% CI 0.6310-0.6315, valid repeats=2)
#   t=36 m: C=0.6139 (95% CI 0.6138-0.6139, valid repeats=2)
#   t=60 m: C=0.6087 (95% CI 0.6086-0.6088, valid repeats=2)
# ============================================================
#Total runtime: 148.87 sec (2.48 min)

results_death_shap <- evaluate_mi_cv_safe_strata(
  formula = formula_shap_death, 
  imputed_list = py_corrected_datasets, 
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  k = 5,
  n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7367 (95% CI 0.7366-0.7369)
# Pooled UnoC (survAUC::UnoC): 0.8059 (95% CI 0.8026-0.8093)
# Pooled IBS: 0.0356 (95% CI 0.0356-0.0356)
# Pooled IBS (IPCW-train): 0.0903 (95% CI 0.0902-0.0904)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7377 (95% CI 0.7375-0.7379, valid repeats=2)
#   t=36 m: C=0.7495 (95% CI 0.7492-0.7499, valid repeats=2)
#   t=60 m: C=0.7500 (95% CI 0.7497-0.7502, valid repeats=2)
# ============================================================

results_readmit_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_shap_readmit_clean,
    imputed_list = py_corrected_datasets, # List of data.frames
    time_col = "readmit_time_from_disch_m", 
    event_col = "readmit_event",
    tau = 99,                # Evaluate up to 60 months
    B = 500,
    cpus = cpus,
    verbose = TRUE,
    min_strata_events = 1
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1584 - 0.1586)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_shap_death,
    imputed_list = py_corrected_datasets, # List of data.frames
    time_col = "death_time_from_disch_m", 
    event_col = "death_event",
    tau = 107,                # Evaluate up to 60 months
    B = 500,
    cpus = cpus,
    verbose = TRUE,
    min_strata_events = 1
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0363 (95% CI: 0.0362 - 0.0363)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================

aicc_readmit_shap <- compute_mi_aicc(formula_shap_readmit_clean, py_corrected_datasets)
#282719.30
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_readmit_shap$mean_aic - aicc_readmit_upd2$mean_aic))
#Difference in Mean AIC: 10170.15
aicc_death_shap <- compute_mi_aicc(formula_shap_death, py_corrected_datasets)
#54321.99
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_death_shap$mean_aic - aicc_death_upd2$mean_aic))
#Difference in Mean AIC: 2825.75
cat("Comparing readmission w/o collinear var and only one strata(plan)\n")
compare_mi_models_d1(formula_readmit_updated,formula_shap_readmit_clean, py_corrected_datasets)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test statistic df1      df2 dfcom      p.value        riv
#  1 ~~ 2  2.705153  42 11680.04 15183 1.835904e-08 0.05960578
# ============================================================

compare_mi_models_d1(formula_death_updated,formula_shap_death, py_corrected_datasets)
# ============================================================
# MODEL COMPARISON RESULT (D1 Statistic)
# ============================================================
#    test statistic df1      df2 dfcom     p.value        riv
#  1 ~~ 2  9.059786  52 2901.688  2975 6.69787e-63 0.04231255
# ============================================================

cal_readmit_shap<- 
  calibrate_cph_grouped(
    formula = formula_shap_readmit_clean, # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    validation = "cv",
    n_repeats = 20,
    n_imputations = 5,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
  )
print(cal_readmit_shap)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00367  0.00245   0.991     5
# 2          36  0.00635  0.00384   0.988     5
# 3          60  0.00691  0.00478   0.984     5
# 4          96  0.00708  0.00434   0.976     5

cal_death_shap<- 
    calibrate_cph_grouped(
    formula = formula_shap_death, # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    validation = "cv",
    n_repeats = 20,
    n_imputations = 5,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
  )

print(cal_death_shap)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00151  0.00148    1.01     5
# 2          36  0.00230  0.00203    1.02     5
# 3          60  0.00327  0.00305    1.03     5
# 4          96  0.00336  0.00312    1.06     5
```


```{r}
#| label: shap-criteria-nph

cox.zph(coxph(formula_shap_death, data= py_corrected_datasets[[1]]))
cox.zph(coxph(formula_shap_readmit_clean, data= py_corrected_datasets[[1]]))
#tr_outcome_adm_discharge_rule_violation_undet          58.6456  1 1.9e-14

```


```{r}
#| label: shap-criteria-additional-strata

# Residuals: discharge to readmission
res <- resid(coxph(formula_shap_readmit_clean, data= py_corrected_datasets[[1]]), "scaledsch")
time <- as.numeric(dimnames(res)[[1]])
z2 <- loess(res[,"tr_outcome_adm_discharge_rule_violation_undet"] ~ time, span=0.50)

plot(time, fitted(z2), type="l", col="blue",
     main="Sch. res. – Completion (Violation of tules of the center)",
     xlab="Time", ylab="Residuals")
lines(supsmu(time, res[,"tr_outcome_adm_discharge_rule_violation_undet"]), lty=2, col="darkgray")


formula_shap_readmit_clean_updated <- update(
  formula_shap_readmit_clean,
  . ~ . - tr_outcome_adm_discharge_rule_violation_undet + strat(tr_outcome_adm_discharge_rule_violation_undet)
)
```

```{r}
#| label: shap-criteria-additional-strata2

results_readmit_shap2 <- evaluate_mi_cv_safe_strata(
  formula = as.formula(
  gsub(
    "strat\\(",
    "strata(",
    paste(deparse(formula_shap_readmit_clean_updated), collapse = " ")
  )
), 
  imputed_list = py_corrected_datasets, 
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  k = 5,
  n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6056 (95% CI 0.6054-0.6057)
# Pooled UnoC (survAUC::UnoC): 0.6610 (95% CI 0.6374-0.6845)
# Pooled IBS: 0.1599 (95% CI 0.1599-0.1599)
# Pooled IBS (IPCW-train): 0.3479 (95% CI 0.3478-0.3480)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6316 (95% CI 0.6314-0.6319, valid repeats=2)
#   t=36 m: C=0.6137 (95% CI 0.6136-0.6138, valid repeats=2)
#   t=60 m: C=0.6085 (95% CI 0.6084-0.6086, valid repeats=2)
# ============================================================
results_death_shap <- evaluate_mi_cv_safe_strata(
  formula =  as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_death))), 
  imputed_list = py_corrected_datasets, 
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  k = 5,
  n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7367 (95% CI 0.7366-0.7369)
# Pooled UnoC (survAUC::UnoC): 0.8059 (95% CI 0.8026-0.8093)
# Pooled IBS: 0.0356 (95% CI 0.0356-0.0356)
# Pooled IBS (IPCW-train): 0.0903 (95% CI 0.0902-0.0904)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7377 (95% CI 0.7375-0.7379, valid repeats=2)
#   t=36 m: C=0.7495 (95% CI 0.7492-0.7499, valid repeats=2)
#   t=60 m: C=0.7500 (95% CI 0.7497-0.7502, valid repeats=2)
# ============================================================
results_readmit_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_readmit_clean_updated))),
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "readmit_time_from_disch_m", 
  event_col = "readmit_event",
  tau = 99,                # Evaluate up to 60 months
  B = 500,
  cpus = cpus,
  verbose = TRUE,
  min_strata_events = 1
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.1585 (95% CI: 0.1584 - 0.1586)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
results_death_shap_ibs <- cph_evaluate_boot_oob_mi_corrected(
  formula = as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_death))),
  imputed_list = py_corrected_datasets, # List of data.frames
  time_col = "death_time_from_disch_m", 
  event_col = "death_event",
  tau = 107,                # Evaluate up to 60 months
  B = 500,
  cpus = cpus,
  verbose = TRUE,
  min_strata_events = 1
)
# ============================================================
# FINAL POOLED RESULTS (OOB Bootstrap + MI)
# ============================================================
# Pooled OOB IBS: 0.0363 (95% CI: 0.0362 - 0.0363)
# Valid bootstrap replicates: 2 / 2
# Event value used: 1
# ============================================================
aicc_readmit_shap <- compute_mi_aicc(as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_readmit_clean_updated))), py_corrected_datasets)
#Mean AICc: 282719.30
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_readmit_shap$mean_aic - aicc_readmit_upd2$mean_aic))
#Difference in Mean AIC: 10170.15
aicc_death_shap <- compute_mi_aicc(as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_death))), py_corrected_datasets)
#Mean AICc: 54321.99
cat(sprintf("\nDifference in Mean AIC: %.2f\n", aicc_death_shap$mean_aic - aicc_death_upd2$mean_aic))
#Difference in Mean AIC: 2825.75

cal_readmit_shap2<- 
  calibrate_cph_grouped(
    formula = as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_readmit_clean_updated))),  # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    validation = "cv",
    n_repeats = 20,
    n_imputations = 5,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
  )
print(cal_readmit_shap2)
#  time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00369  0.00281   0.991     5
# 2          36  0.00646  0.00449   0.988     5
# 3          60  0.00736  0.00500   0.985     5
# 4          96  0.00718  0.00405   0.977     5

cal_death_shap2<- 
    calibrate_cph_grouped(
    formula = as.formula(gsub("strat\\(", "strata(", deparse(formula_shap_death))),  # Original formula with months variable
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    validation = "cv",
    n_repeats = 20,
    n_imputations = 5,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
  )

print(cal_death_shap2)
#   time_months ece_mean ici_mean eo_mean n_imp
#         <dbl>    <dbl>    <dbl>   <dbl> <int>
# 1          12  0.00151  0.00148    1.01     5
# 2          36  0.00230  0.00203    1.02     5
# 3          60  0.00327  0.00305    1.03     5
# 4          96  0.00336  0.00312    1.06     5
```

With stratification, individuals in different strata have different baseline hazards, and therefore different predicted survival probabilities. This creates some discrimination purely from the stratum membership.

For the Wald test (𝐷1) to be valid, the ‘Reduced’ model must be a perfect subset of the ‘Full’ model. You cannot change the underlying structure (strata).


## Interactions

```{r}
#| label: shap-criteria-int

global_df <- read_csv(
  find_latest_file(
    file.path(project_root, "cons", "_out"),
    "xgb11_dual_interactions_global_top_enriched"),
  show_col_types = FALSE
)
time_dep_df <- read_csv(
  find_latest_file(
    file.path(project_root, "cons", "_out"), 
    "xgb11_dual_interactions_time_dependent"),
  show_col_types = FALSE
)

# 2. Merge on unique keys
merged_df <- global_df %>%
  left_join(time_dep_df, by = c("outcome", "main_feature", "interactor"))

# 3. Apply the Effect-Size & Stability Rule
cox_discriminative_interactions <- merged_df %>%
  filter(
    # Ignore FDR completely. Use joint clinical/linear magnitude:
    abs_corr >= 0.10,
    !is.na(smd_binarized) & abs(smd_binarized) >= 0.15,
    
    # Enforce standard Cox Proportional Hazards assumptions:
    time_dependent_flag == FALSE,
    direction_flip_flag == FALSE
  ) %>%
  select(
    outcome, 
    main_feature, 
    interactor, 
    abs_corr, 
    smd_binarized, 
    direction_q4_q1
  ) %>%
  arrange(outcome, desc(abs_corr))

# Print results
cat("Number of highly discriminative candidates for Cox model:", nrow(cox_discriminative_interactions), "\n")
knitr::kable(cox_discriminative_interactions, "markdown", caption = "Highly discriminative interactions for Cox models based on Effect-Size & Stability Rule")


```

```{r}
#| label: shap-criteria-int2
formula_shap_readmit_interact <- stats::update(
  formula_shap_readmit_clean_updated,
  . ~ . +
    tr_outcome_referral +
    primary_sub_mod_cocaine_paste:polysubstance_strict +
    eva_consumo_logro_intermedio:tr_outcome_referral +
    eva_consumo_logro_minimo:tr_outcome_referral
)
formula_shap_death_interact <- stats::update(
  formula_shap_death,
  . ~ . +
    polysubstance_strict +
    cohabitation_with_couple_children +
    occupation_condition_corr24_inactive +
    primary_sub_mod_cocaine_paste:polysubstance_strict +
    cohabitation_with_couple_children:adm_age_rec3 +
    occupation_condition_corr24_unemployed:adm_age_rec3 +
    sex_rec_woman:occupation_condition_corr24_inactive +
    primary_sub_mod_cocaine_paste:first_sub_used_alcohol
)

# Convertir a fórmula
formula_shap_readmit_interact <- as.formula(gsub("strat\\(", "strata(", paste(deparse(formula_shap_readmit_interact), collapse = " ")))
formula_shap_death_interact <- as.formula(
  gsub("strat\\(", "strata(", 
       paste(deparse(formula_shap_death_interact), collapse = " "))
)


results_readmit_int_shap <- evaluate_mi_cv_safe_strata(
    formula = formula_shap_readmit_interact, 
    imputed_list = py_corrected_datasets, 
    time_col = "readmit_time_from_disch_m", 
    event_col = "readmit_event",
    k = 5,
    n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6066 (95% CI 0.6064-0.6068)
# Pooled UnoC (survAUC::UnoC): 0.6583 (95% CI 0.6350-0.6817)
# Pooled IBS: 0.1599 (95% CI 0.1598-0.1599)
# Pooled IBS (IPCW-train): 0.3476 (95% CI 0.3475-0.3477)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6327 (95% CI 0.6324-0.6331, valid repeats=2)
#   t=36 m: C=0.6146 (95% CI 0.6145-0.6148, valid repeats=2)
#   t=60 m: C=0.6096 (95% CI 0.6095-0.6097, valid repeats=2)
# ============================================================
# 
# Total runtime: 127.10 sec (2.12 min)
results_death_int_shap <- evaluate_mi_cv_safe_strata(
    formula = formula_shap_death_interact, 
    imputed_list = py_corrected_datasets, 
    time_col = "death_time_from_disch_m", 
    event_col = "death_event",
    k = 5,
    n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7400 (95% CI 0.7399-0.7400)
# Pooled UnoC (survAUC::UnoC): 0.8173 (95% CI 0.8114-0.8232)
# Pooled IBS: 0.0352 (95% CI 0.0352-0.0352)
# Pooled IBS (IPCW-train): 0.0893 (95% CI 0.0892-0.0894)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7431 (95% CI 0.7430-0.7431, valid repeats=2)
#   t=36 m: C=0.7542 (95% CI 0.7539-0.7545, valid repeats=2)
#   t=60 m: C=0.7549 (95% CI 0.7547-0.7551, valid repeats=2)
# ============================================================
# 
# Total runtime: 123.76 sec (2.06 min)
results_readmit_shap_int_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_shap_readmit_interact,
    imputed_list = py_corrected_datasets, # List of data.frames
    time_col = "readmit_time_from_disch_m", 
    event_col = "readmit_event",
    tau = 99,                # Evaluate up to 60 months
    B = 500,
    cpus = cpus,
    verbose = TRUE,
    min_strata_events = 1
)
results_death_shap_int_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_shap_death_interact,
    imputed_list = py_corrected_datasets, # List of data.frames
    time_col = "death_time_from_disch_m", 
    event_col = "death_event",
    tau = 107,                # Evaluate up to 60 months
    B = 500,
    cpus = cpus,
    verbose = TRUE,
    min_strata_events = 1
)

# ====
aicc_readmit_shap_int <- compute_mi_aicc(formula_shap_readmit_interact, py_corrected_datasets)
aicc_readmit_shap_int
aicc_death_shap_int <- compute_mi_aicc(formula_shap_death_interact, py_corrected_datasets)
aicc_death_shap_int
cal_readmit_shap_int<- 
    calibrate_cph_grouped(
    formula = formula_shap_readmit_interact,
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    validation = "cv",
    n_repeats = 20,
    n_imputations = 5,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
  )

print(cal_readmit_shap_int)
cal_death_shap_int<- 
    calibrate_cph_grouped(
    formula = formula_shap_death_interact,
    data = py_corrected_datasets, # Original data (months)
    times = c(12, 36, 60, 96),
    seed = 2125,
    validation = "cv",
    n_repeats = 20,
    n_imputations = 5,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
  )

print(cal_death_shap_int)
```

## Functional forms

```{r}
#| label: shap-criteria-func-form

#Build spline versions of your current final formulas

# helper: replace term inside formula text
replace_term_in_formula <- function(formula_obj, old_term, new_term) {
  f_txt <- paste(deparse(formula_obj), collapse = " ")
  f_txt <- base::gsub(
    pattern = paste0("(?<![[:alnum:]_])", old_term, "(?![[:alnum:]_])"),
    replacement = new_term,
    x = f_txt,
    perl = TRUE
  )
  stats::as.formula(f_txt)
}
add_boot_splines <- function(df) {
  a <- splines::ns(df$adm_age_rec3, df = 3)
  colnames(a) <- paste0("adm_age_ns", 1:3)
  d <- splines::ns(df$dit_m, df = 4)
  colnames(d) <- paste0("dit_ns", 1:4)
  p <- splines::ns(df$porc_pobr, df = 3)
  colnames(p) <- paste0("porc_pobr_ns", 1:3)
  cbind(df, as.data.frame(a), as.data.frame(d), as.data.frame(p))
}
py_corrected_datasets_boot <- lapply(py_corrected_datasets, add_boot_splines)

#-----------------------------
# READMISSION
#-----------------------------
# READMISSION: only main effects need replacement
formula_readmit_int_spline <- update(
  formula_shap_readmit_interact,
  . ~ . -
    adm_age_rec3 -
    dit_m -
    porc_pobr +
    adm_age_ns1 + adm_age_ns2 + adm_age_ns3 +
    dit_ns1 + dit_ns2 + dit_ns3 + dit_ns4 +
    porc_pobr_ns1 + porc_pobr_ns2 + porc_pobr_ns3
)
# optional extra interaction
formula_readmit_int_spline_i1 <- update(
  formula_readmit_int_spline,
  . ~ .
)

#-----------------------------
# MORTALITY
#-----------------------------
# DEATH: main effect + both age interactions must be expanded
formula_death_int_spline <- update(
  formula_shap_death_interact,
  . ~ . -
    adm_age_rec3 -
    adm_age_rec3:cohabitation_with_couple_children -
    adm_age_rec3:occupation_condition_corr24_unemployed +
    adm_age_ns1 + adm_age_ns2 + adm_age_ns3 +
    adm_age_ns1:cohabitation_with_couple_children +
    adm_age_ns2:cohabitation_with_couple_children +
    adm_age_ns3:cohabitation_with_couple_children +
    adm_age_ns1:occupation_condition_corr24_unemployed +
    adm_age_ns2:occupation_condition_corr24_unemployed +
    adm_age_ns3:occupation_condition_corr24_unemployed
)
```

# Final SHAP model

```{r }
#| label: final-shap-build-splines

gc()
```

```{r }
#| label: final-shap-rank-models

#2. Rank models across your imputed dataset list
safe_cox <- function(formula, data) {
  base::tryCatch(
    survival::coxph(
      formula,
      data = data,
      ties = "efron",
      x = TRUE,
      y = TRUE
    ),
    error = function(e) NULL
  )
}
rank_models <- function(forms, data_list) {
  purrr::map_dfr(names(forms), function(nm) {
    fits <- purrr::map(data_list, ~ safe_cox(forms[[nm]], .x))
    aics <- purrr::map_dbl(fits,~ if (base::is.null(.x)) NA_real_ else stats::AIC(.x))
    tibble::tibble(
      model = nm,
      n_fit = base::sum(!base::is.na(aics)),
      median_AIC = stats::median(aics, na.rm = TRUE),
      mean_AIC = base::mean(aics, na.rm = TRUE)
    )
  }) |>
    dplyr::arrange(median_AIC)
}
```


```{r }
#| label: final-shap-plug-to-cv
#4. Plug directly into your existing CV evaluation

readm_forms <- list(
    base_shap_int = formula_shap_readmit_interact,
    spline_main   = formula_readmit_int_spline#,
    #spline_i1     = formula_readmit_int_spline_i1
)


death_forms <- list(
    base_shap_int = formula_shap_death_interact,
    spline_main   = formula_death_int_spline
)

rank_models(readm_forms, py_corrected_datasets_boot)
#   model         n_fit median_AIC mean_AIC
#   <chr>         <int>      <dbl>    <dbl>
# 1 spline_main       5    272430.  272432.
# 2 spline_i1         5    272444.  272444.
# 3 base_shap_int     5    272560.  272560.

rank_models(death_forms, py_corrected_datasets_boot)
#   model         n_fit median_AIC mean_AIC
#   <chr>         <int>      <dbl>    <dbl>
# 1 spline_main       5     54159.   54156.
# 2 base_shap_int     5     54188.   54186.

```

```{r }
#| label: final-shap-int-spline-eval
options(future.globals.maxSize = 12 * 1024^3)

#formula_death_int_spline
results_readmit_int_shap_spl <- evaluate_mi_cv_safe_strata(
    formula = formula_readmit_int_spline, 
    imputed_list = py_corrected_datasets_boot, 
    time_col = "readmit_time_from_disch_m", 
    event_col = "readmit_event",
    k = 5,
    n_repeats = 20
)#
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 99.10 months
# Pooled Uno's C-Index: 0.6093 (95% CI 0.6092-0.6094)
# Pooled UnoC (survAUC::UnoC): 0.6526 (95% CI 0.6303-0.6749)
# Pooled IBS: 0.1602 (95% CI 0.1602-0.1603)
# Pooled IBS (IPCW-train): 0.3479 (95% CI 0.3478-0.3480)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.6334 (95% CI 0.6332-0.6336, valid repeats=2)
#   t=36 m: C=0.6166 (95% CI 0.6166-0.6167, valid repeats=2)
#   t=60 m: C=0.6123 (95% CI 0.6122-0.6124, valid repeats=2)
# ============================================================
results_death_int_shap_spl <- evaluate_mi_cv_safe_strata(
    formula = formula_death_int_spline, 
    imputed_list = py_corrected_datasets_boot, 
    time_col = "death_time_from_disch_m", 
    event_col = "death_event",
    k = 5,
    n_repeats = 20
)
# ============================================================
# FINAL POOLED RESULTS (Cross-Validated + Multiply Imputed)
# ============================================================
# Global Evaluation Time (tau): 107.94 months
# Pooled Uno's C-Index: 0.7409 (95% CI 0.7408-0.7410)
# Pooled UnoC (survAUC::UnoC): 0.8152 (95% CI 0.8083-0.8220)
# Pooled IBS: 0.0352 (95% CI 0.0352-0.0352)
# Pooled IBS (IPCW-train): 0.0892 (95% CI 0.0891-0.0893)
# RMST-based Uno C-index by horizon:
#   t=12 m: C=0.7439 (95% CI 0.7439-0.7440, valid repeats=2)
#   t=36 m: C=0.7550 (95% CI 0.7547-0.7553, valid repeats=2)
#   t=60 m: C=0.7559 (95% CI 0.7557-0.7561, valid repeats=2)
# ============================================================
results_readmit_shap_int_spl_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_readmit_int_spline,
    imputed_list = py_corrected_datasets_boot, # List of data.frames
    time_col = "readmit_time_from_disch_m", 
    event_col = "readmit_event",
    tau = 99,                # Evaluate up to 60 months
    B = 500,
    cpus = cpus,
    verbose = TRUE,
    min_strata_events = 1
)
results_death_shap_int_spl_ibs <- cph_evaluate_boot_oob_mi_corrected(
    formula = formula_death_int_spline,
    imputed_list = py_corrected_datasets_boot, # List of data.frames
    time_col = "death_time_from_disch_m", 
    event_col = "death_event",
    tau = 107,                # Evaluate up to 60 months
    B = 500,
    cpus = cpus,
    verbose = TRUE,
    min_strata_events = 1
)
# ====
aicc_readmit_shap_int_spl <- compute_mi_aicc(formula_readmit_int_spline, py_corrected_datasets_boot)
aicc_readmit_shap_int_spl
aicc_death_shap_int_spl <- compute_mi_aicc(formula_death_int_spline, py_corrected_datasets_boot)
aicc_death_shap_int_spl
```


```{r }
#| label: calib-shap
gc()
cal_readmit_aj_rcv <- calibrate_cscox_aj_grouped(
    formula_readmit = formula_readmit_int_spline,
    formula_death = formula_death_int_spline,
    data = py_corrected_datasets_boot,
    times = c(12, 36, 60),
    n_imputations = 5,
    validation = "cv",
    folds = 5,
    n_repeats = 20,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE,
    g = 10,
    min_bin_n = 20,
    seed = 2125
)
print(cal_readmit_aj_rcv)
plot(cal_readmit_aj_rcv)


cal_death_ici <- calibrate_cph_grouped(
    formula = formula_death_int_spline,
    data = py_corrected_datasets_boot,
    times = c(12, 36, 60),
    validation = "cv",
    folds = 5,
    n_repeats = 20,
    compute_ici = TRUE,
    ici_times = 60,
    make_plots = TRUE
)
print(cal_death_ici)
plot(cal_death_ici)

# Methods
# We fitted cause-specific Cox models for readmission and death, and combined them to estimate the predicted cumulative incidence of readmission at 12, 36, and 60 months. Calibration was assessed by comparing grouped predicted risks against observed Aalen–Johansen estimates of readmission, thereby accounting for the competing risk of death. To reduce optimism, calibration was evaluated using 5-fold cross-validation across imputations.
# 
# Results
# The cause-specific Cox models showed excellent calibration for readmission risk at all evaluated horizons. Cross-validated expected-to-observed ratios were 0.99 at 12 months, 0.99 at 36 months, and 0.98 at 60 months, indicating minimal miscalibration in the large. Grouped expected calibration error was also very low across horizons (12 months: 0.0029; 36 months: 0.0041; 60 months: 0.0040).
# One thing: in your message, the printed output corresponds to one fitted object, but the second call with formula_readmit_int_spline2 and formula_death_int_spline does not show its printed result yet. So I can interpret the first object with certainty; for the second one, I’d need the printed summary too.
```


# Session info

```{r}
#| label: session-info
#| echo: true
#| error: true
#| message: true
#| paged.print: true
message(paste0("R library: ", Sys.getenv("R_LIBS_USER")))
message(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))
message(paste0("Editor context: ", path))
cat("quarto version: "); quarto::quarto_version()
sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
)|> 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('R packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}")))
```
```{r}
#| label: session-info-python
#| echo: true
#| error: true
#| message: true
#| paged.print: true
#| class-output: center-table
reticulate::py_list_packages()|> 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Package' = 2, 'Version'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Python packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '70%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",
        "}"))) 
```

Save

```{r}
#| label: save
#| echo: true
#| error: true
#| warning: true
#| message: true
#| paged.print: true
#| results: "hold"
#| eval: true
wdpath<-
paste0(gsub("/cons","",gsub("cons","",paste0(getwd(),"/cons"))))
envpath<- if(regmatches(wdpath, regexpr("[A-Za-z]+", wdpath))=="G"){"G:/Mi unidad/Alvacast/SISTRAT 2023/"}else{"E:/Mi unidad/Alvacast/SISTRAT 2023/"}
paste0(getwd(),"/cons")
file.path(paste0(wdpath,"data/20241015_out"))
file.path(paste0(envpath,"data/20241015_out"))
# Save
rdata_path <- file.path(wdpath, "data/20241015_out", paste0("pred22_ndp_", format(Sys.time(), "%Y_%m_%d"), ".Rdata"))
save.image(rdata_path)
cat("Saved in:",
    rdata_path)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
  password <- Sys.getenv("PASSWORD_SECRET")
} else {
  if (interactive()) {
    utils::savehistory(tempfile())
    Sys.setenv(PASSWORD_SECRET = readLines(paste0(wdpath, "secret.txt"), warn = FALSE))
    utils::loadhistory()
  }
  Sys.setenv(PASSWORD_SECRET = readLines(paste0(wdpath, "secret.txt"), warn = FALSE))
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
save.image(paste0(rdata_path,".enc"))
# Encriptar el archivo en el mismo lugar
httr2::secret_encrypt_file(path = paste0(rdata_path,".enc"), key = "PASSWORD_SECRET")
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
cat("Copy renv lock into cons folder\n")
if (Sys.getenv("RSTUDIO_SESSION_TYPE") == "server" || file.exists("/.dockerenv")) {
  message("Running on RStudio Server or inside Docker. Folder copy skipped.")
} else {
    
  source_folder <- 
  destination_folder <- paste0(wdpath,"cons/renv")
  
  # Copy the folder recursively
    file.copy(paste0(wdpath,"renv.lock"), paste0(wdpath,"cons/renv.lock"), overwrite = TRUE)
  
  message("Renv lock copy performed.")
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
time_after_dedup2<-Sys.time()

```