---
title: "harmonization_c1_c2_6_top"
description: |
  Harmonization data from SENDAs patient (c1, c2-c6 and top).
date: "`r withr::with_locale(new = c('LC_TIME' = 'C'), code =format(Sys.time(),'%B %d, %Y'))`"
author: "Andrés González Santa Cruz & Amaru Agüero Jiménez"
format: 
  html:
    code-fold: true
    theme: 
      - journal #https://quarto.org/docs/output-formats/html-themes-more.html
      #- style.scss
editor: source
---

`r message(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))`

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {

$('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');

// onClick function for all plots (img's)

$('img:not(.zoomImg)').click(function() {
$('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
$('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
});

// onClick function for zoomImg

$('img.zoomImg').click(function() {
$('.zoomDiv').css({opacity: '0', width: '0%'});
});
});
```

```{css hideOutput-lib-src, echo = FALSE}
<script src="https://github.com/AGSCL/gine_brechas_pandemia/blob/main/hideOutput.js"></script>
```

```{js hideOutput, echo = FALSE}
$(document).ready(function() {

\$chunks = \$('.fold');

\$chunks.each(function () { // add button to source code chunks
if ( \$(this).hasClass('s') ) {
    \$('pre.r', this).prepend("\<div class=\\"showopt\\"\>Show Source\</div\>\<br style=\\"line-height:22px;\\"/\>");
    \$('pre.r', this).children('code').attr('class', 'folded');
    } // add button to output chunks

    if ( \$(this).hasClass('o') ) {
        \$('pre:not(.r)', this).has('code').prepend("\<div class=\\"showopt\\"\>Show Output\</div\>\<br style=\\"line-height:22px;\\"/\>");
        \$('pre:not(.r)', this).children('code:not(r)').addClass('folded'); // add button to plots
        \$(this).find('img').wrap('\<pre class=\\"plot\\"\>\</pre\>');
        \$('pre.plot', this).prepend("\<div class=\\"showopt\\"\>Show Plot\</div\>\<br style=\\"line-height:22px;\\"/\>");
        \$('pre.plot', this).children('img').addClass('folded');
        }
}); // hide all chunks when document is loaded

\$('.folded').css('display', 'none') // function to toggle the visibility
\$('.showopt').click(function() {
        var label = \$(this).html();
        if (label.indexOf("Show") \>= 0) {
            \$(this).html(label.replace("Show", "Hide"));
        } else {
        \$(this).html(label.replace("Hide", "Show"));
        }

\$(this).siblings('code, img').slideToggle('fast', 'swing');
});
});
```

<style type="text/css">

.showopt {

background-color: #004c93; color: #FFFFFF; width: 100px; height: 20px; text-align: center; vertical-align: middle !important; float: right; font-family: sans-serif; border-radius: 8px;

}

.showopt:hover {
background-color: #dfe4f2;
color: #004c93;

}

pre.plot {
background-color: white !important;
}

.tablelines table, .tablelines td, .tablelines th {
border: 1px solid black;
}

.centrado {
text-align: center;
}

.table.center {
margin-left:auto;
margin-right:auto;
}

/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/

/* Applies to outputs that are not code other than R*/

pre {
overflow-x: auto !important;
}

pre code {
word-wrap: normal !important;
white-space: pre !important;
}

/*
pre:not(.sourceCode) {
white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence */
font-size: 10px !important;
line-height: 50% !important;
}
body{ /* Normal */
text-align: justify;
}
.superbigimage{
overflow-y:scroll;
height:350px;
white-space: nowrap;
overflow-x: auto;
width:100%;
}
.superbigimage img{
overflow-y: scroll;
overflow-x: hidden;
}
.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}
.pandoc-table { /* Should add !important; but it seems no necessary */
margin-left:auto; /* To center */
margin-right:auto;
border-collapse: collapse;
table-layout: auto;
font-size: 11px;
overflow-y: auto;
max-height:450px !important;
white-space: nowrap;
overflow-x: auto;
width:450px;
}
.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}
.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}
.pandoc-table caption {
text-align: left !important;
font-size: 11px !important;
}

.center-table {
text-align: left !important;
font-size: 9px;
overflow-y:scroll;
height:450px;
overflow-x: scroll;
}

.controlly{
overflow-y:scroll;
height:350px;
overflow-x: scroll;

}
</style>

```{css, echo=F}
h1 {
    color: var(--heading-color);
    font-size: 2rem;
    margin-bottom: 1vh;
}

p {
  font-size: 1.1rem;
  line-height: 1.6rem;
}

a {
  color: var(--primary-color);
  text-decoration: none;
  border-bottom: 3px solid transparent;
  font-weight: bold;
  &:hover, &:focus {
      border-bottom: 3px solid currentColor;
  }
}

section {
  margin: 0 auto;
}

.post-meta {
  font-size: 1rem;
  font-style: italic;
  display: block;
  margin-bottom: 4vh;
  color: var(--secondary-color);
}

nav {
  display: flex;
  justify-content: flex-end;
  padding: 20px 0;
}

/*slider switch css */
.theme-switch-wrapper {
  display: flex;
  align-items: center;
  
  em {
    margin-left: 10px;
    font-size: 1rem;
  }
}
.theme-switch {
  display: inline-block;
  height: 34px;
  position: relative;
  width: 60px;
}

.theme-switch input {
  display:none;
}

.slider {
  background-color: #ccc;
  bottom: 0;
  cursor: pointer;
  left: 0;
  position: absolute;
  right: 0;
  top: 0;
  transition: .4s;
}

.slider:before {
  background-color: #fff;
  bottom: 4px;
  content: "";
  height: 26px;
  left: 4px;
  position: absolute;
  transition: .4s;
  width: 26px;
}

input:checked + .slider {
  background-color: #66bb6a;
}

input:checked + .slider:before {
  transform: translateX(26px);
}

.slider.round {
  border-radius: 34px;
}

.slider.round:before {
  border-radius: 50%;
}


```

<script>
function myFunction1() {
var x = document.getElementById("myDIV");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>

<script>
function myFunction2() {
var x = document.getElementById("myDIV2");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>

<br>

```{r pre-setup}
rm(list = ls()) 
if(!grepl("4.4.0",R.version.string)){stop("Versión diferente (debe ser 4.4.0)")}

pacman::p_unlock(lib.loc = .libPaths()) #para no tener problemas reinstalando paquetes
if(!require(devtools)){install.packages("devtools", type = "win.binary", dependencies=T)}

# 
# paste0("Sys time:", format(Sys.time(), "%Y_%m_%d_%H_00_00"))
# 
# 
# path_in <- paste0(gsub("cons","", getwd()), "data/25082023_data_output")
# full_path_in <- paste0(path_in,paste0("/1_ndp_", format(Sys.time(), "%Y_%m_%d"), ".Rdata"))
# 
# unlink("import_c2_c6_files", recursive=T)
# 
# load(full_path_in)
# paste0("Loaded: ",
#        full_path_in
#        )
```

```{r}
#To manipulate data
if(!require(tidyverse)){install.packages("tidyverse")}
#To manipulate big data
if(!require(polars)){install.packages("polars", repos = "https://rpolars.r-universe.dev")}
#code completion
try(polars_code_completion_activate())

if(!require(nanoparquet)){install.packages("nanoparquet")}

# devtools::install_github("bokeh/rbokeh")
# 
if (!require("pacman")) install.packages("pacman")

# pacman::p_load(
#   altair, arrow, biostat3, car, caret, chilemapas, choroplethr, choroplethrAdmin1,
#   choroplethrMaps, codebook, compareGroups, DiagrammeR, DiagrammeRsvg, DT, epiR, epitools,
#   factoextra, FactoMineR, finalfit, flexsurv, fmsb, ggfortify, ggiraph, ggiraphExtra,
#   ggpubr, glca, gridExtra, here, Hmisc, installr, janitor, kableExtra, lsmeans, matrixStats,
#   Metrics, muhaz, naniar, neuralnet, NeuralNetTools, panelr, patchwork, pdp, plotly, plyr,
#   polycor, pROC, psych, radiant, rateratio.test, reshape, reshape2,
#   reticulate, rio, ROCit, rsvg, sf, sjPlot, sqldf, Statamarkdown, survminer, survMisc,
#   tableone, tidylog, treemapify, VIM, webshot, zoo, install=T
# )
```

# Bring DBs


```{r}

sistrat <- "E:/Mi unidad/Alvacast/SISTRAT 2022 (github)/"

#fuentes: 
#https://rpubs.com/georgy_makarov/897844
load(paste0(sistrat,"1_ndp_2023_07_05_19_00_00.RData"))

#load c1 and top data
load(paste0(gsub("cons","",getwd()), "data/25082023_data_output",paste0("/1_ndp_", format(Sys.time(), "%Y_%m_%d"),".Rdata")))
#load c2-c6 data
load(paste0(gsub("cons","",getwd()), "data/25082023_data_output",paste0("/2_ndp_", format(Sys.time(), "%Y_%m_%d"),".Rdata")))
```

```{r}

```
